---
title: "00_Upload_Env_Data"
output: html_document
date: "2024-02-28"
---
load packages
```{r}
library(openxlsx)
library(dplyr)
library(lubridate)
library(readxl)
```



Download data from NANOOS website:
I first went to https://nvs.nanoos.org/CruiseSalish and downloaded all relevant files. I moved them to the NANOOS-files folder. I unzipped them manually. 
```{bash}
#find current directory
pwd
```

move all upcast files to the upcast folder
```{bash}
#move directories
cd /Users/hailaschultz/Dropbox/Schultz_Dissertation/Data_Analysis/Schultz_dissertation-2/data/NANOOS_files

find . -name '*labupcast.xlsx' -exec mv {} ../NANOOS_files/upcasts/ \;
```
The 2014 and 2015 upcasts were not yet in NANOOS, so I had to download them from NCEI: https://www.ncei.noaa.gov/access/ocean-carbon-acidification-data-system/oceans/SalishCruise_DataPackage.html After downloading, I moved them directly to the upcasts folder and converted from csv to excel

import and merge excel files

```{r}
excel_dir<-"/Users/hailaschultz/Dropbox/Schultz_Dissertation/Data_Analysis/Schultz_dissertation-2/data/NANOOS_files/upcasts"

excel_files <- list.files(path = excel_dir, pattern = "\\.xlsx$", full.names = TRUE)

# Initialize an empty list to store the data frames
dfs <- list()

# Loop through each Excel file and read it into a data frame
for (file in excel_files) {
  # Read the Excel file into a data frame
  df <- read_excel(file)
  
  # Store the data frame in the list
  dfs[[length(dfs) + 1]] <- df
}
```

example of how to access dataframes
```{r}
dfs[[1]]
```

See which columns all of the dataframes have in common
```{r}
# Get column names of the first data frame
common_columns <- names(dfs[[1]])

# Loop through the remaining data frames and find common column names
for (i in 2:length(dfs)) {
  # Get column names of the current data frame
  current_columns <- names(dfs[[i]])
  
  # Find common column names with previous data frames
  common_columns <- intersect(common_columns, current_columns)
}

# 'common_columns' now contains the column names that are common across all data frames
print(common_columns)
```
convert date local format
```{r}
# Loop through each dataframe in the list
for (i in seq_along(dfs)) {
  # Check if the dataframe has a column labeled "DATE_LOCAL"
  if ("DATE_LOCAL" %in% names(dfs[[i]])) {
    # Convert the "DATE_LOCAL" column to character
    dfs[[i]]$DATE_LOCAL <- as.character(dfs[[i]]$DATE_LOCAL)
  }
}

for (i in seq_along(dfs)) {
  # Check if the dataframe has a column labeled "DATE_LOCAL"
  if ("DATE_LOCAL" %in% names(dfs[[i]])) {
    # Convert the "DATE_LOCAL" column to character
    dfs[[i]]$DATE_UTC <- as.character(dfs[[i]]$DATE_UTC)
  }
}

for (i in seq_along(dfs)) {
  # Check if the dataframe has a column labeled "DATE_LOCAL"
  if ("DATE_LOCAL" %in% names(dfs[[i]])) {
    # Convert the "DATE_LOCAL" column to character
    dfs[[i]]$TIME_UTC <- as.character(dfs[[i]]$TIME_UTC)
  }
}

for (i in seq_along(dfs)) {
  # Check if the dataframe has a column labeled "DATE_LOCAL"
  if ("DATE_LOCAL" %in% names(dfs[[i]])) {
    # Convert the "DATE_LOCAL" column to character
    dfs[[i]]$TIME_LOCAL <- as.character(dfs[[i]]$TIME_LOCAL)
  }
}
```

subset to common columns
```{r}
# Loop through each data frame in the list
for (i in seq_along(dfs)) {
  # Subset the data frame to only the common columns
  dfs[[i]] <- dfs[[i]][, common_columns, drop = FALSE]
}
```


combine datasheets vertically
```{r}
combined_df <- do.call(rbind, dfs)
```


convert date from character to date
```{r}
combined_df$Date <- ymd(combined_df$DATE_LOCAL)
```

extract month and year
```{r}
combined_df$Month <- month(combined_df$Date)
combined_df$Year <- year(combined_df$Date)
unique(combined_df$Month)
```

recode months
```{r}
combined_df$Month <- recode_factor(combined_df$Month, 
                              '4' = "APR",'3'="APR", '7' = "JUL", 
                              '6' = "JUL", '9' = "SEP", '10' = "SEP")

```

Subset stations
```{r}
unique(combined_df$STATION_NO)
# Define the seven named stations
named_stations <- c("4", "8", "12", "28", "38", "402", "22")

# Subset the dataframe based on the named stations
subset_df <- combined_df[combined_df$STATION_NO %in% named_stations, ]

#add P to station numbers
subset_df$STATION_NO <- paste0("P", subset_df$STATION_NO)
```
Depth categories
