---
title: "00_Upload_Env_Data"
output: html_document
date: "2024-02-28"
---
load packages
```{r}
library(openxlsx)
library(dplyr)
library(lubridate)
library(readxl)
```



Download data from NANOOS website:
I first went to https://nvs.nanoos.org/CruiseSalish and downloaded all relevant files. I moved them to the NANOOS-files folder. I unzipped them manually. 
```{bash}
#find current directory
pwd
```

# Upcast Bottle Measurements
move all upcast files to the upcast folder
```{bash}
#move directories
cd /Users/hailaschultz/Dropbox/Schultz_Dissertation/Data_Analysis/Schultz_dissertation-2/data/NANOOS_files

find . -name '*labupcast.xlsx' -exec mv {} ../NANOOS_files/upcasts/ \;
```
The 2014 and 2015 upcasts were not yet in NANOOS, so I had to download them from NCEI: https://www.ncei.noaa.gov/access/ocean-carbon-acidification-data-system/oceans/SalishCruise_DataPackage.html After downloading, I moved them directly to the upcasts folder and converted from csv to excel

import and merge excel files

```{r}
excel_dir<-"/Users/hailaschultz/Dropbox/Schultz_Dissertation/Data_Analysis/Schultz_dissertation-2/data/NANOOS_files/upcasts"

excel_files <- list.files(path = excel_dir, pattern = "\\.xlsx$", full.names = TRUE)

# Initialize an empty list to store the data frames
dfs <- list()

# Loop through each Excel file and read it into a data frame
for (file in excel_files) {
  # Read the Excel file into a data frame
  df <- read_excel(file)
  
  # Store the data frame in the list
  dfs[[length(dfs) + 1]] <- df
}
```

example of how to access dataframes
```{r}
dfs[[1]]
```

See which columns all of the dataframes have in common
```{r}
# Get column names of the first data frame
common_columns <- names(dfs[[1]])

# Loop through the remaining data frames and find common column names
for (i in 2:length(dfs)) {
  # Get column names of the current data frame
  current_columns <- names(dfs[[i]])
  
  # Find common column names with previous data frames
  common_columns <- intersect(common_columns, current_columns)
}

# 'common_columns' now contains the column names that are common across all data frames
print(common_columns)
```
convert date local format
```{r}
# Loop through each dataframe in the list
for (i in seq_along(dfs)) {
  # Check if the dataframe has a column labeled "DATE_LOCAL"
  if ("DATE_LOCAL" %in% names(dfs[[i]])) {
    # Convert the "DATE_LOCAL" column to character
    dfs[[i]]$DATE_LOCAL <- as.character(dfs[[i]]$DATE_LOCAL)
  }
}

for (i in seq_along(dfs)) {
  # Check if the dataframe has a column labeled "DATE_LOCAL"
  if ("DATE_LOCAL" %in% names(dfs[[i]])) {
    # Convert the "DATE_LOCAL" column to character
    dfs[[i]]$DATE_UTC <- as.character(dfs[[i]]$DATE_UTC)
  }
}

for (i in seq_along(dfs)) {
  # Check if the dataframe has a column labeled "DATE_LOCAL"
  if ("DATE_LOCAL" %in% names(dfs[[i]])) {
    # Convert the "DATE_LOCAL" column to character
    dfs[[i]]$TIME_UTC <- as.character(dfs[[i]]$TIME_UTC)
  }
}

for (i in seq_along(dfs)) {
  # Check if the dataframe has a column labeled "DATE_LOCAL"
  if ("DATE_LOCAL" %in% names(dfs[[i]])) {
    # Convert the "DATE_LOCAL" column to character
    dfs[[i]]$TIME_LOCAL <- as.character(dfs[[i]]$TIME_LOCAL)
  }
}
```

subset to common columns
```{r}
# Loop through each data frame in the list
for (i in seq_along(dfs)) {
  # Subset the data frame to only the common columns
  dfs[[i]] <- dfs[[i]][, common_columns, drop = FALSE]
}
```


combine datasheets vertically
```{r}
combined_df <- do.call(rbind, dfs)
```


convert date from character to date
```{r}
combined_df$Date <- ymd(combined_df$DATE_LOCAL)
```

extract month and year
```{r}
combined_df$Month <- month(combined_df$Date)
combined_df$Year <- year(combined_df$Date)
unique(combined_df$Month)
```

recode months
```{r}
combined_df$Month <- recode_factor(combined_df$Month, 
                              '4' = "APR",'3'="APR", '7' = "JUL", 
                              '6' = "JUL", '9' = "SEP", '10' = "SEP")

```

Subset stations
```{r}
unique(combined_df$STATION_NO)
# Define the seven named stations
named_stations <- c("4", "8", "12", "28", "38", "402", "22")

# Subset the dataframe based on the named stations
subset_df <- combined_df[combined_df$STATION_NO %in% named_stations, ]

#add P to station numbers
subset_df$STATION_NO <- paste0("P", subset_df$STATION_NO)
```
Depth categories

# Downcast CTD data

import and merge excel files(2016-2022)
```{r}
downcast_dir<-"/Users/hailaschultz/Dropbox/Schultz_Dissertation/Data_Analysis/Schultz_dissertation-2/data/NANOOS_files/woac_downcasts"

downcast_excel_files <- list.files(path = downcast_dir, pattern = "\\.xlsx$", full.names = TRUE)

# Initialize an empty list to store the data frames
dfs <- list()

# Loop through each Excel file and read it into a data frame
for (file in downcast_excel_files) {
  # Read the Excel file into a data frame
  df <- read_excel(file)
  
  # Store the data frame in the list
  dfs[[length(dfs) + 1]] <- df
}
```
See which columns all of the dataframes have in common
```{r}
# Get column names of the first data frame
common_columns <- names(dfs[[1]])

# Loop through the remaining data frames and find common column names
for (i in 2:length(dfs)) {
  # Get column names of the current data frame
  current_columns <- names(dfs[[i]])
  
  # Find common column names with previous data frames
  common_columns <- intersect(common_columns, current_columns)
}

# 'common_columns' now contains the column names that are common across all data frames
print(common_columns)
```

subset to common columns
```{r}
# Loop through each data frame in the list
for (i in seq_along(dfs)) {
  # Subset the data frame to only the common columns
  dfs[[i]] <- dfs[[i]][, common_columns, drop = FALSE]
}
```


combine datasheets vertically
```{r}
combined_df <- do.call(rbind, dfs)
```

remove units rows
```{r}
combined_df<-subset(combined_df,NMEAtimeUTC!="[]")
```

convert all dates to the correct format
```{r}
library(dplyr)

# Define a helper function to process each entry
convert_NMEAtime <- function(x) {
  if (grepl("^[0-9]+\\.[0-9]+$", x)) {
    # Convert Excel numeric date to POSIXct
    as.POSIXct(as.numeric(x) * 86400, origin = "1899-12-30", tz = "UTC") # Excel epoch starts on 1899-12-30
  } else {
    # Parse human-readable datetime
    as.POSIXct(x, format = "%b %d %Y %H:%M:%S", tz = "UTC")
  }
}


# Apply the conversion function to standardize the column
combined_df <- combined_df %>%
  mutate(
    # Convert all entries in NMEAtimeUTC to POSIXct format
    NMEAtimeUTC = sapply(NMEAtimeUTC, convert_NMEAtime)
  ) %>%
  # Convert POSIXct to desired character format
  mutate(
    NMEAtimeUTC = format(as.POSIXct(NMEAtimeUTC), "%b %d %Y %H:%M:%S")
  )
```

extract month
```{r}
combined_df$Month <- substr(combined_df$NMEAtimeUTC, 1, 3)
unique(combined_df$Month)
combined_df$Month <- recode_factor(combined_df$Month, 
                              'Apr' = "APR", 'Jul' = "JUL", 'Jun'="JUL",
                              '6' = "JUL", 'Sep' = "SEP")
```

extract year
```{r}
combined_df$Year <- substr(combined_df$NMEAtimeUTC, 8, 11)
unique(combined_df$Year)
```

Subset stations
```{r}
unique(combined_df$Station)
# Define the seven named stations
named_stations <- c("P4", "P8", "P12", "P28", "P38", "P402", "P22")

# Subset the dataframe based on the named stations
subset_df <- combined_df[combined_df$Station %in% named_stations, ]
unique(subset_df$Station)
```


imoort and merge csv files (2014-2015)
```{r}
# Directory containing the CSV files
downcast_dir <- "/Users/hailaschultz/Dropbox/Schultz_Dissertation/Data_Analysis/Schultz_dissertation-2/data/NANOOS_files/CSV_files"

# List all CSV files in the directory
downcast_csv_files <- list.files(path = downcast_dir, pattern = "\\.csv$", full.names = TRUE)

# Initialize an empty list to store the data frames
dfs <- list()

# Loop through each CSV file and read it into a data frame
for (file in downcast_csv_files) {
  # Read the CSV file into a data frame
  df <- read.csv(file, stringsAsFactors = FALSE)
  
  # Store the data frame in the list
  dfs[[length(dfs) + 1]] <- df
}
```

See which columns all of the dataframes have in common
```{r}
# Get column names of the first data frame
common_columns <- names(dfs[[1]])

# Loop through the remaining data frames and find common column names
for (i in 2:length(dfs)) {
  # Get column names of the current data frame
  current_columns <- names(dfs[[i]])
  
  # Find common column names with previous data frames
  common_columns <- intersect(common_columns, current_columns)
}

# 'common_columns' now contains the column names that are common across all data frames
print(common_columns)
```

subset to common columns
```{r}
# Loop through each data frame in the list
for (i in seq_along(dfs)) {
  # Subset the data frame to only the common columns
  dfs[[i]] <- dfs[[i]][, common_columns, drop = FALSE]
}
```


combine datasheets vertically
```{r}
csv_combined_df <- do.call(rbind, dfs)
```

remove units rows
```{r}
csv_combined_df<-subset(csv_combined_df,Pressure!="CTD")
csv_combined_df<-subset(csv_combined_df,Pressure!="[db]")
```

extract month
```{r}
csv_combined_df$Month <- substr(csv_combined_df$UTC.Time, 1, 3)
unique(csv_combined_df$Month)
csv_combined_df$Month <- recode_factor(csv_combined_df$Month, 
                              'Apr' = "APR", 'Jul' = "JUL", 'Jun'="JUL",
                              '6' = "JUL", 'Sep' = "SEP",'Oct'="OCT")
```

extract year
```{r}
csv_combined_df$Year <- substr(csv_combined_df$UTC.Time, 8, 11)
unique(csv_combined_df$Year)
```

Subset stations
```{r}
unique(csv_combined_df$Station)
# Define the seven named stations
named_stations <- c("P4", "P8", "P12", "P28", "P38", "P402", "P22")

# Subset the dataframe based on the named stations
csv_subset_df <- csv_combined_df[csv_combined_df$Station %in% named_stations, ]
unique(csv_subset_df$Station)
```


## merge two datasets
```{r}
colnames(subset_df)
colnames(csv_subset_df)
```
rename columns from first dataset
```{r}
subset_df<-subset_df %>% rename(Cruise.ID = CruiseID, 
                                Depth = "depSM: Depth",
                                Potential.Temperature="potemp090C: Potential Temperature",
                                Salinity="sal00: Salinity  Practical",
                                Oxygen.Concentration.MG="sbeox0Mg/L: Oxygen  SBE 43",
                                Chlorophyll.Fluorescence="flECO-AFL: Fluorescence  WET Labs ECO-AFL/FL")
```

```{r}
# Find common columns
common_columns <- intersect(names(subset_df), names(csv_subset_df))

# Subset each data frame to only the common columns
subset_df_common <- subset_df[, common_columns, drop = FALSE]
csv_subset_df_common <- csv_subset_df[, common_columns, drop = FALSE]

# Combine the data frames vertically
downcasts_combined <- rbind(subset_df_common, csv_subset_df_common)
```

