---
title: "01_data-setup"
date: "2023-09-20"
output:
  html_document:
    toc: true
    toc_depth: 2
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,         # Display code chunks
  eval = TRUE,        # Evaluate code chunks
  warning = FALSE,     # Hide warnings
  message = FALSE,     # Hide messages
  fig.width = 10,       # Set plot width in inches
  fig.height = 6,      # Set plot height in inches
  fig.align = "center" # Align plots to the center
)
```

This code converts datasets from multiple sources into one workable file.


# Import Data

## WOAC water chemistry data
This dataset was acquired from BethElLee Herrmann and is a compilation of mesurements from bottle samples and from the CTD collected at each depth where bottle water was collected. This sheet is the second tab of the raw data spreadsheet. This dataset incluedes data on:
- pH
- pCO2
- CTD pressure
- CTD temperature
- CTD salinity
- CTD oxygen
- oxygen
- DIC
- total alkalinity
- phosphate
- silicate
- aragonite saturation
- note that this dataset does not have nitrogen-based nutrients.

Read data into R
```{r}
Environmental <- read.csv("/Users/hailaschultz/Dropbox/Schultz_Dissertation/Data_Analysis/Schultz_dissertation-2/data/WOAC_Chem_data_2014-2022_all-niskins.csv")
```

## Zooplankton Data
This dataset is from our WOAC zooplankton database and includes morphological abundance and biomass data for each species identified in each sample collected. Vertical and oblique tows are included and represent samples that are integrated through the water column.

Read data into R
```{r}
Zooplankton <- read.csv("/Users/hailaschultz/Dropbox/Schultz_Dissertation/Data_Analysis/Schultz_dissertation-2/data/WOAC_Species_Densities_2014-2022.csv")
```