---
title: "01_Env-Explore"
date: "2023-09-20"
output:
  html_document:
    toc: true
    toc_depth: 2
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,         # Display code chunks
  eval = TRUE,        # Evaluate code chunks
  warning = FALSE,     # Hide warnings
  message = FALSE,     # Hide messages
  fig.width = 10,       # Set plot width in inches
  fig.height = 6,      # Set plot height in inches
  fig.align = "center" # Align plots to the center
)
```

This code is mostly exploratory analysis of WOAC environmental data at bio stations between 2014-2022.

The report is here: https://rpubs.com/HailaSchultz/Chap2Env-Explore

# Load Packages
```{r}
library(dplyr)
library(ggplot2)
library(tidyr)
library(forcats)
library(reshape2)
library(xts)
library(cowplot)
```


# Import Data

WOAC water chemistry data
This dataset was acquired from BethElLee Herrmann and is a compilation of mesurements from bottle samples and from the CTD collected at each depth where bottle water was collected. This sheet is the second tab of the raw data spreadsheet. This dataset incluedes data on:
- pH
- pCO2
- CTD pressure
- CTD oxygenerature
- CTD salinity
- CTD oxygen
- oxygen
- DIC
- total alkalinity
- phosphate
- silicate
- aragonite saturation
- note that this dataset does not have nitrogen-based nutrients.

Read data into R
```{r}
Environmental <- read.csv("/Users/hailaschultz/Dropbox/Schultz_Dissertation/Data_Analysis/Schultz_dissertation-2/data/WOAC_Chem_data_2014-2022_all-niskins.csv")
```

# temperature Plot
summarize by depth category
```{r}
temp<-Environmental %>%
  group_by(Station,Date,Depth) %>%
  summarise(
    temp = mean(CTDTMP_DEG_C_ITS90))
```

convert date format
```{r}
temp <- temp %>% 
  mutate(date_convert = as.Date(Date, format = "%m/%d/%y"))
```


make plot
```{r,fig.width=10,fig.height=16}
temp_plot<-ggplot(data = temp, aes(date_convert, temp,color=Depth)) +
  geom_line()+ 
  geom_point()+
  facet_grid(Station ~ .) +
  theme_classic()+
  xlab("Year") + ylab("temperature")+
  scale_x_date(date_breaks = "1 year", date_labels = "%Y")
temp_plot
```
save plot
```{r, eval=FALSE}
setwd("/Users/hailaschultz/Dropbox/Schultz_Dissertation/Data_Analysis/Schultz_dissertation-2/output")
ggsave(filename = "time-series_temperature.png", plot = temp_plot, width = 10, height = 17, device='png', dpi=700)
```
It looks like P12 and P402 have the highest variability in temperatures, and have especially high summer temperatures. This makes sense becasue these are the hood canal stations.

## Try anomaly analysis
```{r}
library(anomalize)
library(tibble)
```

subset to station, depth
```{r}
temp_5m<- subset(temp, Depth=='Deep')
temp_5m_P402<- subset(temp_5m, Station=='P38')
temp_5m_P402<- subset(temp_5m_P402, date_convert!='2022-07-01')
temp_5m_P402<- subset(temp_5m_P402, date_convert!='2022-04-29')
temp_5m_P402 <- temp_5m_P402[ -c(1:3) ]
temp_5m_P402<-temp_5m_P402[order(temp_5m_P402$date_convert),]
```

```{r}
P402_ts <- temp_5m_P402 %>% as.tibble() 
```

decomposition
```{r,fig.width=5, fig.height=10}
P402_ts %>% 
  time_decompose(temp, method = "stl", frequency = "auto", trend = "auto") %>%
  anomalize(remainder, method = "gesd",  max_anoms = 0.2) %>%
  plot_anomaly_decomposition()
```

## Average across stations
```{r}
#select only surface samples
temp_surf<- subset(temp, Depth=='Surface')
#extract year
temp_surf$year <- format(as.Date(temp_surf$Date, format="%m/%d/%y"),"%Y")
#extract month
temp_surf$month <- format(as.Date(temp_surf$Date, format="%m/%d/%y"),"%m")
#make season column
temp_surf$season<-temp_surf$month
temp_surf$season <- recode(temp_surf$season, 
                                 "04" = "spring",
                                 "05" = "spring",
                                 "06" = "summer",
                                 "07" = "summer",
                           "08" = "summer",
                                "09" = "autumn",
                                 "10"= "autumn",
                                 "11"= "autumn")
#average by year and month
temp_surf<-temp_surf %>%
  group_by(year,season) %>%
  summarise(
    mean = mean(temp),
    sd=sd(temp))
```

plot
```{r}
temp_surf$year<-as.numeric(temp_surf$year)
temp_ave<-ggplot(data = temp_surf, aes(year,mean)) +
  geom_point()+
  theme_classic()+
  xlab("Year") + ylab("temperature")+ 
  geom_line()+
  geom_pointrange(aes(ymin=mean-sd, ymax=mean+sd))+
  facet_grid(season ~ .)
temp_ave
```


# pH Plot
remove missing rows and remove weird values
```{r}
pH<-Environmental %>% drop_na(pH)
pH <- subset(pH, pH != 2130.2)
```

summarize by depth category
```{r}
pH<-pH %>%
  group_by(Station,Date,Depth) %>%
  summarise(
    pH = mean(pH))
```

convert date format
```{r}
pH <- pH %>% 
  mutate(date_convert = as.Date(Date, format = "%m/%d/%y"))
```


make plot
```{r,fig.width=10,fig.height=16}
pH_plot<-ggplot(data = pH, aes(date_convert, pH,color=Depth)) +
  geom_line()+ 
  geom_point()+
  facet_grid(Station ~ .) +
  theme_classic()+
  xlab("Year") + ylab("pH")+
  scale_x_date(date_breaks = "1 year", date_labels = "%Y")
pH_plot
```

save plot
```{r, eval=FALSE}
setwd("/Users/hailaschultz/Dropbox/Schultz_Dissertation/Data_Analysis/Schultz_dissertation-2/output")
ggsave(filename = "time-series_pH.png", plot = pH_plot, width = 10, height = 17, device='png', dpi=700)
```

## Try anomaly analysis
```{r}
library(anomalize)
library(tibble)
```

subset to station, depth
```{r}
pH_5m<- subset(pH, Depth=='5-m')
pH_5m_P402<- subset(pH_5m, Station=='P402')
pH_5m_P402 <- pH_5m_P402[ -c(1:3) ]
pH_5m_P402<-pH_5m_P402[order(pH_5m_P402$date_convert),]
```

```{r}
P402_ts <- pH_5m_P402 %>% as.tibble() 
```

decomposition
```{r,fig.width=5, fig.height=10}
P402_ts %>% 
  time_decompose(pH, method = "stl", frequency = "auto", trend = "auto") %>%
  anomalize(remainder, method = "gesd", alpha = 0.05, max_anoms = 0.2) %>%
  plot_anomaly_decomposition()
```
```{r}
P402_ts %>% 
  time_decompose(pH) %>%
  anomalize(remainder) %>%
  time_recompose() %>%
  plot_anomalies(time_recomposed = TRUE, ncol = 3, alpha_dots = 0.5)
```




## Average across stations
```{r}
#select only surface samples
pH_surf<- subset(pH, Depth=='Surface')
#extract year
pH_surf$year <- format(as.Date(pH_surf$Date, format="%m/%d/%y"),"%Y")
#extract month
pH_surf$month <- format(as.Date(pH_surf$Date, format="%m/%d/%y"),"%m")
#make season column
pH_surf$season<-pH_surf$month
pH_surf$season <- recode(pH_surf$season, 
                                 "04" = "spring",
                                 "05" = "spring",
                                 "06" = "summer",
                                 "07" = "summer",
                           "08" = "summer",
                                "09" = "autumn",
                                 "10"= "autumn",
                                 "11"= "autumn")
#average by year and month
pH_surf<-pH_surf %>%
  group_by(year,season) %>%
  summarise(
    mean = mean(pH),
    sd=sd(pH))
```

plot
```{r}
pH_surf$year<-as.numeric(pH_surf$year)
pH_ave<-ggplot(data = pH_surf, aes(year,mean)) +
  geom_point()+
  theme_classic()+
  xlab("Year") + ylab("pH")+ 
  geom_line()+
  geom_pointrange(aes(ymin=mean-sd, ymax=mean+sd))+
  facet_grid(season ~ .)
pH_ave
```

# O2 Plot
remove missing rows and remove weird values
```{r}
Oxygen<-Environmental %>% drop_na(O2.in.mg.l)
```

summarize by depth category
```{r}
Oxygen<-Oxygen %>%
  group_by(Station,Date,Depth) %>%
  summarise(
    Oxygen = mean(O2.in.mg.l))
```

convert date format
```{r}
Oxygen <- Oxygen %>% 
  mutate(date_convert = as.Date(Date, format = "%m/%d/%y"))
```


make plot
```{r,fig.width=10,fig.height=16}
Oxygen_plot<-ggplot(data = Oxygen, aes(date_convert, Oxygen,color=Depth)) +
  geom_line()+ 
  geom_point()+
  facet_grid(Station ~ .) +
  theme_classic()+
  xlab("Year") + ylab("Oxygen")+
  scale_x_date(date_breaks = "1 year", date_labels = "%Y")
Oxygen_plot
```

save plot
```{r, eval=FALSE}
setwd("/Users/hailaschultz/Dropbox/Schultz_Dissertation/Data_Analysis/Schultz_dissertation-2/output")
ggsave(filename = "time-series_Oxygen.png", plot = Oxygen_plot, width = 10, height = 17, device='png', dpi=700)
```
## Average across stations
```{r}
#select only surface samples
oxygen_surf<- subset(Oxygen, Depth=='Surface')
#extract year
oxygen_surf$year <- format(as.Date(oxygen_surf$Date, format="%m/%d/%y"),"%Y")
#extract month
oxygen_surf$month <- format(as.Date(oxygen_surf$Date, format="%m/%d/%y"),"%m")
#make season column
oxygen_surf$season<-oxygen_surf$month
oxygen_surf$season <- recode(oxygen_surf$season, 
                                 "04" = "spring",
                                 "05" = "spring",
                                 "06" = "summer",
                                 "07" = "summer",
                           "08" = "summer",
                                "09" = "autumn",
                                 "10"= "autumn",
                                 "11"= "autumn")
#average by year and month
oxygen_surf<-oxygen_surf %>%
  group_by(year,season) %>%
  summarise(
    mean = mean(Oxygen),
    sd=sd(Oxygen))
```

plot
```{r}
oxygen_surf$year<-as.numeric(oxygen_surf$year)
oxygen_ave<-ggplot(data = oxygen_surf, aes(year,mean)) +
  geom_point()+
  theme_classic()+
  xlab("Year") + ylab("oxygen")+ 
  geom_line()+
  geom_pointrange(aes(ymin=mean-sd, ymax=mean+sd))+
  facet_grid(season ~ .)
oxygen_ave
```

#Time Series Analysis
I'm struggling with this because of the missing NA values, so I am just going to average over the years
## detrend temperature
```{r}
#select only surface samples
temp_timeseries<- subset(temp, Depth=='Surface')
#extract month
temp_timeseries$month <- format(as.Date(temp_timeseries$Date, format="%m/%d/%y"),"%m")
```
add month column
```{r}
temp_timeseries$month<-fct_recode(temp_timeseries$month, "Oct"="10", "Apr"="04","Jul"="07","Sep"="09","May"="05","Jun"="06","Oct"="11")
```

```{r}
#extract year
temp_timeseries$year <- format(as.Date(temp_timeseries$Date, format="%m/%d/%y"),"%Y")
```

subset to one station
```{r}
temp_timeseries_p12<-subset(temp_timeseries,Station=="P12")
#remove extra row
temp_timeseries_p12<-temp_timeseries_p12[-c(2), ]
```

convert to time series object
```{r}
#remove excess columns
temp_timeseries_p12 <- temp_timeseries_p12[ -c(1:3,5) ]
p12<-dcast(temp_timeseries_p12, year ~ month, value.var = "temp")
#add other months
p12$Jan<-NA
p12$Feb<-NA
p12$Mar<-NA
p12$May<-NA
p12$Aug<-NA
p12$Nov<-NA
p12$Dec<-NA
#reorder columns
p12 <- p12[, c(1,7,8,9,2,10,3,4,11,5,6,12,13)]
#convert back to long format
p12<-melt(p12, id.vars=c("year"))
#sort chronologically
p12<-p12[
  order( p12[,1], p12[,2] ),
]
#remove year and month columns
p12 <- p12[ -c(1:2) ]
str(p12)
p12_ts<-ts(p12, start=c(2014,1),freq=12)
```

Because there are missing values, they must be imputed I followed the tutorial here: https://jtr13.github.io/EDAVold/missingTS.html

```{r}
p12_ts

```

subset to one station
```{r}
temp_timeseries_p12<-subset(temp_timeseries,Station=="P12")
#remove extra row
temp_timeseries_p12<-temp_timeseries_p12[-c(2), ]
```

convert to time series object
```{r}
#remove excess columns
temp_timeseries_p12 <- temp_timeseries_p12[ -c(1:3,6:7) ]
xt1 <- xts(temp_timeseries_p12$temp, order.by = as.Date(temp_timeseries_p12$date_convert))
plot(xt1)
```

# Individual station plots

## P402

remove missing rows and remove weird values
```{r}
Environmental_sum<-Environmental %>% drop_na(pH)
Environmental_sum <- subset(Environmental_sum, pH != 2130.2)
```

```{r}
Environmental_sum<-Environmental_sum %>% drop_na(O2.in.mg.l)
```

summarize by depth category
```{r}
Environmental_sum<-Environmental_sum %>%
  group_by(Station,Date,Depth) %>%
  summarise(
    temp = mean(CTDTMP_DEG_C_ITS90),
    pH = mean(pH),
    Oxygen = mean(O2.in.mg.l))
```

convert date format
```{r}
Environmental_sum <- Environmental_sum %>% 
  mutate(date_convert = as.Date(Date, format = "%m/%d/%y"))
```

subset to station and depth
```{r}
Environmental_sum_P402<-subset(Environmental_sum,Station=="P402")
Environmental_sum_P402<-subset(Environmental_sum_P402,Depth!="Surface")
Environmental_sum_P402<-subset(Environmental_sum_P402,Depth!="Mid")
```

convert to long format
```{r}
Environmental_sum_P402<-melt(Environmental_sum_P402, na.rm = FALSE,  id = c("Station","Date","Depth","date_convert"))
```

make plot
```{r,fig.width=5,fig.height=2.5}
#change factor order
Environmental_sum_P402$Depth <- factor(Environmental_sum_P402$Depth, levels = c("5-m","Deep"))

P402_temp<-subset(Environmental_sum_P402,variable=="temp")

P402_temp<-ggplot(data = P402_temp, aes(date_convert, value,color=Depth)) +
  geom_line()+ 
  geom_point() +
  scale_y_continuous(breaks = c(5,10,15,20), limits = c(5,20), expand = c(0,0))+
  theme_classic()+
  xlab("Year") +
  ylab("Temperature (C)")+
  scale_x_date(date_breaks = "1 year", date_labels = "%Y")+
    theme(axis.text=element_text(size=11))+
  theme(axis.title.x = element_blank())
P402_temp

P402_pH<-subset(Environmental_sum_P402,variable=="pH")

P402_pH<-ggplot(data = P402_pH, aes(date_convert, value,color=Depth)) +
  geom_line()+ 
  geom_point() +
  scale_y_continuous(breaks = c(7,7.5,8,8.5), limits = c(7,8.5), expand = c(0,0))+
  theme_classic()+
  xlab("Year") +
  ylab("pH")+
  scale_x_date(date_breaks = "1 year", date_labels = "%Y")+
    theme(axis.text=element_text(size=11))+
  theme(axis.title.x = element_blank())
P402_pH

P402_Oxy<-subset(Environmental_sum_P402,variable=="Oxygen")

P402_Oxy<-ggplot(data = P402_Oxy, aes(date_convert, value,color=Depth)) +
  geom_line()+ 
  geom_point() +
  scale_y_continuous(breaks = c(0,5,10,15), limits = c(0,15), expand = c(0,0))+
  theme_classic()+
  xlab("Year") +
  ylab("Oxygen mg/l")+
  scale_x_date(date_breaks = "1 year", date_labels = "%Y")+
    theme(axis.text=element_text(size=11))+
  theme(axis.title.x = element_blank())
P402_Oxy
```

save plots
```{r, eval=FALSE}
setwd("/Users/hailaschultz/Dropbox/Schultz_Dissertation/Data_Analysis/Schultz_dissertation-2/output")
ggsave(filename = "P402_temp.png", plot = P402_temp, height = 2.5, width = 5, units="in", device='png', dpi=600)

ggsave(filename = "P402_pH.png", plot = P402_pH, height = 2.5, width = 5, units="in", device='png', dpi=600)

ggsave(filename = "P402_Oxy.png", plot = P402_Oxy, height = 2.5, width = 5, units="in", device='png', dpi=600)
```


## P22

remove missing rows and remove weird values
```{r}
Environmental_sum<-Environmental %>% drop_na(pH)
Environmental_sum <- subset(Environmental_sum, pH != 2130.2)
```

```{r}
Environmental_sum<-Environmental_sum %>% drop_na(O2.in.mg.l)
```

summarize by depth category
```{r}
Environmental_sum<-Environmental_sum %>%
  group_by(Station,Date,Depth) %>%
  summarise(
    temp = mean(CTDTMP_DEG_C_ITS90),
    pH = mean(pH),
    Oxygen = mean(O2.in.mg.l))
```

convert date format
```{r}
Environmental_sum <- Environmental_sum %>% 
  mutate(date_convert = as.Date(Date, format = "%m/%d/%y"))
```

subset to station and depth
```{r}
Environmental_sum_P22<-subset(Environmental_sum,Station=="P22")
Environmental_sum_P22<-subset(Environmental_sum_P22,Depth!="Surface")
Environmental_sum_P22<-subset(Environmental_sum_P22,Depth!="Mid")
```

convert to long format
```{r}
Environmental_sum_P22<-melt(Environmental_sum_P22, na.rm = FALSE,  id = c("Station","Date","Depth","date_convert"))
```

make plot
```{r,fig.width=5,fig.height=2.5}
#change factor order
Environmental_sum_P22$Depth <- factor(Environmental_sum_P22$Depth, levels = c("5-m","Deep"))

P22_temp<-subset(Environmental_sum_P22,variable=="temp")

P22_temp<-ggplot(data = P22_temp, aes(date_convert, value,color=Depth)) +
  geom_line()+ 
  geom_point() +
  scale_y_continuous(breaks = c(5,10,15,20), limits = c(5,20), expand = c(0,0))+
  theme_classic()+
  xlab("Year") +
  ylab("Temperature (C)")+
  scale_x_date(date_breaks = "1 year", date_labels = "%Y")+
    theme(axis.text=element_text(size=11))+
  theme(axis.title.x = element_blank())
P22_temp

P22_pH<-subset(Environmental_sum_P22,variable=="pH")

P22_pH<-ggplot(data = P22_pH, aes(date_convert, value,color=Depth)) +
  geom_line()+ 
  geom_point() +
  scale_y_continuous(breaks = c(7,7.5,8,8.5), limits = c(7,8.5), expand = c(0,0))+
  theme_classic()+
  xlab("Year") +
  ylab("pH")+
  scale_x_date(date_breaks = "1 year", date_labels = "%Y")+
    theme(axis.text=element_text(size=11))+
  theme(axis.title.x = element_blank())
P22_pH

P22_Oxy<-subset(Environmental_sum_P22,variable=="Oxygen")

P22_Oxy<-ggplot(data = P22_Oxy, aes(date_convert, value,color=Depth)) +
  geom_line()+ 
  geom_point() +
  scale_y_continuous(breaks = c(0,5,10,15), limits = c(0,15), expand = c(0,0))+
  theme_classic()+
  xlab("Year") +
  ylab("Oxygen mg/l")+
  scale_x_date(date_breaks = "1 year", date_labels = "%Y")+
    theme(axis.text=element_text(size=11))+
  theme(axis.title.x = element_blank())
P22_Oxy
```

save plots
```{r, eval=FALSE}
setwd("/Users/hailaschultz/Dropbox/Schultz_Dissertation/Data_Analysis/Schultz_dissertation-2/output")
ggsave(filename = "P22_temp.png", plot = P22_temp, height = 2.5, width = 5, units="in", device='png', dpi=600)

ggsave(filename = "P22_pH.png", plot = P22_pH, height = 2.5, width = 5, units="in", device='png', dpi=600)

ggsave(filename = "P22_Oxy.png", plot = P22_Oxy, height = 2.5, width = 5, units="in", device='png', dpi=600)
```


## P28

remove missing rows and remove weird values
```{r}
Environmental_sum<-Environmental %>% drop_na(pH)
Environmental_sum <- subset(Environmental_sum, pH != 2130.2)
```

```{r}
Environmental_sum<-Environmental_sum %>% drop_na(O2.in.mg.l)
```

summarize by depth category
```{r}
Environmental_sum<-Environmental_sum %>%
  group_by(Station,Date,Depth) %>%
  summarise(
    temp = mean(CTDTMP_DEG_C_ITS90),
    pH = mean(pH),
    Oxygen = mean(O2.in.mg.l))
```

convert date format
```{r}
Environmental_sum <- Environmental_sum %>% 
  mutate(date_convert = as.Date(Date, format = "%m/%d/%y"))
```

subset to station and depth
```{r}
Environmental_sum_P28<-subset(Environmental_sum,Station=="P28")
Environmental_sum_P28<-subset(Environmental_sum_P28,Depth!="Surface")
Environmental_sum_P28<-subset(Environmental_sum_P28,Depth!="Mid")
```

convert to long format
```{r}
Environmental_sum_P28<-melt(Environmental_sum_P28, na.rm = FALSE,  id = c("Station","Date","Depth","date_convert"))
```

make plot
```{r,fig.width=5,fig.height=2.5}
#change factor order
Environmental_sum_P28$Depth <- factor(Environmental_sum_P28$Depth, levels = c("5-m","Deep"))

P28_temp<-subset(Environmental_sum_P28,variable=="temp")

P28_temp<-ggplot(data = P28_temp, aes(date_convert, value,color=Depth)) +
  geom_line()+ 
  geom_point() +
  scale_y_continuous(breaks = c(5,10,15,20), limits = c(5,20), expand = c(0,0))+
  theme_classic()+
  xlab("Year") +
  ylab("Temperature (C)")+
  scale_x_date(date_breaks = "1 year", date_labels = "%Y")+
    theme(axis.text=element_text(size=11))+
  theme(axis.title.x = element_blank())
P28_temp

P28_pH<-subset(Environmental_sum_P28,variable=="pH")

P28_pH<-ggplot(data = P28_pH, aes(date_convert, value,color=Depth)) +
  geom_line()+ 
  geom_point() +
  scale_y_continuous(breaks = c(7,7.5,8,8.5), limits = c(7,8.5), expand = c(0,0))+
  theme_classic()+
  xlab("Year") +
  ylab("pH")+
  scale_x_date(date_breaks = "1 year", date_labels = "%Y")+
    theme(axis.text=element_text(size=11))+
  theme(axis.title.x = element_blank())
P28_pH

P28_Oxy<-subset(Environmental_sum_P28,variable=="Oxygen")

P28_Oxy<-ggplot(data = P28_Oxy, aes(date_convert, value,color=Depth)) +
  geom_line()+ 
  geom_point() +
  scale_y_continuous(breaks = c(0,5,10,15), limits = c(0,15), expand = c(0,0))+
  theme_classic()+
  xlab("Year") +
  ylab("Oxygen mg/l")+
  scale_x_date(date_breaks = "1 year", date_labels = "%Y")+
    theme(axis.text=element_text(size=11))+
  theme(axis.title.x = element_blank())
P28_Oxy
```

save plots
```{r, eval=FALSE}
setwd("/Users/hailaschultz/Dropbox/Schultz_Dissertation/Data_Analysis/Schultz_dissertation-2/output")
ggsave(filename = "P28_temp.png", plot = P28_temp, height = 2.5, width = 5, units="in", device='png', dpi=600)

ggsave(filename = "P28_pH.png", plot = P28_pH, height = 2.5, width = 5, units="in", device='png', dpi=600)

ggsave(filename = "P28_Oxy.png", plot = P28_Oxy, height = 2.5, width = 5, units="in", device='png', dpi=600)
```


### P402 April

remove missing rows and remove weird values
```{r}
Environmental_sum<-Environmental %>% drop_na(pH)
Environmental_sum <- subset(Environmental_sum, pH != 2130.2)
```

```{r}
Environmental_sum<-Environmental_sum %>% drop_na(O2.in.mg.l)
```

summarize by depth category
```{r}
Environmental_sum<-Environmental_sum %>%
  group_by(Station,Date,Depth, Month) %>%
  summarise(
    temp = mean(CTDTMP_DEG_C_ITS90),
    pH = mean(pH),
    Oxygen = mean(O2.in.mg.l))
```

convert date format
```{r}
Environmental_sum <- Environmental_sum %>% 
  mutate(date_convert = as.Date(Date, format = "%m/%d/%y"))
```

subset to station and depth
```{r}
Environmental_sum_P402<-subset(Environmental_sum,Station=="P402")
Environmental_sum_P402<-subset(Environmental_sum_P402,Depth!="Surface")
Environmental_sum_P402<-subset(Environmental_sum_P402,Depth!="Mid")
Environmental_sum_P402<-subset(Environmental_sum_P402,Month=="APR")
```

convert to long format
```{r}
Environmental_sum_P402<-melt(Environmental_sum_P402, na.rm = FALSE,  id = c("Station","Date","Depth","date_convert","Month"))
```

make plot
```{r,fig.width=5,fig.height=2.5}
#change factor order
Environmental_sum_P402$Depth <- factor(Environmental_sum_P402$Depth, levels = c("5-m","Deep"))

P402_temp<-subset(Environmental_sum_P402,variable=="temp")

P402_temp<-ggplot(data = P402_temp, aes(date_convert, value,color=Depth)) +
  geom_line()+ 
  geom_point() +
  scale_y_continuous(breaks = c(5,10,15,20), limits = c(5,20), expand = c(0,0))+
  theme_classic()+
  xlab("Year") +
  ylab("Temperature (C)")+
  scale_x_date(date_breaks = "1 year", date_labels = "%Y")+
    theme(axis.text=element_text(size=11))+
  theme(axis.title.x = element_blank())
P402_temp

P402_pH<-subset(Environmental_sum_P402,variable=="pH")

P402_pH<-ggplot(data = P402_pH, aes(date_convert, value,color=Depth)) +
  geom_line()+ 
  geom_point() +
  scale_y_continuous(breaks = c(7,7.5,8,8.5), limits = c(7,8.5), expand = c(0,0))+
  theme_classic()+
  xlab("Year") +
  ylab("pH")+
  scale_x_date(date_breaks = "1 year", date_labels = "%Y")+
    theme(axis.text=element_text(size=11))+
  theme(axis.title.x = element_blank())
P402_pH

P402_Oxy<-subset(Environmental_sum_P402,variable=="Oxygen")

P402_Oxy<-ggplot(data = P402_Oxy, aes(date_convert, value,color=Depth)) +
  geom_line()+ 
  geom_point() +
  scale_y_continuous(breaks = c(0,5,10,15), limits = c(0,15), expand = c(0,0))+
  theme_classic()+
  xlab("Year") +
  ylab("Oxygen mg/l")+
  scale_x_date(date_breaks = "1 year", date_labels = "%Y")+
    theme(axis.text=element_text(size=11))+
  theme(axis.title.x = element_blank())
P402_Oxy
```


### P28 July

remove missing rows and remove weird values
```{r}
Environmental_sum<-Environmental %>% drop_na(pH)
Environmental_sum <- subset(Environmental_sum, pH != 2130.2)
```

```{r}
Environmental_sum<-Environmental_sum %>% drop_na(O2.in.mg.l)
```

summarize by depth category
```{r}
Environmental_sum<-Environmental_sum %>%
  group_by(Station,Date,Depth, Month) %>%
  summarise(
    temp = mean(CTDTMP_DEG_C_ITS90),
    pH = mean(pH),
    Oxygen = mean(O2.in.mg.l))
```

convert date format
```{r}
Environmental_sum <- Environmental_sum %>% 
  mutate(date_convert = as.Date(Date, format = "%m/%d/%y"))
```

subset to station and depth
```{r}
Environmental_sum_P402<-subset(Environmental_sum,Station=="P402")
Environmental_sum_P402<-subset(Environmental_sum_P402,Depth!="Surface")
Environmental_sum_P402<-subset(Environmental_sum_P402,Depth!="Mid")
Environmental_sum_P402<-subset(Environmental_sum_P402,Month=="JUL")
```

convert to long format
```{r}
Environmental_sum_P402<-melt(Environmental_sum_P402, na.rm = FALSE,  id = c("Station","Date","Depth","date_convert","Month"))
```

make plot
```{r,fig.width=5,fig.height=2.5}
#change factor order
Environmental_sum_P402$Depth <- factor(Environmental_sum_P402$Depth, levels = c("5-m","Deep"))

P402_temp<-subset(Environmental_sum_P402,variable=="temp")

P402_temp<-ggplot(data = P402_temp, aes(date_convert, value,color=Depth)) +
  geom_line()+ 
  geom_point() +
  scale_y_continuous(breaks = c(5,10,15,20), limits = c(5,20), expand = c(0,0))+
  theme_classic()+
  xlab("Year") +
  ylab("Temperature (C)")+
  scale_x_date(date_breaks = "1 year", date_labels = "%Y")+
    theme(axis.text=element_text(size=11))+
  theme(axis.title.x = element_blank())
P402_temp

P402_pH<-subset(Environmental_sum_P402,variable=="pH")

P402_pH<-ggplot(data = P402_pH, aes(date_convert, value,color=Depth)) +
  geom_line()+ 
  geom_point() +
  scale_y_continuous(breaks = c(7,7.5,8,8.5), limits = c(7,8.5), expand = c(0,0))+
  theme_classic()+
  xlab("Year") +
  ylab("pH")+
  scale_x_date(date_breaks = "1 year", date_labels = "%Y")+
    theme(axis.text=element_text(size=11))+
  theme(axis.title.x = element_blank())
P402_pH

P402_Oxy<-subset(Environmental_sum_P402,variable=="Oxygen")

P402_Oxy<-ggplot(data = P402_Oxy, aes(date_convert, value,color=Depth)) +
  geom_line()+ 
  geom_point() +
  scale_y_continuous(breaks = c(0,5,10,15), limits = c(0,15), expand = c(0,0))+
  theme_classic()+
  xlab("Year") +
  ylab("Oxygen mg/l")+
  scale_x_date(date_breaks = "1 year", date_labels = "%Y")+
    theme(axis.text=element_text(size=11))+
  theme(axis.title.x = element_blank())
P402_Oxy
```

### P402 Sep

remove missing rows and remove weird values
```{r}
Environmental_sum<-Environmental %>% drop_na(pH)
Environmental_sum <- subset(Environmental_sum, pH != 2130.2)
```

```{r}
Environmental_sum<-Environmental_sum %>% drop_na(O2.in.mg.l)
```

summarize by depth category
```{r}
Environmental_sum<-Environmental_sum %>%
  group_by(Station,Date,Depth, Month) %>%
  summarise(
    temp = mean(CTDTMP_DEG_C_ITS90),
    pH = mean(pH),
    Oxygen = mean(O2.in.mg.l))
```

convert date format
```{r}
Environmental_sum <- Environmental_sum %>% 
  mutate(date_convert = as.Date(Date, format = "%m/%d/%y"))
```

subset to station and depth
```{r}
Environmental_sum_P402<-subset(Environmental_sum,Station=="P402")
Environmental_sum_P402<-subset(Environmental_sum_P402,Depth!="Surface")
Environmental_sum_P402<-subset(Environmental_sum_P402,Depth!="Mid")
Environmental_sum_P402<-subset(Environmental_sum_P402,Month=="SEP")
```

convert to long format
```{r}
Environmental_sum_P402<-melt(Environmental_sum_P402, na.rm = FALSE,  id = c("Station","Date","Depth","date_convert","Month"))
```

make plot
```{r,fig.width=5,fig.height=2.5}
#change factor order
Environmental_sum_P402$Depth <- factor(Environmental_sum_P402$Depth, levels = c("5-m","Deep"))

P402_temp<-subset(Environmental_sum_P402,variable=="temp")

P402_temp<-ggplot(data = P402_temp, aes(date_convert, value,color=Depth)) +
  geom_line()+ 
  geom_point() +
  scale_y_continuous(breaks = c(5,10,15,20), limits = c(5,20), expand = c(0,0))+
  theme_classic()+
  xlab("Year") +
  ylab("Temperature (C)")+
  scale_x_date(date_breaks = "1 year", date_labels = "%Y")+
    theme(axis.text=element_text(size=11))+
  theme(axis.title.x = element_blank())
P402_temp

P402_pH<-subset(Environmental_sum_P402,variable=="pH")

P402_pH<-ggplot(data = P402_pH, aes(date_convert, value,color=Depth)) +
  geom_line()+ 
  geom_point() +
  scale_y_continuous(breaks = c(7,7.5,8,8.5), limits = c(7,8.5), expand = c(0,0))+
  theme_classic()+
  xlab("Year") +
  ylab("pH")+
  scale_x_date(date_breaks = "1 year", date_labels = "%Y")+
    theme(axis.text=element_text(size=11))+
  theme(axis.title.x = element_blank())
P402_pH

P402_Oxy<-subset(Environmental_sum_P402,variable=="Oxygen")

P402_Oxy<-ggplot(data = P402_Oxy, aes(date_convert, value,color=Depth)) +
  geom_line()+ 
  geom_point() +
  scale_y_continuous(breaks = c(0,5,10,15), limits = c(0,15), expand = c(0,0))+
  theme_classic()+
  xlab("Year") +
  ylab("Oxygen mg/l")+
  scale_x_date(date_breaks = "1 year", date_labels = "%Y")+
    theme(axis.text=element_text(size=11))+
  theme(axis.title.x = element_blank())
P402_Oxy
```