---
title: "02_Explore-Zoop"
date: "2023-09-25"
output:
  html_document:
    toc: true
    toc_depth: 2
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,         # Display code chunks
  eval = TRUE,        # Evaluate code chunks
  warning = FALSE,     # Hide warnings
  message = FALSE,     # Hide messages
  fig.width = 10,       # Set plot width in inches
  fig.height = 6,      # Set plot height in inches
  fig.align = "center" # Align plots to the center
)
```

load packages
```{r}
library(vegan)
library(EcolUtils)
```


# Load Zooplankton Data
This dataset is from our WOAC zooplankton database and includes morphological abundance and biomass data for each species identified in each sample collected. Vertical and oblique tows are included and represent samples that are integrated through the water column.

Read data into R
```{r}
Zooplankton <- read.csv("/Users/hailaschultz/Dropbox/Schultz_Dissertation/Data_Analysis/Schultz_dissertation-2/data/WOAC_Species_Densities_2014-2022.csv")
```

# NMDS plots
NMDS is chosen based on advice on this webpage:https://www.researchgate.net/post/How-to-choose-ordination-method-such-as-PCA-CA-PCoA-and-NMDS

"If you have a dataset that include null values (e.g. most dataset from genotyping using fingerprinting methods include null values, when for example a bacterial OTU is present in some samples and not in others), I would advise you to use Bray-Curtis similarity matrix and nMDS ordination. Bray-Curtis distance is chosen because it is not affected by the number of null values between samples like Euclidean distance, and nMDS is chosen because you can choose any similarity matrix, not like PCA."

Since the data have null values, I am using nMDS with Bray-Curtis distance

subset to stations sampled during WOAC cruises
```{r}
unique(Zooplankton$Station)
Zoop_sub<-subset(Zooplankton,Station!="PEFS1d")
Zoop_sub<-subset(Zoop_sub,Station!="PEFN1d")
Zoop_sub<-subset(Zoop_sub,Station!="PEFN2")
Zoop_sub<-subset(Zoop_sub,Station!="PEFS2")
Zoop_sub<-subset(Zoop_sub,Station!="P381")
Zoop_sub<-subset(Zoop_sub,Station!="P105")
Zoop_sub<-subset(Zoop_sub,Station!="P136")
Zoop_sub<-subset(Zoop_sub,Station!="P7")
Zoop_sub<-subset(Zoop_sub,Station!="P123")
Zoop_sub<-subset(Zoop_sub,Station!="P132")
unique(Zoop_sub$Station)
#remove october and november samples
Zoop_sub<-subset(Zoop_sub,Month!="NOV")
Zoop_sub<-subset(Zoop_sub,Month!="OCT")
Zoop_sub<-subset(Zoop_sub,Month!="MAR")
#recode p4
Zoop_sub$Station <- recode(Zoop_sub$Station, 
                                 "p4" = "P4")
#recode month names to match
unique(Zoop_sub$Month)
Zoop_sub$Month <- recode(Zoop_sub$Month, 
                                 "Jul" = "JUL",
                         "Sep"="SEP")
```

examine taxa present
```{r}
unique(Zoop_sub$Genus.species)
```
add up multiple lines per station
```{r}
Zoop_sub <- Zoop_sub %>%
  group_by(Code,Station,Basin,Year,Month,Genus.species) %>%
  summarise(
    zoop_density = sum(Density....m3.))
```

change from long to wide format
```{r}
colnames(Zoop_sub)
Zoop_sub_wide<-dcast(Zoop_sub, Code+Station+Basin+Year+Month~ Genus.species,value.var = "zoop_density")
#make sure factors are formatted right
Zoop_sub_wide$Year=as.factor(Zoop_sub_wide$Year)
Zoop_sub_wide$Station=as.factor(Zoop_sub_wide$Station)
Zoop_sub_wide$Month=as.factor(Zoop_sub_wide$Month)
Zoop_sub_wide$Basin=as.factor(Zoop_sub_wide$Basin)
#remove station without data
Zoop_sub_wide <- Zoop_sub_wide[-c(195), ]
```

remove non-data columns, convert to proportionas, and arcsine sqrt transformation
```{r}
RE2<- Zoop_sub_wide[,6:ncol(Zoop_sub_wide)]
#replace N/A with 0
RE2[is.na(RE2)] <- 0
#convert to proportions
RE2<-RE2/rowSums(RE2)
#arcsine sqrt transformation
RE2<-asin(sqrt(RE2))
RE3<-as.matrix(RE2)
```

## PERMANOVA
```{r}
dist<-vegdist(RE2, method='bray')
dist
perm<-adonis2(dist~Year*Month*Station, data=Zoop_sub_wide, permutations = 999, method="bray")
perm
```
See pairwise differences
```{r}
adonis.pair(vegdist(RE2),Zoop_sub_wide$Station)
adonis.pair(vegdist(RE2),Zoop_sub_wide$Year)
adonis.pair(vegdist(RE2),Zoop_sub_wide$Month)
```

## PERMDISPR
Station: check for dispersion
```{r}
Zoop.bd <- betadisper(dist, Zoop_sub_wide$Station)
Zoop.bd
anova(Zoop.bd)
permutest(Zoop.bd)
```
yes, differences in dispersion

Year: check for dispersion
```{r}
Zoop.bd <- betadisper(dist, Zoop_sub_wide$Year)
Zoop.bd
anova(Zoop.bd)
permutest(Zoop.bd)
```
yes, differences in dispersion

Month: check for dispersion
```{r}
Zoop.bd <- betadisper(dist, Zoop_sub_wide$Month)
Zoop.bd
anova(Zoop.bd)
permutest(Zoop.bd)
```
yes, differences in dispersion

## Run NMDS
```{r}
NMDSmodel <- metaMDS(RE2, distance = "bray",trymax = 200)
NMDSmodel
```

```{r}
plot(NMDSmodel)
```

get datascores
```{r}
data.scores<-scores(NMDSmodel,display="sites")
data.scores<- as.data.frame(data.scores)
```

```{r}
data.scores$Station = Zoop_sub_wide$Station
data.scores$Year = Zoop_sub_wide$Year
data.scores$Month = Zoop_sub_wide$Month
```

plot data scores
```{r}
xx <- ggplot(data.scores, aes(x = NMDS1, y = NMDS2,colour = Station)) +stat_ellipse()+geom_point(aes(colour = Station))+theme_bw()+theme(panel.grid.major = element_blank(),
                                                                                           panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))
xx
```

```{r}
xx <- ggplot(data.scores, aes(x = NMDS1, y = NMDS2,colour = Year)) +stat_ellipse()+geom_point(aes(colour = Year)+theme_bw()+theme(panel.grid.major = element_blank(),panel.grid.minor = element_blank(), axis.line = element_line(colour = "black")))
xx
```

```{r}
xx <- ggplot(data.scores, aes(x = NMDS1, y = NMDS2,colour = Month)) +stat_ellipse()+geom_point(aes(colour = Month)+theme_bw()+theme(panel.grid.major = element_blank(),panel.grid.minor = element_blank(), axis.line = element_line(colour = "black")))
xx
```

## Species fitting
```{r}
vf <- envfit(NMDSmodel, RE3, perm = 999)
vf
spp.scrs <- as.data.frame(scores(vf, display = "vectors"))
spp.scrs <- cbind(spp.scrs, Species = rownames(spp.scrs))
####for ggplot
arrow_factor <- ordiArrowMul(vf)
spp.scrs <- as.data.frame(scores(vf, display = "vectors")) * arrow_factor
spp.scrs <- cbind(spp.scrs, Species = rownames(spp.scrs), Pvalues = vf$vectors$pvals, R_squared = vf$vectors$r)

# select significance similarly to `plot(vf, p.max = 0.01)`
spp.scrs <- subset(spp.scrs, Pvalues < 0.05)


```