---
title: "04_hypothesis-1"
date: "2024-02-08"
output:
  html_document:
    toc: true
    toc_depth: 2
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,         # Display code chunks
  eval = TRUE,        # Evaluate code chunks
  warning = FALSE,     # Hide warnings
  message = FALSE,     # Hide messages
  fig.width = 10,       # Set plot width in inches
  fig.height = 6,      # Set plot height in inches
  fig.align = "center" # Align plots to the center
)
```

This markdown file details analyses done for hypothesis 1 of chapter 2 of my dissertation: Gelatinous zooplankton will be more abundant than crustaceans during marine heatwave years.

load packages
```{r}
library(ggrepel)
library("scales")   
library(dplyr)
library(tidyr)
library(data.table)
library(ggeffects)
```

Read zooplanton data into R
```{r}
Zooplankton <- read.csv("/Users/hailaschultz/Dropbox/Schultz_Dissertation/Data_Analysis/Schultz_dissertation-2/data/WOAC_Species_Densities_2014-2022.csv")
```
Read environmental data into R
```{r}
Environmental <- read.csv("/Users/hailaschultz/Dropbox/Schultz_Dissertation/Data_Analysis/Schultz_dissertation-2/data/WOAC_Chem_data_2014-2022_all-niskins.csv")
```

# Environmental Data Summary
recode months to match with zooplankton data
```{r}
unique(Environmental$Month)
Environmental$Month <- recode(Environmental$Month, 
                                 "Jul" = "JUL",
                         "Sep"="SEP",
                         "Oct"="SEP",
                         "Apr"="APR",
                         "May"="MAY",
                         "Nov"="NOV",
                         "Aug"="AUG",
                         "Jun"="JUL",
                         "OCT"="SEP")
#remove unnecessary sample dates
unique(Environmental$Date)
unique(Zooplankton$Date)
Environmental<-subset(Environmental,Date!="10/30/14")
Environmental<-subset(Environmental,Date!="10/31/14")
Environmental<-subset(Environmental,Date!="10/23/14")
Environmental<-subset(Environmental,Date!="10/29/14")
Environmental<-subset(Environmental,Date!="5/24/15")
Environmental<-subset(Environmental,Date!="11/18/15")
Environmental<-subset(Environmental,Date!="11/16/15")
Environmental<-subset(Environmental,Date!="11/17/15")
Environmental<-subset(Environmental,Date!="10/27/16")
Environmental<-subset(Environmental,Date!="5/3/17")
Environmental<-subset(Environmental,Date!="5/4/17")
Environmental<-subset(Environmental,Date!="10/18/17")
Environmental<-subset(Environmental,Date!="5/24/18")
Environmental<-subset(Environmental,Date!="10/18/18")
Environmental<-subset(Environmental,Date!="10/19/18")
Environmental<-subset(Environmental,Date!="5/23/19")
Environmental<-subset(Environmental,Date!="8/31/20")
Environmental<-subset(Environmental,Date!="10/1/21")
Environmental<-subset(Environmental,Date!="5/6/22")
Environmental<-subset(Environmental,Date!="10/14/22")
Environmental<-subset(Environmental,Date!="10/15/22")
Environmental<-subset(Environmental,Date!="9/29/21")
Environmental<-subset(Environmental,Date!="7/4/20")
Environmental<-subset(Environmental,Date!="9/30/20")
Environmental<-subset(Environmental,Date!="9/18/17")
```

Temp: summarize by depth category
```{r}
temp<-Environmental %>%
  group_by(Station,Date,Depth,Year,Month) %>%
  summarise(
    temp = mean(CTDTMP_DEG_C_ITS90))

#subset to surface samples
temp_surf<-subset(temp,Depth=="Surface")
#make unique column
temp_surf$code<-paste(temp_surf$Station, temp_surf$Year,temp_surf$Month,sep="-")
```

pH: remove missing rows and remove weird values
```{r}
pH<-Environmental %>% drop_na(pH)
pH <- subset(pH, pH != 2130.2)
```
summarize by depth category
```{r}
pH<-pH %>%
  group_by(Station,Date,Depth,Year,Month) %>%
  summarise(
    pH = mean(pH))
#subset to surface samples
pH_surf<-subset(pH,Depth=="Surface")
#make unique column
pH_surf$code<-paste(pH_surf$Station, pH_surf$Year,pH_surf$Month,sep="-")
```

O2: remove missing rows and remove weird values
```{r}
Oxygen<-Environmental %>% drop_na(O2.in.mg.l)
```

summarize by depth category
```{r}
Oxygen<-Oxygen %>%
  group_by(Station,Date,Depth,Year,Month) %>%
  summarise(
    Oxygen = mean(O2.in.mg.l))
#subset to surface samples
Oxygen_surf<-subset(Oxygen,Depth=="Surface")
#make unique column
Oxygen_surf$code<-paste(Oxygen_surf$Station, Oxygen_surf$Year,Oxygen_surf$Month,sep="-")
```


Merge environmental tables
```{r}
joined <- merge(temp_surf, pH_surf, by.x = "code", 
             by.y = "code", all.x = TRUE, all.y = TRUE)
joined <- merge(joined, Oxygen_surf, by.x = "code", 
             by.y = "code", all.x = TRUE, all.y = TRUE)
#remove excess columns
env_joined <- joined[ -c(8:12,14:18) ]
```


# Zooplankton Data
subset to stations sampled during WOAC cruises
```{r}
unique(Zooplankton$Station)
Zoop_sub<-subset(Zooplankton,Station!="PEFS1d")
Zoop_sub<-subset(Zoop_sub,Station!="PEFN1d")
Zoop_sub<-subset(Zoop_sub,Station!="PEFN2")
Zoop_sub<-subset(Zoop_sub,Station!="PEFS2")
Zoop_sub<-subset(Zoop_sub,Station!="P381")
Zoop_sub<-subset(Zoop_sub,Station!="P105")
Zoop_sub<-subset(Zoop_sub,Station!="P136")
Zoop_sub<-subset(Zoop_sub,Station!="P7")
Zoop_sub<-subset(Zoop_sub,Station!="P123")
Zoop_sub<-subset(Zoop_sub,Station!="P132")
unique(Zoop_sub$Date)
#remove unnecessary dates
Zoop_sub<-subset(Zoop_sub,Date!="10/30/14")
Zoop_sub<-subset(Zoop_sub,Date!="10/31/14")
Zoop_sub<-subset(Zoop_sub,Date!="10/29/14")
Zoop_sub<-subset(Zoop_sub,Date!="9/14/22")
Zoop_sub<-subset(Zoop_sub,Date!="9/16/22")
Zoop_sub<-subset(Zoop_sub,Date!="9/13/22")
Zoop_sub<-subset(Zoop_sub,Date!="9/12/22")
Zoop_sub<-subset(Zoop_sub,Date!="11/16/15")
Zoop_sub<-subset(Zoop_sub,Date!="3/17/16")
Zoop_sub<-subset(Zoop_sub,Date!="3/18/16")
Zoop_sub<-subset(Zoop_sub,Date!="11/18/15")
#recode p4
Zoop_sub$Station <- recode(Zoop_sub$Station, 
                                 "p4" = "P4")
#recode month names to match
unique(Zoop_sub$Month)
Zoop_sub$Month <- recode(Zoop_sub$Month, 
                                 "Jul" = "JUL",
                         "Sep"="SEP",
                         "OCT"="SEP")
#remove ethanol samples
Zoop_sub<-Zoop_sub[!grepl('_EtOH', Zoop_sub$Code),]
#remove oblique samples
Zoop_sub<-subset(Zoop_sub,Tow.Type!="Oblique")
```

add up multiple lines per station, categorize by crustaceans and gelatinous
```{r}
Zoop_sub <- Zoop_sub %>%
  group_by(Code,Station,Basin,Year,Month,Crustaceans.vs.Gelatinous) %>%
  summarise(
    zoop_density = sum(Density....m3.))
```

change from long to wide format
```{r}
colnames(Zoop_sub)
Zoop_sub_wide<-dcast(Zoop_sub, Code+Station+Basin+Year+Month~ Crustaceans.vs.Gelatinous,value.var = "zoop_density")
#make sure factors are formatted right
Zoop_sub_wide$Year=as.factor(Zoop_sub_wide$Year)
Zoop_sub_wide$Station=as.factor(Zoop_sub_wide$Station)
Zoop_sub_wide$Month=as.factor(Zoop_sub_wide$Month)
Zoop_sub_wide$Basin=as.factor(Zoop_sub_wide$Basin)
```

```{r}
#make unique column
Zoop_sub_wide$code<-paste(Zoop_sub_wide$Station, Zoop_sub_wide$Year,Zoop_sub_wide$Month,sep="-")
unique(Zoop_sub$Code)
```


# Combine env and zoop
```{r}
Zoop_env <- merge(env_joined,Zoop_sub_wide, by.x = "code", 
             by.y = "code", all.x = TRUE, all.y = TRUE)
#remove excess columns
Zoop_env <- Zoop_env[ -c(11,13:14) ]
#remove rows with NA
Zoop_env<-Zoop_env[!is.na(Zoop_env$Year.x),]
```

## Proportions
remove non-data columns, convert to proportions
```{r}
RE2<- Zoop_env[,12:ncol(Zoop_env)]
#replace N/A with 0
RE2[is.na(RE2)] <- 0
#convert to proportions
RE2<-RE2/rowSums(RE2)
```

add columns back in
```{r}
RE2$code = Zoop_env$code
RE2$Station.x = Zoop_env$Station.x
RE2$Year.x = Zoop_env$Year.x
RE2$Month.x = Zoop_env$Month.x
RE2$pH = Zoop_env$pH
RE2$temp = Zoop_env$temp
RE2$Oxygen = Zoop_env$Oxygen
```

```{r}
Zoop_env_prop<-RE2
```


### change format
```{r}
Zoop_env_long<-melt(Zoop_env_prop, na.rm = FALSE, c("code", "Station.x","Year.x","Month.x","temp","pH","Oxygen"))
```

```{r}
Zoop_env_long$value[Zoop_env_long$value == 0] <- NA

Zoop_env_long<-Zoop_env_long[complete.cases(Zoop_env_long), ]
```

### subset to gelatinous
```{r}
Gelatinous<-subset(Zoop_env_long,variable=="Gelatinous")
Gelatinous$Year.x<-as.factor(Gelatinous$Year.x)
```

```{r}
#change factor order
Gelatinous$Station.x <- factor(Gelatinous$Station.x, levels = c("P22", "P4", "P8","P28","P12","P402","P38"))
```

```{r,fig.width=4,fig.height=3}
Gelatinous_plot<-ggplot(Gelatinous, aes(x=temp, y=value))+geom_point(aes(colour = Station.x))+xlab("Temperature")+ylab("Gelatinous Zooplankton Relative Abundance")+theme_bw()+theme(panel.grid.major = element_blank(),                panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))+ scale_colour_manual(values=c("#F8766D","#C49A00","#53B400","#00C094","#00B6EB","#A69AFF","#FB61D7"))
Gelatinous_plot
```

matrix of plots - surface values
```{r}
pairs(Gelatinous[,c(5:7,9)], pch = 19)
```

### subset to crustacean
```{r}
Crustacean<-subset(Zoop_env_long,variable=="Crustaceans")
Crustacean$Year.x<-as.factor(Crustacean$Year.x)
```

```{r}
#change factor order
Crustacean$Station.x <- factor(Crustacean$Station.x, levels = c("P22", "P4", "P8","P28","P12","P402","P38"))
```

```{r,fig.width=4,fig.height=3}
Crustacean_plot<-ggplot(Crustacean, aes(x=temp, y=value))+geom_point(aes(colour = Station.x))+xlab("Temperature")+ylab("Crustacean Zooplankton Relative Abundance")+theme_bw()+theme(panel.grid.major = element_blank(),                panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))+ scale_colour_manual(values=c("#F8766D","#C49A00","#53B400","#00C094","#00B6EB","#A69AFF","#FB61D7"))
Crustacean_plot
```
matrix of plots - surface values
```{r}
pairs(Crustacean[,c(5:7,9)], pch = 19)
```

## Abundance
### change format
```{r}
#subset columns
Zoop_env<-select(Zoop_env, Gelatinous,Crustaceans, code, Station.x, Year.x, Month.x, temp, pH, Oxygen)


Zoop_env_long<-melt(Zoop_env, na.rm = FALSE, c("code", "Station.x","Year.x","Month.x","temp","pH","Oxygen"))
```

```{r}
#remove NA rows
Zoop_env_long$value[Zoop_env_long$value == 0] <- NA

Zoop_env_long<-Zoop_env_long[complete.cases(Zoop_env_long), ]
```

# subset to gelatinous
```{r}
Gelatinous<-subset(Zoop_env_long,variable=="Gelatinous")
Gelatinous$Year.x<-as.factor(Gelatinous$Year.x)
```

```{r}
#change factor order
Gelatinous$Station.x <- factor(Gelatinous$Station.x, levels = c("P22", "P4", "P8","P28","P12","P402","P38"))
```

```{r,fig.width=4,fig.height=3}
Gelatinous_plot<-ggplot(Gelatinous, aes(x=temp, y=value))+geom_point(aes(colour = Station.x))+xlab("Temperature")+ylab("Gelatinous Zooplankton Abundance")+theme_bw()+theme(panel.grid.major = element_blank(),                panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))+ scale_colour_manual(values=c("#F8766D","#C49A00","#53B400","#00C094","#00B6EB","#A69AFF","#FB61D7"))
Gelatinous_plot
```

matrix of plots - surface values
```{r}
pairs(Gelatinous[,c(5:7,9)], pch = 19)
```


# subset to crustacean
```{r}
Crustacean<-subset(Zoop_env_long,variable=="Crustaceans")
Crustacean$Year.x<-as.factor(Crustacean$Year.x)
```

```{r}
#change factor order
Crustacean$Station.x <- factor(Crustacean$Station.x, levels = c("P22", "P4", "P8","P28","P12","P402","P38"))
```

```{r,fig.width=4,fig.height=3}
Crustacean_plot<-ggplot(Crustacean, aes(x=temp, y=value))+geom_point(aes(colour = Station.x))+xlab("Temperature")+ylab("Crustacean Zooplankton Abundance")+theme_bw()+theme(panel.grid.major = element_blank(),                panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))+ scale_colour_manual(values=c("#F8766D","#C49A00","#53B400","#00C094","#00B6EB","#A69AFF","#FB61D7"))
Crustacean_plot
```

matrix of plots - surface values
```{r}
pairs(Crustacean[,c(5:7,9)], pch = 19)
```



## Calculate temperature anomalies
Taken from BethElLee's methods: 
Average by season, subtract datapoint from average, divide by standard deviaiton of the average to standardize

### Average over all stations
Deep temperatures
```{r}
temp<-Environmental %>%
  group_by(Depth,Year,Month,DateCat) %>%
  summarise(
    temp = mean(CTDTMP_DEG_C_ITS90))
#subset to surface samples
temp_surf<-subset(temp,Depth=="Deep")
```
month averages
```{r}
temp_ave<-Environmental %>%
  group_by(Month,Depth) %>%
  summarise(
    temp = mean(CTDTMP_DEG_C_ITS90),
    stdev = sd(CTDTMP_DEG_C_ITS90))

#subset to surface samples
temp_surf_ave<-subset(temp_ave,Depth=="Deep")
```
anomalies
```{r}
#add average column 
temp_surf$Ave <- temp_surf_ave$temp[match(temp_surf$Month, temp_surf_ave$Month)]

#add statndard deviation column
temp_surf$stdev <- temp_surf_ave$stdev[match(temp_surf$Month, temp_surf_ave$Month)]

# calculate anomalies
temp_surf$anomaly<-(temp_surf$temp-temp_surf$Ave)/temp_surf$stdev
```

convert date format
```{r}
temp_surf <- temp_surf %>% 
  mutate(date_convert = as.Date(DateCat, format = "%m/%d/%y"))
```



plot
```{r}
temp_plot<-ggplot(data = temp_surf, aes(date_convert, anomaly)) +
  geom_bar(aes(fill = anomaly < 0),stat='identity',width=30)+
  theme_classic()+
  xlab("Year") + ylab("temperature anomaly")+
  scale_x_date(date_breaks = "1 year", date_labels = "%Y")+
  scale_fill_manual(guide = FALSE, breaks = c(TRUE, FALSE), values=c("blue", "red"))
temp_plot
```
regime shift detection
```{r}
RSI_data<-Rodionov(temp_surf, "anomaly", "date_convert",3, merge=TRUE)
```
visualize
```{r}
graph<-RSI_graph(RSI_data, "anomaly", "date_convert", "...10")
graph
```




### Surface

month and station averages
```{r}
temp_ave<-Environmental %>%
  group_by(Depth,Month,Station) %>%
  summarise(
    temp = mean(CTDTMP_DEG_C_ITS90),
    stdev = sd(CTDTMP_DEG_C_ITS90))

#subset to surface samples
temp_surf_ave<-subset(temp_ave,Depth=="Surface")

#unique station-month combo
temp_surf_ave$station_month<-paste(temp_surf_ave$Station, temp_surf_ave$Month,sep="-")
```

temperatures
```{r}
temp<-Environmental %>%
  group_by(Station,Date,Depth,Year,Month) %>%
  summarise(
    temp = mean(CTDTMP_DEG_C_ITS90))
#subset to surface samples
temp_surf<-subset(temp,Depth=="Surface")
#unique station-month combo
temp_surf$station_month<-paste(temp_surf$Station, temp_surf$Month,sep="-")
```

anomalies
```{r}
#add average column 
temp_surf$Ave <- temp_surf_ave$temp[match(temp_surf$station_month, temp_surf_ave$station_month)]

#add statndard deviation column
temp_surf$stdev <- temp_surf_ave$stdev[match(temp_surf$station_month, temp_surf_ave$station_month)]

# calculate anomalies
temp_surf$anomaly<-(temp_surf$temp-temp_surf$Ave)/temp_surf$stdev
```

convert date format
```{r}
temp_surf <- temp_surf %>% 
  mutate(date_convert = as.Date(Date, format = "%m/%d/%y"))
```

plot
```{r,fig.width=5,fig.height=9}
temp_plot<-ggplot(data = temp_surf, aes(date_convert, anomaly)) +
  geom_bar(aes(fill = anomaly < 0),stat='identity',width=30)+
  facet_grid(Station ~ .) +
  theme_classic()+
  xlab("Year") + ylab("temperature anomaly")+
  scale_x_date(date_breaks = "1 year", date_labels = "%Y")+
  scale_fill_manual(guide = FALSE, breaks = c(TRUE, FALSE), values=c("blue", "red"))
temp_plot
```


### 5-m

month and station averages
```{r}
temp_ave<-Environmental %>%
  group_by(Depth,Month,Station) %>%
  summarise(
    temp = mean(CTDTMP_DEG_C_ITS90),
    stdev = sd(CTDTMP_DEG_C_ITS90))

#subset to surface samples
temp_surf_ave<-subset(temp_ave,Depth=="5-m")

#unique station-month combo
temp_surf_ave$station_month<-paste(temp_surf_ave$Station, temp_surf_ave$Month,sep="-")
```

temperatures
```{r}
temp<-Environmental %>%
  group_by(Station,Date,Depth,Year,Month) %>%
  summarise(
    temp = mean(CTDTMP_DEG_C_ITS90))
#subset to surface samples
temp_surf<-subset(temp,Depth=="5-m")
#unique station-month combo
temp_surf$station_month<-paste(temp_surf$Station, temp_surf$Month,sep="-")
```

anomalies
```{r}
#add average column 
temp_surf$Ave <- temp_surf_ave$temp[match(temp_surf$station_month, temp_surf_ave$station_month)]

#add statndard deviation column
temp_surf$stdev <- temp_surf_ave$stdev[match(temp_surf$station_month, temp_surf_ave$station_month)]

# calculate anomalies
temp_surf$anomaly<-(temp_surf$temp-temp_surf$Ave)/temp_surf$stdev
```

convert date format
```{r}
temp_surf <- temp_surf %>% 
  mutate(date_convert = as.Date(Date, format = "%m/%d/%y"))
```

plot
```{r,fig.width=5,fig.height=9}
temp_plot<-ggplot(data = temp_surf, aes(date_convert, anomaly)) +
  geom_bar(aes(fill = anomaly < 0),stat='identity',width=30)+
  facet_grid(Station ~ .) +
  theme_classic()+
  xlab("Year") + ylab("temperature anomaly")+
  scale_x_date(date_breaks = "1 year", date_labels = "%Y")+
  scale_fill_manual(guide = FALSE, breaks = c(TRUE, FALSE), values=c("blue", "red"))
temp_plot
```

### Deep

month and station averages
```{r}
temp_ave<-Environmental %>%
  group_by(Depth,Month,Station) %>%
  summarise(
    temp = mean(CTDTMP_DEG_C_ITS90),
    stdev = sd(CTDTMP_DEG_C_ITS90))

#subset to surface samples
temp_surf_ave<-subset(temp_ave,Depth=="Deep")

#unique station-month combo
temp_surf_ave$station_month<-paste(temp_surf_ave$Station, temp_surf_ave$Month,sep="-")
```

temperatures
```{r}
temp<-Environmental %>%
  group_by(Station,Date,Depth,Year,Month) %>%
  summarise(
    temp = mean(CTDTMP_DEG_C_ITS90))
#subset to surface samples
temp_surf<-subset(temp,Depth=="Deep")
#unique station-month combo
temp_surf$station_month<-paste(temp_surf$Station, temp_surf$Month,sep="-")
```

anomalies
```{r}
#add average column 
temp_surf$Ave <- temp_surf_ave$temp[match(temp_surf$station_month, temp_surf_ave$station_month)]

#add statndard deviation column
temp_surf$stdev <- temp_surf_ave$stdev[match(temp_surf$station_month, temp_surf_ave$station_month)]

# calculate anomalies
temp_surf$anomaly<-(temp_surf$temp-temp_surf$Ave)/temp_surf$stdev
```

convert date format
```{r}
temp_surf <- temp_surf %>% 
  mutate(date_convert = as.Date(Date, format = "%m/%d/%y"))
```

plot
```{r,fig.width=5,fig.height=9}
temp_plot<-ggplot(data = temp_surf, aes(date_convert, anomaly)) +
  geom_bar(aes(fill = anomaly < 0),stat='identity',width=30)+
  facet_grid(Station ~ .) +
  theme_classic()+
  xlab("Year") + ylab("temperature anomaly")+
  scale_x_date(date_breaks = "1 year", date_labels = "%Y")+
  scale_fill_manual(guide = FALSE, breaks = c(TRUE, FALSE), values=c("blue", "red"))
temp_plot
```


# GLM Gelatinous
```{r}
hist(Gelatinous$value) 
```
Dependent variable is heavily zero-skewed

```{r}
hist(Gelatinous$temp) 
hist(Gelatinous$pH) 
hist(Gelatinous$Oxygen)
```
Independent variables are normally distributed

```{r}
ggplot(Gelatinous, aes(x=temp, y=value)) + geom_point()
```


Gaussian distribution
```{r}
gelatinous_gaussian<-glm(value~temp, data= Gelatinous, family="gaussian")
summary(gelatinous_gaussian)
plot(gelatinous_gaussian)
```
pretty bad outliers

Gamma distribution
```{r}
gelatinous_gamma<-glm(value~temp, data= Gelatinous, family="Gamma")
summary(gelatinous_gamma)
plot(gelatinous_gamma)
```
better but not perfect

Gamma with log link
```{r}
gelatinous_gamma_log<-glm(value~temp, data= Gelatinous, family="Gamma"(link="log"))
summary(gelatinous_gamma_log)
plot(gelatinous_gamma_log)
```
slightly better

check for multicollinearity
```{r}
DAData1 = subset(Gelatinous, select = c(temp, pH,Oxygen,value) )
pairs(DAData1 , panel = panel.smooth)

```
Oxygen and pH are correlated, choose only one

Gamma, log link, month, temp, oxygen
```{r}
Full_mod_1<-glm(value~Month.x+temp+Oxygen, data= Gelatinous, family="Gamma"(link="log"))
summary(Full_mod_1)
plot(Full_mod_1)
```
Month is not significant

Gamma, log link, month,station, temp, oxygen
```{r}
Full_mod_1<-glm(value~Month.x+Station.x+temp+Oxygen, data= Gelatinous, family="Gamma"(link="log"))
summary(Full_mod_1)
plot(Full_mod_1)
```
looks much better! Include station.


Gamma, log link,station, temp, oxygen
```{r}
Full_mod_1<-glm(value~Station.x+temp+Oxygen, data= Gelatinous, family="Gamma"(link="log"))
summary(Full_mod_1)
plot(Full_mod_1)
```
AIC is the same without month

Gamma, log link,station, temp
```{r}
Full_mod_1<-glm(value~Station.x+temp, data= Gelatinous, family="Gamma"(link="log"))
summary(Full_mod_1)
plot(Full_mod_1)
```
AIC is even better when I take out oxygen. Slightly better with month in, but since month is insignificant, I don't think it's necessary
P28 and P12 are significantly different than the other months



Use predict function to predict model values

```{r}
predictmod<-ggpredict(
  Full_mod_1,
  terms=c("temp","Station.x"),
  ci.lvl = 0.95,
  type = "fe",
  typical = "mean",
  condition = NULL,
  back.transform = TRUE,
  ppd = FALSE,
  vcov.fun = NULL,
  vcov.type = NULL,
  vcov.args = NULL,
  interval = "confidence")
```

```{r}
Plot<-plot(predictmod,add.data = TRUE,dot.size=1.5,dot.alpha=0.65)+
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),panel.background = element_blank(), axis.line = element_line(colour = "black"))+theme_classic() +
  geom_line(linewidth=1.5) +
  geom_ribbon( aes(ymin = conf.low, ymax = conf.high), alpha = .15)+
  xlab(bquote('Temperature'))+ylab(bquote('Gelatinous Zooplankton Density'))+
  theme(axis.text=element_text(size=15,colour="black"),
        legend.position="none",
        axis.title=element_text(size=15,colour="black"),
        axis.line = element_line(colour = "black"))+theme(plot.title = element_blank())+
  theme(plot.margin=margin(0.7, 0.7, 0.7, 0.7, unit = "cm"))+
  facet_grid(group ~ .)
Plot
```

