---
title: "06_env-cluster"
date: "2024-02-20"
output:
  html_document:
    toc: true
    toc_depth: 2
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,         # Display code chunks
  eval = TRUE,        # Evaluate code chunks
  warning = FALSE,     # Hide warnings
  message = FALSE,     # Hide messages
  fig.width = 10,       # Set plot width in inches
  fig.height = 6,      # Set plot height in inches
  fig.align = "center" # Align plots to the center
)
```

The purpose of this code is to conduct cluster analysis on the WOAC stations based on their environmental variables. I am interested in seeing if there are groups of stations that are environmentally similar which I can cluster together. This analysis is based on the workflow documented here: https://jtr13.github.io/cc21fall2/tutorial-on-cluster-analysis-1.html


load packages
```{r}
library(ggrepel)
library("scales")   
library(corrplot)
library(factoextra)
library(dplyr)
library(tidyr)
library(data.table)
library(tidyverse)
library(stats)
```


Read environmental data into R
```{r}
Environmental <- read.csv("/Users/hailaschultz/Dropbox/Schultz_Dissertation/Data_Analysis/Schultz_dissertation-2/data/WOAC_Chem_data_2014-2022_all-niskins.csv")
```

# Environmental Data Summary
recode months to match with zooplankton data
```{r}
unique(Environmental$Month)
Environmental$Month <- recode(Environmental$Month, 
                                 "Jul" = "JUL",
                         "Sep"="SEP",
                         "Oct"="SEP",
                         "Apr"="APR",
                         "May"="MAY",
                         "Nov"="NOV",
                         "Aug"="AUG",
                         "Jun"="JUL",
                         "OCT"="SEP")
#remove unnecessary sample dates
unique(Environmental$Date)
unique(Zooplankton$Date)
Environmental<-subset(Environmental,Date!="10/30/14")
Environmental<-subset(Environmental,Date!="10/31/14")
Environmental<-subset(Environmental,Date!="10/23/14")
Environmental<-subset(Environmental,Date!="10/29/14")
Environmental<-subset(Environmental,Date!="5/24/15")
Environmental<-subset(Environmental,Date!="11/18/15")
Environmental<-subset(Environmental,Date!="11/16/15")
Environmental<-subset(Environmental,Date!="11/17/15")
Environmental<-subset(Environmental,Date!="10/27/16")
Environmental<-subset(Environmental,Date!="5/3/17")
Environmental<-subset(Environmental,Date!="5/4/17")
Environmental<-subset(Environmental,Date!="10/18/17")
Environmental<-subset(Environmental,Date!="5/24/18")
Environmental<-subset(Environmental,Date!="10/18/18")
Environmental<-subset(Environmental,Date!="10/19/18")
Environmental<-subset(Environmental,Date!="5/23/19")
Environmental<-subset(Environmental,Date!="8/31/20")
Environmental<-subset(Environmental,Date!="10/1/21")
Environmental<-subset(Environmental,Date!="5/6/22")
Environmental<-subset(Environmental,Date!="10/14/22")
Environmental<-subset(Environmental,Date!="10/15/22")
Environmental<-subset(Environmental,Date!="9/29/21")
Environmental<-subset(Environmental,Date!="7/4/20")
Environmental<-subset(Environmental,Date!="9/30/20")
Environmental<-subset(Environmental,Date!="9/18/17")
```

get variable names
```{r}
colnames(Environmental)
```



## Surface

Temp: summarize by depth category
```{r}
temp<-Environmental %>%
  group_by(Station,Date,Depth,Year,Month) %>%
  summarise(
    temp = mean(CTDTMP_DEG_C_ITS90))

#subset to surface and deep
temp<-subset(temp,Depth!="5-m")
temp<-subset(temp,Depth!="Mid")
#make unique column
temp$code<-paste(temp$Station, temp$Year,temp$Month,temp$Depth,sep="-")
```


pH: remove missing rows and remove weird values
```{r}
pH<-Environmental %>% drop_na(pH)
pH <- subset(pH, pH != 2130.2)
```
summarize by depth category
```{r}
pH<-pH %>%
  group_by(Station,Date,Depth,Year,Month) %>%
  summarise(
    pH = mean(pH))
#subset 
pH<-subset(pH,Depth!="5-m")
pH<-subset(pH,Depth!="Mid")
#make unique column
pH$code<-paste(pH$Station, pH$Year,pH$Month,pH$Depth,sep="-")
```


O2: remove missing rows and remove weird values
```{r}
Oxygen<-Environmental %>% drop_na(O2.in.mg.l)
```

summarize by depth category
```{r}
Oxygen<-Oxygen %>%
  group_by(Station,Date,Depth,Year,Month) %>%
  summarise(
    Oxygen = mean(O2.in.mg.l))
#subset 
Oxygen<-subset(Oxygen,Depth!="5-m")
Oxygen<-subset(Oxygen,Depth!="Mid")
#make unique column
Oxygen$code<-paste(Oxygen$Station, Oxygen$Year,Oxygen$Month,Salinity$Depth,sep="-")
```

Salinity: remove missing rows and remove weird values
```{r}
Salinity<-Environmental %>% drop_na(CTD.Salinity..PSU.)
```

summarize by depth category
```{r}
Salinity<-Salinity %>%
  group_by(Station,Date,Depth,Year,Month) %>%
  summarise(
    Salinity = mean(CTD.Salinity..PSU.))
#subset to surface samples
Salinity<-subset(Salinity,Depth!="5-m")
Salinity<-subset(Salinity,Depth!="Mid")
#make unique column
Salinity$code<-paste(Salinity$Station, Salinity$Year,Salinity$Month,Salinity$Depth,sep="-")
```



Merge environmental tables
```{r}
joined <- merge(temp, pH, by.x = "code", 
             by.y = "code", all.x = TRUE, all.y = TRUE)
joined <- merge(joined, Oxygen, by.x = "code", 
             by.y = "code", all.x = TRUE, all.y = TRUE)
joined <- merge(joined, Salinity, by.x = "code", 
             by.y = "code", all.x = TRUE, all.y = TRUE)

#remove excess columns and na rows
env_joined <- joined[ -c(3,8:12,14:18,20:24) ]
#remove deep or joined
env_joined$code <- gsub("-Deep$", "", env_joined$code)
env_joined$code <- gsub("-Surface$", "", env_joined$code)
#make surface and deep variables separate
env_joined <- env_joined %>%
  pivot_wider(names_from = Depth.x, values_from = c(temp, pH, Oxygen, Salinity)) 
```

Note: unfortunately, not all stations have nutrient data, so when nutrients are included, a lot of stations are dropped. I have decided to proceed, omitting the nutrients.


```{r}
#drop pH column
env_joined<-subset(env_joined, select = -c(pH_Deep,pH_Surface))
#drop na
env_joined <- env_joined[complete.cases(env_joined), ]
```

# Anomaly analysis
month and station averages
```{r}
env_ave<-env_joined %>%
  group_by(Month.x,Station.x) %>%
  summarise(across("temp_Deep":"Salinity_Surface", list(mean)))
```

standard deviation
```{r}
env_stdev<-env_joined %>%
  group_by(Month.x,Station.x) %>%
  summarise(across("temp_Deep":"Salinity_Surface", list(sd)))
```

mean anomalies
```{r}
#add averages column
#make station-month column
env_joined$stationmonth<-paste(env_joined$Station.x, env_joined$Month.x,sep="-")
env_ave$stationmonth<-paste(env_ave$Station.x, env_ave$Month.x,sep="-")
env_stdev$stationmonth<-paste(env_stdev$Station.x, env_stdev$Month.x,sep="-")

env_joined_ave <- merge(env_joined,env_ave, by="stationmonth", all.x=TRUE)
env_joined_stdev <- merge(env_joined_ave,env_stdev, by="stationmonth", all.x=TRUE)

#calculate anomalies
env_joined_anom <- (env_joined_stdev[, 6:11] - env_joined_stdev[, 14:19])/env_joined_stdev[, 22:27]
```

```{r}
#convert station column to rownames
env_joined_anom$code=env_joined$code
env_joined_anom<-env_joined_anom %>% remove_rownames %>% column_to_rownames(var="code")
cor_matrix <- cor(env_joined_anom)
corrplot(cor_matrix, method = "number", type = "lower", tl.pos = 'd')
```


### Centroid Based
gap statistic: do this to determine number of clusters
```{r}
res.km<-eclust(env_joined_anom, FUNcluster = "kmeans", hc_metric = "euclidean", graph = FALSE)
fviz_gap_stat(res.km$gap_stat)

#silhouette plot
fviz_silhouette(res.km)
```
this determined that 10 clusters were optimal

elbow method
```{r}
fviz_nbclust(env_joined_anom, FUNcluster = kmeans, method = "wss") + labs(subtitle = "Elbow method") 
```
looks like there is an elbow at 2, 4, 6 and 9

cluster plot
```{r}
eclust(env_joined_anom, FUNcluster = "kmeans", k = 2, hc_metric = "euclidean")
```

### Heirarchical
```{r,fig.width=6,fig.height=8}
res.hclust <- eclust(env_joined_anom, FUNcluster = "hclust", hc_metric = "euclidean")
env_dend<-fviz_dend(res.hclust, rect = TRUE,repel=phylogenic)
env_dend
fviz_cluster(res.hclust, labelsize = 10)
```
detected 9 clusters

save dendogram
```{r, eval=FALSE}
setwd("/Users/hailaschultz/Dropbox/Schultz_Dissertation/Data_Analysis/Schultz_dissertation-2/figures")
ggsave(filename = "env_heirarchical_dend.png", plot = env_dend, width = 30, height = 17, device='png', dpi=700)
```

### Agglomerative clustering
```{r,fig.width=6,fig.height=8}
res.aclust <- eclust(env_joined_anom, FUNcluster = "hclust", hc_metric = "euclidean", hc_method = "ward.D2")
fviz_dend(res.aclust, rect = TRUE)
fviz_cluster(res.aclust, labelsize = 10)
```
the last two methods gave the same result

### visualize the strength of the variables
```{r}
df_clusters <- res.km$centers
res.km
```

```{r}
melt_df <- melt(df_clusters)

heatmap <- ggplot(melt_df, aes(Var2, Var1)) +
  scale_fill_continuous(type = "viridis", direction = -1) +
  geom_tile(aes(fill = value)) +
  geom_text(aes(label = round(value, 1))) +
  theme_bw() +
  ggtitle("Strength of Each of the Variables in the Clusters") +
  theme(plot.title = element_text(hjust = 0.5)) +
  labs(x="Variable", y="Cluster")
heatmap
```


### list of statons and clusters
```{r}
env_joined["Cluster"] <- res.km$cluster
df_out <- env_joined[order(-env_joined$Cluster), ]
knitr::kable(df_out)
```
save table

```{r, eval=FALSE}
write.csv(df_out, "/Users/hailaschultz/Dropbox/Schultz_Dissertation/Data_Analysis/Schultz_dissertation-2/output/env_cluster.csv", row.names=TRUE)
```


