---
title: "07_regime-shift_detection"
date: "2024-02-22"
output:
  html_document:
    toc: true
    toc_depth: 2
    toc_float: true
---

The purpose of this code is to attempt to detect temperature regime shifts in WOAC data

Read environmental data into R
```{r}
Environmental <- read.csv("/Users/hailaschultz/Dropbox/Schultz_Dissertation/Data_Analysis/Schultz_dissertation-2/data/WOAC_Chem_data_2014-2022_all-niskins.csv")
```

# Install packages
```{r}
library(rshift)
```

# Environmental Data Summary
recode months to match with zooplankton data
```{r}
unique(Environmental$Month)
Environmental$Month <- recode(Environmental$Month, 
                                 "Jul" = "JUL",
                         "Sep"="SEP",
                         "Oct"="SEP",
                         "Apr"="APR",
                         "May"="MAY",
                         "Nov"="NOV",
                         "Aug"="AUG",
                         "Jun"="JUL",
                         "OCT"="SEP")
#remove unnecessary sample dates
unique(Environmental$Date)
unique(Zooplankton$Date)
Environmental<-subset(Environmental,Date!="10/30/14")
Environmental<-subset(Environmental,Date!="10/31/14")
Environmental<-subset(Environmental,Date!="10/23/14")
Environmental<-subset(Environmental,Date!="10/29/14")
Environmental<-subset(Environmental,Date!="5/24/15")
Environmental<-subset(Environmental,Date!="11/18/15")
Environmental<-subset(Environmental,Date!="11/16/15")
Environmental<-subset(Environmental,Date!="11/17/15")
Environmental<-subset(Environmental,Date!="10/27/16")
Environmental<-subset(Environmental,Date!="5/3/17")
Environmental<-subset(Environmental,Date!="5/4/17")
Environmental<-subset(Environmental,Date!="10/18/17")
Environmental<-subset(Environmental,Date!="5/24/18")
Environmental<-subset(Environmental,Date!="10/18/18")
Environmental<-subset(Environmental,Date!="10/19/18")
Environmental<-subset(Environmental,Date!="5/23/19")
Environmental<-subset(Environmental,Date!="8/31/20")
Environmental<-subset(Environmental,Date!="10/1/21")
Environmental<-subset(Environmental,Date!="5/6/22")
Environmental<-subset(Environmental,Date!="10/14/22")
Environmental<-subset(Environmental,Date!="10/15/22")
Environmental<-subset(Environmental,Date!="9/29/21")
Environmental<-subset(Environmental,Date!="7/4/20")
Environmental<-subset(Environmental,Date!="9/30/20")
Environmental<-subset(Environmental,Date!="9/18/17")
```
# Get temperature data

summarize by depth category
```{r}
temp<-Environmental %>%
  group_by(Station,Date,Depth,Year,Month) %>%
  summarise(
    temp = mean(CTDTMP_DEG_C_ITS90))

#subset to deep samples
temp_deep<-subset(temp,Depth=="Deep")
#make unique column
temp_deep$code<-paste(temp_deep$Station, temp_deep$Year,temp_deep$Month,sep="-")
```

test out on P28
```{r}
temp_deep_P28<-subset(temp_deep,Station=="P28")
#convert date
temp_deep_P28 <- temp_deep_P28 %>% 
  mutate(date_convert = as.Date(Date, format = "%m/%d/%y"))
#order by date
temp_deep_P28 <- temp_deep_P28[order(temp_deep_P28$date_convert),]
```


Regime shift detection
```{r}
RSI_data<-Rodionov(temp_deep_P28, "temp", "date_convert",10, merge=TRUE)
```
visualize
```{r}
graph<-RSI_graph(RSI_data, "temp", "date_convert", "...10")
graph
```
It looks like I will need to look at anomalies due to the large variation in temperature

# Anomaly analysis
month and station averages
```{r}
temp_ave<-Environmental %>%
  group_by(Depth,Month,Station) %>%
  summarise(
    temp = mean(CTDTMP_DEG_C_ITS90),
    stdev = sd(CTDTMP_DEG_C_ITS90))

#subset to deep samples
temp_deep_ave<-subset(temp_ave,Depth=="Deep")

```

subset to P28
```{r}
temp_deep_ave_P28<-subset(temp_deep_ave,Station=="P28")
```

anomalies
```{r}
#add average column 
temp_deep_P28$Ave <- temp_deep_ave_P28$temp[match(temp_deep_P28$Month, temp_deep_ave_P28$Month)]

#add statndard deviation column
temp_deep_P28$stdev <- temp_deep_ave_P28$stdev[match(temp_deep_P28$Month, temp_deep_ave_P28$Month)]

# calculate anomalies
temp_deep_P28$anomaly<-(temp_deep_P28$temp-temp_deep_P28$Ave)/temp_deep_P28$stdev
```

Regime shift detection
```{r}
RSI_data<-Rodionov(temp_deep_P28, "anomaly", "date_convert",5, merge=TRUE)
```
visualize
```{r}
graph<-RSI_graph(RSI_data, "anomaly", "date_convert", "...13")
graph
```