---
title: "08_zoop-cluster"
date: "2024-02-22"
output:
  html_document:
    toc: true
    toc_depth: 2
    toc_float: true
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,         # Display code chunks
  eval = TRUE,        # Evaluate code chunks
  warning = FALSE,     # Hide warnings
  message = FALSE,     # Hide messages
  fig.width = 10,       # Set plot width in inches
  fig.height = 6,      # Set plot height in inches
  fig.align = "center" # Align plots to the center
)
```

load packages
```{r}
library(ggrepel)
library("scales")   
library(corrplot)
library(factoextra)
```

Read zooplanton data into R
```{r}
Zooplankton <- read.csv("/Users/hailaschultz/Dropbox/Schultz_Dissertation/Data_Analysis/Schultz_dissertation-2/data/WOAC_Species_Densities_2014-2022.csv")
```

# Zooplankton Data
subset to stations sampled during WOAC cruises
```{r}
unique(Zooplankton$Station)
Zoop_sub<-subset(Zooplankton,Station!="PEFS1d")
Zoop_sub<-subset(Zoop_sub,Station!="PEFN1d")
Zoop_sub<-subset(Zoop_sub,Station!="PEFN2")
Zoop_sub<-subset(Zoop_sub,Station!="PEFS2")
Zoop_sub<-subset(Zoop_sub,Station!="P381")
Zoop_sub<-subset(Zoop_sub,Station!="P105")
Zoop_sub<-subset(Zoop_sub,Station!="P136")
Zoop_sub<-subset(Zoop_sub,Station!="P7")
Zoop_sub<-subset(Zoop_sub,Station!="P123")
Zoop_sub<-subset(Zoop_sub,Station!="P132")
unique(Zoop_sub$Date)
#remove unnecessary dates
Zoop_sub<-subset(Zoop_sub,Date!="10/30/14")
Zoop_sub<-subset(Zoop_sub,Date!="10/31/14")
Zoop_sub<-subset(Zoop_sub,Date!="10/29/14")
Zoop_sub<-subset(Zoop_sub,Date!="9/14/22")
Zoop_sub<-subset(Zoop_sub,Date!="9/16/22")
Zoop_sub<-subset(Zoop_sub,Date!="9/13/22")
Zoop_sub<-subset(Zoop_sub,Date!="9/12/22")
Zoop_sub<-subset(Zoop_sub,Date!="11/16/15")
Zoop_sub<-subset(Zoop_sub,Date!="3/17/16")
Zoop_sub<-subset(Zoop_sub,Date!="3/18/16")
Zoop_sub<-subset(Zoop_sub,Date!="11/18/15")
#recode p4
Zoop_sub$Station <- recode(Zoop_sub$Station, 
                                 "p4" = "P4")
#recode month names to match
unique(Zoop_sub$Month)
Zoop_sub$Month <- recode(Zoop_sub$Month, 
                                 "Jul" = "JUL",
                         "Sep"="SEP",
                         "OCT"="SEP")
#remove ethanol samples
Zoop_sub<-Zoop_sub[!grepl('_EtOH', Zoop_sub$Code),]
#remove oblique samples
Zoop_sub<-subset(Zoop_sub,Tow.Type!="Oblique")
```

add up multiple lines per station
```{r}
Zoop_sub <- Zoop_sub %>%
  group_by(Code,Station,Basin,Year,Month,Genus.species) %>%
  summarise(
    zoop_density = sum(Density....m3.))
```

change from long to wide format
```{r}
colnames(Zoop_sub)
Zoop_sub_wide<-dcast(Zoop_sub, Code+Station+Basin+Year+Month~ Genus.species,value.var = "zoop_density")
#make sure factors are formatted right
Zoop_sub_wide$Year=as.factor(Zoop_sub_wide$Year)
Zoop_sub_wide$Station=as.factor(Zoop_sub_wide$Station)
Zoop_sub_wide$Month=as.factor(Zoop_sub_wide$Month)
Zoop_sub_wide$Basin=as.factor(Zoop_sub_wide$Basin)

#make unique column
Zoop_sub_wide$code<-paste(Zoop_sub_wide$Station, Zoop_sub_wide$Year,Zoop_sub_wide$Month,sep="-")
unique(Zoop_sub$Code)
```
remove non-data columns, convert to proportionas, and arcsine sqrt transformation
```{r}
RE2<- Zoop_sub_wide[,6:ncol(Zoop_sub_wide)]
RE2 <- RE2[ -c(211) ]
#replace N/A with 0
RE2[is.na(RE2)] <- 0
#convert to proportions
RE2<-RE2/rowSums(RE2)
#arcsine sqrt transformation
RE2<-asin(sqrt(RE2))
RE3<-as.matrix(RE2)
```
add back in code as rownames
```{r}
RE2$code = Zoop_sub_wide$code
#remove duplicate rows
RE2<-RE2[-c(167,91,168),]
RE2<-RE2 %>% remove_rownames %>% column_to_rownames(var="code")
#remove columns with all zeroes
RE2<-RE2[,-(which(colSums(RE2)==0))]
#remove columns with fewer than 9 rows with nonzero numbers
non_zero_counts <- colSums(RE2 != 0)
RE2 <- RE2[, non_zero_counts >= 9]

```

### Centroid Based
gap statistic: do this to determine number of clusters
```{r}
res.km<-eclust(RE2, FUNcluster = "kmeans", hc_metric = "euclidean", graph = FALSE)
fviz_gap_stat(res.km$gap_stat)

#silhouette plot
fviz_silhouette(res.km)
```
It looks like nine clusters are identified
elbow method
```{r}
fviz_nbclust(RE2, FUNcluster = kmeans, method = "wss") + labs(subtitle = "Elbow method") 
```
There are elbows at 2, 5 and 9

cluster plot
```{r}
eclust(env_joined_standard, FUNcluster = "kmeans", k = 9, hc_metric = "euclidean")
```
### Heirarchical
```{r,fig.width=6,fig.height=8}
res.hclust <- eclust(RE2, FUNcluster = "hclust", hc_metric = "euclidean")
zoop_dend<-fviz_dend(res.hclust, rect = TRUE,repel=phylogenic)
zoop_dend
fviz_cluster(res.hclust, labelsize = 10)

```
save dendogram
```{r, eval=FALSE}
setwd("/Users/hailaschultz/Dropbox/Schultz_Dissertation/Data_Analysis/Schultz_dissertation-2/figures")
ggsave(filename = "zoop_heirarchical_dend.png", plot = zoop_dend, width = 30, height = 17, device='png', dpi=700)
```
### Agglomerative clustering
```{r,fig.width=6,fig.height=8}
res.aclust <- eclust(RE2, FUNcluster = "hclust", hc_metric = "euclidean", hc_method = "ward.D2")
fviz_dend(res.aclust, rect = TRUE)
fviz_cluster(res.aclust, labelsize = 10)
```
same results

### visualize the strength of the variables
```{r}
df_clusters <- res.km$centers
res.km
```

```{r}
melt_df <- melt(df_clusters)
melt_df
heatmap <- ggplot(melt_df, aes(Var2, Var1)) +
  scale_fill_continuous(type = "viridis", direction = -1) +
  geom_tile(aes(fill = value))  +
  theme_bw() +
  ggtitle("Strength of Each of the Variables in the Clusters") +
  theme(plot.title = element_text(hjust = 0.5)) +
  labs(x="Variable", y="Cluster")
heatmap
```

### list of statons and clusters
```{r}
RE2["Cluster"] <- res.km$cluster
df_out <- RE2[order(-RE2$Cluster), ]
knitr::kable(df_out)
```
save table

```{r, eval=FALSE}
write.csv(df_out, "/Users/hailaschultz/Dropbox/Schultz_Dissertation/Data_Analysis/Schultz_dissertation-2/output/zoop_cluster.csv", row.names=TRUE)
```


