---
title: "10_anomaly_NMDS"
date: "2024-03-05"
output:
  html_document:
    toc: true
    toc_depth: 2
    toc_float: true
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,         # Display code chunks
  eval = TRUE,        # Evaluate code chunks
  warning = FALSE,     # Hide warnings
  message = FALSE,     # Hide messages
  fig.width = 10,       # Set plot width in inches
  fig.height = 6,      # Set plot height in inches
  fig.align = "center" # Align plots to the center
)
```
load packages
```{r}
library(ggrepel)
library("scales") 
library(dplyr)
library(tidyr)
library(reshape2)
library(vegan)
library(EcolUtils)
```

Read zooplanton data into R
```{r}
Zooplankton <- read.csv("/Users/hailaschultz/Dropbox/Schultz_Dissertation/Data_Analysis/Schultz_dissertation-2/data/WOAC_Species_Densities_2014-2022.csv")
```
Read environmental data into R
```{r}
Environmental <- read.csv("/Users/hailaschultz/Dropbox/Schultz_Dissertation/Data_Analysis/Schultz_dissertation-2/data/WOAC_Chem_data_2014-2022_all-niskins.csv")
```

# Environmental Data Summary
recode months to match with zooplankton data
```{r}
unique(Environmental$Month)
Environmental$Month <- recode(Environmental$Month, 
                                 "Jul" = "JUL",
                         "Sep"="SEP",
                         "Oct"="SEP",
                         "Apr"="APR",
                         "May"="MAY",
                         "Nov"="NOV",
                         "Aug"="AUG",
                         "Jun"="JUL",
                         "OCT"="SEP")
#remove unnecessary sample dates
unique(Environmental$Date)
unique(Zooplankton$Date)
Environmental<-subset(Environmental,Date!="10/30/14")
Environmental<-subset(Environmental,Date!="10/31/14")
Environmental<-subset(Environmental,Date!="10/23/14")
Environmental<-subset(Environmental,Date!="10/29/14")
Environmental<-subset(Environmental,Date!="5/24/15")
Environmental<-subset(Environmental,Date!="11/18/15")
Environmental<-subset(Environmental,Date!="11/16/15")
Environmental<-subset(Environmental,Date!="11/17/15")
Environmental<-subset(Environmental,Date!="10/27/16")
Environmental<-subset(Environmental,Date!="5/3/17")
Environmental<-subset(Environmental,Date!="5/4/17")
Environmental<-subset(Environmental,Date!="10/18/17")
Environmental<-subset(Environmental,Date!="5/24/18")
Environmental<-subset(Environmental,Date!="10/18/18")
Environmental<-subset(Environmental,Date!="10/19/18")
Environmental<-subset(Environmental,Date!="5/23/19")
Environmental<-subset(Environmental,Date!="8/31/20")
Environmental<-subset(Environmental,Date!="10/1/21")
Environmental<-subset(Environmental,Date!="5/6/22")
Environmental<-subset(Environmental,Date!="10/14/22")
Environmental<-subset(Environmental,Date!="10/15/22")
Environmental<-subset(Environmental,Date!="9/29/21")
Environmental<-subset(Environmental,Date!="7/4/20")
Environmental<-subset(Environmental,Date!="9/30/20")
Environmental<-subset(Environmental,Date!="9/18/17")
```

Temp: summarize by depth category
```{r}
temp<-Environmental %>%
  group_by(Station,Date,Depth,Year,Month) %>%
  summarise(
    temp = mean(CTDTMP_DEG_C_ITS90))

#subset to surface samples
temp_surf<-subset(temp,Depth=="Surface")
#make unique column
temp_surf$code<-paste(temp_surf$Station, temp_surf$Year,temp_surf$Month,sep="-")
```

pH: remove missing rows and remove weird values
```{r}
pH<-Environmental %>% drop_na(pH)
pH <- subset(pH, pH != 2130.2)
```
summarize by depth category
```{r}
pH<-pH %>%
  group_by(Station,Date,Depth,Year,Month) %>%
  summarise(
    pH = mean(pH))
#subset to surface samples
pH_surf<-subset(pH,Depth=="Surface")
#make unique column
pH_surf$code<-paste(pH_surf$Station, pH_surf$Year,pH_surf$Month,sep="-")
```

O2: remove missing rows and remove weird values
```{r}
Oxygen<-Environmental %>% drop_na(O2.in.mg.l)
```

summarize by depth category
```{r}
Oxygen<-Oxygen %>%
  group_by(Station,Date,Depth,Year,Month) %>%
  summarise(
    Oxygen = mean(O2.in.mg.l))
#subset to surface samples
Oxygen_surf<-subset(Oxygen,Depth=="Surface")
#make unique column
Oxygen_surf$code<-paste(Oxygen_surf$Station, Oxygen_surf$Year,Oxygen_surf$Month,sep="-")
```

Merge environmental tables
```{r}
joined <- merge(temp_surf, pH_surf, by.x = "code", 
             by.y = "code", all.x = TRUE, all.y = TRUE)
joined <- merge(joined, Oxygen_surf, by.x = "code", 
             by.y = "code", all.x = TRUE, all.y = TRUE)
#remove excess columns
env_joined <- joined[ -c(8:12,14:18) ]
```


# Zooplankton Data
subset to stations sampled during WOAC cruises
```{r}
unique(Zooplankton$Station)
Zoop_sub<-subset(Zooplankton,Station!="PEFS1d")
Zoop_sub<-subset(Zoop_sub,Station!="PEFN1d")
Zoop_sub<-subset(Zoop_sub,Station!="PEFN2")
Zoop_sub<-subset(Zoop_sub,Station!="PEFS2")
Zoop_sub<-subset(Zoop_sub,Station!="P381")
Zoop_sub<-subset(Zoop_sub,Station!="P105")
Zoop_sub<-subset(Zoop_sub,Station!="P136")
Zoop_sub<-subset(Zoop_sub,Station!="P7")
Zoop_sub<-subset(Zoop_sub,Station!="P383")
Zoop_sub<-subset(Zoop_sub,Station!="P132")
unique(Zoop_sub$Date)
#remove unnecessary dates
Zoop_sub<-subset(Zoop_sub,Date!="10/30/14")
Zoop_sub<-subset(Zoop_sub,Date!="10/31/14")
Zoop_sub<-subset(Zoop_sub,Date!="10/29/14")
Zoop_sub<-subset(Zoop_sub,Date!="9/14/22")
Zoop_sub<-subset(Zoop_sub,Date!="9/16/22")
Zoop_sub<-subset(Zoop_sub,Date!="9/13/22")
Zoop_sub<-subset(Zoop_sub,Date!="9/12/22")
Zoop_sub<-subset(Zoop_sub,Date!="11/16/15")
Zoop_sub<-subset(Zoop_sub,Date!="3/17/16")
Zoop_sub<-subset(Zoop_sub,Date!="3/18/16")
Zoop_sub<-subset(Zoop_sub,Date!="11/18/15")
#recode p4
Zoop_sub$Station <- recode(Zoop_sub$Station, 
                                 "p4" = "P4")
#recode month names to match
unique(Zoop_sub$Month)
Zoop_sub$Month <- recode(Zoop_sub$Month, 
                                 "Jul" = "JUL",
                         "Sep"="SEP",
                         "OCT"="SEP")
#remove ethanol samples
Zoop_sub<-Zoop_sub[!grepl('_EtOH', Zoop_sub$Code),]
#remove oblique samples
Zoop_sub<-subset(Zoop_sub,Tow.Type!="Oblique")
```

add up multiple lines per station
```{r}
Zoop_sub <- Zoop_sub %>%
  group_by(Code,Station,Basin,Year,Month,Genus.species) %>%
  summarise(
    zoop_density = sum(Density....m3.))
```


change from long to wide format
```{r}
colnames(Zoop_sub)
Zoop_sub_wide<-dcast(Zoop_sub, Code+Station+Basin+Year+Month~ Genus.species,value.var = "zoop_density")
#make sure factors are formatted right
Zoop_sub_wide$Year=as.factor(Zoop_sub_wide$Year)
Zoop_sub_wide$Station=as.factor(Zoop_sub_wide$Station)
Zoop_sub_wide$Month=as.factor(Zoop_sub_wide$Month)
Zoop_sub_wide$Basin=as.factor(Zoop_sub_wide$Basin)
```


```{r}
#make unique column
Zoop_sub_wide$code<-paste(Zoop_sub_wide$Station, Zoop_sub_wide$Year,Zoop_sub_wide$Month,sep="-")
unique(Zoop_sub$Code)
```

# Combine env and zoop
```{r}
Zoop_env <- merge(env_joined,Zoop_sub_wide, by.x = "code", 
             by.y = "code", all.x = TRUE, all.y = TRUE)
#remove excess columns
Zoop_env <- Zoop_env[ -c(11,13:14) ]
#remove rows with NA
Zoop_env<-Zoop_env[!is.na(Zoop_env$Year.x),]
```


# NMDS
remove non-data columns, convert to proportionas, and arcsine sqrt transformation
```{r}
#subset to only cluster 1 stations (P4, P12, P402)
Zoop_env<-subset(Zoop_env, Station.x != "P22")
Zoop_env<-subset(Zoop_env, Station.x != "P38")
Zoop_env<-subset(Zoop_env, Station.x != "P8")
Zoop_env<-subset(Zoop_env, Station.x != "P28")
RE2<- Zoop_env[,12:ncol(Zoop_env)]
#replace N/A with 0
RE2[is.na(RE2)] <- 0
#remove columns with fewer than 9 rows with nonzero numbers
non_zero_counts <- colSums(RE2 != 0)
RE2 <- RE2[, non_zero_counts >= 9]

RE3<-as.matrix(RE2)
```


# Anomaly analysis
month and station averages
```{r}
Zoop_env[is.na(Zoop_env)] <- 0
zoop_ave<-Zoop_env %>%
  group_by(Month.x,Station.x) %>%
  summarise(across("ACARTIA":"XENACANTHOMYSIS PSEUDOMACROPSIS", list(mean)))
```


anomalies
```{r}
#add averages column
#make station-month column
Zoop_env$stationmonth<-paste(Zoop_env$Station.x, Zoop_env$Month.x,sep="-")
zoop_ave$stationmonth<-paste(zoop_ave$Station.x, zoop_ave$Month.x,sep="-")

Zoop_env_ave <- merge(Zoop_env,zoop_ave, by="stationmonth", all.x=TRUE)


####need to figure out how to calculate anomalies still...

temp_deep_P28$Ave <- temp_deep_ave_P28$temp[match(temp_deep_P28$Month, temp_deep_ave_P28$Month)]

#add statndard deviation column
temp_deep_P28$stdev <- temp_deep_ave_P28$stdev[match(temp_deep_P28$Month, temp_deep_ave_P28$Month)]

# calculate anomalies
temp_deep_P28$anomaly<-(temp_deep_P28$temp-temp_deep_P28$Ave)/temp_deep_P28$stdev
```
