---
title: "10_Mol16S-diversity"
date: "2024-03-07"
output:
  html_document:
    toc: true
    toc_depth: 2
    toc_float: true
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,         # Display code chunks
  eval = TRUE,        # Evaluate code chunks
  warning = FALSE,     # Hide warnings
  message = FALSE,     # Hide messages
  fig.width = 10,       # Set plot width in inches
  fig.height = 6,      # Set plot height in inches
  fig.align = "center" # Align plots to the center
)
```

```{r}
library(stringr)
```

read data into R
note: I first had to convert the taxonomy table from text into csv for this to work
```{r}
MOL <- read.csv("/Users/hailaschultz/Dropbox/Schultz_Dissertation/Data_Analysis/Schultz_dissertation-2/data/Barcode_2018-2020/ASV_count_tables/Mol16S_ASVs_counts.tsv", sep = "\t")

MOL_tax <- read.csv("/Users/hailaschultz/Dropbox/Schultz_Dissertation/Data_Analysis/Schultz_dissertation-2/data/Barcode_2018-2020/ASV2Taxonomy_Tables/RV_Mol16S_asvTaxonomyTable.csv")
```

add taxonomy assignments to MOL dataframe
```{r}
#rename column
MOL<-MOL %>% 
  rename(
    ASV = x,
    )

MOL <- merge(MOL_tax,MOL, by="ASV", all.x=FALSE)
```

Convert to present (1) vs absent (0)
```{r}
#remove non-species rows
MOL_complete<- MOL[complete.cases(MOL), ]
#remove non-zooplankton taxa
unique(MOL_complete$Phylum)
MOL_complete <- droplevels(MOL_complete[!MOL_complete$Phylum == 'Chlorophyta',])
MOL_complete <- droplevels(MOL_complete[!MOL_complete$Phylum == 'Bacillariophyta',])
MOL_complete <- droplevels(MOL_complete[!MOL_complete$Phylum == 'Hemichordata',])
```

```{r}
#remove excess columns including oblique tows
MOL_pa <- MOL_complete[ -c(1:2,4:7) ]
MOL_pa<-melt(MOL_pa, id.vars=c("Phylum","Species"))
```

```{r}
#add up multiple lines per station
MOL_pa <- MOL_pa %>%
  group_by(Phylum,Species,variable) %>%
  summarise(
    value = sum(value))

MOL_pa<-MOL_pa %>%
  mutate(across(where(is.numeric), ~ifelse(.x >0, 1, 0)))
```
split code string up
```{r}
MOL_pa$code<-MOL_pa$variable
MOL_pa$variable<-str_split_fixed(MOL_pa$variable,"_",4)
```

remove oblique and other replicates
```{r}
MOL_pa<-subset(MOL_pa,variable[,4]!="Ob")
MOL_pa<-subset(MOL_pa,variable[,4]!="ObA")
MOL_pa<-subset(MOL_pa,variable[,4]!="ObB")
MOL_pa<-subset(MOL_pa,variable[,4]!="1B")
MOL_pa<-subset(MOL_pa,variable[,4]!="2B")
MOL_pa<-subset(MOL_pa,variable[,4]!="3B")
MOL_pa<-subset(MOL_pa,variable[,4]!="")


MOL_pa<-subset(MOL_pa,variable[,3]!="P105")
MOL_pa<-subset(MOL_pa,variable[,3]!="P123")
MOL_pa<-subset(MOL_pa,variable[,3]!="P136")
MOL_pa<-subset(MOL_pa,variable[,3]!="P381")
MOL_pa<-subset(MOL_pa,variable[,3]!="P7")

MOL_pa<-subset(MOL_pa,variable[,2]!="2010")
```

merge replicates
```{r}
MOL_pa$sample<-paste(MOL_pa$variable[,2],MOL_pa$variable[,3],sep="")

MOL_pa <- MOL_pa %>%
  group_by(Phylum,Species,sample) %>%
  summarise(
    value = sum(value))

MOL_pa<-MOL_pa %>%
  mutate(across(where(is.numeric), ~ifelse(.x >0, 1, 0)))
```

summarize by phylum
```{r}
MOL_pa_sum<-MOL_pa %>%
  group_by(Phylum, sample) %>%
  summarise(
    ASVs = sum(value))
```

```{r}
MOL_pa_sum<-transform(MOL_pa_sum, Year = substr(sample, 1, 2), Month = substr(sample, 3, 4),Station=substr(sample, 5,7))
```

make bar graph
```{r}
MOL_diversity<-ggplot(MOL_pa_sum, aes(x = Station, y = ASVs, fill = Phylum)) +
  geom_bar(stat = "identity") +
  labs(title = "Mol16S all species",
       x = "Station",
       y = "number of species") +
  theme_minimal()+
  scale_x_discrete(guide = guide_axis(angle = 90))+ facet_wrap(~ Month*Year) 
MOL_diversity
```
Average by station

```{r}
MOL_pa_ave<-MOL_pa_sum %>%
  group_by(Phylum, Station) %>%
  summarise(
    ASVs = mean(ASVs),
    sd= sd(ASVs))
```

```{r}
MOL_diversity_ave<-ggplot(MOL_pa_ave, aes(x = Station, y = ASVs, fill = Phylum)) +
  geom_bar(stat = "identity") +
  labs(title = "Mol all species",
       x = "Station",
       y = "number of species") +
  theme_minimal()+
  scale_x_discrete(guide = guide_axis(angle = 90)) 
MOL_diversity_ave
```
save plot
```{r, eval=FALSE}
setwd("/Users/hailaschultz/Dropbox/Schultz_Dissertation/Data_Analysis/Schultz_dissertation-2/figures")
ggsave(filename = "MOL_diversity.png", plot = MOL_diversity, width = 10, height = 8, device='png', dpi=700)
```


average per station and month

subset
```{r}
MOL_pa_gastropod<-subset(MOL_pa_sum,Phylum=="Mollusca")
MOL_pa_gastropod$Station<-as.character(MOL_pa_gastropod$Station)
```


```{r}
MOL_pa_gastropod_mean<-MOL_pa_gastropod %>%
  group_by(Station,Month) %>%
  summarise(
    mean = mean(ASVs),
    sd= sd(ASVs))
```
```{r}
ggplot(MOL_pa_gastropod_mean) +
    geom_bar( aes(x=Station, y=mean), stat="identity", fill="skyblue", alpha=0.7) +
    geom_errorbar( aes(x=Station, ymin=mean-sd, ymax=mean+sd), width=0.4, alpha=0.9, size=1.3)+facet_wrap(~ Month) 
```
```{r}
library(ggpubr)
ggboxplot(MOL_pa_gastropod, x = "Month", y = "ASVs", color = "Station")

ggline(MOL_pa_gastropod, x = "Station", y = "ASVs", color = "Month",
       add = c("mean_se", "dotplot"))
```
```{r}
res.aov2 <- aov(ASVs ~ Month+Station+Year, data = MOL_pa_gastropod)
summary(res.aov2)
```
no statistical difference in diversity



```{r}
MOL_pa_gastropod_station<-MOL_pa_gastropod %>%
  group_by(Station) %>%
  summarise(
    mean = mean(ASVs),
    sd= sd(ASVs))
```

```{r}
ggplot(MOL_pa_gastropod_station) +
    geom_bar( aes(x=Station, y=mean), stat="identity", fill="skyblue", alpha=0.7) +
    geom_errorbar( aes(x=Station, ymin=mean-sd, ymax=mean+sd), width=0.4, alpha=0.9, size=1.3)
```

recode factors
```{r}
MOL_pa_gastropod$Year <- recode(MOL_pa_gastropod$Year, 
                                 "20" = "2020",
                         "19"="2019",
                         "18"="2018")

MOL_pa_gastropod$Month <- recode(MOL_pa_gastropod$Month, 
                                 "04" = "APR",
                         "07"="JUL",
                         "09"="SEP")

MOL_pa_gastropod$Station <- recode(MOL_pa_gastropod$Station, 
                                 "P40" = "P402")
MOL_pa_gastropod$code<-paste(MOL_pa_gastropod$Station,MOL_pa_gastropod$Year,MOL_pa_gastropod$Month,sep="-")
```

add in environmental
```{r}
MOL_pa_gastropod <- merge(Env_subset,MOL_pa_gastropod, by.x = "code", 
             by.y = "code", all.x = TRUE, all.y = TRUE)

#remove rows with NA
MOL_pa_gastropod<-MOL_pa_gastropod[!is.na(MOL_pa_gastropod$sample),]
MOL_pa_gastropod<-MOL_pa_gastropod[!is.na(MOL_pa_gastropod$"Date"),]
```

```{r}
colnames(MOL_pa_gastropod)
ggplot(MOL_pa_gastropod,aes(x=OXYGEN_MG_L_1_Deep,y=ASVs))+geom_point()+
  stat_smooth(method = "lm")+
  labs(
       x = "Deep Disolved Oxygen (mg L-1)",
       y = "Number of Mollusk Species")+theme_bw()+theme(panel.grid.major = element_blank(),
                                                                                           panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))

model2 <- lm(ASVs ~ OXYGEN_MG_L_1_Deep, data = MOL_pa_gastropod)
summary(model2)
plot(model2, panel = panel.smooth)
```

barplot
just July 2020
```{r}
Mol_pa_jul2020<-subset(MOL_pa_sum,Year=="20")
Mol_pa_jul2020<-subset(Mol_pa_jul2020,Month=="09")
```
summarize by phylum
```{r}
Mol_pa_jul2020<-Mol_pa_jul2020 %>%
  group_by(Phylum, Station) %>%
  summarise(
    ASVs = sum(ASVs))
```

```{r}
#change factor order
Mol_pa_jul2020$Station <- factor(Mol_pa_jul2020$Station, levels = c("P22", "P4", "P8","P28","P12","P40","P38"))
Mol_pa_jul2020$Station <- recode(Mol_pa_jul2020$Station, 
                                 "P22" = "Juan de Fuca",
                         "P4"="Whidbey Basin",
                         "P8"="HC Sill",
                         "P28"="Main Basin",
                         "P12"="Hood Canal 1",
                         "P40"="Hood Canal 2",
                         "P38"="South Sound")

```

```{r}

Mol_barplot<-ggplot(Mol_pa_jul2020, aes(x = Station, y = ASVs, fill = Phylum)) +
  geom_bar(stat = "identity") +
  labs(
       x = "Station",
       y = "Number of Species") +
  theme_minimal()+
  scale_x_discrete(guide = guide_axis(angle = 90)) +theme_bw()+theme(panel.grid.major = element_blank(),
                                                                                           panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))+ scale_colour_manual(values=c("#F8766D", "#E68613", "#CD9600", "#7CAE00", "#0CB702",  "#00C19A", "#00BFC4", "#00A9FF", "#8494FF")) +
  scale_fill_manual(values=c("#F8766D", "#E68613", "#CD9600", "#7CAE00", "#0CB702",  "#00C19A", "#00BFC4", "#00A9FF", "#8494FF"))+ylim(0,50)+theme(axis.text=element_text(size=15),
        axis.title=element_text(size=15,face="bold"))
Mol_barplot
```

save plot
```{r, eval=FALSE}
setwd("/Users/hailaschultz/Dropbox/Schultz_Dissertation/Data_Analysis/Schultz_dissertation-2/figures")
ggsave(filename = "Mol_barplot.png", plot = Mol_barplot, width = 10, height = 8, device='png', dpi=700)
```