---
title: "11_MORPH-diversity"
date: "2024-03-07"
output:
  html_document:
    toc: true
    toc_depth: 2
    toc_float: true
---

# Keister lab data

Read zooplanton data into R
```{r}
Zooplankton <- read.csv("/Users/hailaschultz/Dropbox/Schultz_Dissertation/Data_Analysis/Schultz_dissertation-2/data/WOAC_Species_Densities_2014-2022.csv")
```

subset to stations sampled during WOAC cruises
```{r}
unique(Zooplankton$Station)
Zoop_sub<-subset(Zooplankton,Station!="PEFS1d")
Zoop_sub<-subset(Zoop_sub,Station!="PEFN1d")
Zoop_sub<-subset(Zoop_sub,Station!="PEFN2")
Zoop_sub<-subset(Zoop_sub,Station!="PEFS2")
Zoop_sub<-subset(Zoop_sub,Station!="P381")
Zoop_sub<-subset(Zoop_sub,Station!="P105")
Zoop_sub<-subset(Zoop_sub,Station!="P136")
Zoop_sub<-subset(Zoop_sub,Station!="P7")
Zoop_sub<-subset(Zoop_sub,Station!="P123")
Zoop_sub<-subset(Zoop_sub,Station!="P132")
unique(Zoop_sub$Date)
#remove unnecessary dates
Zoop_sub<-subset(Zoop_sub,Date!="10/30/14")
Zoop_sub<-subset(Zoop_sub,Date!="10/31/14")
Zoop_sub<-subset(Zoop_sub,Date!="10/29/14")
Zoop_sub<-subset(Zoop_sub,Date!="9/14/22")
Zoop_sub<-subset(Zoop_sub,Date!="9/16/22")
Zoop_sub<-subset(Zoop_sub,Date!="9/13/22")
Zoop_sub<-subset(Zoop_sub,Date!="9/12/22")
Zoop_sub<-subset(Zoop_sub,Date!="11/16/15")
Zoop_sub<-subset(Zoop_sub,Date!="3/17/16")
Zoop_sub<-subset(Zoop_sub,Date!="3/18/16")
Zoop_sub<-subset(Zoop_sub,Date!="11/18/15")
#recode p4
Zoop_sub$Station <- recode(Zoop_sub$Station, 
                                 "p4" = "P4")
#recode month names to match
unique(Zoop_sub$Month)
Zoop_sub$Month <- recode(Zoop_sub$Month, 
                                 "Jul" = "JUL",
                         "Sep"="SEP",
                         "OCT"="SEP")
#remove ethanol samples
Zoop_sub<-Zoop_sub[!grepl('_EtOH', Zoop_sub$Code),]
#remove oblique samples
Zoop_sub<-subset(Zoop_sub,Tow.Type!="Oblique")
```

add up multiple lines per station
```{r}
Zoop_sub <- Zoop_sub %>%
  group_by(Code,Station,Basin,Year,Month,Broad.Group,Genus.species) %>%
  summarise(
    zoop_density = sum(Density....m3.))
```

convert to 1 and o
```{r}
MORPH_pa<-Zoop_sub %>%
  mutate(across(where(is.numeric), ~ifelse(.x >0, 1, 0)))
```

summarize by broad group
```{r}
MORPH_pa_sum<-MORPH_pa %>%
  group_by(Broad.Group, Year,Month,Station) %>%
  summarise(
    species = sum(zoop_density))
```

subset to 2018
```{r}
MORPH_pa_sum<-subset(MORPH_pa_sum,Year!="2014")
MORPH_pa_sum<-subset(MORPH_pa_sum,Year!="2015")
MORPH_pa_sum<-subset(MORPH_pa_sum,Year!="2016")
MORPH_pa_sum<-subset(MORPH_pa_sum,Year!="2017")
MORPH_pa_sum<-subset(MORPH_pa_sum,Year!="2021")
MORPH_pa_sum<-subset(MORPH_pa_sum,Year!="2022")
```


make bar graph
```{r}
MORPH_diversity<-ggplot(MORPH_pa_sum, aes(x = Station, y = species, fill = Broad.Group)) +
  geom_bar(stat = "identity") +
  labs(title = "Mol16S all species",
       x = "Station",
       y = "number of species") +
  theme_minimal()+
  scale_x_discrete(guide = guide_axis(angle = 90)) + facet_wrap(~ Month*Year)
MORPH_diversity
```


# data from Sean's run
read data into R
note: I first had to convert the taxonomy table from text into csv for this to work
```{r}
Morph <- read.csv("/Users/hailaschultz/Dropbox/Schultz_Dissertation/Data_Analysis/Schultz_dissertation-2/data/Barcode_2018-2020/ASV_count_tables/Morphdensity_ASVs_counts.tsv", sep = "\t")

Morph_tax <- read.csv("/Users/hailaschultz/Dropbox/Schultz_Dissertation/Data_Analysis/Schultz_dissertation-2/data/Barcode_2018-2020/ASV2Taxonomy_Tables/morphology_asvTaxonomyTable.csv")
```

add taxonomy assignments to COI dataframe
```{r}
#rename column
Morph<-Morph %>% 
  rename(
    ASV = x,
    )

Morph <- merge(Morph_tax,Morph, by="ASV", all.x=FALSE)
```


Convert to present (1) vs absent (0)
```{r}
#remove non-species rows
Morph_complete<- Morph[complete.cases(Morph), ]
#remove non-zooplankton taxa
unique(Morph_complete$Phylum)

Morph_pa<-Morph_complete %>%
  mutate(across(where(is.numeric), ~ifelse(.x >0, 1, 0)))

```

wide to long format
```{r}
#remove exces columns
Morph_pa <- Morph_pa[ -c(2,4:7) ]
Morph_pa<-melt(Morph_pa, id.vars=c("Phylum", "Species","ASV"))
```
split code string up
```{r}
Morph_pa$code<-Morph_pa$variable
Morph_pa$variable<-str_split_fixed(Morph_pa$variable,"_",4)
```
remove oblique and other replicates
```{r}
Morph_pa<-subset(Morph_pa,variable[,4]!="Ob")
Morph_pa<-subset(Morph_pa,variable[,4]!="ObA")
Morph_pa<-subset(Morph_pa,variable[,4]!="ObB")
Morph_pa<-subset(Morph_pa,variable[,4]!="1B")
Morph_pa<-subset(Morph_pa,variable[,4]!="2B")
Morph_pa<-subset(Morph_pa,variable[,4]!="3B")
Morph_pa<-subset(Morph_pa,variable[,4]!="")


Morph_pa<-subset(Morph_pa,variable[,3]!="P105")
Morph_pa<-subset(Morph_pa,variable[,3]!="P123")
Morph_pa<-subset(Morph_pa,variable[,3]!="P136")
Morph_pa<-subset(Morph_pa,variable[,3]!="P381")
Morph_pa<-subset(Morph_pa,variable[,3]!="P7")

Morph_pa<-subset(Morph_pa,variable[,2]!="2010")
```

summarize by phylum
```{r}
Morph_pa_sum<-Morph_pa %>%
  group_by(Phylum, variable) %>%
  summarise(
    ASVs = sum(value))
```
```{r}
Morph_pa_sum<-transform(Morph_pa_sum, Year = substr(variable[,2], 1, 2), Month = substr(variable[,2], 3, 4))
Morph_pa_sum$Station<-Morph_pa_sum$variable[,3]
```


make bar graph
```{r}
Morph_diversity<-ggplot(Morph_pa_sum, aes(x = Station, y = ASVs, fill = Phylum)) +
  geom_bar(stat = "identity") +
  labs(title = "Morphology all species",
       x = "Station",
       y = "number of species") +
  theme_minimal()+
  scale_x_discrete(guide = guide_axis(angle = 90)) + facet_wrap(~ Month*Year)
Morph_diversity
```

save plot
```{r, eval=FALSE}
setwd("/Users/hailaschultz/Dropbox/Schultz_Dissertation/Data_Analysis/Schultz_dissertation-2/figures")
ggsave(filename = "Morph_diversity.png", plot = Morph_diversity, width = 10, height = 8, device='png', dpi=700)
```


Average by station

```{r}
Morph_pa_ave<-Morph_pa_sum %>%
  group_by(Phylum, Station) %>%
  summarise(
    ASVs = mean(ASVs),
    sd= sd(ASVs))
```
```{r}
Morph_diversity_ave<-ggplot(Morph_pa_ave, aes(x = Station, y = ASVs, fill = Phylum)) +
  geom_bar(stat = "identity") +
  labs(title = "Morphology all species",
       x = "Station",
       y = "number of species") +
  theme_minimal()+
  scale_x_discrete(guide = guide_axis(angle = 90)) 
Morph_diversity_ave
```
just July 2020
```{r}
Morph_pa_jul2020<-subset(Morph_pa,variable[,2]=="2009")
```
summarize by phylum
```{r}
Morph_pa_jul2020<-Morph_pa_jul2020 %>%
  group_by(Phylum, variable) %>%
  summarise(
    ASVs = sum(value))
```
```{r}
#change factor order
Morph_pa_jul2020$Station<-Morph_pa_jul2020$variable[,3]
Morph_pa_jul2020$Station <- factor(Morph_pa_jul2020$Station, levels = c("P22", "P4", "P8","P28","P12","P402","P38"))
Morph_pa_jul2020$Station <- recode(Morph_pa_jul2020$Station, 
                                 "P22" = "Juan de Fuca",
                         "P4"="Whidbey Basin",
                         "P8"="HC Sill",
                         "P28"="Main Basin",
                         "P12"="Hood Canal 1",
                         "P402"="Hood Canal 2",
                         "P38"="South Sound")

Morph_barplot<-ggplot(Morph_pa_jul2020, aes(x = Station, y = ASVs, fill = Phylum)) +
  geom_bar(stat = "identity") +
  labs(
       x = "Station",
       y = "Number of Species") +
  theme_minimal()+
  scale_x_discrete(guide = guide_axis(angle = 90)) +theme_bw()+theme(panel.grid.major = element_blank(),
                                                                                           panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))+ scale_colour_manual(values=c("#F8766D", "#E68613","#0CB702", "#00BE67","#00BFC4")) +
  scale_fill_manual(values=c("#F8766D", "#E68613","#0CB702", "#00BE67","#00BFC4"))+theme(axis.text=element_text(size=15),
        axis.title=element_text(size=15,face="bold"))+ylim(0,50)
Morph_barplot
```

save plot
```{r, eval=FALSE}
setwd("/Users/hailaschultz/Dropbox/Schultz_Dissertation/Data_Analysis/Schultz_dissertation-2/figures")
ggsave(filename = "Morph_barplot.png", plot = Morph_barplot, width = 10, height = 8, device='png', dpi=700)
```