---
title: "12_zoop-seasonal-NMDS"
date: "2024-03-09"
output:
  html_document:
    toc: true
    toc_depth: 2
    toc_float: true
---
```{r}
library(tidyr)
library(data.table)
library(vegan)
library(EcolUtils)
library(ggplot2)
library(corrplot)
library(ggrepel)
```

# Environmental Data
the environmental data was created by 00_Upload_Env_Data

subset environmental data to only shallow and deep niskins
```{r}

#make unique code
subset_df$code<-paste(subset_df$STATION_NO, subset_df$Year,subset_df$Month,sep="-")

# Sort the dataframe by sample and then by depth
subset_df <- subset_df %>% arrange(code, NISKIN_NO)
```

remove unnecessary columns
```{r}
Env_subset <- subset_df[ -c(1:7,10,12,14:17,19:25,27:32) ]
```

remove rows with missing data
```{r}
Env_subset<-Env_subset %>% drop_na(OXYGEN_MG_L_1)
```


```{r}
# Group by sample and filter the shallowest and deepest rows for each sample
Env_subset <- Env_subset %>%
  group_by(code) %>%
  mutate(row_num = row_number()) %>%
  filter(row_num == 1 | row_num == n())

# Create a new column to label them as shallow or deep
Env_subset <- Env_subset %>%
  mutate(Depth = ifelse(row_num == 1, "Deep","Surface" )) %>%
  select(-row_num) # Remove the temporary row number column
```

separate surface and deep samples
```{r}
#remove niskin no
Env_subset <- Env_subset[ -c(2) ]
#make surface and deep variables separate
Env_subset <- Env_subset %>%
  pivot_wider(names_from = Depth, values_from = c(CTDTMP_DEG_C_ITS90, CTDSAL_PSS78, OXYGEN_MG_L_1, NITRATE_UMOL_L)) 
```

replace negative values with NA
```{r}
Env_subset[Env_subset < 0] <- NA
#remove rows with NA
Env_subset<-Env_subset[!is.na(Env_subset$NITRATE_UMOL_L_Surface),]
Env_subset<-Env_subset[!is.na(Env_subset$CTDSAL_PSS78_Surface),]
```

anomalies

month and station averages
```{r}
colnames (Env_subset)
env_ave<-Env_subset %>%
  group_by(Month,STATION_NO) %>%
  summarise(across("CTDTMP_DEG_C_ITS90_Deep":"NITRATE_UMOL_L_Surface", list(mean)))
```

standard deviation
```{r}
env_stdev<- Env_subset  %>%
  group_by(Month,STATION_NO) %>%
  summarise(across("CTDTMP_DEG_C_ITS90_Deep":"NITRATE_UMOL_L_Surface", list(sd)))
```

mean anomalies
```{r}
#add averages column
#make station-month column
Env_subset$stationmonth<-paste(Env_subset$STATION_NO, Env_subset$Month,sep="-")
env_ave$stationmonth<-paste(env_ave$STATION_NO, env_ave$Month,sep="-")
env_stdev$stationmonth<-paste(env_stdev$STATION_NO, env_stdev$Month,sep="-")

env_joined_ave <- merge(Env_subset,env_ave, by="stationmonth", all.x=TRUE)
env_joined_stdev <- merge(env_joined_ave,env_stdev, by="stationmonth", all.x=TRUE)

#calculate anomalies
env_joined_anom <- (env_joined_stdev[, 7:14] - env_joined_stdev[, 17:24])/env_joined_stdev[, 27:34]
```

add anomalies to data
```{r}
Env_subset<-cbind(Env_subset, env_joined_anom)
```


# Zooplankton data
Read zooplanton data into R
```{r}
Zooplankton <- read.csv("/Users/hailaschultz/Dropbox/Schultz_Dissertation/Data_Analysis/Schultz_dissertation-2/data/WOAC_Species_Densities_2014-2022.csv")
```

subset to stations sampled during WOAC cruises
```{r}
unique(Zooplankton$Station)
Zoop_sub<-subset(Zooplankton,Station!="PEFS1d")
Zoop_sub<-subset(Zoop_sub,Station!="PEFN1d")
Zoop_sub<-subset(Zoop_sub,Station!="PEFN2")
Zoop_sub<-subset(Zoop_sub,Station!="PEFS2")
Zoop_sub<-subset(Zoop_sub,Station!="P381")
Zoop_sub<-subset(Zoop_sub,Station!="P105")
Zoop_sub<-subset(Zoop_sub,Station!="P136")
Zoop_sub<-subset(Zoop_sub,Station!="P7")
Zoop_sub<-subset(Zoop_sub,Station!="P383")
Zoop_sub<-subset(Zoop_sub,Station!="P132")
unique(Zoop_sub$Date)
#remove unnecessary dates
Zoop_sub<-subset(Zoop_sub,Date!="10/30/14")
Zoop_sub<-subset(Zoop_sub,Date!="10/31/14")
Zoop_sub<-subset(Zoop_sub,Date!="10/29/14")
Zoop_sub<-subset(Zoop_sub,Date!="9/14/22")
Zoop_sub<-subset(Zoop_sub,Date!="9/16/22")
Zoop_sub<-subset(Zoop_sub,Date!="9/13/22")
Zoop_sub<-subset(Zoop_sub,Date!="9/12/22")
Zoop_sub<-subset(Zoop_sub,Date!="11/16/15")
Zoop_sub<-subset(Zoop_sub,Date!="3/17/16")
Zoop_sub<-subset(Zoop_sub,Date!="3/18/16")
Zoop_sub<-subset(Zoop_sub,Date!="11/18/15")
#recode p4
Zoop_sub$Station <- recode(Zoop_sub$Station, 
                                 "p4" = "P4")
#recode month names to match
unique(Zoop_sub$Month)
Zoop_sub$Month <- recode(Zoop_sub$Month, 
                                 "Jul" = "JUL",
                         "Sep"="SEP",
                         "OCT"="SEP")
#remove ethanol samples
Zoop_sub<-Zoop_sub[!grepl('_EtOH', Zoop_sub$Code),]
#remove oblique samples
Zoop_sub<-subset(Zoop_sub,Tow.Type!="Oblique")
```
remove non-species columns
```{r}
Zoop_sub$Genus.species <- recode(Zoop_sub$Genus.species, 
                                 "Cancridae Lg" = "Cancridae",
                         "Cancridae Sm"="Cancridae",
                         "PSEUDOCALANUS Lg"="PSEUDOCALANUS",
                         "PSEUDOCALANUS Sm"="PSEUDOCALANUS")
```


add up multiple lines per station
```{r}
Zoop_sub <- Zoop_sub %>%
  group_by(Code,Station,Basin,Year,Month,Genus.species) %>%
  summarise(
    zoop_density = sum(Density....m3.))
```

change from long to wide format
```{r}
colnames(Zoop_sub)
Zoop_sub_wide<-dcast(Zoop_sub, Code+Station+Basin+Year+Month~ Genus.species,value.var = "zoop_density")
#make sure factors are formatted right
Zoop_sub_wide$Year=as.factor(Zoop_sub_wide$Year)
Zoop_sub_wide$Station=as.factor(Zoop_sub_wide$Station)
Zoop_sub_wide$Month=as.factor(Zoop_sub_wide$Month)
Zoop_sub_wide$Basin=as.factor(Zoop_sub_wide$Basin)
```

```{r}
#make unique column
Zoop_sub_wide$code<-paste(Zoop_sub_wide$Station, Zoop_sub_wide$Year,Zoop_sub_wide$Month,sep="-")
unique(Zoop_sub$Code)
```

# Combine env and zoop
```{r}
Zoop_env <- merge(Env_subset,Zoop_sub_wide, by.x = "code", 
             by.y = "code", all.x = TRUE, all.y = TRUE)

#remove rows with NA
Zoop_env<-Zoop_env[!is.na(Zoop_env$Year.x),]
```




#remove rows with all zoooplankton NAs
```{r}
rows_to_keep <- apply(Zoop_env[, 28:235], 1, function(row) !all(is.na(row)))

Zoop_env <- Zoop_env[rows_to_keep, ]
```

# NMDS

## prep data
remove non-data columns, convert to proportionas, and arcsine sqrt transformation
```{r}
RE2<- Zoop_env[,28:ncol(Zoop_env)]
#replace N/A with 0
RE2[is.na(RE2)] <- 0
#convert to proportions
RE2<-RE2/rowSums(RE2)

#arcsine sqrt transformation
RE2<-asin(sqrt(RE2))

#remove columns with fewer than 9 rows with nonzero numbers
non_zero_counts <- colSums(RE2 != 0)
RE2 <- RE2[, non_zero_counts >= 9]
```

indicator species
```{r}
inv = multipatt(RE2, Zoop_env$STATION_NO, func = "r.g", control = how(nperm=999))
summary(inv)
```


## PERMANOVA
```{r}
dist<-vegdist(RE2, method='bray')
perm<-adonis2(dist~Year.x*STATION_NO*Month.x, data=Zoop_env, permutations = 999, method="bray")
perm
```
It looks like year, station, and month are significant. Additionally, the interaction between year and month and station and month are different.

See pairwise differences
```{r}
#make sure factors are formatted right
Zoop_env$Year.x=as.factor(Zoop_env$Year.x)
Zoop_env$STATION_NO=as.factor(Zoop_env$STATION_NO)
Zoop_env$Month.x=as.factor(Zoop_env$Month.x)
adonis.pair(vegdist(RE2),Zoop_env$STATION_NO)
adonis.pair(vegdist(RE2),Zoop_env$Year.x)
adonis.pair(vegdist(RE2),Zoop_env$Month.x)
```

all stations are different from one another

2014 is similar to 2015
2014 is similar to 2020
2014 is similar to 2021
2015 is similar to 2016
2015 is similar to 2017
2015 is similar to 2021
2016 is similar to 2017
2016 is similar to 2018
2016 is similar to 2019
2017 is similar to 2018
2017 is similar to 2019
2018 is similar to 2019
2018 is similar to 2020
2018 is similar to 2021
2019 is similar to 2020
2019 is similar to 2021
2019 is similar to 2022
2020 is similar to 2021
2020 is similar to 2022

all months are different from one another


check for dispersion
```{r}
#station
Zoop.bd <- betadisper(dist, Zoop_env$STATION_NO)
Zoop.bd
anova(Zoop.bd)
permutest(Zoop.bd)

#year
Zoop.bd <- betadisper(dist, Zoop_env$Year.x)
Zoop.bd
anova(Zoop.bd)
permutest(Zoop.bd)

#month
Zoop.bd <- betadisper(dist, Zoop_env$Month.x)
Zoop.bd
anova(Zoop.bd)
permutest(Zoop.bd)
```

station and year are significant

## Run NMDS
```{r}
NMDSmodel <- metaMDS(RE2, distance = "bray",trymax = 200)
NMDSmodel
```

get datascores
```{r}
data.scores<-scores(NMDSmodel,display="sites")
data.scores<- as.data.frame(data.scores)
```

add station, month, year back in
```{r}
data.scores$Station = Zoop_env$STATION_NO
data.scores$Year = Zoop_env$Year.x
data.scores$Month = Zoop_env$Month.x
```

add environmental variables back in
```{r}
colnames(Zoop_env)
data.scores$temp_deep = Zoop_env$CTDTMP_DEG_C_ITS90_Deep...6
data.scores$temp_surface = Zoop_env$CTDTMP_DEG_C_ITS90_Surface...7
data.scores$salinity_deep = Zoop_env$CTDSAL_PSS78_Deep...8
data.scores$salinity_surface = Zoop_env$CTDSAL_PSS78_Surface...9
data.scores$oxygen_deep = Zoop_env$OXYGEN_MG_L_1_Deep...10
data.scores$oxygen_surface = Zoop_env$OXYGEN_MG_L_1_Surface...11
data.scores$nitrate_deep = Zoop_env$NITRATE_UMOL_L_Deep...12
data.scores$nitrate_surface = Zoop_env$NITRATE_UMOL_L_Surface...13
data.scores$temp_surf_anom = Zoop_env$CTDTMP_DEG_C_ITS90_Surface...16
data.scores$temp_deep_anom = Zoop_env$CTDTMP_DEG_C_ITS90_Deep...15
data.scores$oxy_deep_anom = Zoop_env$OXYGEN_MG_L_1_Deep...19
```

plot data scores
```{r}
xx <- ggplot(data.scores, aes(x = NMDS1, y = NMDS2,colour=Station)) +stat_ellipse()+geom_point(aes(colour = Station))+theme_bw()+theme(panel.grid.major = element_blank(),
                                                                                           panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))
xx

xx <- ggplot(data.scores, aes(x = NMDS1, y = NMDS2,colour = Year)) +stat_ellipse()+geom_point(aes(colour = Year))+theme_bw()+theme(panel.grid.major = element_blank(),
                                                                                           panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))
xx

xx <- ggplot(data.scores, aes(x = NMDS1, y = NMDS2,colour = Month)) +stat_ellipse()+geom_point(aes(colour = Month))+theme_bw()+theme(panel.grid.major = element_blank(),
                                                                                           panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))
xx
```

April is pretty clearly different than July and September

## environmental fitting
create environmetal fit table
```{r}
#check for correlation among environmental variables.
#remove nutrients other than nitrate because of autocorrelation

data.scores.cor<- data.scores[ -c(1:5) ]
cor_matrix <- cor(data.scores.cor)
corrplot(cor_matrix, method = "number", type = "lower", tl.pos = 'd')

Zoo.env <- data.frame(matrix(ncol = 0, nrow=151))
colnames(data.scores)
Zoo.env$temp_deep<-as.numeric(data.scores$temp_deep)
Zoo.env$temp_surface<-as.numeric(data.scores$temp_surface)
Zoo.env$salinity_deep<-as.numeric(data.scores$salinity_deep)
Zoo.env$salinity_surface<-as.numeric(data.scores$salinity_surface)
Zoo.env$oxygen_deep<-as.numeric(data.scores$oxygen_deep)
Zoo.env$oxygen_surface<-as.numeric(data.scores$oxygen_surface)
Zoo.env$temp_deep_anom<-as.numeric(data.scores$temp_deep_anom)
Zoo.env$temp_surf_anom<-as.numeric(data.scores$temp_surf_anom)
Zoo.env$oxy_deep_anom<-as.numeric(data.scores$oxy_deep_anom)

env <- envfit(NMDSmodel, Zoo.env, na.rm = TRUE,perm = 999)
env


env.scrs <- as.data.frame(scores(env, display = "vectors")) 
env.scrs <- cbind(env.scrs, Species = rownames(env.scrs), Pvalues = env$vectors$pvals, R_squared = env$vectors$r)
env.scrs <- subset(env.scrs, Pvalues < 0.05)


env.scrs
```

## Species fitting
```{r}
vf <- envfit(NMDSmodel, RE2, perm = 999)

spp.scrs <- as.data.frame(scores(vf, display = "vectors"))
spp.scrs <- cbind(spp.scrs, Species = rownames(spp.scrs))
####for ggplot
arrow_factor <- ordiArrowMul(vf)
spp.scrs <- as.data.frame(scores(vf, display = "vectors")) * arrow_factor
spp.scrs <- cbind(spp.scrs, Species = rownames(spp.scrs), Pvalues = vf$vectors$pvals, R_squared = vf$vectors$r)

# select significance similarly to `plot(vf, p.max = 0.01)`
spp.scrs <- subset(spp.scrs, Pvalues < 0.05)
spp.scrs <- subset(spp.scrs, R_squared > 0.3)
spp.scrs

```
### all data with species
extract hulls
```{r}
P12 <- data.scores[data.scores$Station == "P12", ][chull(data.scores[data.scores$Station ==                                                           "P12", c("NMDS1", "NMDS2")]), ]  
P22 <- data.scores[data.scores$Station == "P22", ][chull(data.scores[data.scores$Station ==                                                           "P22", c("NMDS1", "NMDS2")]), ]  
P28 <- data.scores[data.scores$Station == "P28", ][chull(data.scores[data.scores$Station ==                                                           "P28", c("NMDS1", "NMDS2")]), ]
P38 <- data.scores[data.scores$Station == "P38", ][chull(data.scores[data.scores$Station ==                                                           "P38", c("NMDS1", "NMDS2")]), ]  
P4 <- data.scores[data.scores$Station == "P4", ][chull(data.scores[data.scores$Station ==                                                           "P4", c("NMDS1", "NMDS2")]), ]  
P402 <- data.scores[data.scores$Station == "P402", ][chull(data.scores[data.scores$Station ==                                                           "P402", c("NMDS1", "NMDS2")]), ]  
P8 <- data.scores[data.scores$Station == "P8", ][chull(data.scores[data.scores$Station ==                                                           "P8", c("NMDS1", "NMDS2")]), ]  
```
get hull data
```{r}
hull.data <- rbind(P12, P22, P28,P38,P4,P402,P8)  #combine grp.a and grp.b
hull.data

```
```{r,fig.width=6.5,fig.height=5, dpi=600}

Species_plot <- ggplot(data.scores, aes(x = NMDS1, y = NMDS2)) +geom_point(aes(colour = Station))+theme_bw()+theme(panel.grid.major = element_blank(),
                                                                                           panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))+ 
  geom_polygon(data=hull.data,aes(x=NMDS1,y=NMDS2,fill=Station,group=Station),alpha=0.20, size=0.1, linetype=1, colour="black") +
  annotate("text", label = "2D Stress: 0.16", x = 0.5, y = 0.57, size = 5, colour = "black")+ scale_colour_manual(values=c("#F8766D","#C49A00","#53B400","#00C094","#00B6EB","#A69AFF","#FB61D7")) +
  scale_fill_manual(values=c("#F8766D","#C49A00","#53B400","#00C094","#00B6EB","#A69AFF","#FB61D7"))+ geom_segment(data = spp.scrs,size=0.2,
               aes(x = 0, xend = NMDS1, y = 0, yend = NMDS2),
               arrow = arrow(length = unit(0.08, "cm")), colour = "black")+
  geom_label_repel(data = spp.scrs, aes(x = NMDS1, y = NMDS2, label = Species),
                  size = 1.5, fontface="bold", fill="white", label.padding = unit(0.15, "lines"), box.padding = unit(0.16, "lines"), label.size = 0.05)
Species_plot
```
```{r,fig.width=6.5,fig.height=5, dpi=600}

Species_plot <- ggplot(data.scores, aes(x = NMDS1, y = NMDS2)) +geom_point(aes(colour = temp_deep_anom))+theme_bw()+theme(panel.grid.major = element_blank(),
                                                                                           panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))+ 
  geom_polygon(data=hull.data,aes(x=NMDS1,y=NMDS2,group=Station),alpha=0.05, size=0.1, linetype=1, colour="black") +
  annotate("text", label = "2D Stress: 0.16", x = 0.5, y = 0.57, size = 5, colour = "black")+ geom_segment(data = spp.scrs,size=0.2,
               aes(x = 0, xend = NMDS1, y = 0, yend = NMDS2),
               arrow = arrow(length = unit(0.08, "cm")), colour = "black")+
  geom_label_repel(data = spp.scrs, aes(x = NMDS1, y = NMDS2, label = Species),
                  size = 1.5, fontface="bold", fill="white", label.padding = unit(0.15, "lines"), box.padding = unit(0.16, "lines"), label.size = 0.05)+scale_color_gradient(low="blue", high="red")+ 
  geom_polygon(data=apr.hull.data,aes(x=NMDS1,y=NMDS2,group=Station),alpha=0.20, size=0.1, linetype=1, colour="black")
Species_plot
```

## plot only April

subset data scores
```{r}
APR_datascores<-subset(data.scores,Month=="APR")
```
extract hulls
```{r}
P12 <- APR_datascores[APR_datascores$Station == "P12", ][chull(APR_datascores[APR_datascores$Station ==                                                           "P12", c("NMDS1", "NMDS2")]), ]  
P22 <- APR_datascores[APR_datascores$Station == "P22", ][chull(APR_datascores[APR_datascores$Station ==                                                           "P22", c("NMDS1", "NMDS2")]), ]  
P28 <- APR_datascores[APR_datascores$Station == "P28", ][chull(APR_datascores[APR_datascores$Station ==                                                           "P28", c("NMDS1", "NMDS2")]), ]
P38 <- APR_datascores[APR_datascores$Station == "P38", ][chull(APR_datascores[APR_datascores$Station ==                                                           "P38", c("NMDS1", "NMDS2")]), ]  
P4 <- APR_datascores[APR_datascores$Station == "P4", ][chull(APR_datascores[APR_datascores$Station ==                                                           "P4", c("NMDS1", "NMDS2")]), ]  
P402 <- APR_datascores[APR_datascores$Station == "P402", ][chull(APR_datascores[APR_datascores$Station ==                                                           "P402", c("NMDS1", "NMDS2")]), ]  
P8 <- APR_datascores[APR_datascores$Station == "P8", ][chull(APR_datascores[APR_datascores$Station ==                                                           "P8", c("NMDS1", "NMDS2")]), ]  
```

get hull data
```{r}
apr.hull.data <- rbind(P12, P22, P28,P38,P4,P402,P8)  #combine grp.a and grp.b
apr.hull.data

```
plot vectors
```{r,fig.width=5,fig.height=5.5, dpi=600}

Blank <- ggplot(APR_datascores, aes(x = NMDS1, y = NMDS2)) +theme_bw()+theme(panel.grid.major = element_blank(),
                                                                                           panel.grid.minor = element_blank(), axis.line = element_line(colour = "black")) +
  annotate("text", label = "2D Stress: 0.18", x = 0.7, y = 1, size = 5, colour = "black")+ scale_colour_manual(values=c("#F8766D","#C49A00","#53B400","#00C094","#00B6EB","#A69AFF","#FB61D7")) +
  scale_fill_manual(values=c("#F8766D","#C49A00","#53B400","#00C094","#00B6EB","#A69AFF","#FB61D7"))+
  theme(axis.text=element_text(size=12),legend.text=element_text(size=12),legend.title=element_text(size=12))+
  geom_segment(data = env.scrs,size=0.5,
               aes(x = 0, xend = NMDS1, y = 0, yend = NMDS2),
               arrow = arrow(length = unit(0.5, "cm")), colour = "black")+xlim(-0.8, 1)+ylim(-1.3,1)
Blank
```
save dendogram
```{r, eval=FALSE}
setwd("/Users/hailaschultz/Dropbox/Schultz_Dissertation/Data_Analysis/Schultz_dissertation-2/figures")
ggsave(filename = "NMDS_blank.png", plot = Blank, width = 5, height = 5.5, device='png', dpi=700)
```


plot data scores
```{r,fig.width=5,fig.height=5.5, dpi=600}
#change factor order
APR_datascores$Station <- factor(APR_datascores$Station, levels = c("P22", "P4", "P8","P28","P12","P402","P38"))
apr.hull.data$Station <- factor(apr.hull.data$Station, levels = c("P22", "P4", "P8","P28","P12","P402","P38"))

April <- ggplot(APR_datascores, aes(x = NMDS1, y = NMDS2)) +geom_point(aes(colour = Station),size=2.5)+theme_bw()+theme(panel.grid.major = element_blank(),
                                                                                           panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))+ 
  geom_polygon(data=apr.hull.data,aes(x=NMDS1,y=NMDS2,fill=Station,group=Station),alpha=0.20, size=0.1, linetype=1, colour="black") +
  annotate("text", label = "2D Stress: 0.18", x = 0.7, y = 1, size = 5, colour = "black")+ scale_colour_manual(values=c("#F8766D","#C49A00","#53B400","#00C094","#00B6EB","#A69AFF","#FB61D7")) +
  scale_fill_manual(values=c("#F8766D","#C49A00","#53B400","#00C094","#00B6EB","#A69AFF","#FB61D7"))+
  theme(axis.text=element_text(size=12))+
  geom_segment(data = env.scrs,size=0.5,
               aes(x = 0, xend = NMDS1, y = 0, yend = NMDS2),
               arrow = arrow(length = unit(0.5, "cm")), colour = "black")+xlim(-0.8, 1)+ylim(-1.3,1)+ theme(legend.position = "none")
April
```
save dendogram
```{r, eval=FALSE}
setwd("/Users/hailaschultz/Dropbox/Schultz_Dissertation/Data_Analysis/Schultz_dissertation-2/figures")
ggsave(filename = "NMDS_April.png", plot = April, width = 5, height = 5.5, device='png', dpi=700)
```


color by anomaly
```{r,fig.width=6.5,fig.height=5, dpi=600}
#change factor order
APR_datascores$Station <- factor(APR_datascores$Station, levels = c("P22", "P4", "P8","P28","P12","P402","P38"))
apr.hull.data$Station <- factor(apr.hull.data$Station, levels = c("P22", "P4", "P8","P28","P12","P402","P38"))

April <- ggplot(APR_datascores, aes(x = NMDS1, y = NMDS2)) +geom_point(aes(colour = temp_deep_anom),size=2.5)+theme_bw()+theme(panel.grid.major = element_blank(),
                                                                                           panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))+
  annotate("text", label = "2D Stress: 0.16", x = 0.5, y = 0.57, size = 5, colour = "black")+ 
  theme(axis.text=element_text(size=12),legend.text=element_text(size=12),legend.title=element_text(size=12))+
  geom_segment(data = env.scrs,size=0.5,
               aes(x = 0, xend = NMDS1, y = 0, yend = NMDS2),
               arrow = arrow(length = unit(0.5, "cm")), colour = "black")+
  geom_label_repel(data = env.scrs, aes(x = NMDS1, y = NMDS2, label = Species),
                  size = 5, fontface="bold", fill="white", label.padding = unit(0.15, "lines"), point.padding = unit(15, "lines"), label.size = 0.05)+scale_color_gradient(low="blue", high="red")+ 
  geom_polygon(data=apr.hull.data,aes(x=NMDS1,y=NMDS2,group=Station),alpha=0.20, size=0.1, linetype=1, colour="black")
April
```


## plot only July

subset data scores
```{r}
JUL_datascores<-subset(data.scores,Month=="JUL")
```
extract hulls
```{r}
P12 <- JUL_datascores[JUL_datascores$Station == "P12", ][chull(JUL_datascores[JUL_datascores$Station ==                                                           "P12", c("NMDS1", "NMDS2")]), ]  
P22 <- JUL_datascores[JUL_datascores$Station == "P22", ][chull(JUL_datascores[JUL_datascores$Station ==                                                           "P22", c("NMDS1", "NMDS2")]), ]  
P28 <- JUL_datascores[JUL_datascores$Station == "P28", ][chull(JUL_datascores[JUL_datascores$Station ==                                                           "P28", c("NMDS1", "NMDS2")]), ]
P38 <- JUL_datascores[JUL_datascores$Station == "P38", ][chull(JUL_datascores[JUL_datascores$Station ==                                                           "P38", c("NMDS1", "NMDS2")]), ]  
P4 <- JUL_datascores[JUL_datascores$Station == "P4", ][chull(JUL_datascores[JUL_datascores$Station ==                                                           "P4", c("NMDS1", "NMDS2")]), ]  
P402 <- JUL_datascores[JUL_datascores$Station == "P402", ][chull(JUL_datascores[JUL_datascores$Station ==                                                           "P402", c("NMDS1", "NMDS2")]), ]  
P8 <- JUL_datascores[JUL_datascores$Station == "P8", ][chull(JUL_datascores[JUL_datascores$Station ==                                                           "P8", c("NMDS1", "NMDS2")]), ]  
```

get hull data
```{r}
jul.hull.data <- rbind(P12, P22, P28,P38,P4,P402,P8)  #combine grp.a and grp.b
jul.hull.data
```

plot data scores
```{r,fig.width=5,fig.height=5.5, dpi=600}
#change factor order
JUL_datascores$Station <- factor(JUL_datascores$Station, levels = c("P22", "P4", "P8","P28","P12","P402","P38"))
jul.hull.data$Station <- factor(jul.hull.data$Station, levels = c("P22", "P4", "P8","P28","P12","P402","P38"))

July <- ggplot(JUL_datascores, aes(x = NMDS1, y = NMDS2)) +geom_point(aes(colour = Station),size=2.5)+theme_bw()+theme(panel.grid.major = element_blank(),
                                                                                           panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))+ 
  geom_polygon(data=jul.hull.data,aes(x=NMDS1,y=NMDS2,fill=Station,group=Station),alpha=0.20, size=0.1, linetype=1, colour="black") +
  annotate("text", label = "2D Stress: 0.18", x = 0.7, y = 1, size = 5, colour = "black")+ scale_colour_manual(values=c("#F8766D","#C49A00","#53B400","#00C094","#00B6EB","#A69AFF","#FB61D7")) +
  scale_fill_manual(values=c("#F8766D","#C49A00","#53B400","#00C094","#00B6EB","#A69AFF","#FB61D7"))+
  theme(axis.text=element_text(size=12))+
  geom_segment(data = env.scrs,size=0.5,
               aes(x = 0, xend = NMDS1, y = 0, yend = NMDS2),
               arrow = arrow(length = unit(0.5, "cm")), colour = "black")+xlim(-0.8, 1)+ylim(-1.3,1)+ theme(legend.position = "none")
July
```
save dendogram
```{r, eval=FALSE}
setwd("/Users/hailaschultz/Dropbox/Schultz_Dissertation/Data_Analysis/Schultz_dissertation-2/figures")
ggsave(filename = "NMDS_July.png", plot = July, width = 5, height = 5.5, device='png', dpi=700)
```

## plot only September

subset data scores
```{r}
SEP_datascores<-subset(data.scores,Month=="SEP")
```
extract hulls
```{r}
P12 <- SEP_datascores[SEP_datascores$Station == "P12", ][chull(SEP_datascores[SEP_datascores$Station ==                                                           "P12", c("NMDS1", "NMDS2")]), ]  
P22 <- SEP_datascores[SEP_datascores$Station == "P22", ][chull(SEP_datascores[SEP_datascores$Station ==                                                           "P22", c("NMDS1", "NMDS2")]), ]  
P28 <- SEP_datascores[SEP_datascores$Station == "P28", ][chull(SEP_datascores[SEP_datascores$Station ==                                                           "P28", c("NMDS1", "NMDS2")]), ]
P38 <- SEP_datascores[SEP_datascores$Station == "P38", ][chull(SEP_datascores[SEP_datascores$Station ==                                                           "P38", c("NMDS1", "NMDS2")]), ]  
P4 <- SEP_datascores[SEP_datascores$Station == "P4", ][chull(SEP_datascores[SEP_datascores$Station ==                                                           "P4", c("NMDS1", "NMDS2")]), ]  
P402 <- SEP_datascores[SEP_datascores$Station == "P402", ][chull(SEP_datascores[SEP_datascores$Station ==                                                           "P402", c("NMDS1", "NMDS2")]), ]  
P8 <- SEP_datascores[SEP_datascores$Station == "P8", ][chull(SEP_datascores[SEP_datascores$Station ==                                                           "P8", c("NMDS1", "NMDS2")]), ]  
```

get hull data
```{r}
SEP.hull.data <- rbind(P12, P22, P28,P38,P4,P402,P8)  #combine grp.a and grp.b
SEP.hull.data
```

plot data scores
```{r,fig.width=5,fig.height=5.5, dpi=600}
#change factor order
SEP_datascores$Station <- factor(SEP_datascores$Station, levels = c("P22", "P4", "P8","P28","P12","P402","P38"))
SEP.hull.data$Station <- factor(SEP.hull.data$Station, levels = c("P22", "P4", "P8","P28","P12","P402","P38"))

September <- ggplot(SEP_datascores, aes(x = NMDS1, y = NMDS2)) +geom_point(aes(colour = Station),size=2.5)+theme_bw()+theme(panel.grid.major = element_blank(),
                                                                                           panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))+ 
  geom_polygon(data=SEP.hull.data,aes(x=NMDS1,y=NMDS2,fill=Station,group=Station),alpha=0.20, size=0.1, linetype=1, colour="black") +
  annotate("text", label = "2D Stress: 0.18", x = 0.7, y = 1, size = 5, colour = "black")+ scale_colour_manual(values=c("#F8766D","#C49A00","#53B400","#00C094","#00B6EB","#A69AFF","#FB61D7")) +
  scale_fill_manual(values=c("#F8766D","#C49A00","#53B400","#00C094","#00B6EB","#A69AFF","#FB61D7"))+
  theme(axis.text=element_text(size=12))+
  geom_segment(data = env.scrs,size=0.5,
               aes(x = 0, xend = NMDS1, y = 0, yend = NMDS2),
               arrow = arrow(length = unit(0.5, "cm")), colour = "black")+xlim(-0.8, 1)+ylim(-1.3,1)+ theme(legend.position = "none")
September
```
save dendogram
```{r, eval=FALSE}
setwd("/Users/hailaschultz/Dropbox/Schultz_Dissertation/Data_Analysis/Schultz_dissertation-2/figures")
ggsave(filename = "NMDS_September.png", plot = September, width = 5, height = 5.5, device='png', dpi=700)
```
species scores
```{r,fig.width=5,fig.height=5.5, dpi=600}


September_sp <- ggplot(SEP_datascores, aes(x = NMDS1, y = NMDS2)) +geom_point(aes(colour = Station),size=2.5)+theme_bw()+theme(panel.grid.major = element_blank(),
                                                                                           panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))+ 
  geom_polygon(data=SEP.hull.data,aes(x=NMDS1,y=NMDS2,fill=Station,group=Station),alpha=0.20, size=0.1, linetype=1, colour="black") +
  annotate("text", label = "2D Stress: 0.18", x = 0.7, y = 1, size = 5, colour = "black")+ scale_colour_manual(values=c("#F8766D","#C49A00","#53B400","#00C094","#00B6EB","#A69AFF","#FB61D7")) +
  scale_fill_manual(values=c("#F8766D","#C49A00","#53B400","#00C094","#00B6EB","#A69AFF","#FB61D7"))+
  theme(axis.text=element_text(size=12),legend.text=element_text(size=12),legend.title=element_text(size=12))+
  geom_segment(data = spp.scrs,size=0.4,
               aes(x = 0, xend = NMDS1, y = 0, yend = NMDS2),
               arrow = arrow(length = unit(0.2, "cm")), colour = "darkgrey")+xlim(-1, 1)+ylim(-1.3,1)+ theme(legend.position = "none")+
  geom_segment(data = env.scrs,size=0.4,
               aes(x = 0, xend = NMDS1, y = 0, yend = NMDS2),
               arrow = arrow(length = unit(0.2, "cm")), colour = "black")

September_sp
```

save dendogram
```{r, eval=FALSE}
setwd("/Users/hailaschultz/Dropbox/Schultz_Dissertation/Data_Analysis/Schultz_dissertation-2/figures")
ggsave(filename = "NMDS_September_Sp.png", plot = September_sp, width = 5, height = 5.5, device='png', dpi=700)
```

### try group 1
remove non-data columns, convert to proportionas, and arcsine sqrt transformation
```{r}
group1<-subset(Zoop_env,STATION_NO!="P22")
group1<-subset(group1,STATION_NO!="P8")
group1<-subset(group1,STATION_NO!="P38")
group1<-subset(group1,STATION_NO!="P28")
RE2<- group1[,28:ncol(group1)]
#replace N/A with 0
RE2[is.na(RE2)] <- 0
#convert to proportions
RE2<-RE2/rowSums(RE2)

#arcsine sqrt transformation
RE2<-asin(sqrt(RE2))
#remove columns with fewer than 9 rows with nonzero numbers
non_zero_counts <- colSums(RE2 != 0)
RE2 <- RE2[, non_zero_counts >= 9]
```

## PERMANOVA
```{r}
dist<-vegdist(RE2, method='bray')
perm<-adonis2(dist~Station*Month.x, data=group1, permutations = 999, method="bray")
perm
```
See pairwise differences
```{r}
#make sure factors are formatted right
group1$STATION_NO=as.factor(group1$Station)
group1$Month.x=as.factor(group1$Month.x)
adonis.pair(vegdist(RE2),group1$Station)
adonis.pair(vegdist(RE2),group1$Month.x)
```
## Run NMDS
```{r}
NMDSmodel <- metaMDS(RE2, distance = "bray",trymax = 200)
NMDSmodel
```



get datascores
```{r}
data.scores<-scores(NMDSmodel,display="sites")
data.scores<- as.data.frame(data.scores)
```

add station, month, year back in
```{r}
data.scores$Station = group1$STATION_NO
data.scores$Year = group1$Year.x
data.scores$Month = group1$Month.x
```

add environmental variables back in
```{r}
data.scores$temp_deep = group1$CTDTMP_DEG_C_ITS90_Deep...6
data.scores$temp_surface = group1$CTDTMP_DEG_C_ITS90_Surface...7
data.scores$salinity_deep = group1$CTDSAL_PSS78_Deep...8
data.scores$salinity_surface = group1$CTDSAL_PSS78_Surface...9
data.scores$oxygen_deep = group1$OXYGEN_MG_L_1_Deep...10
data.scores$oxygen_surface = group1$OXYGEN_MG_L_1_Surface...11
data.scores$nitrate_deep = group1$NITRATE_UMOL_L_Deep...12
data.scores$nitrate_surface = group1$NITRATE_UMOL_L_Surface...13
data.scores$temp_surf_anom = group1$CTDTMP_DEG_C_ITS90_Surface...16
data.scores$temp_deep_anom = group1$CTDTMP_DEG_C_ITS90_Deep...15
data.scores$oxy_deep_anom = group1$OXYGEN_MG_L_1_Deep...19
```

plot data scores
```{r}
xx <- ggplot(data.scores, aes(x = NMDS1, y = NMDS2,colour=Station)) +stat_ellipse()+geom_point(aes(colour = Station))+theme_bw()+theme(panel.grid.major = element_blank(),
                                                                                           panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))
xx

xx <- ggplot(data.scores, aes(x = NMDS1, y = NMDS2,colour = Year)) +stat_ellipse()+geom_point(aes(colour = Year))+theme_bw()+theme(panel.grid.major = element_blank(),
                                                                                           panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))
xx

xx <- ggplot(data.scores, aes(x = NMDS1, y = NMDS2,colour = Month)) +stat_ellipse()+geom_point(aes(colour = Month))+theme_bw()+theme(panel.grid.major = element_blank(),
                                                                                           panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))
xx

xx <- ggplot(data.scores, aes(x = NMDS1, y = NMDS2))+geom_point(aes(colour = temp_surf_anom))+theme_bw()+theme(panel.grid.major = element_blank(),
                                                                                           panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))+scale_color_gradient(low="blue", high="red")
xx
```
