---
title: "12_zoop-seasonal-NMDS"
Sample.Date: "2024-03-09"
output:
  html_document:
    toc: true
    toc_depth: 2
    toc_float: true
editor_options: 
  chunk_output_type: inline
---
```{r}
library(tidyr)
library(data.table)
library(vegan)
library(EcolUtils)
library(ggplot2)
library(corrplot)
library(ggrepel)
library(indicspecies)
library(dplyr)
```

# Environmental Data
the environmental data was created by 00_Upload_Env_Data

subset environmental data to only shallow and deep niskins
```{r}
#make unique Sample.Code
subset_df$Sample.Code<-paste(subset_df$STATION_NO, subset_df$Year,subset_df$Month,sep="-")

# Sort the dataframe by sample and then by depth
subset_df <- subset_df %>% arrange(Sample.Code, NISKIN_NO)
```

remove unnecessary columns
```{r}
Env_subset <- subset_df[ -c(1:7,10,12,14:17,19:25,27:32) ]
```

remove rows with missing data
```{r}
Env_subset<-Env_subset %>% drop_na(OXYGEN_MG_L_1)
```


```{r}
# Group by sample and filter the shallowest and deepest rows for each sample
Env_subset <- Env_subset %>%
  group_by(Sample.Code) %>%
  mutate(row_num = row_number()) %>%
  filter(row_num == 1 | row_num == n())

# Create a new column to label them as shallow or deep
Env_subset <- Env_subset %>%
  mutate(Depth = ifelse(row_num == 1, "Deep","Surface" )) 
#remove row no
Env_subset <- Env_subset[ -c(11) ]
```

separate surface and deep samples
```{r}
#remove niskin no
Env_subset <- Env_subset[ -c(2) ]
#make surface and deep variables separate
Env_subset <- Env_subset %>%
  pivot_wider(names_from = Depth, values_from = c(CTDTMP_DEG_C_ITS90, CTDSAL_PSS78, OXYGEN_MG_L_1, NITRATE_UMOL_L)) 
```

replace negative values with NA
```{r}
Env_subset[Env_subset < 0] <- NA
#remove rows with NA
Env_subset<-Env_subset[!is.na(Env_subset$NITRATE_UMOL_L_Surface),]
Env_subset<-Env_subset[!is.na(Env_subset$CTDSAL_PSS78_Surface),]
```



Sample.Month and station averages
```{r}
colnames (Env_subset)
env_ave<-Env_subset %>%
  group_by(Month,STATION_NO) %>%
  summarise(across("CTDTMP_DEG_C_ITS90_Deep":"NITRATE_UMOL_L_Surface", list(mean)))
```

standard deviation
```{r}
env_stdev<- Env_subset  %>%
  group_by(Month,STATION_NO) %>%
  summarise(across("CTDTMP_DEG_C_ITS90_Deep":"NITRATE_UMOL_L_Surface", list(sd)))
```

mean anomalies
```{r}
#add averages column
#make stationmonth column
Env_subset$stationmonth<-paste(Env_subset$STATION_NO, Env_subset$Month,sep="-")
env_ave$stationmonth<-paste(env_ave$STATION_NO, env_ave$Month,sep="-")
env_stdev$stationmonth<-paste(env_stdev$STATION_NO, env_stdev$Month,sep="-")

env_joined_ave <- merge(Env_subset,env_ave, by="stationmonth", all.x=TRUE)
env_joined_stdev <- merge(env_joined_ave,env_stdev, by="stationmonth", all.x=TRUE)

#calculate anomalies
env_joined_anom <- (env_joined_stdev[, 7:14] - env_joined_stdev[, 17:24])/env_joined_stdev[, 27:34]
```

add anomalies to data
```{r}
Env_subset<-cbind(Env_subset, env_joined_anom)
```

add in fluorescence
make sample code
```{r}
downcasts_combined$Sample.Code<-paste(downcasts_combined$Station,downcasts_combined$Year,downcasts_combined$Month, sep="-")
```

```{r}
downcasts_combined$Chlorophyll.Fluorescence <- as.numeric(downcasts_combined$Chlorophyll.Fluorescence)
```

get max fluorescence
```{r}
colnames(downcasts_combined)
downcast <- downcasts_combined %>%
  group_by(Sample.Code) %>%
  summarise(
    chla = max(Chlorophyll.Fluorescence))
```

add to other environmental data
```{r}
Env_subset <- merge(Env_subset,downcast, by.x = "Sample.Code", 
             by.y = "Sample.Code", all.x = TRUE, all.y = TRUE)

```



# Zooplankton data
Read zooplanton data into R
```{r}
Zooplankton <- read.csv("/Users/hailaschultz/Dropbox/Schultz_Dissertation/Data_Analysis/Schultz_dissertation-2/data/WOAC_2014-2022_Species_Data_Keister_Lab.csv")
```
remove low quality zooplankton data
```{r}
unique(Zooplankton$Confidence)
Zoop_sub<-subset(Zooplankton,Confidence!="low- flow high")
Zoop_sub<-subset(Zoop_sub,Confidence!="LOW")
Zoop_sub<-subset(Zoop_sub,Confidence!="Low-low flow")
```
subset to stations sampled during WOAC cruises
```{r}
Zoop_sub<-subset(Zoop_sub,Station!="PEFS1d")
Zoop_sub<-subset(Zoop_sub,Station!="PEFN1d")
Zoop_sub<-subset(Zoop_sub,Station!="PEFN2")
Zoop_sub<-subset(Zoop_sub,Station!="PEFS2")
Zoop_sub<-subset(Zoop_sub,Station!="P381")
Zoop_sub<-subset(Zoop_sub,Station!="P105")
Zoop_sub<-subset(Zoop_sub,Station!="P136")
Zoop_sub<-subset(Zoop_sub,Station!="P7")
Zoop_sub<-subset(Zoop_sub,Station!="P383")
Zoop_sub<-subset(Zoop_sub,Station!="P132")
unique(Zoop_sub$Sample.Date)
#remove unnecessary Sample.Dates
Zoop_sub<-subset(Zoop_sub,Sample.Date!="10/30/14")
Zoop_sub<-subset(Zoop_sub,Sample.Date!="10/31/14")
Zoop_sub<-subset(Zoop_sub,Sample.Date!="10/29/14")
Zoop_sub<-subset(Zoop_sub,Sample.Date!="9/14/22")
Zoop_sub<-subset(Zoop_sub,Sample.Date!="9/16/22")
Zoop_sub<-subset(Zoop_sub,Sample.Date!="9/13/22")
Zoop_sub<-subset(Zoop_sub,Sample.Date!="9/12/22")
Zoop_sub<-subset(Zoop_sub,Sample.Date!="11/16/15")
Zoop_sub<-subset(Zoop_sub,Sample.Date!="3/17/16")
Zoop_sub<-subset(Zoop_sub,Sample.Date!="3/18/16")
Zoop_sub<-subset(Zoop_sub,Sample.Date!="11/18/15")
#recode p4
Zoop_sub$Station <- recode(Zoop_sub$Station, 
                                 "p4" = "P4")
#recode Sample.Month names to match
unique(Zoop_sub$Sample.Month)
Zoop_sub$Sample.Month <- recode(Zoop_sub$Sample.Month, 
                                 "Jul" = "JUL",
                         "Sep"="SEP",
                         "OCT"="SEP")
#remove ethanol samples
Zoop_sub<-Zoop_sub[!grepl('_EtOH', Zoop_sub$Sample.Code),]
#remove oblique samples
Zoop_sub<-subset(Zoop_sub,Tow.Type!="Oblique")
#exclude weird taxa
Zoop_sub<-subset(Zoop_sub,WOAC.Group!="Exclude")
```
remove non-species columns
```{r}
Zoop_sub$Genus.species <- recode(Zoop_sub$Genus.species, 
                                 "Cancridae Lg" = "Cancridae",
                         "Cancridae Sm"="Cancridae",
                         "PSEUDOCALANUS Lg"="PSEUDOCALANUS",
                         "PSEUDOCALANUS Sm"="PSEUDOCALANUS")
```


add up multiple lines per station
```{r}
Zoop_sub <- Zoop_sub %>%
  group_by(Sample.Code,Station,Basin,Sample.Year,Sample.Month,Genus.species) %>%
  summarise(
    zoop_density = sum(Density....m3.))
```

change from long to wide format
```{r}
colnames(Zoop_sub)
Zoop_sub_wide<-dcast(Zoop_sub, Sample.Code+Station+Basin+Sample.Year+Sample.Month~ Genus.species,value.var = "zoop_density")
#make sure factors are formatted right
Zoop_sub_wide$Sample.Year=as.factor(Zoop_sub_wide$Sample.Year)
Zoop_sub_wide$Station=as.factor(Zoop_sub_wide$Station)
Zoop_sub_wide$Sample.Month=as.factor(Zoop_sub_wide$Sample.Month)
Zoop_sub_wide$Basin=as.factor(Zoop_sub_wide$Basin)
```

```{r}
#make unique column
Zoop_sub_wide$Sample.Code<-paste(Zoop_sub_wide$Station, Zoop_sub_wide$Sample.Year,Zoop_sub_wide$Sample.Month,sep="-")
unique(Zoop_sub_wide$Sample.Code)
unique(Env_subset$Sample.Code)
```


# Combine env and zoop
```{r}
Zoop_env <- merge(Env_subset,Zoop_sub_wide, by.x = "Sample.Code", 
             by.y = "Sample.Code", all.x = TRUE, all.y = TRUE)

#remove rows with NA
colnames(Zoop_env)
Zoop_env<-Zoop_env[!is.na(Zoop_env$STATION_NO),]
```




#remove rows with all zoooplankton NAs
```{r}
rows_to_keep <- apply(Zoop_env[, 27:234], 1, function(row) !all(is.na(row)))

Zoop_env <- Zoop_env[rows_to_keep, ]
```
remove rows with chla NA
```{r}
Zoop_env<-Zoop_env[!is.na(Zoop_env$chla),]
```



# NMDS

## prep data
remove non-data columns, convert to proportionas, and arcsine sqrt transformation
```{r}
RE2<- Zoop_env[,28:ncol(Zoop_env)]
#replace N/A with 0
RE2[is.na(RE2)] <- 0
#convert to proportions
RE2<-RE2/rowSums(RE2)

#arcsine sqrt transformation
RE2<-asin(sqrt(RE2))

#remove columns with fewer than 9 rows with nonzero numbers
non_zero_counts <- colSums(RE2 != 0)
RE2 <- RE2[, non_zero_counts >= 9]
```

indicator species
```{r}
inv = multipatt(RE2, Zoop_env$Sample.Month.x, func = "r.g", control = how(nperm=999))
summary(inv)
```
## Clustering

gap statistic: do this to determine number of clusters
```{r}
library(factoextra)
res.km<-eclust(RE2, FUNcluster = "hclust", hc_metric = "euclidean", graph = FALSE)
fviz_gap_stat(res.km$gap_stat)

#silhouette plot
fviz_silhouette(res.km)
```
9 is the ultimate number of clusters


bray-curtis heirarchical clustering
```{r}
library(vegan)
dist<-vegdist(RE2, method='bray',trymax = 200)
cluster <- hclust(dist, method = "average")
plot(cluster)
grp <- cutree(cluster, 3)
grp
```
I'm choosing 5 clusters, cutting off at 55% similarity

nmds
```{r}
ord <- metaMDS(RE2)
```

```{r}
col <- c("red2", "green4", "mediumblue","pink","purple","lightblue","black","lightgreen","yellow","orange")
col[grp]
```

plot
```{r}
plot(ord, type = "n", display = "sites")
points(ord, col = col[grp], bg = col[grp], pch = 21)
legend("topright", legend = paste("Cluster", 1:3),
       col = col, pt.bg = col, bty = "n", pch = 21)
```



## PERMANOVA
```{r}
dist<-vegdist(RE2, method='bray')
Zoop_env$STATION_NO<-as.factor(Zoop_env$STATION_NO)
Zoop_env$Year<-as.factor(Zoop_env$Year)
Zoop_env$Month<-as.factor(Zoop_env$Month)
perm<-adonis(dist~Year*STATION_NO*Month, data=Zoop_env, permutations = 999, method="bray")
perm
```
It looks like Sample.Year, station, and Sample.Month are significant. Additionally, the interaction between Sample.Year and Sample.Month and station and Sample.Month are different, need to split up

See pairwise differences
```{r}
#make sure factors are formatted right
Zoop_env$Year=as.factor(Zoop_env$Year)
Zoop_env$STATION_NO=as.factor(Zoop_env$STATION_NO)
Zoop_env$Month=as.factor(Zoop_env$Month)
adonis.pair(vegdist(RE2),Zoop_env$STATION_NO)
adonis.pair(vegdist(RE2),Zoop_env$Year)
adonis.pair(vegdist(RE2),Zoop_env$Month)
```

all stations are different from one another

2014 is similar to 2015
2014 is similar to 2020
2014 is similar to 2021
2015 is similar to 2016
2015 is similar to 2017
2015 is similar to 2021
2016 is similar to 2017
2016 is similar to 2018
2016 is similar to 2019
2017 is similar to 2018
2017 is similar to 2019
2018 is similar to 2019
2018 is similar to 2020
2018 is similar to 2021
2019 is similar to 2020
2019 is similar to 2021
2019 is similar to 2022
2020 is similar to 2021
2020 is similar to 2022

all Sample.Months are different from one another


check for dispersion
```{r}
#station
Zoop.bd <- betadisper(dist, Zoop_env$STATION_NO)
Zoop.bd
anova(Zoop.bd)
permutest(Zoop.bd)

#Sample.Year
Zoop.bd <- betadisper(dist, Zoop_env$Year)
Zoop.bd
anova(Zoop.bd)
permutest(Zoop.bd)

#Sample.Month
Zoop.bd <- betadisper(dist, Zoop_env$Month)
Zoop.bd
anova(Zoop.bd)
permutest(Zoop.bd)
```

only year is significant

### All Sample.Months NMDS


```{r}
NMDSmodel <- metaMDS(RE2, distance = "bray",trymax = 200)
NMDSmodel
```

get datascores
```{r}
data.scores<-scores(NMDSmodel,display="sites")
data.scores<- as.data.frame(data.scores)
```

add group, station, Sample.Month, Sample.Year back in
```{r}
data.scores$Station = Zoop_env$STATION_NO
data.scores$Sample.Year = Zoop_env$Year
data.scores$Sample.Month = Zoop_env$Month
data.scores$group = as.factor(grp)
```

add environmental variables back in
```{r}
colnames(Zoop_env)
data.scores$temp_deep = Zoop_env$CTDTMP_DEG_C_ITS90_Deep...6
data.scores$temp_surface = Zoop_env$CTDTMP_DEG_C_ITS90_Surface...7
data.scores$salinity_deep = Zoop_env$CTDSAL_PSS78_Deep...8
data.scores$salinity_surface = Zoop_env$CTDSAL_PSS78_Surface...9
data.scores$oxygen_deep = Zoop_env$OXYGEN_MG_L_1_Deep...10
data.scores$oxygen_surface = Zoop_env$OXYGEN_MG_L_1_Surface...11
data.scores$nitrate_deep = Zoop_env$NITRATE_UMOL_L_Deep...12
data.scores$nitrate_surface = Zoop_env$NITRATE_UMOL_L_Surface...13
data.scores$temp_surf_anom = Zoop_env$CTDTMP_DEG_C_ITS90_Surface...16
data.scores$temp_deep_anom = Zoop_env$CTDTMP_DEG_C_ITS90_Deep...15
data.scores$oxy_deep_anom = Zoop_env$OXYGEN_MG_L_1_Deep...19
data.scores$chla = Zoop_env$chla
```

plot data scores
```{r}
xx <- ggplot(data.scores, aes(x = NMDS1, y = NMDS2,colour=Station)) +stat_ellipse()+geom_point(aes(colour = Station))+theme_bw()+theme(panel.grid.major = element_blank(),
                                                                                           panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))
xx

xx <- ggplot(data.scores, aes(x = NMDS1, y = NMDS2,colour = Sample.Year)) +stat_ellipse()+geom_point(aes(colour = Sample.Year))+theme_bw()+theme(panel.grid.major = element_blank(),
                                                                                           panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))
xx

xx <- ggplot(data.scores, aes(x = NMDS1, y = NMDS2,colour = Sample.Month)) +stat_ellipse()+geom_point(aes(colour = Sample.Month))+theme_bw()+theme(panel.grid.major = element_blank(),
                                                                                           panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))
xx

xx <- ggplot(data.scores, aes(x = NMDS1, y = NMDS2,colour = group)) +stat_ellipse()+geom_point(aes(colour = group))+theme_bw()+theme(panel.grid.major = element_blank(),  panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))
xx
```

April is pretty clearly different than July and September

## environmental fitting
create environmetal fit table
```{r}
#check for correlation among environmental variables.
#remove nutrients other than nitrate because of autocorrelation

data.scores.cor<- data.scores[ -c(1:6) ]
cor_matrix <- cor(data.scores.cor)
corrplot(cor_matrix, method = "number", type = "lower", tl.pos = 'd')

Zoo.env <- data.frame(matrix(ncol = 0, nrow=139))
colnames(data.scores)
Zoo.env$temp_deep<-as.numeric(data.scores$temp_deep)
Zoo.env$temp_surface<-as.numeric(data.scores$temp_surface)
Zoo.env$salinity_deep<-as.numeric(data.scores$salinity_deep)
Zoo.env$salinity_surface<-as.numeric(data.scores$salinity_surface)
Zoo.env$oxygen_deep<-as.numeric(data.scores$oxygen_deep)
Zoo.env$oxygen_surface<-as.numeric(data.scores$oxygen_surface)
Zoo.env$temp_deep_anom<-as.numeric(data.scores$temp_deep_anom)
Zoo.env$temp_surf_anom<-as.numeric(data.scores$temp_surf_anom)
Zoo.env$oxy_deep_anom<-as.numeric(data.scores$oxy_deep_anom)
Zoo.env$chla<-as.numeric(data.scores$chla)

env <- envfit(NMDSmodel, Zoo.env, na.rm = TRUE,perm = 999)
env


env.scrs <- as.data.frame(scores(env, display = "vectors")) 
env.scrs <- cbind(env.scrs, Species = rownames(env.scrs), Pvalues = env$vectors$pvals, R_squared = env$vectors$r)
env.scrs <- subset(env.scrs, Pvalues < 0.05)


env.scrs
```

## Species fitting
```{r}
vf <- envfit(NMDSmodel, RE2, perm = 999)

spp.scrs <- as.data.frame(scores(vf, display = "vectors"))
spp.scrs <- cbind(spp.scrs, Species = rownames(spp.scrs))
####for ggplot
arrow_factor <- ordiArrowMul(vf)
spp.scrs <- as.data.frame(scores(vf, display = "vectors")) * arrow_factor
spp.scrs <- cbind(spp.scrs, Species = rownames(spp.scrs), Pvalues = vf$vectors$pvals, R_squared = vf$vectors$r)

# select significance similarly to `plot(vf, p.max = 0.01)`
spp.scrs <- subset(spp.scrs, Pvalues < 0.05)
spp.scrs <- subset(spp.scrs, R_squared > 0.3)
spp.scrs

```
### all data with species
extract month hulls
```{r}
G1 <- data.scores[data.scores$group == "1", ][chull(data.scores[data.scores$group ==                                                           "1", c("NMDS1", "NMDS2")]), ]  
G2 <- data.scores[data.scores$group == "2", ][chull(data.scores[data.scores$group ==                                                           "2", c("NMDS1", "NMDS2")]), ]  
G3 <- data.scores[data.scores$group == "3", ][chull(data.scores[data.scores$group ==                                                           "3", c("NMDS1", "NMDS2")]), ]
```
get hull data
```{r}
hull.data <- rbind(G1,G2,G3)
hull.data

```

```{r,fig.width=6.5,fig.height=5, dpi=600}

Species_plot <- ggplot(data.scores, aes(x = NMDS1, y = NMDS2)) +geom_point(aes(colour = Sample.Month))+theme_bw()+theme(panel.grid.major = element_blank(),
                                                                                           panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))+ 
  geom_polygon(data=hull.data,aes(x=NMDS1,y=NMDS2,group=group),alpha=0.20, size=0.1, linetype=1, colour="black") +
  annotate("text", label = "2D Stress: 0.18", x = 0.8, y = 1.3, size = 5, colour = "black")+ geom_segment(data = spp.scrs,size=0.2,
               aes(x = 0, xend = NMDS1, y = 0, yend = NMDS2),
               arrow = arrow(length = unit(0.08, "cm")), colour = "black")+
  geom_label_repel(data = spp.scrs, aes(x = NMDS1, y = NMDS2, label = Species),
                  size = 1.5, fontface="bold", fill="white", label.padding = unit(0.15, "lines"), box.padding = unit(0.16, "lines"), label.size = 0.05)
Species_plot
```

```{r,fig.width=6.5,fig.height=5, dpi=600}

Species_plot <- ggplot(data.scores, aes(x = NMDS1, y = NMDS2)) +geom_point(aes(colour = temp_deep_anom))+theme_bw()+theme(panel.grid.major = element_blank(),
                                                                                           panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))+ 
  geom_polygon(data=hull.data,aes(x=NMDS1,y=NMDS2,group=Station),alpha=0.05, size=0.1, linetype=1, colour="black") +
  annotate("text", label = "2D Stress: 0.18", x = 1, y = 1, size = 5, colour = "black")+ geom_segment(data = spp.scrs,size=0.2,
               aes(x = 0, xend = NMDS1, y = 0, yend = NMDS2),
               arrow = arrow(length = unit(0.08, "cm")), colour = "black")+
  geom_label_repel(data = spp.scrs, aes(x = NMDS1, y = NMDS2, label = Species),
                  size = 1.5, fontface="bold", fill="white", label.padding = unit(0.15, "lines"), box.padding = unit(0.16, "lines"), label.size = 0.05)+scale_color_gradient(low="blue", high="red")+ 
  geom_polygon(data=apr.hull.data,aes(x=NMDS1,y=NMDS2,group=Station),alpha=0.20, size=0.1, linetype=1, colour="black")
Species_plot
```


color by group
```{r,fig.width=6.5,fig.height=5, dpi=600}

Species_plot <- ggplot(data.scores, aes(x = NMDS1, y = NMDS2)) +geom_point(aes(colour = group))+theme_bw()+theme(panel.grid.major = element_blank(),
                                                                                           panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))+ 
  geom_polygon(data=hull.data,aes(x=NMDS1,y=NMDS2,fill=Station,group=Station),alpha=0.20, size=0.1, linetype=1, colour="black") +
  annotate("text", label = "2D Stress: 0.18", x = 1, y = 1, size = 5, colour = "black")+ scale_colour_manual(values=c("#F8766D","#C49A00","#53B400","#00C094","#00B6EB","#A69AFF","#FB61D7","yellow","red","orange")) +
  scale_fill_manual(values=c("#F8766D","#C49A00","#53B400","#00C094","#00B6EB","#A69AFF","#FB61D7","yellow","red","orange"))+ geom_segment(data = spp.scrs,size=0.2,
               aes(x = 0, xend = NMDS1, y = 0, yend = NMDS2),
               arrow = arrow(length = unit(0.08, "cm")), colour = "black")+
  geom_label_repel(data = spp.scrs, aes(x = NMDS1, y = NMDS2, label = Species),
                  size = 1.5, fontface="bold", fill="white", label.padding = unit(0.15, "lines"), box.padding = unit(0.16, "lines"), label.size = 0.05)
Species_plot
```


## plot only April

subset data scores
```{r}
APR_datascores<-subset(data.scores,Sample.Month=="APR")
```
extract hulls
```{r}
P12 <- APR_datascores[APR_datascores$Station == "P12", ][chull(APR_datascores[APR_datascores$Station ==                                                           "P12", c("NMDS1", "NMDS2")]), ]  
P22 <- APR_datascores[APR_datascores$Station == "P22", ][chull(APR_datascores[APR_datascores$Station ==                                                           "P22", c("NMDS1", "NMDS2")]), ]  
P28 <- APR_datascores[APR_datascores$Station == "P28", ][chull(APR_datascores[APR_datascores$Station ==                                                           "P28", c("NMDS1", "NMDS2")]), ]
P38 <- APR_datascores[APR_datascores$Station == "P38", ][chull(APR_datascores[APR_datascores$Station ==                                                           "P38", c("NMDS1", "NMDS2")]), ]  
P4 <- APR_datascores[APR_datascores$Station == "P4", ][chull(APR_datascores[APR_datascores$Station ==                                                           "P4", c("NMDS1", "NMDS2")]), ]  
P402 <- APR_datascores[APR_datascores$Station == "P402", ][chull(APR_datascores[APR_datascores$Station ==                                                           "P402", c("NMDS1", "NMDS2")]), ]  
P8 <- APR_datascores[APR_datascores$Station == "P8", ][chull(APR_datascores[APR_datascores$Station ==                                                           "P8", c("NMDS1", "NMDS2")]), ]  
```

get hull data
```{r}
apr.hull.data <- rbind(P12, P22, P28,P38,P4,P402,P8)  #combine grp.a and grp.b
apr.hull.data

```
plot vectors
```{r,fig.width=5,fig.height=5.5, dpi=600}

Blank <- ggplot(APR_datascores, aes(x = NMDS1, y = NMDS2)) +theme_bw()+theme(panel.grid.major = element_blank(),
                                                                                           panel.grid.minor = element_blank(), axis.line = element_line(colour = "black")) +
  annotate("text", label = "2D Stress: 0.18", x = 0.7, y = 1, size = 5, colour = "black")+ scale_colour_manual(values=c("#F8766D","#C49A00","#53B400","#00C094","#00B6EB","#A69AFF","#FB61D7")) +
  scale_fill_manual(values=c("#F8766D","#C49A00","#53B400","#00C094","#00B6EB","#A69AFF","#FB61D7"))+
  theme(axis.text=element_text(size=12),legend.text=element_text(size=12),legend.title=element_text(size=12))+
  geom_segment(data = env.scrs,size=0.5,
               aes(x = 0, xend = NMDS1, y = 0, yend = NMDS2),
               arrow = arrow(length = unit(0.5, "cm")), colour = "black")+xlim(-0.8, 1)+ylim(-1.3,1)
Blank
```
save dendogram
```{r, eval=FALSE}
setwd("/Users/hailaschultz/Dropbox/Schultz_Dissertation/Data_Analysis/Schultz_dissertation-2/figures")
ggsave(filename = "NMDS_blank.png", plot = Blank, width = 5, height = 5.5, device='png', dpi=700)
```


plot data scores
```{r,fig.width=5,fig.height=5.5, dpi=600}
#change factor order
APR_datascores$Station <- factor(APR_datascores$Station, levels = c("P22", "P4", "P8","P28","P12","P402","P38"))
apr.hull.data$Station <- factor(apr.hull.data$Station, levels = c("P22", "P4", "P8","P28","P12","P402","P38"))

April <- ggplot(APR_datascores, aes(x = NMDS1, y = NMDS2)) +geom_point(aes(colour = Station),size=2.5)+theme_bw()+theme(panel.grid.major = element_blank(),
                                                                                           panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))+ 
  geom_polygon(data=apr.hull.data,aes(x=NMDS1,y=NMDS2,fill=Station,group=Station),alpha=0.20, size=0.1, linetype=1, colour="black") +
  annotate("text", label = "2D Stress: 0.18", x = 0.7, y = 1, size = 5, colour = "black")+ scale_colour_manual(values=c("#F8766D","#C49A00","#53B400","#00C094","#00B6EB","#A69AFF","#FB61D7")) +
  scale_fill_manual(values=c("#F8766D","#C49A00","#53B400","#00C094","#00B6EB","#A69AFF","#FB61D7"))+
  theme(axis.text=element_text(size=12))+
  geom_segment(data = env.scrs,size=0.5,
               aes(x = 0, xend = NMDS1, y = 0, yend = NMDS2),
               arrow = arrow(length = unit(0.5, "cm")), colour = "black")+xlim(-0.8, 1)+ylim(-1.3,1)+ theme(legend.position = "none")
April
```
save dendogram
```{r, eval=FALSE}
setwd("/Users/hailaschultz/Dropbox/Schultz_Dissertation/Data_Analysis/Schultz_dissertation-2/figures")
ggsave(filename = "NMDS_April.png", plot = April, width = 5, height = 5.5, device='png', dpi=700)
```


color by anomaly
```{r,fig.width=6.5,fig.height=5, dpi=600}
#change factor order
APR_datascores$Station <- factor(APR_datascores$Station, levels = c("P22", "P4", "P8","P28","P12","P402","P38"))
apr.hull.data$Station <- factor(apr.hull.data$Station, levels = c("P22", "P4", "P8","P28","P12","P402","P38"))

April <- ggplot(APR_datascores, aes(x = NMDS1, y = NMDS2)) +geom_point(aes(colour = temp_deep_anom),size=2.5)+theme_bw()+theme(panel.grid.major = element_blank(),
                                                                                           panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))+
  annotate("text", label = "2D Stress: 0.16", x = 0.5, y = 0.57, size = 5, colour = "black")+ 
  theme(axis.text=element_text(size=12),legend.text=element_text(size=12),legend.title=element_text(size=12))+
  geom_segment(data = env.scrs,size=0.5,
               aes(x = 0, xend = NMDS1, y = 0, yend = NMDS2),
               arrow = arrow(length = unit(0.5, "cm")), colour = "black")+
  geom_label_repel(data = env.scrs, aes(x = NMDS1, y = NMDS2, label = Species),
                  size = 5, fontface="bold", fill="white", label.padding = unit(0.15, "lines"), point.padding = unit(15, "lines"), label.size = 0.05)+scale_color_gradient(low="blue", high="red")+ 
  geom_polygon(data=apr.hull.data,aes(x=NMDS1,y=NMDS2,group=Station),alpha=0.20, size=0.1, linetype=1, colour="black")
April
```


## plot only July

subset data scores
```{r}
JUL_datascores<-subset(data.scores,Sample.Month=="JUL")
```
extract hulls
```{r}
P12 <- JUL_datascores[JUL_datascores$Station == "P12", ][chull(JUL_datascores[JUL_datascores$Station ==                                                           "P12", c("NMDS1", "NMDS2")]), ]  
P22 <- JUL_datascores[JUL_datascores$Station == "P22", ][chull(JUL_datascores[JUL_datascores$Station ==                                                           "P22", c("NMDS1", "NMDS2")]), ]  
P28 <- JUL_datascores[JUL_datascores$Station == "P28", ][chull(JUL_datascores[JUL_datascores$Station ==                                                           "P28", c("NMDS1", "NMDS2")]), ]
P38 <- JUL_datascores[JUL_datascores$Station == "P38", ][chull(JUL_datascores[JUL_datascores$Station ==                                                           "P38", c("NMDS1", "NMDS2")]), ]  
P4 <- JUL_datascores[JUL_datascores$Station == "P4", ][chull(JUL_datascores[JUL_datascores$Station ==                                                           "P4", c("NMDS1", "NMDS2")]), ]  
P402 <- JUL_datascores[JUL_datascores$Station == "P402", ][chull(JUL_datascores[JUL_datascores$Station ==                                                           "P402", c("NMDS1", "NMDS2")]), ]  
P8 <- JUL_datascores[JUL_datascores$Station == "P8", ][chull(JUL_datascores[JUL_datascores$Station ==                                                           "P8", c("NMDS1", "NMDS2")]), ]  
```

get hull data
```{r}
jul.hull.data <- rbind(P12, P22, P28,P38,P4,P402,P8)  #combine grp.a and grp.b
jul.hull.data
```

plot data scores
```{r,fig.width=5,fig.height=5.5, dpi=600}
#change factor order
JUL_datascores$Station <- factor(JUL_datascores$Station, levels = c("P22", "P4", "P8","P28","P12","P402","P38"))
jul.hull.data$Station <- factor(jul.hull.data$Station, levels = c("P22", "P4", "P8","P28","P12","P402","P38"))

July <- ggplot(JUL_datascores, aes(x = NMDS1, y = NMDS2)) +geom_point(aes(colour = Station),size=2.5)+theme_bw()+theme(panel.grid.major = element_blank(),
                                                                                           panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))+ 
  geom_polygon(data=jul.hull.data,aes(x=NMDS1,y=NMDS2,fill=Station,group=Station),alpha=0.20, size=0.1, linetype=1, colour="black") +
  annotate("text", label = "2D Stress: 0.18", x = 0.7, y = 1, size = 5, colour = "black")+ scale_colour_manual(values=c("#F8766D","#C49A00","#53B400","#00C094","#00B6EB","#A69AFF","#FB61D7")) +
  scale_fill_manual(values=c("#F8766D","#C49A00","#53B400","#00C094","#00B6EB","#A69AFF","#FB61D7"))+
  theme(axis.text=element_text(size=12))+
  geom_segment(data = env.scrs,size=0.5,
               aes(x = 0, xend = NMDS1, y = 0, yend = NMDS2),
               arrow = arrow(length = unit(0.5, "cm")), colour = "black")+xlim(-0.8, 1)+ylim(-1.3,1)+ theme(legend.position = "none")
July
```
save dendogram
```{r, eval=FALSE}
setwd("/Users/hailaschultz/Dropbox/Schultz_Dissertation/Data_Analysis/Schultz_dissertation-2/figures")
ggsave(filename = "NMDS_July.png", plot = July, width = 5, height = 5.5, device='png', dpi=700)
```

## plot only September

subset data scores
```{r}
SEP_datascores<-subset(data.scores,Sample.Month=="SEP")
```
extract hulls
```{r}
P12 <- SEP_datascores[SEP_datascores$Station == "P12", ][chull(SEP_datascores[SEP_datascores$Station ==                                                           "P12", c("NMDS1", "NMDS2")]), ]  
P22 <- SEP_datascores[SEP_datascores$Station == "P22", ][chull(SEP_datascores[SEP_datascores$Station ==                                                           "P22", c("NMDS1", "NMDS2")]), ]  
P28 <- SEP_datascores[SEP_datascores$Station == "P28", ][chull(SEP_datascores[SEP_datascores$Station ==                                                           "P28", c("NMDS1", "NMDS2")]), ]
P38 <- SEP_datascores[SEP_datascores$Station == "P38", ][chull(SEP_datascores[SEP_datascores$Station ==                                                           "P38", c("NMDS1", "NMDS2")]), ]  
P4 <- SEP_datascores[SEP_datascores$Station == "P4", ][chull(SEP_datascores[SEP_datascores$Station ==                                                           "P4", c("NMDS1", "NMDS2")]), ]  
P402 <- SEP_datascores[SEP_datascores$Station == "P402", ][chull(SEP_datascores[SEP_datascores$Station ==                                                           "P402", c("NMDS1", "NMDS2")]), ]  
P8 <- SEP_datascores[SEP_datascores$Station == "P8", ][chull(SEP_datascores[SEP_datascores$Station ==                                                           "P8", c("NMDS1", "NMDS2")]), ]  
```

get hull data
```{r}
SEP.hull.data <- rbind(P12, P22, P28,P38,P4,P402,P8)  #combine grp.a and grp.b
SEP.hull.data
```

plot data scores
```{r,fig.width=5,fig.height=5.5, dpi=600}
#change factor order
SEP_datascores$Station <- factor(SEP_datascores$Station, levels = c("P22", "P4", "P8","P28","P12","P402","P38"))
SEP.hull.data$Station <- factor(SEP.hull.data$Station, levels = c("P22", "P4", "P8","P28","P12","P402","P38"))

September <- ggplot(SEP_datascores, aes(x = NMDS1, y = NMDS2)) +geom_point(aes(colour = Station),size=2.5)+theme_bw()+theme(panel.grid.major = element_blank(),
                                                                                           panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))+ 
  geom_polygon(data=SEP.hull.data,aes(x=NMDS1,y=NMDS2,fill=Station,group=Station),alpha=0.20, size=0.1, linetype=1, colour="black") +
  annotate("text", label = "2D Stress: 0.18", x = 0.7, y = 1, size = 5, colour = "black")+ scale_colour_manual(values=c("#F8766D","#C49A00","#53B400","#00C094","#00B6EB","#A69AFF","#FB61D7")) +
  scale_fill_manual(values=c("#F8766D","#C49A00","#53B400","#00C094","#00B6EB","#A69AFF","#FB61D7"))+
  theme(axis.text=element_text(size=12))+
  geom_segment(data = env.scrs,size=0.5,
               aes(x = 0, xend = NMDS1, y = 0, yend = NMDS2),
               arrow = arrow(length = unit(0.5, "cm")), colour = "black")+xlim(-0.8, 1)+ylim(-1.3,1)+ theme(legend.position = "none")
September
```
save dendogram
```{r, eval=FALSE}
setwd("/Users/hailaschultz/Dropbox/Schultz_Dissertation/Data_Analysis/Schultz_dissertation-2/figures")
ggsave(filename = "NMDS_September.png", plot = September, width = 5, height = 5.5, device='png', dpi=700)
```
species scores
```{r,fig.width=5,fig.height=5.5, dpi=600}


September_sp <- ggplot(SEP_datascores, aes(x = NMDS1, y = NMDS2)) +geom_point(aes(colour = Station),size=2.5)+theme_bw()+theme(panel.grid.major = element_blank(),
                                                                                           panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))+ 
  geom_polygon(data=SEP.hull.data,aes(x=NMDS1,y=NMDS2,fill=Station,group=Station),alpha=0.20, size=0.1, linetype=1, colour="black") +
  annotate("text", label = "2D Stress: 0.18", x = 0.7, y = 1, size = 5, colour = "black")+ scale_colour_manual(values=c("#F8766D","#C49A00","#53B400","#00C094","#00B6EB","#A69AFF","#FB61D7")) +
  scale_fill_manual(values=c("#F8766D","#C49A00","#53B400","#00C094","#00B6EB","#A69AFF","#FB61D7"))+
  theme(axis.text=element_text(size=12),legend.text=element_text(size=12),legend.title=element_text(size=12))+
  geom_segment(data = spp.scrs,size=0.4,
               aes(x = 0, xend = NMDS1, y = 0, yend = NMDS2),
               arrow = arrow(length = unit(0.2, "cm")), colour = "darkgrey")+xlim(-1, 1)+ylim(-1.3,1)+ theme(legend.position = "none")+
  geom_segment(data = env.scrs,size=0.4,
               aes(x = 0, xend = NMDS1, y = 0, yend = NMDS2),
               arrow = arrow(length = unit(0.2, "cm")), colour = "black")

September_sp
```

save dendogram
```{r, eval=FALSE}
setwd("/Users/hailaschultz/Dropbox/Schultz_Dissertation/Data_Analysis/Schultz_dissertation-2/figures")
ggsave(filename = "NMDS_September_Sp.png", plot = September_sp, width = 5, height = 5.5, device='png', dpi=700)
```

### try group 1
remove non-data columns, convert to proportionas, and arcsine sqrt transformation
```{r}
Zoop_env$group = as.factor(grp)
group1<-subset(Zoop_env,group=="1")

RE2<- group1[,28:ncol(group1)]
RE2 <- RE2[, -ncol(RE2)]
#replace N/A with 0
RE2[is.na(RE2)] <- 0
#convert to proportions
RE2<-RE2/rowSums(RE2)

#arcsine sqrt transformation
RE2<-asin(sqrt(RE2))
#remove columns with fewer than 9 rows with nonzero numbers
non_zero_counts <- colSums(RE2 != 0)
RE2 <- RE2[, non_zero_counts >= 9]
```

## PERMANOVA
```{r}
colnames(group1)
dist<-vegdist(RE2, method='bray')
perm<-adonis2(dist~STATION_NO*Year, data=group1, permutations = 999, method="bray")
perm
```
See pairwise differences
```{r}
#make sure factors are formatted right
group1$STATION_NO<-as.factor(group1$STATION_NO)
group1$Year=as.factor(group1$Year)
adonis.pair(vegdist(RE2),group1$STATION_NO)
adonis.pair(vegdist(RE2),group1$Year)
```

## Clustering

gap statistic: do this to determine number of clusters
```{r}
library(factoextra)
res.km<-eclust(RE2, FUNcluster = "hclust", hc_metric = "euclidean", graph = FALSE)
fviz_gap_stat(res.km$gap_stat)

#silhouette plot
fviz_silhouette(res.km)
```
9 is the ultimate number of clusters

bray-curtis heirarchical clustering
```{r}
dist<-vegdist(RE2, method='bray',trymax = 200)
cluster <- hclust(dist, method = "average")
plot(cluster)
grp <- cutree(cluster, 9)
grp
```

nmds
```{r}
ord <- metaMDS(RE2)
```

```{r}
col <- c("red2", "green4", "mediumblue","pink","purple","lightblue","black","orange","yellow")
col[grp]
```

plot
```{r}
plot(ord, type = "n", display = "sites")
points(ord, col = col[grp], bg = col[grp], pch = 21)
legend("topright", legend = paste("Cluster", 1:3),
       col = col, pt.bg = col, bty = "n", pch = 21)
```

## Run NMDS
```{r}
NMDSmodel <- metaMDS(RE2, distance = "bray",trymax = 200)
NMDSmodel
```



get datascores
```{r}
data.scores<-scores(NMDSmodel,display="sites")
data.scores<- as.data.frame(data.scores)
```

add station, Sample.Month, Sample.Year back in
```{r}
data.scores$Station = group1$STATION_NO
data.scores$Sample.Year = group1$Sample.Year.x
data.scores$Sample.Month = group1$Sample.Month.x
data.scores$group = grp
```

add environmental variables back in
```{r}
data.scores$temp_deep = group1$CTDTMP_DEG_C_ITS90_Deep...6
data.scores$temp_surface = group1$CTDTMP_DEG_C_ITS90_Surface...7
data.scores$salinity_deep = group1$CTDSAL_PSS78_Deep...8
data.scores$salinity_surface = group1$CTDSAL_PSS78_Surface...9
data.scores$oxygen_deep = group1$OXYGEN_MG_L_1_Deep...10
data.scores$oxygen_surface = group1$OXYGEN_MG_L_1_Surface...11
data.scores$nitrate_deep = group1$NITRATE_UMOL_L_Deep...12
data.scores$nitrate_surface = group1$NITRATE_UMOL_L_Surface...13
data.scores$temp_surf_anom = group1$CTDTMP_DEG_C_ITS90_Surface...16
data.scores$temp_deep_anom = group1$CTDTMP_DEG_C_ITS90_Deep...15
data.scores$oxy_deep_anom = group1$OXYGEN_MG_L_1_Deep...19
data.scores$chla = group1$chla
```

plot data scores
```{r}
xx <- ggplot(data.scores, aes(x = NMDS1, y = NMDS2,colour=Station)) +stat_ellipse()+geom_point(aes(colour = Station))+theme_bw()+theme(panel.grid.major = element_blank(),
                                                                                           panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))
xx

xx <- ggplot(data.scores, aes(x = NMDS1, y = NMDS2,colour = Year)) +stat_ellipse()+geom_point(aes(colour = Year))+theme_bw()+theme(panel.grid.major = element_blank(),
                                                                                           panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))
xx

xx <- ggplot(data.scores, aes(x = NMDS1, y = NMDS2,colour = Sample.Month)) +stat_ellipse()+geom_point(aes(colour = Sample.Month))+theme_bw()+theme(panel.grid.major = element_blank(),
                                                                                           panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))
xx

xx <- ggplot(data.scores, aes(x = NMDS1, y = NMDS2))+geom_point(aes(colour = temp_surf_anom))+theme_bw()+theme(panel.grid.major = element_blank(),
                                                                                           panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))+scale_color_gradient(low="blue", high="red")
xx
```

### all data with species
extract hulls
```{r}
G1 <- data.scores[data.scores$group == "1", ][chull(data.scores[data.scores$group ==                                                           "1", c("NMDS1", "NMDS2")]), ]  
G2 <- data.scores[data.scores$group == "2", ][chull(data.scores[data.scores$group ==                                                           "2", c("NMDS1", "NMDS2")]), ]  
G3 <- data.scores[data.scores$group == "3", ][chull(data.scores[data.scores$group ==                                                           "3", c("NMDS1", "NMDS2")]), ]
G4 <- data.scores[data.scores$group == "4", ][chull(data.scores[data.scores$group ==                                                           "4", c("NMDS1", "NMDS2")]), ]  
G5 <- data.scores[data.scores$group == "5", ][chull(data.scores[data.scores$group ==                                                           "5", c("NMDS1", "NMDS2")]), ]  
G6 <- data.scores[data.scores$group == "6", ][chull(data.scores[data.scores$group ==                                                           "6", c("NMDS1", "NMDS2")]), ]  
G7 <- data.scores[data.scores$group == "7", ][chull(data.scores[data.scores$group ==                                                           "7", c("NMDS1", "NMDS2")]), ] 
G8 <- data.scores[data.scores$group == "8", ][chull(data.scores[data.scores$group ==                                                           "8", c("NMDS1", "NMDS2")]), ]
G9 <- data.scores[data.scores$group == "9", ][chull(data.scores[data.scores$group ==                                                           "9", c("NMDS1", "NMDS2")]), ]
```
get hull data
```{r}
hull.data <- rbind(G1,G2,G3,G4,G5,G6,G7,G8,G9)  #combine grp.a and grp.b
hull.data

```

```{r,fig.width=6.5,fig.height=5, dpi=600}

Species_plot <- ggplot(data.scores, aes(x = NMDS1, y = NMDS2)) +geom_point(aes(colour = Station))+theme_bw()+theme(panel.grid.major = element_blank(),
                                                                                           panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))+ 
  geom_polygon(data=hull.data,aes(x=NMDS1,y=NMDS2,fill=as.factor(group),group=as.factor(group)),alpha=0.20, size=0.1, linetype=1, colour="black") +
  annotate("text", label = "2D Stress: 0.16", x = 0.5, y = 0.57, size = 5, colour = "black")+ geom_segment(data = spp.scrs,size=0.2,
               aes(x = 0, xend = NMDS1, y = 0, yend = NMDS2),
               arrow = arrow(length = unit(0.08, "cm")), colour = "black")+
  geom_label_repel(data = spp.scrs, aes(x = NMDS1, y = NMDS2, label = Species),
                  size = 1.5, fontface="bold", fill="white", label.padding = unit(0.15, "lines"), box.padding = unit(0.16, "lines"), label.size = 0.05)
Species_plot
```




## NMDS: Subset to April
subset to April
```{r}
Zoop_env_april<-subset(Zoop_env,Month=="APR")
```

## prep data
remove non-data columns, convert to proportionas, and arcsine sqrt transformation
```{r}
RE2<- Zoop_env_april[,28:ncol(Zoop_env_april)]
#replace N/A with 0
RE2[is.na(RE2)] <- 0
#convert to proportions
RE2<-RE2/rowSums(RE2)

#arcsine sqrt transformation
RE2<-asin(sqrt(RE2))

#remove columns with fewer than 9 rows with nonzero numbers
non_zero_counts <- colSums(RE2 != 0)
RE2 <- RE2[, non_zero_counts >= 9]
```


## Clustering

gap statistic: do this to determine number of clusters
```{r}
library(factoextra)
res.km<-eclust(RE2, FUNcluster = "hclust", hc_metric = "euclidean", graph = FALSE)
fviz_gap_stat(res.km$gap_stat)

#silhouette plot
fviz_silhouette(res.km)
```
4 is the ultimate number of clusters

bray-curtis heirarchical clustering
```{r}
dist<-vegdist(RE2, method='bray',trymax = 200)
cluster <- hclust(dist, method = "average")
plot(cluster)
grp <- cutree(cluster, 4)
grp
```


nmds
```{r}
ord <- metaMDS(RE2)
```

```{r}
col <- c("red2", "green4", "mediumblue","pink","purple","lightblue","black")
col[grp]
```

plot
```{r}
plot(ord, type = "n", display = "sites")
points(ord, col = col[grp], bg = col[grp], pch = 21)
legend("topright", legend = paste("Cluster", 1:3),
       col = col, pt.bg = col, bty = "n", pch = 21)
```


## PERMANOVA
```{r}
dist<-vegdist(RE2, method='bray')
perm<-adonis(dist~STATION_NO, data=Zoop_env_april, permutations = 999, method="bray")
perm
```
It looks like Sample.Year, station, and Sample.Month are significant. Additionally, the interaction between Sample.Year and Sample.Month and station and Sample.Month are different.

See pairwise differences
```{r}
#make sure factors are formatted right
Zoop_env_april$STATION_NO=as.factor(Zoop_env_april$STATION_NO)
adonis.pair(vegdist(RE2),Zoop_env_april$STATION_NO)
```

all stations are different from one another


check for dispersion
```{r}
#station
Zoop.bd <- betadisper(dist, Zoop_env_april$STATION_NO)
Zoop.bd
anova(Zoop.bd)
permutest(Zoop.bd)
```
no dispersion issues!

NMDS
```{r}
NMDSmodel <- metaMDS(RE2, distance = "bray",trymax = 200)
NMDSmodel
```

get datascores
```{r}
data.scores<-scores(NMDSmodel,display="sites")
data.scores<- as.data.frame(data.scores)
```

add group, station, Sample.Month, Sample.Year back in
```{r}
data.scores$Station = Zoop_env_april$STATION_NO
data.scores$Sample.Year = Zoop_env_april$Year
data.scores$Sample.Month = Zoop_env_april$Month
data.scores$group = as.factor(grp)
```

add environmental variables back in
```{r}
colnames(Zoop_env_april)
data.scores$temp_deep = Zoop_env_april$CTDTMP_DEG_C_ITS90_Deep...6
data.scores$temp_surface = Zoop_env_april$CTDTMP_DEG_C_ITS90_Surface...7
data.scores$salinity_deep = Zoop_env_april$CTDSAL_PSS78_Deep...8
data.scores$salinity_surface = Zoop_env_april$CTDSAL_PSS78_Surface...9
data.scores$oxygen_deep = Zoop_env_april$OXYGEN_MG_L_1_Deep...10
data.scores$oxygen_surface = Zoop_env_april$OXYGEN_MG_L_1_Surface...11
data.scores$nitrate_deep = Zoop_env_april$NITRATE_UMOL_L_Deep...12
data.scores$nitrate_surface = Zoop_env_april$NITRATE_UMOL_L_Surface...13
data.scores$temp_surf_anom = Zoop_env_april$CTDTMP_DEG_C_ITS90_Surface...16
data.scores$temp_deep_anom = Zoop_env_april$CTDTMP_DEG_C_ITS90_Deep...15
data.scores$oxy_deep_anom = Zoop_env_april$OXYGEN_MG_L_1_Deep...19
data.scores$chla = Zoop_env_april$chla
```

plot data scores
```{r}
xx <- ggplot(data.scores, aes(x = NMDS1, y = NMDS2,colour=Station)) +stat_ellipse()+geom_point(aes(colour = Station))+theme_bw()+theme(panel.grid.major = element_blank(),
                                                                                           panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))
xx

xx <- ggplot(data.scores, aes(x = NMDS1, y = NMDS2,colour = Sample.Year)) +stat_ellipse()+geom_point(aes(colour = Sample.Year))+theme_bw()+theme(panel.grid.major = element_blank(),
                                                                                           panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))
xx

```


## environmental fitting
create environmetal fit table
```{r}
#check for correlation among environmental variables.
#remove nutrients other than nitrate because of autocorrelation
#remove temperature deep because of correlation

data.scores.cor<- data.scores[ -c(1:6) ]
cor_matrix <- cor(data.scores.cor)
corrplot(cor_matrix, method = "number", type = "lower", tl.pos = 'd')

Zoo.env <- data.frame(matrix(ncol = 0, nrow=38))
colnames(data.scores)
Zoo.env$temp_surface<-as.numeric(data.scores$temp_surface)
Zoo.env$salinity_deep<-as.numeric(data.scores$salinity_deep)
Zoo.env$salinity_surface<-as.numeric(data.scores$salinity_surface)
Zoo.env$oxygen_deep<-as.numeric(data.scores$oxygen_deep)
Zoo.env$oxygen_surface<-as.numeric(data.scores$oxygen_surface)
Zoo.env$temp_deep_anom<-as.numeric(data.scores$temp_deep_anom)
Zoo.env$temp_surf_anom<-as.numeric(data.scores$temp_surf_anom)
Zoo.env$oxy_deep_anom<-as.numeric(data.scores$oxy_deep_anom)
Zoo.env$chla<-as.numeric(data.scores$chla)

env <- envfit(NMDSmodel, Zoo.env, na.rm = TRUE,perm = 999)
env


env.scrs <- as.data.frame(scores(env, display = "vectors")) 
env.scrs <- cbind(env.scrs, Species = rownames(env.scrs), Pvalues = env$vectors$pvals, R_squared = env$vectors$r)
env.scrs <- subset(env.scrs, Pvalues < 0.05)


env.scrs
```

## Species fitting
```{r}
vf <- envfit(NMDSmodel, RE2, perm = 999)

spp.scrs <- as.data.frame(scores(vf, display = "vectors"))
spp.scrs <- cbind(spp.scrs, Species = rownames(spp.scrs))
####for ggplot
arrow_factor <- ordiArrowMul(vf)
spp.scrs <- as.data.frame(scores(vf, display = "vectors")) * arrow_factor
spp.scrs <- cbind(spp.scrs, Species = rownames(spp.scrs), Pvalues = vf$vectors$pvals, R_squared = vf$vectors$r)

# select significance similarly to `plot(vf, p.max = 0.01)`
spp.scrs <- subset(spp.scrs, Pvalues < 0.05)
spp.scrs <- subset(spp.scrs, R_squared > 0.3)
spp.scrs

```
### all data with species

extract hulls

```{r}
P12 <- data.scores[data.scores$Station == "P12", ][chull(data.scores[data.scores$Station ==                                                           "P12", c("NMDS1", "NMDS2")]), ]  
P22 <- data.scores[data.scores$Station == "P22", ][chull(data.scores[data.scores$Station ==                                                           "P22", c("NMDS1", "NMDS2")]), ]  
P28 <- data.scores[data.scores$Station == "P28", ][chull(data.scores[data.scores$Station ==                                                           "P28", c("NMDS1", "NMDS2")]), ]
P38 <- data.scores[data.scores$Station == "P38", ][chull(data.scores[data.scores$Station ==                                                           "P38", c("NMDS1", "NMDS2")]), ]  
P4 <- data.scores[data.scores$Station == "P4", ][chull(data.scores[data.scores$Station ==                                                           "P4", c("NMDS1", "NMDS2")]), ]  
P402 <- data.scores[data.scores$Station == "P402", ][chull(data.scores[data.scores$Station ==                                                           "P402", c("NMDS1", "NMDS2")]), ]  
P8 <- data.scores[data.scores$Station == "P8", ][chull(data.scores[data.scores$Station ==                                                           "P8", c("NMDS1", "NMDS2")]), ]  
```


get hull data
```{r}
apr.hull.data <- rbind(P12, P22, P28,P38,P4,P402,P8)  #combine grp.a and grp.b
apr.hull.data
```

```{r,fig.width=6.5,fig.height=5, dpi=600}

April_Species <- ggplot(data.scores, aes(x = NMDS1, y = NMDS2)) +geom_point(aes(colour = Station))+theme_bw()+theme(panel.grid.major = element_blank(),
                                                                                           panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))+ 
  geom_polygon(data=apr.hull.data,aes(x=NMDS1,y=NMDS2,group=Station, fill=Station),alpha=0.20, size=0.1, linetype=1, colour = "black") +
  annotate("text", label = "2D Stress: 0.15", x = 0.66, y = 0.66, size = 4, colour = "black")+ geom_segment(data = env.scrs,size=0.2,
               aes(x = 0, xend = NMDS1, y = 0, yend = NMDS2),
               arrow = arrow(length = unit(0.08, "cm")), colour = "black")+
  geom_label_repel(data = env.scrs, aes(x = NMDS1, y = NMDS2, label = Species),
                  size = 2.5, fontface="bold", fill="white", label.padding = unit(0.15, "lines"), box.padding = unit(0.16, "lines"), label.size = 0.05)+ scale_colour_manual(values=c("#00B6EB","#F8766D","#00C094","#FB61D7","#C49A00","#A69AFF","#53B400")) +
  scale_fill_manual(values=c("#00B6EB","#F8766D","#00C094","#FB61D7","#C49A00","#A69AFF","#53B400"))
April_Species

```
save dendogram
```{r, eval=FALSE}
setwd("/Users/hailaschultz/Dropbox/Schultz_Dissertation/Data_Analysis/Schultz_dissertation-2/figures")
ggsave(filename = "April_Species_month.png", plot = April_Species, width = 6.5, height = 5.5, device='png', dpi=700)
```

```{r}
G1 <- data.scores[data.scores$group == "1", ][chull(data.scores[data.scores$group ==                                                           "1", c("NMDS1", "NMDS2")]), ]  
G2 <- data.scores[data.scores$group == "2", ][chull(data.scores[data.scores$group ==                                                           "2", c("NMDS1", "NMDS2")]), ]  
G3 <- data.scores[data.scores$group == "3", ][chull(data.scores[data.scores$group ==                                                           "3", c("NMDS1", "NMDS2")]), ]
G4 <- data.scores[data.scores$group == "4", ][chull(data.scores[data.scores$group ==                                                           "4", c("NMDS1", "NMDS2")]), ]  

```
get hull data
```{r}
hull.data <- rbind(G1,G2,G3,G4)  #combine grp.a and grp.b
hull.data

```


```{r,fig.width=6.5,fig.height=5, dpi=600}

April_Species_vect <- ggplot(data.scores, aes(x = NMDS1, y = NMDS2)) +geom_point(aes(colour = Station))+theme_bw()+theme(panel.grid.major = element_blank(),
                                                                                           panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))+ 
  geom_polygon(data=hull.data,aes(x=NMDS1,y=NMDS2,group=group),alpha=0.20, size=0.1, linetype=1, fill=NA, colour = "black") +
  annotate("text", label = "2D Stress: 0.15", x = 0.66, y = 0.66, size = 4, colour = "black")+ geom_segment(data = spp.scrs,size=0.2,
               aes(x = 0, xend = NMDS1, y = 0, yend = NMDS2),
               arrow = arrow(length = unit(0.08, "cm")), colour = "gray")+
  geom_label_repel(data = spp.scrs, aes(x = NMDS1, y = NMDS2, label = Species),
                  size = 2, fontface="bold", fill="white", label.padding = unit(0.15, "lines"), box.padding = unit(0.16, "lines"), label.size = 0.05)+ scale_colour_manual(values=c("#00B6EB","#F8766D","#00C094","#FB61D7","#C49A00","#A69AFF","#53B400"))+ geom_segment(data = env.scrs,size=0.2,
               aes(x = 0, xend = NMDS1, y = 0, yend = NMDS2),
               arrow = arrow(length = unit(0.08, "cm")), colour = "red")
April_Species_vect

```
save
```{r, eval=FALSE}
setwd("/Users/hailaschultz/Dropbox/Schultz_Dissertation/Data_Analysis/Schultz_dissertation-2/figures")
ggsave(filename = "April_Species_vect.png", plot = April_Species_vect, width = 6.5, height = 5.5, device='png', dpi=700)
```

```{r,fig.width=6.5,fig.height=5, dpi=600}

April_Species_plot <- ggplot(data.scores, aes(x = NMDS1, y = NMDS2)) +geom_point(aes(colour = Station))+theme_bw()+theme(panel.grid.major = element_blank(),
                                                                                           panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))+ 
  geom_polygon(data=hull.data,aes(x=NMDS1,y=NMDS2,group=group),alpha=0.20, size=0.1, linetype=1, fill=NA, colour = "black") +
  annotate("text", label = "2D Stress: 0.15", x = 0.66, y = 0.66, size = 4, colour = "black")+ geom_segment(data = env.scrs,size=0.2,
               aes(x = 0, xend = NMDS1, y = 0, yend = NMDS2),
               arrow = arrow(length = unit(0.08, "cm")), colour = "black")+
  geom_label_repel(data = env.scrs, aes(x = NMDS1, y = NMDS2, label = Species),
                  size = 2.5, fontface="bold", fill="white", label.padding = unit(0.15, "lines"), box.padding = unit(0.16, "lines"), label.size = 0.05)+ 
  geom_text_repel(aes(label = Sample.Year), size = 2.5)+ scale_colour_manual(values=c("#00B6EB","#F8766D","#00C094","#FB61D7","#C49A00","#A69AFF","#53B400"))
April_Species_plot

```

save
```{r, eval=FALSE}
setwd("/Users/hailaschultz/Dropbox/Schultz_Dissertation/Data_Analysis/Schultz_dissertation-2/figures")
ggsave(filename = "April_Species_cluster.png", plot = April_Species_plot, width = 6.5, height = 5.5, device='png', dpi=700)
```


## NMDS: Subset to July
subset to July
```{r}
Zoop_env_april<-subset(Zoop_env,Month=="JUL")
```

## prep data
remove non-data columns, convert to proportionas, and arcsine sqrt transformation
```{r}
RE2<- Zoop_env_april[,28:ncol(Zoop_env_april)]
#replace N/A with 0
RE2[is.na(RE2)] <- 0
#convert to proportions
RE2<-RE2/rowSums(RE2)

#arcsine sqrt transformation
RE2<-asin(sqrt(RE2))

#remove columns with fewer than 9 rows with nonzero numbers
non_zero_counts <- colSums(RE2 != 0)
RE2 <- RE2[, non_zero_counts >= 9]
```

indicator species
```{r}
inv = multipatt(RE2, Zoop_env_april$Month, func = "r.g", control = how(nperm=999))
summary(inv)
```
## Clustering

gap statistic: do this to determine number of clusters
```{r}
library(factoextra)
res.km<-eclust(RE2, FUNcluster = "hclust", hc_metric = "euclidean", graph = FALSE)
fviz_gap_stat(res.km$gap_stat)

#silhouette plot
fviz_silhouette(res.km)
```
6 is the ultimate number of clusters

bray-curtis heirarchical clustering
```{r}
dist<-vegdist(RE2, method='bray',trymax = 200)
cluster <- hclust(dist, method = "average")
plot(cluster)
grp <- cutree(cluster, 6)
grp
```


nmds
```{r}
ord <- metaMDS(RE2)
```

```{r}
col <- c("red2", "green4", "mediumblue","pink","purple","lightblue","black")
col[grp]
```

plot
```{r}
plot(ord, type = "n", display = "sites")
points(ord, col = col[grp], bg = col[grp], pch = 21)
legend("topright", legend = paste("Cluster", 1:3),
       col = col, pt.bg = col, bty = "n", pch = 21)
```


## PERMANOVA
```{r}
dist<-vegdist(RE2, method='bray')
perm<-adonis(dist~Year*STATION_NO, data=Zoop_env_april, permutations = 999, method="bray")
perm
```
It looks like Sample.Year, station, and Sample.Month are significant. Additionally, the interaction between Sample.Year and Sample.Month and station and Sample.Month are different.

See pairwise differences
```{r}
#make sure factors are formatted right
Zoop_env_april$Year=as.factor(Zoop_env_april$Year)
Zoop_env_april$STATION_NO=as.factor(Zoop_env_april$STATION_NO)
adonis.pair(vegdist(RE2),Zoop_env_april$STATION_NO)
adonis.pair(vegdist(RE2),Zoop_env_april$Year)
```

all stations are different from one another

years are similar


check for dispersion
```{r}
#station
Zoop.bd <- betadisper(dist, Zoop_env_april$STATION_NO)
Zoop.bd
anova(Zoop.bd)
permutest(Zoop.bd)

#Sample.Year
Zoop.bd <- betadisper(dist, Zoop_env_april$Year)
Zoop.bd
anova(Zoop.bd)
permutest(Zoop.bd)

```
no dispersion issues!

NMDS
```{r}
NMDSmodel <- metaMDS(RE2, distance = "bray",trymax = 200)
NMDSmodel
```

get datascores
```{r}
data.scores<-scores(NMDSmodel,display="sites")
data.scores<- as.data.frame(data.scores)
```

add group, station, Sample.Month, Sample.Year back in
```{r}
data.scores$Station = Zoop_env_april$STATION_NO
data.scores$Sample.Year = Zoop_env_april$Year
data.scores$Sample.Month = Zoop_env_april$Month
data.scores$group = as.factor(grp)
```

add environmental variables back in
```{r}

data.scores$temp_deep = Zoop_env_april$CTDTMP_DEG_C_ITS90_Deep...6
data.scores$temp_surface = Zoop_env_april$CTDTMP_DEG_C_ITS90_Surface...7
data.scores$salinity_deep = Zoop_env_april$CTDSAL_PSS78_Deep...8
data.scores$salinity_surface = Zoop_env_april$CTDSAL_PSS78_Surface...9
data.scores$oxygen_deep = Zoop_env_april$OXYGEN_MG_L_1_Deep...10
data.scores$oxygen_surface = Zoop_env_april$OXYGEN_MG_L_1_Surface...11
data.scores$nitrate_deep = Zoop_env_april$NITRATE_UMOL_L_Deep...12
data.scores$nitrate_surface = Zoop_env_april$NITRATE_UMOL_L_Surface...13
data.scores$temp_surf_anom = Zoop_env_april$CTDTMP_DEG_C_ITS90_Surface...16
data.scores$temp_deep_anom = Zoop_env_april$CTDTMP_DEG_C_ITS90_Deep...15
data.scores$oxy_deep_anom = Zoop_env_april$OXYGEN_MG_L_1_Deep...19
data.scores$chla = Zoop_env_april$chla
```

plot data scores
```{r}
xx <- ggplot(data.scores, aes(x = NMDS1, y = NMDS2,colour=Station)) +stat_ellipse()+geom_point(aes(colour = Station))+theme_bw()+theme(panel.grid.major = element_blank(),
                                                                                           panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))
xx

xx <- ggplot(data.scores, aes(x = NMDS1, y = NMDS2,colour = Sample.Year)) +stat_ellipse()+geom_point(aes(colour = Sample.Year))+theme_bw()+theme(panel.grid.major = element_blank(),
                                                                                           panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))
xx

```


## environmental fitting
create environmetal fit table
```{r}
#check for correlation among environmental variables.
#remove nutrients other than nitrate because of autocorrelation

data.scores.cor<- data.scores[ -c(1:6) ]
cor_matrix <- cor(data.scores.cor)
corrplot(cor_matrix, method = "number", type = "lower", tl.pos = 'd')

Zoo.env <- data.frame(matrix(ncol = 0, nrow=44))
colnames(data.scores)
Zoo.env$temp_deep<-as.numeric(data.scores$temp_deep)
Zoo.env$temp_surface<-as.numeric(data.scores$temp_surface)
Zoo.env$salinity_deep<-as.numeric(data.scores$salinity_deep)
Zoo.env$salinity_surface<-as.numeric(data.scores$salinity_surface)
Zoo.env$oxygen_deep<-as.numeric(data.scores$oxygen_deep)
Zoo.env$oxygen_surface<-as.numeric(data.scores$oxygen_surface)
Zoo.env$temp_deep_anom<-as.numeric(data.scores$temp_deep_anom)
Zoo.env$temp_surf_anom<-as.numeric(data.scores$temp_surf_anom)
Zoo.env$oxy_deep_anom<-as.numeric(data.scores$oxy_deep_anom)
Zoo.env$chla<-as.numeric(data.scores$chla)

env <- envfit(NMDSmodel, Zoo.env, na.rm = TRUE,perm = 999)
env


env.scrs <- as.data.frame(scores(env, display = "vectors")) 
env.scrs <- cbind(env.scrs, Species = rownames(env.scrs), Pvalues = env$vectors$pvals, R_squared = env$vectors$r)
env.scrs <- subset(env.scrs, Pvalues < 0.05)


env.scrs
```

## Species fitting
```{r}
vf <- envfit(NMDSmodel, RE2, perm = 999)

spp.scrs <- as.data.frame(scores(vf, display = "vectors"))
spp.scrs <- cbind(spp.scrs, Species = rownames(spp.scrs))
####for ggplot
arrow_factor <- ordiArrowMul(vf)
spp.scrs <- as.data.frame(scores(vf, display = "vectors")) * arrow_factor
spp.scrs <- cbind(spp.scrs, Species = rownames(spp.scrs), Pvalues = vf$vectors$pvals, R_squared = vf$vectors$r)

# select significance similarly to `plot(vf, p.max = 0.01)`
spp.scrs <- subset(spp.scrs, Pvalues < 0.05)
spp.scrs <- subset(spp.scrs, R_squared > 0.3)
spp.scrs

```
### all data with species
extract hulls

```{r}
P12 <- data.scores[data.scores$Station == "P12", ][chull(data.scores[data.scores$Station ==                                                           "P12", c("NMDS1", "NMDS2")]), ]  
P22 <- data.scores[data.scores$Station == "P22", ][chull(data.scores[data.scores$Station ==                                                           "P22", c("NMDS1", "NMDS2")]), ]  
P28 <- data.scores[data.scores$Station == "P28", ][chull(data.scores[data.scores$Station ==                                                           "P28", c("NMDS1", "NMDS2")]), ]
P38 <- data.scores[data.scores$Station == "P38", ][chull(data.scores[data.scores$Station ==                                                           "P38", c("NMDS1", "NMDS2")]), ]  
P4 <- data.scores[data.scores$Station == "P4", ][chull(data.scores[data.scores$Station ==                                                           "P4", c("NMDS1", "NMDS2")]), ]  
P402 <- data.scores[data.scores$Station == "P402", ][chull(data.scores[data.scores$Station ==                                                           "P402", c("NMDS1", "NMDS2")]), ]  
P8 <- data.scores[data.scores$Station == "P8", ][chull(data.scores[data.scores$Station ==                                                           "P8", c("NMDS1", "NMDS2")]), ]  
```


get hull data
```{r}
sep.hull.data <- rbind(P12, P22, P28,P38,P4,P402,P8)  #combine grp.a and grp.b
sep.hull.data
```

```{r,fig.width=6.5,fig.height=5, dpi=600}

July_Species <- ggplot(data.scores, aes(x = NMDS1, y = NMDS2)) +geom_point(aes(colour = Station))+theme_bw()+theme(panel.grid.major = element_blank(),
                                                                                           panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))+ 
  geom_polygon(data=sep.hull.data,aes(x=NMDS1,y=NMDS2,group=Station, fill=Station),alpha=0.20, size=0.1, linetype=1, colour = "black") +
  annotate("text", label = "2D Stress: 0.18", x = 0.5, y = 0.66, size = 4, colour = "black")+ geom_segment(data = env.scrs,size=0.2,
               aes(x = 0, xend = NMDS1, y = 0, yend = NMDS2),
               arrow = arrow(length = unit(0.08, "cm")), colour = "black")+
  geom_label_repel(data = env.scrs, aes(x = NMDS1, y = NMDS2, label = Species),
                  size = 2.5, fontface="bold", fill="white", label.padding = unit(0.15, "lines"), box.padding = unit(0.16, "lines"), label.size = 0.05)+ scale_colour_manual(values=c("#00B6EB","#F8766D","#00C094","#FB61D7","#C49A00","#A69AFF","#53B400")) +
  scale_fill_manual(values=c("#00B6EB","#F8766D","#00C094","#FB61D7","#C49A00","#A69AFF","#53B400"))
July_Species

```
save 
```{r, eval=FALSE}
setwd("/Users/hailaschultz/Dropbox/Schultz_Dissertation/Data_Analysis/Schultz_dissertation-2/figures")
ggsave(filename = "July_Species_month.png", plot = July_Species, width = 6.5, height = 5.5, device='png', dpi=700)
```
extract hulls
```{r}
G1 <- data.scores[data.scores$group == "1", ][chull(data.scores[data.scores$group ==                                                           "1", c("NMDS1", "NMDS2")]), ]  
G2 <- data.scores[data.scores$group == "2", ][chull(data.scores[data.scores$group ==                                                           "2", c("NMDS1", "NMDS2")]), ]  
G3 <- data.scores[data.scores$group == "3", ][chull(data.scores[data.scores$group ==                                                           "3", c("NMDS1", "NMDS2")]), ]
G4 <- data.scores[data.scores$group == "4", ][chull(data.scores[data.scores$group ==                                                           "4", c("NMDS1", "NMDS2")]), ]  
G5 <- data.scores[data.scores$group == "5", ][chull(data.scores[data.scores$group ==                                                           "5", c("NMDS1", "NMDS2")]), ] 
G6 <- data.scores[data.scores$group == "6", ][chull(data.scores[data.scores$group ==                                                           "6", c("NMDS1", "NMDS2")]), ] 

```
get hull data
```{r}
hull.data <- rbind(G1,G2,G3,G4,G5,G6)  #combine grp.a and grp.b
hull.data

```
```{r,fig.width=6.5,fig.height=5, dpi=600}

July_Species_plot <- ggplot(data.scores, aes(x = NMDS1, y = NMDS2)) +geom_point(aes(colour = Station))+theme_bw()+theme(panel.grid.major = element_blank(),
                                                                                           panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))+ 
  geom_polygon(data=hull.data,aes(x=NMDS1,y=NMDS2,group=group),alpha=0.20, size=0.1, linetype=1, fill=NA, colour = "black") +
  annotate("text", label = "2D Stress: 0.18", x = 0.5, y = 0.66, size = 4, colour = "black")+ geom_segment(data = env.scrs,size=0.2,
               aes(x = 0, xend = NMDS1, y = 0, yend = NMDS2),
               arrow = arrow(length = unit(0.08, "cm")), colour = "black")+
  geom_label_repel(data = env.scrs, aes(x = NMDS1, y = NMDS2, label = Species),
                  size = 2.5, fontface="bold", fill="white", label.padding = unit(0.15, "lines"), box.padding = unit(0.16, "lines"), label.size = 0.05)+ scale_colour_manual(values=c("#00B6EB","#F8766D","#00C094","#FB61D7","#C49A00","#A69AFF","#53B400"))+ geom_segment(data = spp.scrs,size=0.2,
               aes(x = 0, xend = NMDS1, y = 0, yend = NMDS2),
               arrow = arrow(length = unit(0.08, "cm")), colour = "gray")+
  geom_label_repel(data = spp.scrs, aes(x = NMDS1, y = NMDS2, label = Species),
                  size = 2, fontface="bold", fill="white", label.padding = unit(0.15, "lines"), box.padding = unit(0.16, "lines"), label.size = 0.05)
July_Species_plot

```
save
```{r, eval=FALSE}
setwd("/Users/hailaschultz/Dropbox/Schultz_Dissertation/Data_Analysis/Schultz_dissertation-2/figures")
ggsave(filename = "July_Species_cluster.png", plot =July_Species_plot, width = 6.5, height = 5.5, device='png', dpi=700)
```

```{r,fig.width=6.5,fig.height=5, dpi=600}

July_Species_year <- ggplot(data.scores, aes(x = NMDS1, y = NMDS2)) +geom_point(aes(colour = Station))+theme_bw()+theme(panel.grid.major = element_blank(),
                                                                                           panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))+ 
  geom_polygon(data=hull.data,aes(x=NMDS1,y=NMDS2,group=group),alpha=0.20, size=0.1, linetype=1, fill=NA, colour = "black") +
  annotate("text", label = "2D Stress: 0.18", x = 0.5, y = 0.66, size = 4, colour = "black")+ 
  geom_text_repel(aes(label = Sample.Year), size = 2.5)+ scale_colour_manual(values=c("#00B6EB","#F8766D","#00C094","#FB61D7","#C49A00","#A69AFF","#53B400"))
July_Species_year

```
save
```{r, eval=FALSE}
setwd("/Users/hailaschultz/Dropbox/Schultz_Dissertation/Data_Analysis/Schultz_dissertation-2/figures")
ggsave(filename = "July_Species_year.png", plot =July_Species_year, width = 6.5, height = 5.5, device='png', dpi=700)
```





## NMDS: Subset to September
subset to september
```{r}
Zoop_env_september<-subset(Zoop_env,Month=="SEP")
```

## prep data
remove non-data columns, convert to proportionas, and arcsine sqrt transformation
```{r}
RE2<- Zoop_env_september[,28:ncol(Zoop_env_september)]
#replace N/A with 0
RE2[is.na(RE2)] <- 0
#convert to proportions
RE2<-RE2/rowSums(RE2)

#arcsine sqrt transformation
RE2<-asin(sqrt(RE2))

#remove columns with fewer than 9 rows with nonzero numbers
non_zero_counts <- colSums(RE2 != 0)
RE2 <- RE2[, non_zero_counts >= 9]
```


## Clustering

gap statistic: do this to determine number of clusters
```{r}
library(factoextra)
res.km<-eclust(RE2, FUNcluster = "hclust", hc_metric = "euclidean", graph = FALSE)
fviz_gap_stat(res.km$gap_stat)

#silhouette plot
fviz_silhouette(res.km)
```
7 is the ultimate number of clusters

bray-curtis heirarchical clustering
```{r}
dist<-vegdist(RE2, method='bray',trymax = 200)
cluster <- hclust(dist, method = "average")
plot(cluster)
grp <- cutree(cluster, 7)
grp
```


nmds
```{r}
ord <- metaMDS(RE2)
```

```{r}
col <- c("red2", "green4", "mediumblue","pink","purple","lightblue","black")
col[grp]
```

plot
```{r}
plot(ord, type = "n", display = "sites")
points(ord, col = col[grp], bg = col[grp], pch = 21)
legend("topright", legend = paste("Cluster", 1:3),
       col = col, pt.bg = col, bty = "n", pch = 21)
```


## PERMANOVA
```{r}
dist<-vegdist(RE2, method='bray')
perm<-adonis(dist~Year*STATION_NO, data=Zoop_env_september, permutations = 999, method="bray")
perm
```
It looks like Sample.Year, station, and Sample.Month are significant. Additionally, the interaction between Sample.Year and Sample.Month and station and Sample.Month are different.

See pairwise differences
```{r}
#make sure factors are formatted right
Zoop_env_september$Year=as.factor(Zoop_env_september$Year)
Zoop_env_september$STATION_NO=as.factor(Zoop_env_september$STATION_NO)
adonis.pair(vegdist(RE2),Zoop_env_september$STATION_NO)
adonis.pair(vegdist(RE2),Zoop_env_september$Year)
```

all stations are different from one another

years are similar


check for dispersion
```{r}
#station
Zoop.bd <- betadisper(dist, Zoop_env_september$STATION_NO)
Zoop.bd
anova(Zoop.bd)
permutest(Zoop.bd)

#Sample.Year
Zoop.bd <- betadisper(dist, Zoop_env_september$Year)
Zoop.bd
anova(Zoop.bd)
permutest(Zoop.bd)

```


NMDS
```{r}
NMDSmodel <- metaMDS(RE2, distance = "bray",trymax = 200)
NMDSmodel
```

get datascores
```{r}
data.scores<-scores(NMDSmodel,display="sites")
data.scores<- as.data.frame(data.scores)
```

add group, station, Sample.Month, Sample.Year back in
```{r}
data.scores$Station = Zoop_env_september$STATION_NO
data.scores$Sample.Year = Zoop_env_september$Year
data.scores$Sample.Month = Zoop_env_september$Month
data.scores$group = as.factor(grp)
```

add environmental variables back in
```{r}

data.scores$temp_deep = Zoop_env_september$CTDTMP_DEG_C_ITS90_Deep...6
data.scores$temp_surface = Zoop_env_september$CTDTMP_DEG_C_ITS90_Surface...7
data.scores$salinity_deep = Zoop_env_september$CTDSAL_PSS78_Deep...8
data.scores$salinity_surface = Zoop_env_september$CTDSAL_PSS78_Surface...9
data.scores$oxygen_deep = Zoop_env_september$OXYGEN_MG_L_1_Deep...10
data.scores$oxygen_surface = Zoop_env_september$OXYGEN_MG_L_1_Surface...11
data.scores$nitrate_deep = Zoop_env_september$NITRATE_UMOL_L_Deep...12
data.scores$nitrate_surface = Zoop_env_september$NITRATE_UMOL_L_Surface...13
data.scores$temp_surf_anom = Zoop_env_september$CTDTMP_DEG_C_ITS90_Surface...16
data.scores$temp_deep_anom = Zoop_env_september$CTDTMP_DEG_C_ITS90_Deep...15
data.scores$oxy_deep_anom = Zoop_env_september$OXYGEN_MG_L_1_Deep...19
data.scores$chla = Zoop_env_september$chla
```

plot data scores
```{r}
xx <- ggplot(data.scores, aes(x = NMDS1, y = NMDS2,colour=Station)) +stat_ellipse()+geom_point(aes(colour = Station))+theme_bw()+theme(panel.grid.major = element_blank(),
                                                                                           panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))
xx

xx <- ggplot(data.scores, aes(x = NMDS1, y = NMDS2,colour = Sample.Year)) +stat_ellipse()+geom_point(aes(colour = Sample.Year))+theme_bw()+theme(panel.grid.major = element_blank(),
                                                                                           panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))
xx

```


## environmental fitting
create environmetal fit table
```{r}
#check for correlation among environmental variables.
#remove nutrients other than nitrate because of autocorrelation

data.scores.cor<- data.scores[ -c(1:6) ]
cor_matrix <- cor(data.scores.cor)
corrplot(cor_matrix, method = "number", type = "lower", tl.pos = 'd')

Zoo.env <- data.frame(matrix(ncol = 0, nrow=57))
colnames(data.scores)
Zoo.env$temp_deep<-as.numeric(data.scores$temp_deep)
Zoo.env$temp_surface<-as.numeric(data.scores$temp_surface)
Zoo.env$salinity_deep<-as.numeric(data.scores$salinity_deep)
Zoo.env$salinity_surface<-as.numeric(data.scores$salinity_surface)
Zoo.env$oxygen_deep<-as.numeric(data.scores$oxygen_deep)
Zoo.env$oxygen_surface<-as.numeric(data.scores$oxygen_surface)
Zoo.env$temp_deep_anom<-as.numeric(data.scores$temp_deep_anom)
Zoo.env$temp_surf_anom<-as.numeric(data.scores$temp_surf_anom)
Zoo.env$oxy_deep_anom<-as.numeric(data.scores$oxy_deep_anom)
Zoo.env$chla<-as.numeric(data.scores$chla)

env <- envfit(NMDSmodel, Zoo.env, na.rm = TRUE,perm = 999)
env


env.scrs <- as.data.frame(scores(env, display = "vectors")) 
env.scrs <- cbind(env.scrs, Species = rownames(env.scrs), Pvalues = env$vectors$pvals, R_squared = env$vectors$r)
env.scrs <- subset(env.scrs, Pvalues < 0.05)


env.scrs
```

## Species fitting
```{r}
vf <- envfit(NMDSmodel, RE2, perm = 999)

spp.scrs <- as.data.frame(scores(vf, display = "vectors"))
spp.scrs <- cbind(spp.scrs, Species = rownames(spp.scrs))
####for ggplot
arrow_factor <- ordiArrowMul(vf)
spp.scrs <- as.data.frame(scores(vf, display = "vectors")) * arrow_factor
spp.scrs <- cbind(spp.scrs, Species = rownames(spp.scrs), Pvalues = vf$vectors$pvals, R_squared = vf$vectors$r)

# select significance similarly to `plot(vf, p.max = 0.01)`
spp.scrs <- subset(spp.scrs, Pvalues < 0.05)
spp.scrs <- subset(spp.scrs, R_squared > 0.3)
spp.scrs

```
### all data with species
extract hulls

```{r}
P12 <- data.scores[data.scores$Station == "P12", ][chull(data.scores[data.scores$Station ==                                                           "P12", c("NMDS1", "NMDS2")]), ]  
P22 <- data.scores[data.scores$Station == "P22", ][chull(data.scores[data.scores$Station ==                                                           "P22", c("NMDS1", "NMDS2")]), ]  
P28 <- data.scores[data.scores$Station == "P28", ][chull(data.scores[data.scores$Station ==                                                           "P28", c("NMDS1", "NMDS2")]), ]
P38 <- data.scores[data.scores$Station == "P38", ][chull(data.scores[data.scores$Station ==                                                           "P38", c("NMDS1", "NMDS2")]), ]  
P4 <- data.scores[data.scores$Station == "P4", ][chull(data.scores[data.scores$Station ==                                                           "P4", c("NMDS1", "NMDS2")]), ]  
P402 <- data.scores[data.scores$Station == "P402", ][chull(data.scores[data.scores$Station ==                                                           "P402", c("NMDS1", "NMDS2")]), ]  
P8 <- data.scores[data.scores$Station == "P8", ][chull(data.scores[data.scores$Station ==                                                           "P8", c("NMDS1", "NMDS2")]), ]  
```


get hull data
```{r}
sep.hull.data <- rbind(P12, P22, P28,P38,P4,P402,P8)  #combine grp.a and grp.b
sep.hull.data
```

```{r,fig.width=6.5,fig.height=5, dpi=600}

sep_Species <- ggplot(data.scores, aes(x = NMDS1, y = NMDS2)) +geom_point(aes(colour = Station))+theme_bw()+theme(panel.grid.major = element_blank(),
                                                                                           panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))+ 
  geom_polygon(data=sep.hull.data,aes(x=NMDS1,y=NMDS2,group=Station, fill=Station),alpha=0.20, size=0.1, linetype=1, colour = "black") +
  annotate("text", label = "2D Stress: 0.16", x = 0.65, y = 0.8, size = 4, colour = "black")+ geom_segment(data = env.scrs,size=0.2,
               aes(x = 0, xend = NMDS1, y = 0, yend = NMDS2),
               arrow = arrow(length = unit(0.08, "cm")), colour = "black")+
  geom_label_repel(data = env.scrs, aes(x = NMDS1, y = NMDS2, label = Species),
                  size = 2.5, fontface="bold", fill="white", label.padding = unit(0.15, "lines"), box.padding = unit(0.16, "lines"), label.size = 0.05)+ scale_colour_manual(values=c("#00B6EB","#F8766D","#00C094","#FB61D7","#C49A00","#A69AFF","#53B400")) +
  scale_fill_manual(values=c("#00B6EB","#F8766D","#00C094","#FB61D7","#C49A00","#A69AFF","#53B400"))
sep_Species

```
save 
```{r, eval=FALSE}
setwd("/Users/hailaschultz/Dropbox/Schultz_Dissertation/Data_Analysis/Schultz_dissertation-2/figures")
ggsave(filename = "Sep_Species_month.png", plot = sep_Species, width = 6.5, height = 5.5, device='png', dpi=700)
```
extract hulls
```{r}
G1 <- data.scores[data.scores$group == "1", ][chull(data.scores[data.scores$group ==                                                           "1", c("NMDS1", "NMDS2")]), ]  
G2 <- data.scores[data.scores$group == "2", ][chull(data.scores[data.scores$group ==                                                           "2", c("NMDS1", "NMDS2")]), ]  
G3 <- data.scores[data.scores$group == "3", ][chull(data.scores[data.scores$group ==                                                           "3", c("NMDS1", "NMDS2")]), ]
G4 <- data.scores[data.scores$group == "4", ][chull(data.scores[data.scores$group ==                                                           "4", c("NMDS1", "NMDS2")]), ]  
G5 <- data.scores[data.scores$group == "5", ][chull(data.scores[data.scores$group ==                                                           "5", c("NMDS1", "NMDS2")]), ] 
G6 <- data.scores[data.scores$group == "6", ][chull(data.scores[data.scores$group ==                                                           "6", c("NMDS1", "NMDS2")]), ] 
G7 <- data.scores[data.scores$group == "7", ][chull(data.scores[data.scores$group ==                                                           "7", c("NMDS1", "NMDS2")]), ] 

```
get hull data
```{r}
hull.data <- rbind(G1,G2,G3,G4,G5,G6,G7)  #combine grp.a and grp.b
hull.data

```
```{r,fig.width=6.5,fig.height=5, dpi=600}

sep_Species_plot <- ggplot(data.scores, aes(x = NMDS1, y = NMDS2)) +geom_point(aes(colour = Station))+theme_bw()+theme(panel.grid.major = element_blank(),
                                                                                           panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))+ 
  geom_polygon(data=hull.data,aes(x=NMDS1,y=NMDS2,group=group),alpha=0.20, size=0.1, linetype=1, fill=NA, colour = "black") +
  annotate("text", label = "2D Stress: 0.16", x = 0.65, y = 0.8, size = 4, colour = "black")+ geom_segment(data = env.scrs,size=0.2,
               aes(x = 0, xend = NMDS1, y = 0, yend = NMDS2),
               arrow = arrow(length = unit(0.08, "cm")), colour = "black")+
  geom_label_repel(data = env.scrs, aes(x = NMDS1, y = NMDS2, label = Species),
                  size = 2.5, fontface="bold", fill="white", label.padding = unit(0.15, "lines"), box.padding = unit(0.16, "lines"), label.size = 0.05)+ scale_colour_manual(values=c("#00B6EB","#F8766D","#00C094","#FB61D7","#C49A00","#A69AFF","#53B400"))
sep_Species_plot

```
save
```{r, eval=FALSE}
setwd("/Users/hailaschultz/Dropbox/Schultz_Dissertation/Data_Analysis/Schultz_dissertation-2/figures")
ggsave(filename = "Sep_Species_cluster.png", plot =sep_Species_plot, width = 6.5, height = 5.5, device='png', dpi=700)
```

```{r,fig.width=6.5,fig.height=5, dpi=600}

sep_Species_year <- ggplot(data.scores, aes(x = NMDS1, y = NMDS2)) +geom_point(aes(colour = Station))+theme_bw()+theme(panel.grid.major = element_blank(),
                                                                                           panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))+ 
  geom_polygon(data=hull.data,aes(x=NMDS1,y=NMDS2,group=group),alpha=0.20, size=0.1, linetype=1, fill=NA, colour = "black") +
  annotate("text", label = "2D Stress: 0.16", x = 0.65, y = 0.8, size = 4, colour = "black")+ 
  geom_text_repel(aes(label = Sample.Year), size = 2.5)+ scale_colour_manual(values=c("#00B6EB","#F8766D","#00C094","#FB61D7","#C49A00","#A69AFF","#53B400"))
sep_Species_year

```
save
```{r, eval=FALSE}
setwd("/Users/hailaschultz/Dropbox/Schultz_Dissertation/Data_Analysis/Schultz_dissertation-2/figures")
ggsave(filename = "Sep_Species_year.png", plot =sep_Species_year, width = 6.5, height = 5.5, device='png', dpi=700)
```



# heatmap April P28
```{r}
unique(Zoop_sub$Station)
Zoop_april<-subset(Zoop_sub,Sample.Month=="APR")
Zoop_april<-subset(Zoop_april,Station=="P28")
Zoop_sub_heat<-dcast(Zoop_april, Sample.Code~ Genus.species,value.var = "zoop_density")
Zoop_sub_heat <- Zoop_sub_heat %>% replace(is.na(.), 0)
#convert to proportions
Zoop_sub_heat_data  = subset(Zoop_sub_heat, select = -c(Sample.Code) )

Zoop_sub_heat_data1<-as.matrix(Zoop_sub_heat_data)
Zoop_sub_heat_data1 <- matrix( 
  as.numeric(Zoop_sub_heat_data1), ncol = 100)  
rownames(Zoop_sub_heat_data1)<-Zoop_sub_heat$Sample.Code
column<-colnames(Zoop_sub_heat_data)
colnames(Zoop_sub_heat_data1)<-column

heatmap(Zoop_sub_heat_data1,show_row_names=TRUE,show_col_names=TRUE,mar = c(15, 15))

heatmap(Zoop_sub_heat_data1, scale="column",show_row_names=TRUE,show_col_names=TRUE,mar = c(15, 15))
```
#heatmap July P28
```{r}
unique(Zoop_sub$Station)
Zoop_april<-subset(Zoop_sub,Sample.Month=="JUL")
Zoop_april<-subset(Zoop_april,Station=="P28")
Zoop_sub_heat<-dcast(Zoop_april, Sample.Code~ Genus.species,value.var = "zoop_density")
Zoop_sub_heat <- Zoop_sub_heat %>% replace(is.na(.), 0)
#convert to proportions
Zoop_sub_heat_data  = subset(Zoop_sub_heat, select = -c(Sample.Code) )

Zoop_sub_heat_data1<-as.matrix(Zoop_sub_heat_data)
Zoop_sub_heat_data1 <- matrix( 
  as.numeric(Zoop_sub_heat_data1), ncol = 92)  
rownames(Zoop_sub_heat_data1)<-Zoop_sub_heat$Sample.Code
column<-colnames(Zoop_sub_heat_data)
colnames(Zoop_sub_heat_data1)<-column

heatmap(Zoop_sub_heat_data1,show_row_names=TRUE,show_col_names=TRUE,mar = c(15, 15))

heatmap(Zoop_sub_heat_data1, scale="column",show_row_names=TRUE,show_col_names=TRUE,mar = c(15, 15))
```

#heatmap September P28
```{r}
unique(Zoop_sub$Station)
Zoop_april<-subset(Zoop_sub,Sample.Month=="SEP")
Zoop_april<-subset(Zoop_april,Station=="P28")
Zoop_sub_heat<-dcast(Zoop_april, Sample.Code~ Genus.species,value.var = "zoop_density")
Zoop_sub_heat <- Zoop_sub_heat %>% replace(is.na(.), 0)
#convert to proportions
Zoop_sub_heat_data  = subset(Zoop_sub_heat, select = -c(Sample.Code) )

Zoop_sub_heat_data1<-as.matrix(Zoop_sub_heat_data)
Zoop_sub_heat_data1 <- matrix( 
  as.numeric(Zoop_sub_heat_data1), ncol = 61)  
rownames(Zoop_sub_heat_data1)<-Zoop_sub_heat$Sample.Code
column<-colnames(Zoop_sub_heat_data)
colnames(Zoop_sub_heat_data1)<-column

heatmap(Zoop_sub_heat_data1,show_row_names=TRUE,show_col_names=TRUE,mar = c(15, 15))

heatmap(Zoop_sub_heat_data1, scale="column",show_row_names=TRUE,show_col_names=TRUE,mar = c(15, 15))
```

#heatmap April P402
```{r}
unique(Zoop_sub$Station)
Zoop_april<-subset(Zoop_sub,Sample.Month=="APR")
Zoop_april<-subset(Zoop_april,Station=="P402")
Zoop_sub_heat<-dcast(Zoop_april, Sample.Code~ Genus.species,value.var = "zoop_density")
Zoop_sub_heat <- Zoop_sub_heat %>% replace(is.na(.), 0)
#convert to proportions
Zoop_sub_heat_data  = subset(Zoop_sub_heat, select = -c(Sample.Code) )

Zoop_sub_heat_data1<-as.matrix(Zoop_sub_heat_data)
Zoop_sub_heat_data1 <- matrix( 
  as.numeric(Zoop_sub_heat_data1), ncol = 103)  
rownames(Zoop_sub_heat_data1)<-Zoop_sub_heat$Sample.Code
column<-colnames(Zoop_sub_heat_data)
colnames(Zoop_sub_heat_data1)<-column

heatmap(Zoop_sub_heat_data1,show_row_names=TRUE,show_col_names=TRUE,mar = c(15, 15))

heatmap(Zoop_sub_heat_data1, scale="column",show_row_names=TRUE,show_col_names=TRUE,mar = c(15, 15))
```

#heatmap July P402
```{r}
unique(Zoop_sub$Station)
Zoop_april<-subset(Zoop_sub,Sample.Month=="JUL")
Zoop_april<-subset(Zoop_april,Station=="P402")
Zoop_sub_heat<-dcast(Zoop_april, Sample.Code~ Genus.species,value.var = "zoop_density")
Zoop_sub_heat <- Zoop_sub_heat %>% replace(is.na(.), 0)
#convert to proportions
Zoop_sub_heat_data  = subset(Zoop_sub_heat, select = -c(Sample.Code) )

Zoop_sub_heat_data1<-as.matrix(Zoop_sub_heat_data)
Zoop_sub_heat_data1 <- matrix( 
  as.numeric(Zoop_sub_heat_data1), ncol = 89)  
rownames(Zoop_sub_heat_data1)<-Zoop_sub_heat$Sample.Code
column<-colnames(Zoop_sub_heat_data)
colnames(Zoop_sub_heat_data1)<-column

heatmap(Zoop_sub_heat_data1,show_row_names=TRUE,show_col_names=TRUE,mar = c(15, 15))

heatmap(Zoop_sub_heat_data1, scale="column",show_row_names=TRUE,show_col_names=TRUE,mar = c(15, 15))
```

#heatmap September P402
```{r}
unique(Zoop_sub$Station)
Zoop_april<-subset(Zoop_sub,Sample.Month=="SEP")
Zoop_april<-subset(Zoop_april,Station=="P402")
Zoop_sub_heat<-dcast(Zoop_april, Sample.Code~ Genus.species,value.var = "zoop_density")
Zoop_sub_heat <- Zoop_sub_heat %>% replace(is.na(.), 0)
#convert to proportions
Zoop_sub_heat_data  = subset(Zoop_sub_heat, select = -c(Sample.Code) )

Zoop_sub_heat_data1<-as.matrix(Zoop_sub_heat_data)
Zoop_sub_heat_data1 <- matrix( 
  as.numeric(Zoop_sub_heat_data1), ncol = 85)  
rownames(Zoop_sub_heat_data1)<-Zoop_sub_heat$Sample.Code
column<-colnames(Zoop_sub_heat_data)
colnames(Zoop_sub_heat_data1)<-column

heatmap(Zoop_sub_heat_data1,show_row_names=TRUE,show_col_names=TRUE,mar = c(15, 15))

heatmap(Zoop_sub_heat_data1, scale="column",show_row_names=TRUE,show_col_names=TRUE,mar = c(15, 15))
```

#heatmap April P12
```{r}
unique(Zoop_sub$Station)
Zoop_april<-subset(Zoop_sub,Sample.Month=="APR")
Zoop_april<-subset(Zoop_april,Station=="P12")
Zoop_sub_heat<-dcast(Zoop_april, Sample.Code~ Genus.species,value.var = "zoop_density")
Zoop_sub_heat <- Zoop_sub_heat %>% replace(is.na(.), 0)
#convert to proportions
Zoop_sub_heat_data  = subset(Zoop_sub_heat, select = -c(Sample.Code) )

Zoop_sub_heat_data1<-as.matrix(Zoop_sub_heat_data)
Zoop_sub_heat_data1 <- matrix( 
  as.numeric(Zoop_sub_heat_data1), ncol = 106)  
rownames(Zoop_sub_heat_data1)<-Zoop_sub_heat$Sample.Code
column<-colnames(Zoop_sub_heat_data)
colnames(Zoop_sub_heat_data1)<-column

heatmap(Zoop_sub_heat_data1,show_row_names=TRUE,show_col_names=TRUE,mar = c(15, 15))

heatmap(Zoop_sub_heat_data1, scale="column",show_row_names=TRUE,show_col_names=TRUE,mar = c(15, 15))
```

#heatmap July P12
```{r}
unique(Zoop_sub$Station)
Zoop_april<-subset(Zoop_sub,Sample.Month=="JUL")
Zoop_april<-subset(Zoop_april,Station=="P12")
Zoop_sub_heat<-dcast(Zoop_april, Sample.Code~ Genus.species,value.var = "zoop_density")
Zoop_sub_heat <- Zoop_sub_heat %>% replace(is.na(.), 0)
#convert to proportions
Zoop_sub_heat_data  = subset(Zoop_sub_heat, select = -c(Sample.Code) )

Zoop_sub_heat_data1<-as.matrix(Zoop_sub_heat_data)
Zoop_sub_heat_data1 <- matrix( 
  as.numeric(Zoop_sub_heat_data1), ncol = 93)  
rownames(Zoop_sub_heat_data1)<-Zoop_sub_heat$Sample.Code
column<-colnames(Zoop_sub_heat_data)
colnames(Zoop_sub_heat_data1)<-column

heatmap(Zoop_sub_heat_data1,show_row_names=TRUE,show_col_names=TRUE,mar = c(15, 15))

heatmap(Zoop_sub_heat_data1, scale="column",show_row_names=TRUE,show_col_names=TRUE,mar = c(15, 15))
```

#heatmap Sep P12
```{r}
unique(Zoop_sub$Station)
Zoop_april<-subset(Zoop_sub,Sample.Month=="SEP")
Zoop_april<-subset(Zoop_april,Station=="P12")
Zoop_sub_heat<-dcast(Zoop_april, Sample.Code~ Genus.species,value.var = "zoop_density")
Zoop_sub_heat <- Zoop_sub_heat %>% replace(is.na(.), 0)
#convert to proportions
Zoop_sub_heat_data  = subset(Zoop_sub_heat, select = -c(Sample.Code) )

Zoop_sub_heat_data1<-as.matrix(Zoop_sub_heat_data)
Zoop_sub_heat_data1 <- matrix( 
  as.numeric(Zoop_sub_heat_data1), ncol = 73)  
rownames(Zoop_sub_heat_data1)<-Zoop_sub_heat$Sample.Code
column<-colnames(Zoop_sub_heat_data)
colnames(Zoop_sub_heat_data1)<-column

heatmap(Zoop_sub_heat_data1,show_row_names=TRUE,show_col_names=TRUE,mar = c(15, 15))

heatmap(Zoop_sub_heat_data1, scale="column",show_row_names=TRUE,show_col_names=TRUE,mar = c(15, 15))
```

#heatmap April P4
```{r}
unique(Zoop_sub$Station)
Zoop_april<-subset(Zoop_sub,Sample.Month=="APR")
Zoop_april<-subset(Zoop_april,Station=="P4")
Zoop_sub_heat<-dcast(Zoop_april, Sample.Code~ Genus.species,value.var = "zoop_density")
Zoop_sub_heat <- Zoop_sub_heat %>% replace(is.na(.), 0)
#convert to proportions
Zoop_sub_heat_data  = subset(Zoop_sub_heat, select = -c(Sample.Code) )

Zoop_sub_heat_data1<-as.matrix(Zoop_sub_heat_data)
Zoop_sub_heat_data1 <- matrix( 
  as.numeric(Zoop_sub_heat_data1), ncol = 86)  
rownames(Zoop_sub_heat_data1)<-Zoop_sub_heat$Sample.Code
column<-colnames(Zoop_sub_heat_data)
colnames(Zoop_sub_heat_data1)<-column

heatmap(Zoop_sub_heat_data1,show_row_names=TRUE,show_col_names=TRUE,mar = c(15, 15))

heatmap(Zoop_sub_heat_data1, scale="column",show_row_names=TRUE,show_col_names=TRUE,mar = c(15, 15))
```

#heatmap July P4
```{r}
unique(Zoop_sub$Station)
Zoop_april<-subset(Zoop_sub,Sample.Month=="JUL")
Zoop_april<-subset(Zoop_april,Station=="P4")
Zoop_sub_heat<-dcast(Zoop_april, Sample.Code~ Genus.species,value.var = "zoop_density")
Zoop_sub_heat <- Zoop_sub_heat %>% replace(is.na(.), 0)
#convert to proportions
Zoop_sub_heat_data  = subset(Zoop_sub_heat, select = -c(Sample.Code) )

Zoop_sub_heat_data1<-as.matrix(Zoop_sub_heat_data)
Zoop_sub_heat_data1 <- matrix( 
  as.numeric(Zoop_sub_heat_data1), ncol = 98)  
rownames(Zoop_sub_heat_data1)<-Zoop_sub_heat$Sample.Code
column<-colnames(Zoop_sub_heat_data)
colnames(Zoop_sub_heat_data1)<-column

heatmap(Zoop_sub_heat_data1,show_row_names=TRUE,show_col_names=TRUE,mar = c(15, 15))

heatmap(Zoop_sub_heat_data1, scale="column",show_row_names=TRUE,show_col_names=TRUE,mar = c(15, 15))
```

#heatmap September P4
```{r}
unique(Zoop_sub$Station)
Zoop_april<-subset(Zoop_sub,Sample.Month=="SEP")
Zoop_april<-subset(Zoop_april,Station=="P4")
Zoop_sub_heat<-dcast(Zoop_april, Sample.Code~ Genus.species,value.var = "zoop_density")
Zoop_sub_heat <- Zoop_sub_heat %>% replace(is.na(.), 0)
#convert to proportions
Zoop_sub_heat_data  = subset(Zoop_sub_heat, select = -c(Sample.Code) )

Zoop_sub_heat_data1<-as.matrix(Zoop_sub_heat_data)
Zoop_sub_heat_data1 <- matrix( 
  as.numeric(Zoop_sub_heat_data1), ncol = 100)  
rownames(Zoop_sub_heat_data1)<-Zoop_sub_heat$Sample.Code
column<-colnames(Zoop_sub_heat_data)
colnames(Zoop_sub_heat_data1)<-column

heatmap(Zoop_sub_heat_data1,show_row_names=TRUE,show_col_names=TRUE,mar = c(15, 15))

heatmap(Zoop_sub_heat_data1, scale="column",show_row_names=TRUE,show_col_names=TRUE,mar = c(15, 15))
```

#heatmap April P8
```{r}
unique(Zoop_sub$Station)
Zoop_april<-subset(Zoop_sub,Sample.Month=="APR")
Zoop_april<-subset(Zoop_april,Station=="P8")
Zoop_sub_heat<-dcast(Zoop_april, Sample.Code~ Genus.species,value.var = "zoop_density")
Zoop_sub_heat <- Zoop_sub_heat %>% replace(is.na(.), 0)
#convert to proportions
Zoop_sub_heat_data  = subset(Zoop_sub_heat, select = -c(Sample.Code) )

Zoop_sub_heat_data1<-as.matrix(Zoop_sub_heat_data)
Zoop_sub_heat_data1 <- matrix( 
  as.numeric(Zoop_sub_heat_data1), ncol = 111)  
rownames(Zoop_sub_heat_data1)<-Zoop_sub_heat$Sample.Code
column<-colnames(Zoop_sub_heat_data)
colnames(Zoop_sub_heat_data1)<-column

heatmap(Zoop_sub_heat_data1,show_row_names=TRUE,show_col_names=TRUE,mar = c(15, 15))

heatmap(Zoop_sub_heat_data1, scale="column",show_row_names=TRUE,show_col_names=TRUE,mar = c(15, 15))
```

#heatmap July P8 organize by phylum
```{r}
unique(Zoop_sub$Station)
Zoop_april<-subset(Zoop_sub,Sample.Month=="JUL")
Zoop_april<-subset(Zoop_april,Station=="P8")
Zoop_sub_heat<-dcast(Zoop_april, Sample.Code~ Genus.species,value.var = "zoop_density")
Zoop_sub_heat <- Zoop_sub_heat %>% replace(is.na(.), 0)
#convert to proportions
Zoop_sub_heat_data  = subset(Zoop_sub_heat, select = -c(Sample.Code) )

Zoop_sub_heat_data1<-as.matrix(Zoop_sub_heat_data)
Zoop_sub_heat_data1 <- matrix( 
  as.numeric(Zoop_sub_heat_data1), ncol = 111)  
rownames(Zoop_sub_heat_data1)<-Zoop_sub_heat$Sample.Code
column<-colnames(Zoop_sub_heat_data)
colnames(Zoop_sub_heat_data1)<-column

heatmap(Zoop_sub_heat_data1,show_row_names=TRUE,show_col_names=TRUE,mar = c(15, 15))

heatmap(Zoop_sub_heat_data1, scale="column",show_row_names=TRUE,show_col_names=TRUE,mar = c(15, 15))
```

