---
title: "13_Metabarcoding_CA"
date: "2024-03-11"
output:
  html_document:
    toc: true
    toc_depth: 2
    toc_float: true
---

correspondence analysis of presence-absence data: https://r.qcbs.ca/workshop09/book-en/correspondence-analysis.html

load packages
```{r}
library(dplyr)
library(reshape2)
library(vegan)
library(stringr)
```


```{r}
COI <- read.csv("/Users/hailaschultz/Dropbox/Schultz_Dissertation/Data_Analysis/Schultz_dissertation-2/data/Barcode_2018/COI/ASVs_counts_mergedOnTaxonomy.tsv", sep = "\t")

COI_tax <- read.csv("/Users/hailaschultz/Dropbox/Schultz_Dissertation/Data_Analysis/Schultz_dissertation-2/data/Barcode_2018/COI/COI_Tax_Assignment.csv")
```

add taxonomy assignments to COI dataframe
```{r}
#rename column
COI<-COI %>% 
  rename(
    ASV = x,
    )

COI <- merge(COI_tax,COI, by="ASV", all.x=FALSE)
```

Convert to present (1) vs absent (0)
```{r}
#remove non-species rows
COI_complete<- COI[complete.cases(COI), ]
#remove non-zooplankton taxa
unique(COI_complete$Phylum)
COI_complete <- droplevels(COI_complete[!COI_complete$Phylum == 'Chlorophyta',])
COI_complete <- droplevels(COI_complete[!COI_complete$Phylum == 'Bacillariophyta',])

COI_pa<-COI_complete %>%
  mutate(across(where(is.numeric), ~ifelse(.x >0, 1, 0)))

```
```{r}
#remove excess columns including oblique tows
COI_pa <- COI_pa[ -c(1:7,11,13,16:18) ]
COI_pa<-melt(COI_pa, id.vars=c("Species"))
```

back to wide format
```{r}
COI_pa_wide<-dcast(COI_pa, variable ~ Species )
```

delete columns with all zeroes
```{r}
#remove first column
COI_pa_wide_x <- COI_pa_wide[ -c(1) ]
COI_pa_wide_x  <- COI_pa_wide_x [,-(which(colSums(COI_pa_wide_x ) == 0))]
COI_pa_wide_y = COI_pa_wide_x
COI_pa_wide_y$sample=COI_pa_wide$variable
```


Run CA
```{r}
spe.ca <- cca(COI_pa_wide_x[, -1])
```

```{r}
# Identify the significant axes
ev <- spe.ca$CA$eig

ev[ev > mean(ev)]
```

```{r}
n = length(ev)
barplot(ev, main = "Eigenvalues", col = "grey", las = 2)
abline(h = mean(ev), col = "red")
legend("topright", "Average eigenvalue", lwd = 1, col = 2, bty = "n")
```
```{r}
summary(spe.ca)
```

```{r,fig.width=20,fig.height=12}
par(mfrow = c(1, 2))

#### scaling 1
plot(spe.ca, scaling = 1, type = "none", main = "CA - biplot scaling 1",
    xlab = c("CA1 (%)", round((spe.ca$CA$eig[1]/sum(spe.ca$CA$eig)) *
        100, 2)), ylab = c("CA2 (%)", round((spe.ca$CA$eig[2]/sum(spe.ca$CA$eig)) *
        100, 2)))

points(scores(spe.ca, display = "sites", choices = c(1, 2), scaling = 1),
    pch = 21, col = "black", bg = "steelblue", cex = 1.2)

text(scores(spe.ca, display = "species", choices = c(1), scaling = 1),
    scores(spe.ca, display = "species", choices = c(2), scaling = 1),
    labels = rownames(scores(spe.ca, display = "species", scaling = 1)),
    col = "red", cex = 0.8)

#### scaling 2
plot(spe.ca, scaling = 1, type = "none", main = "CA - biplot scaling 2",
    xlab = c("CA1 (%)", round((spe.ca$CA$eig[1]/sum(spe.ca$CA$eig)) *
        100, 2)), ylab = c("CA2 (%)", round((spe.ca$CA$eig[2]/sum(spe.ca$CA$eig)) *
        100, 2)), ylim = c(-2, 3))

points(scores(spe.ca, display = "sites", choices = c(1, 2), scaling = 2),
    pch = 21, col = "black", bg = "steelblue", cex = 1.2)
text(scores(spe.ca, display = "species", choices = c(1), scaling = 2),
    scores(spe.ca, display = "species", choices = c(2), scaling = 2),
    labels = rownames(scores(spe.ca, display = "species", scaling = 2)),
    col = "red", cex = 0.8)

```



# NMDS
```{r}
NMDSmodel <- metaMDS(COI_pa_wide_x, distance = "bray",trymax = 200)
NMDSmodel
```


# Mol16S

```{r}
MOL <- read.csv("/Users/hailaschultz/Dropbox/Schultz_Dissertation/Data_Analysis/Schultz_dissertation-2/data/Barcode_2018-2020/ASV_count_tables/Mol16S_ASVs_counts.tsv", sep = "\t")

MOL_tax <- read.csv("/Users/hailaschultz/Dropbox/Schultz_Dissertation/Data_Analysis/Schultz_dissertation-2/data/Barcode_2018-2020/ASV2Taxonomy_Tables/RV_Mol16S_asvTaxonomyTable.csv")
```

add taxonomy assignments to MOL dataframe
```{r}
#rename column
MOL<-MOL %>% 
  rename(
    ASV = x,
    )

MOL <- merge(MOL_tax,MOL, by="ASV", all.x=FALSE)
```

Convert to present (1) vs absent (0)
```{r}
#remove non-species rows
MOL_complete<- MOL[complete.cases(MOL), ]
#remove everything except molluscs, echinoderms
unique(MOL_complete$Phylum)
MOL_complete <- droplevels(MOL_complete[!MOL_complete$Phylum == 'Chlorophyta',])
MOL_complete <- droplevels(MOL_complete[!MOL_complete$Phylum == 'Bacillariophyta',])
MOL_complete <- droplevels(MOL_complete[!MOL_complete$Phylum == 'Nemertea',])
MOL_complete <- droplevels(MOL_complete[!MOL_complete$Phylum == 'Arthropoda',])
MOL_complete <- droplevels(MOL_complete[!MOL_complete$Phylum == 'Chordata',])
MOL_complete <- droplevels(MOL_complete[!MOL_complete$Phylum == 'Cnidaria',])
MOL_complete <- droplevels(MOL_complete[!MOL_complete$Phylum == 'Bryozoa',])
MOL_complete <- droplevels(MOL_complete[!MOL_complete$Phylum == 'Annelida',])
MOL_complete <- droplevels(MOL_complete[!MOL_complete$Phylum == 'Hemichordata',])
MOL_complete <- droplevels(MOL_complete[!MOL_complete$Phylum == 'Phoronida',])

MOL_pa<-MOL_complete %>%
  mutate(across(where(is.numeric), ~ifelse(.x >0, 1, 0)))
unique(MOL_pa$Phylum)
```

```{r}
#remove excess columns including oblique tows
MOL_pa <- MOL_pa[ -c(1:7) ]
MOL_pa<-melt(MOL_pa, id.vars=c("Species"))
```
add up multiple lines per station
```{r}
MOL_pa <- MOL_pa %>%
  group_by(Species,variable) %>%
  summarise(
    value = sum(value))

MOL_pa<-MOL_pa %>%
  mutate(across(where(is.numeric), ~ifelse(.x >0, 1, 0)))
```
split code string up
```{r}
MOL_pa$variable<-str_split_fixed(MOL_pa$variable,"_",4)
```

remove oblique and other replicates
```{r}
MOL_pa<-subset(MOL_pa,variable[,4]!="Ob")
MOL_pa<-subset(MOL_pa,variable[,4]!="ObA")
MOL_pa<-subset(MOL_pa,variable[,4]!="ObB")
MOL_pa<-subset(MOL_pa,variable[,4]!="1B")
MOL_pa<-subset(MOL_pa,variable[,4]!="2B")
MOL_pa<-subset(MOL_pa,variable[,4]!="3B")
MOL_pa<-subset(MOL_pa,variable[,4]!="")


MOL_pa<-subset(MOL_pa,variable[,3]!="P105")
MOL_pa<-subset(MOL_pa,variable[,3]!="P123")
MOL_pa<-subset(MOL_pa,variable[,3]!="P136")
MOL_pa<-subset(MOL_pa,variable[,3]!="P381")
MOL_pa<-subset(MOL_pa,variable[,3]!="P7")

MOL_pa<-subset(MOL_pa,variable[,2]!="2010")
```



merge replicates
```{r}
MOL_pa$sample<-paste(MOL_pa$variable[,2],MOL_pa$variable[,3],sep="")

MOL_pa <- MOL_pa %>%
  group_by(Species,sample) %>%
  summarise(
    value = sum(value))

MOL_pa<-MOL_pa %>%
  mutate(across(where(is.numeric), ~ifelse(.x >0, 1, 0)))
```
back to wide format
```{r}
MOL_pa_wide<-dcast(MOL_pa, sample ~ Species )
```

delete columns with all zeroes
```{r}
#remove first column
MOL_pa_wide_x <- MOL_pa_wide[ -c(1) ]
MOL_pa_wide_x  <- MOL_pa_wide_x [,-(which(colSums(MOL_pa_wide_x ) == 0))]
```

Split up station, year, month
```{r}
MOL_pa_wide_y<-transform(MOL_pa_wide, Year = substr(sample, 1, 2), Month = substr(sample, 3, 4),Station=substr(sample, 5,7))
```

recode factors
```{r}
MOL_pa_wide_y$Year <- recode(MOL_pa_wide_y$Year, 
                                 "20" = "2020",
                         "19"="2019",
                         "18"="2018")

MOL_pa_wide_y$Month <- recode(MOL_pa_wide_y$Month, 
                                 "04" = "APR",
                         "07"="JUL",
                         "09"="SEP")

MOL_pa_wide_y$Station <- recode(MOL_pa_wide_y$Station, 
                                 "P40" = "P402")
MOL_pa_wide_y$code<-paste(MOL_pa_wide_y$Station,MOL_pa_wide_y$Year,MOL_pa_wide_y$Month,sep="-")
```

add in evironmental data
```{r}
MOL_pa_wide_y <- merge(Env_subset,MOL_pa_wide_y, by.x = "code", 
             by.y = "code", all.x = TRUE, all.y = TRUE)

#remove rows with NA
MOL_pa_wide_y<-MOL_pa_wide_y[!is.na(MOL_pa_wide_y$"sample"),]
MOL_pa_wide_y<-MOL_pa_wide_y[!is.na(MOL_pa_wide_y$"Date"),]
MOL_pa_wide_x<-MOL_pa_wide_y[ , 24:89]
```

## PERMANOVA
```{r}
dist<-vegdist(MOL_pa_wide_x, method='bray')
perm<-adonis2(dist~STATION_NO*Year.x*Month.x, data=MOL_pa_wide_y, permutations = 999, method="bray")
perm
```

```{r}
MOL_pa_wide_y$Station=as.factor(MOL_pa_wide_y$Station)
adonis.pair(vegdist(MOL_pa_wide_x),MOL_pa_wide_y$Station)
```
P12 is similar to P28,P38,P4, P402
P22 is similar to P28,P38,P4,P402,P8
P28 is similar to 

```{r}
NMDSmodel <- metaMDS(MOL_pa_wide_x, distance = "bray",trymax = 200)
NMDSmodel
```

get datascores
```{r}
data.scores<-scores(NMDSmodel,display="sites")
data.scores<- as.data.frame(data.scores)
```



add environmental variables back in
```{r}
colnames(MOL_pa_wide_y)
data.scores$Month = MOL_pa_wide_y$Month.x
data.scores$Year = MOL_pa_wide_y$Year.x
data.scores$Station = MOL_pa_wide_y$Station
data.scores$temp_deep = MOL_pa_wide_y$CTDTMP_DEG_C_ITS90_Deep...6
data.scores$temp_surface = MOL_pa_wide_y$CTDTMP_DEG_C_ITS90_Surface...7
data.scores$salinity_deep = MOL_pa_wide_y$CTDSAL_PSS78_Deep...8
data.scores$salinity_surface = MOL_pa_wide_y$CTDSAL_PSS78_Surface...9
data.scores$oxygen_deep = MOL_pa_wide_y$OXYGEN_MG_L_1_Deep...19
data.scores$oxygen_surface = MOL_pa_wide_y$OXYGEN_MG_L_1_Surface...20
data.scores$nitrate_deep = MOL_pa_wide_y$NITRATE_UMOL_L_Deep...21
data.scores$nitrate_surface = MOL_pa_wide_y$NITRATE_UMOL_L_Surface...22
```


plot datascores
```{r}
xx <- ggplot(data.scores, aes(x = NMDS1, y = NMDS2,colour = Station)) +geom_point(aes(colour = Station))+theme_bw()+theme(panel.grid.major = element_blank(),
                                                                                           panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))
xx
```

## environmental fitting
create environmetal fit table
```{r}
#check for correlation among environmental variables.
#remove nutrients other than nitrate because of autocorrelation

data.scores.cor<- data.scores[ -c(1:5,13) ]
cor_matrix <- cor(data.scores.cor)
corrplot(cor_matrix, method = "number", type = "lower", tl.pos = 'd')

Zoo.env <- data.frame(matrix(ncol = 0, nrow=53))
Zoo.env$temp_deep<-as.numeric(data.scores$temp_deep)
Zoo.env$temp_surface<-as.numeric(data.scores$temp_surface)
Zoo.env$salinity_deep<-as.numeric(data.scores$salinity_deep)
Zoo.env$salinity_surface<-as.numeric(data.scores$salinity_surface)
Zoo.env$oxygen_deep<-as.numeric(data.scores$oxygen_deep)
Zoo.env$oxygen_surface<-as.numeric(data.scores$oxygen_surface)

env <- envfit(NMDSmodel, Zoo.env, na.rm = TRUE,perm = 999)
env


env.scrs <- as.data.frame(scores(env, display = "vectors")) 
env.scrs <- cbind(env.scrs, Species = rownames(env.scrs), Pvalues = env$vectors$pvals, R_squared = env$vectors$r)
env.scrs <- subset(env.scrs, Pvalues < 0.05)
env.scrs <- subset(env.scrs, R_squared > 0.3)
env.scrs
```

## Species fitting
```{r}
vf <- envfit(NMDSmodel, MOL_pa_wide_x, perm = 999)

spp.scrs <- as.data.frame(scores(vf, display = "vectors"))
spp.scrs <- cbind(spp.scrs, Species = rownames(spp.scrs))
####for ggplot
arrow_factor <- ordiArrowMul(vf)
spp.scrs <- as.data.frame(scores(vf, display = "vectors")) * arrow_factor
spp.scrs <- cbind(spp.scrs, Species = rownames(spp.scrs), Pvalues = vf$vectors$pvals, R_squared = vf$vectors$r)

# select significance similarly to `plot(vf, p.max = 0.01)`
spp.scrs <- subset(spp.scrs, Pvalues < 0.05)
spp.scrs <- subset(spp.scrs, R_squared > 0.3)
spp.scrs

```
 extract hulls
```{r}
P12 <- data.scores[data.scores$Station == "P12", ][chull(data.scores[data.scores$Station ==                                                           "P12", c("NMDS1", "NMDS2")]), ]  
P22 <- data.scores[data.scores$Station == "P22", ][chull(data.scores[data.scores$Station ==                                                           "P22", c("NMDS1", "NMDS2")]), ]  
P28 <- data.scores[data.scores$Station == "P28", ][chull(data.scores[data.scores$Station ==                                                           "P28", c("NMDS1", "NMDS2")]), ]
P38 <- data.scores[data.scores$Station == "P38", ][chull(data.scores[data.scores$Station ==                                                           "P38", c("NMDS1", "NMDS2")]), ]  
P4 <- data.scores[data.scores$Station == "P4", ][chull(data.scores[data.scores$Station ==                                                           "P4", c("NMDS1", "NMDS2")]), ]  
P402 <- data.scores[data.scores$Station == "P402", ][chull(data.scores[data.scores$Station ==                                                           "P402", c("NMDS1", "NMDS2")]), ]  
P8 <- data.scores[data.scores$Station == "P8", ][chull(data.scores[data.scores$Station ==                                                           "P8", c("NMDS1", "NMDS2")]), ]  
```

get hull data
```{r}
hull.data <- rbind(P12, P22, P28,P38,P4,P402,P8)  #combine grp.a and grp.b
hull.data
```
 
```{r}

data.scores$Station <- factor(data.scores$Station, levels = c("P22", "P4", "P8","P28","P12","P402","P38"))
data.scores$Station <- recode(data.scores$Station, 
                                 "P22" = "Juan de Fuca",
                         "P4"="Whidbey Basin",
                         "P8"="HC Sill",
                         "P28"="Main Basin",
                         "P12"="Hood Canal 1",
                         "P402"="Hood Canal 2",
                         "P38"="South Sound")

hull.data$Station <- factor(hull.data$Station, levels = c("P22", "P4", "P8","P28","P12","P402","P38"))
hull.data$Station <- recode(hull.data$Station, 
                                 "P22" = "Juan de Fuca",
                         "P4"="Whidbey Basin",
                         "P8"="HC Sill",
                         "P28"="Main Basin",
                         "P12"="Hood Canal 1",
                         "P402"="Hood Canal 2",
                         "P38"="South Sound")
```

```{r,fig.width=6.5,fig.height=5, dpi=600}

Species_plot <- ggplot(data.scores, aes(x = NMDS1, y = NMDS2)) +geom_point(aes(colour =  Station))+ scale_colour_manual(values=c("#F8766D","#C49A00","#53B400","#00C094","#00B6EB","#A69AFF","#FB61D7")) +
  scale_fill_manual(values=c("#F8766D","#C49A00","#53B400","#00C094","#00B6EB","#A69AFF","#FB61D7"))+theme_bw()+theme(panel.grid.major = element_blank(),
                                                                                           panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))+ 
  geom_polygon(data=hull.data,aes(x=NMDS1,y=NMDS2,fill=Station,group=Station),alpha=0.20, size=0.1, linetype=1, colour="black") +
  geom_segment(data = spp.scrs,size=0.4,
               aes(x = 0, xend = NMDS1, y = 0, yend = NMDS2),
               arrow = arrow(length = unit(0.2, "cm")), colour = "darkgrey")+
  geom_segment(data = env.scrs,size=0.4,
               aes(x = 0, xend = NMDS1, y = 0, yend = NMDS2),
               arrow = arrow(length = unit(0.2, "cm")), colour = "black")+
  annotate("text", label = "2D Stress: 0.2", x = 0.6, y = 1.5, size = 5, colour = "black")+
  geom_label_repel(data = spp.scrs, aes(x = NMDS1, y = NMDS2, label = Species),
                  size = 3, fontface="bold", fill="white", label.padding = unit(0.15, "lines"), box.padding = unit(0.16, "lines"), label.size = 0.05)
Species_plot
```
save dendogram
```{r, eval=FALSE}
setwd("/Users/hailaschultz/Dropbox/Schultz_Dissertation/Data_Analysis/Schultz_dissertation-2/figures")
ggsave(filename = "PA_Mol_nmds.png", plot = Species_plot, width = 7, height = 5.5, device='png', dpi=700)
```
#cca
Run CA
```{r}
spe.ca <- cca(MOL_pa_wide_x)
```

```{r}
# Identify the significant axes
ev <- spe.ca$CA$eig

ev[ev > mean(ev)]
```

```{r}
n = length(ev)
barplot(ev, main = "Eigenvalues", col = "grey", las = 2)
abline(h = mean(ev), col = "red")
legend("topright", "Average eigenvalue", lwd = 1, col = 2, bty = "n")
```
```{r}
summary(spe.ca)
```

```{r,fig.width=20,fig.height=12}
par(mfrow = c(1, 2))

#### scaling 1
plot(spe.ca, scaling = 1, type = "none", main = "CA - biplot scaling 1",
    xlab = c("CA1 (%)", round((spe.ca$CA$eig[1]/sum(spe.ca$CA$eig)) *
        100, 2)), ylab = c("CA2 (%)", round((spe.ca$CA$eig[2]/sum(spe.ca$CA$eig)) *
        100, 2)))

points(scores(spe.ca, display = "sites", choices = c(1, 2), scaling = 1),
    pch = 21, col = "black", bg = "steelblue", cex = 1.2)

text(scores(spe.ca, display = "species", choices = c(1), scaling = 1),
    scores(spe.ca, display = "species", choices = c(2), scaling = 1),
    labels = rownames(scores(spe.ca, display = "species", scaling = 1)),
    col = "red", cex = 0.8)

#### scaling 2
plot(spe.ca, scaling = 1, type = "none", main = "CA - biplot scaling 2",
    xlab = c("CA1 (%)", round((spe.ca$CA$eig[1]/sum(spe.ca$CA$eig)) *
        100, 2)), ylab = c("CA2 (%)", round((spe.ca$CA$eig[2]/sum(spe.ca$CA$eig)) *
        100, 2)), ylim = c(-2, 3))

points(scores(spe.ca, display = "sites", choices = c(1, 2), scaling = 2),
    pch = 21, col = "black", bg = "steelblue", cex = 1.2)
text(scores(spe.ca, display = "species", choices = c(1), scaling = 2),
    scores(spe.ca, display = "species", choices = c(2), scaling = 2),
    labels = rownames(scores(spe.ca, display = "species", scaling = 2)),
    col = "red", cex = 0.8)

```

plot cca using ggplot
```{r}
# Enable the r-universe repo
options(repos = c(
    fawda123 = 'https://fawda123.r-universe.dev',
    CRAN = 'https://cloud.r-project.org'))

# Install ggord
install.packages('ggord')
library(ggord)
ggord(spe.ca, MOL_pa_wide_y$Station) + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) # looking at the raw code, this is plotting the 'wa scores'
```

