---
title: "14_env-cluster_nanoos"
date: "2024-03-13"
output:
  html_document:
    toc: true
    toc_depth: 2
    toc_float: true
---
the environmental data was created by 00_Upload_Env_Data

subset environmental data to only shallow and deep niskins
```{r}
#add P to station numbers
subset_df$STATION_NO <- paste0("P", subset_df$STATION_NO)
#make unique code
subset_df$code<-paste(subset_df$STATION_NO, subset_df$Year,subset_df$Month,sep="-")

# Sort the dataframe by sample and then by depth
subset_df <- subset_df %>% arrange(sample, NISKIN_NO)
```

remove unnecessary columns
```{r}
Env_subset <- subset_df[ -c(1:7,10,12,14:17,19:25,27:32) ]
```

remove rows with missing data
```{r}
Env_subset<-Env_subset %>% drop_na(OXYGEN_MG_L_1)
```


```{r}
# Group by sample and filter the shallowest and deepest rows for each sample
Env_subset <- Env_subset %>%
  group_by(sample) %>%
  mutate(row_num = row_number()) %>%
  filter(row_num == 1 | row_num == n())

# Create a new column to label them as shallow or deep
Env_subset <- Env_subset %>%
  mutate(Depth = ifelse(row_num == 1, "Deep","Surface" )) %>%
  select(-row_num) # Remove the temporary row number column
```

separate surface and deep samples
```{r}
#remove niskin no
Env_subset <- Env_subset[ -c(2) ]
#make surface and deep variables separate
Env_subset <- Env_subset %>%
  pivot_wider(names_from = Depth, values_from = c(CTDTMP_DEG_C_ITS90, CTDSAL_PSS78, OXYGEN_MG_L_1, NITRATE_UMOL_L)) 
Env_subset [Env_subset  < 0] <- NA
Env_subset <- Env_subset[complete.cases(Env_subset), ]
```


# Station Anomalies

```{r}
colnames(Env_subset)
env_ave<-Env_subset %>%
  group_by(STATION_NO) %>%
  summarise(across("CTDTMP_DEG_C_ITS90_Deep":"NITRATE_UMOL_L_Surface", list(mean)))

```

standard deviation
```{r}
env_stdev<-Env_subset %>%
  group_by(STATION_NO) %>%
  summarise(across("CTDTMP_DEG_C_ITS90_Deep":"NITRATE_UMOL_L_Surface", list(sd)))

```

mean anomalies
```{r}
#add averages column
env_joined_ave <- merge(Env_subset,env_ave, by="STATION_NO", all.x=TRUE)
env_joined_stdev <- merge(env_joined_ave,env_stdev, by="STATION_NO", all.x=TRUE)

#calculate anomalies
env_joined_anom <- (env_joined_stdev[, 7:14] - env_joined_stdev[, 15:22])/env_joined_stdev[, 23:30]
```


```{r,fig.width=10,fig.height=10}
#convert station column to rownames
env_joined_anom$code=Env_subset$code
env_joined_anom<-env_joined_anom %>% remove_rownames %>% column_to_rownames(var="code")
#remove columns and rows with na
env_joined_anom <- env_joined_anom[ -c(8) ]
env_joined_anom <- env_joined_anom[complete.cases(env_joined_anom), ]
cor_matrix <- cor(env_joined_anom)
corrplot(cor_matrix, method = "number", type = "lower", tl.pos = 'd')
```
no crazy correlations

### Centroid Based
gap statistic: do this to determine number of clusters
```{r}
res.km<-eclust(env_joined_anom, FUNcluster = "kmeans", hc_metric = "euclidean", graph = FALSE)
fviz_gap_stat(res.km$gap_stat)

#silhouette plot
fviz_silhouette(res.km)
```
2 groups! 

### Heirarchical
```{r,fig.width=20,fig.height=12}
res.hclust <- eclust(env_joined_anom, FUNcluster = "hclust", hc_metric = "euclidean")
env_dend<-fviz_dend(res.hclust, rect = TRUE,repel=phylogenic)
env_dend
fviz_cluster(res.hclust, labelsize = 10)
```
okay, they are very clearly separating by month now. 

### Agglomerative clustering
```{r,fig.width=15,fig.height=10}
res.aclust <- eclust(env_joined_anom, FUNcluster = "hclust", hc_metric = "euclidean", hc_method = "ward.D2")
fviz_dend(res.aclust, rect = TRUE)
fviz_cluster(res.aclust, labelsize = 10)
```

### visualize the strength of the variables
```{r}
df_clusters <- res.km$centers
res.km
```

```{r}
melt_df <- melt(df_clusters)

heatmap <- ggplot(melt_df, aes(Var2, Var1)) +
  scale_fill_continuous(type = "viridis", direction = -1) +
  geom_tile(aes(fill = value)) +
  geom_text(aes(label = round(value, 1))) +
  theme_bw() +
  ggtitle("Strength of Each of the Variables in the Clusters") +
  theme(plot.title = element_text(hjust = 0.5)) +
  labs(x="Variable", y="Cluster")
heatmap
```
### list of statons and clusters
```{r}
env_joined_anom["Cluster"] <- res.km$cluster
df_out <- env_joined_anom[order(-env_joined_anom$Cluster), ]
knitr::kable(df_out)
```

# Month Anomalies

```{r}
colnames(Env_subset)
env_ave<-Env_subset %>%
  group_by(Month) %>%
  summarise(across("CTDTMP_DEG_C_ITS90_Deep":"NITRATE_UMOL_L_Surface", list(mean)))

```

standard deviation
```{r}
env_stdev<-Env_subset %>%
  group_by(Month) %>%
  summarise(across("CTDTMP_DEG_C_ITS90_Deep":"NITRATE_UMOL_L_Surface", list(sd)))

```

mean anomalies
```{r}
#add averages column
env_joined_ave <- merge(Env_subset,env_ave, by="Month", all.x=TRUE)
env_joined_stdev <- merge(env_joined_ave,env_stdev, by="Month", all.x=TRUE)

#calculate anomalies
env_joined_anom <- (env_joined_stdev[, 7:14] - env_joined_stdev[, 15:22])/env_joined_stdev[, 23:30]
```


```{r,fig.width=10,fig.height=10}
#convert station column to rownames
env_joined_anom$code=Env_subset$code
env_joined_anom<-env_joined_anom %>% remove_rownames %>% column_to_rownames(var="code")
#remove columns and rows with na
env_joined_anom <- env_joined_anom[ -c(8) ]
env_joined_anom <- env_joined_anom[complete.cases(env_joined_anom), ]
cor_matrix <- cor(env_joined_anom)
corrplot(cor_matrix, method = "number", type = "lower", tl.pos = 'd')
```
remove deep nitrate
```{r}
env_joined_anom <- env_joined_anom[ -c(7) ]
```

### Centroid Based
gap statistic: do this to determine number of clusters
```{r}
res.km<-eclust(env_joined_anom, FUNcluster = "kmeans", hc_metric = "euclidean", graph = FALSE)
fviz_gap_stat(res.km$gap_stat)

#silhouette plot
fviz_silhouette(res.km)
```
5 groups (very interesting)
### Heirarchical
```{r,fig.width=20,fig.height=12}
res.hclust <- eclust(env_joined_anom, FUNcluster = "hclust", hc_metric = "euclidean")
env_dend<-fviz_dend(res.hclust, rect = TRUE,repel=phylogenic)
env_dend
fviz_cluster(res.hclust, labelsize = 10)
```
8 groups 

### visualize the strength of the variables
```{r}
df_clusters <- res.km$centers
res.km
```
```{r}
melt_df <- melt(df_clusters)

heatmap <- ggplot(melt_df, aes(Var2, Var1)) +
  scale_fill_continuous(type = "viridis", direction = -1) +
  geom_tile(aes(fill = value)) +
  geom_text(aes(label = round(value, 1))) +
  theme_bw() +
  ggtitle("Strength of Each of the Variables in the Clusters") +
  theme(plot.title = element_text(hjust = 0.5)) +
  labs(x="Variable", y="Cluster")
heatmap
```
### list of statons and clusters
```{r}
env_joined_anom["Cluster"] <- res.km$cluster
df_out <- env_joined_anom[order(-env_joined_anom$Cluster), ]
knitr::kable(df_out)
```

save table

```{r, eval=FALSE}
write.csv(df_out, "/Users/hailaschultz/Dropbox/Schultz_Dissertation/Data_Analysis/Schultz_dissertation-2/output/env_cluster_nanoos.csv", row.names=TRUE)
```


# Non-anomaly clusters
```{r,fig.width=10,fig.height=10}
#convert station column to rownames
Env_full <- Env_subset[ -c(1:5) ]
Env_full<-Env_full %>% remove_rownames %>% column_to_rownames(var="code")
#remove columns and rows with na
Env_full <- Env_full[ -c(8,9) ]
Env_full <- Env_full[complete.cases(Env_full), ]
# Standardize variables
Env_full <- Env_full %>% mutate_all(~(scale(.) %>% as.vector))
cor_matrix <- cor(Env_full)
corrplot(cor_matrix, method = "number", type = "lower", tl.pos = 'd')
```

looks like there are some correlations, remove deep nitrate because it's correlated with deep oxygen
```{r}
Env_full <- Env_full[ -c(7) ]
```

### Centroid Based
gap statistic: do this to determine number of clusters
```{r}
res.km<-eclust(Env_full, FUNcluster = "kmeans", hc_metric = "euclidean", graph = FALSE)
fviz_gap_stat(res.km$gap_stat)

#silhouette plot
fviz_silhouette(res.km)
```
six groups with raw data
cluster plot
```{r,fig.width=20,fig.height=20}
eclust(Env_full, FUNcluster = "kmeans", k = 6, hc_metric = "euclidean")
```
April and P22 pretty clearly separate out

### Heirarchical
```{r,fig.width=20,fig.height=12}
res.hclust <- eclust(Env_full, FUNcluster = "hclust", hc_metric = "euclidean")
env_dend<-fviz_dend(res.hclust, rect = TRUE,repel=phylogenic)
env_dend
fviz_cluster(res.hclust, labelsize = 10)
```

gives 9 groups: larger groups of (P402,P12,P4)
(P22) (P38,P28,P8)


### visualize the strength of the variables
```{r}
df_clusters <- res.km$centers
res.km
```


# Month Anomaly by station group

```{r}
group1<-subset(Env_subset,STATION_NO==c("PP402","PP12","PP4"))
```


```{r}
colnames(Env_subset)
env_ave<-group1 %>%
  group_by(Month) %>%
  summarise(across("CTDTMP_DEG_C_ITS90_Deep":"NITRATE_UMOL_L_Surface", list(mean)))

```

standard deviation
```{r}
env_stdev<-group1 %>%
  group_by(Month) %>%
  summarise(across("CTDTMP_DEG_C_ITS90_Deep":"NITRATE_UMOL_L_Surface", list(sd)))

```

mean anomalies
```{r}
#add averages column
env_joined_ave <- merge(group1,env_ave, by="Month", all.x=TRUE)
env_joined_stdev <- merge(env_joined_ave,env_stdev, by="Month", all.x=TRUE)

#calculate anomalies
env_joined_anom <- (env_joined_stdev[, 7:14] - env_joined_stdev[, 15:22])/env_joined_stdev[, 23:30]
```


```{r,fig.width=10,fig.height=10}
#convert station column to rownames
env_joined_anom$code=group1$code
env_joined_anom<-env_joined_anom %>% remove_rownames %>% column_to_rownames(var="code")
#remove columns and rows with na
env_joined_anom <- env_joined_anom[ -c(8) ]
env_joined_anom <- env_joined_anom[complete.cases(env_joined_anom), ]
cor_matrix <- cor(env_joined_anom)
corrplot(cor_matrix, method = "number", type = "lower", tl.pos = 'd')
```
### Centroid Based
gap statistic: do this to determine number of clusters
```{r}
res.km<-eclust(env_joined_anom, FUNcluster = "kmeans", hc_metric = "euclidean", graph = FALSE)
fviz_gap_stat(res.km$gap_stat)

#silhouette plot
fviz_silhouette(res.km)
```

okay, this doesn't work.