---
title: "15_COI_NMDS"
date: "2024-03-18"
output:
  html_document:
    toc: true
    toc_depth: 2
    toc_float: true
---

```{r}
COI <- read.csv("/Users/hailaschultz/Dropbox/Schultz_Dissertation/Data_Analysis/Schultz_dissertation-2/data/Barcode_2018-2020/ASV_count_tables/COI_ASVs_counts.tsv", sep = "\t")

COI_tax <- read.csv("/Users/hailaschultz/Dropbox/Schultz_Dissertation/Data_Analysis/Schultz_dissertation-2/data/Barcode_2018-2020/ASV2Taxonomy_Tables/RV_COI_asvTaxonomyTable.csv")
```


add taxonomy assignments to dataframe
```{r}
#rename column
COI<-COI %>% 
  rename(
    ASV = x,
    )

COI <- merge(COI_tax,COI, by="ASV", all.x=FALSE)
```

Convert to present (1) vs absent (0)
```{r}
#remove non-species rows
COI_complete<- COI[complete.cases(COI), ]
#remove everything except COIluscs, echinoderms
unique(COI_complete$Phylum)
COI_complete <- droplevels(COI_complete[!COI_complete$Phylum == 'Chlorophyta',])
COI_complete <- droplevels(COI_complete[!COI_complete$Phylum == 'Bacillariophyta',])
COI_complete <- droplevels(COI_complete[!COI_complete$Phylum == 'Rhodophyta',])
COI_complete <- droplevels(COI_complete[!COI_complete$Phylum == 'Cryptophyceae__p',])
COI_complete <- droplevels(COI_complete[!COI_complete$Phylum == 'Haptophyta',])
COI_complete <- droplevels(COI_complete[!COI_complete$Phylum == 'Basidiomycota',])
COI_complete <- droplevels(COI_complete[!COI_complete$Phylum == 'Tubulinea',])
COI_complete <- droplevels(COI_complete[!COI_complete$Phylum == 'Dictyochophyceae__p',])
COI_complete <- droplevels(COI_complete[!COI_complete$Phylum == 'Ascomycota',])
COI_complete <- droplevels(COI_complete[!COI_complete$Phylum == 'Phaeophyceae__p',])
COI_complete <- droplevels(COI_complete[!COI_complete$Phylum == 'Pelagophyceae__p',])

COI_pa<-COI_complete %>%
  mutate(across(where(is.numeric), ~ifelse(.x >0, 1, 0)))
unique(COI_pa$Phylum)
```

```{r}
#remove excess columns including oblique tows
COI_pa <- COI_pa[ -c(1:7) ]
COI_pa<-melt(COI_pa, id.vars=c("Species"))
```

add up multiple lines per station
```{r}
COI_pa <- COI_pa %>%
  group_by(Species,variable) %>%
  summarise(
    value = sum(value))

COI_pa<-COI_pa %>%
  mutate(across(where(is.numeric), ~ifelse(.x >0, 1, 0)))
```
split code string up
```{r}
COI_pa$variable<-str_split_fixed(COI_pa$variable,"_",4)
```

remove oblique and other replicates
```{r}
COI_pa<-subset(COI_pa,variable[,4]!="Ob")
COI_pa<-subset(COI_pa,variable[,4]!="ObA")
COI_pa<-subset(COI_pa,variable[,4]!="ObB")
COI_pa<-subset(COI_pa,variable[,4]!="1B")
COI_pa<-subset(COI_pa,variable[,4]!="2B")
COI_pa<-subset(COI_pa,variable[,4]!="3B")
COI_pa<-subset(COI_pa,variable[,4]!="")


COI_pa<-subset(COI_pa,variable[,3]!="P105")
COI_pa<-subset(COI_pa,variable[,3]!="P123")
COI_pa<-subset(COI_pa,variable[,3]!="P136")
COI_pa<-subset(COI_pa,variable[,3]!="P381")
COI_pa<-subset(COI_pa,variable[,3]!="P7")

COI_pa<-subset(COI_pa,variable[,2]!="2010")
```

merge replicates
```{r}
COI_pa$sample<-paste(COI_pa$variable[,2],COI_pa$variable[,3],sep="")

COI_pa <- COI_pa %>%
  group_by(Species,sample) %>%
  summarise(
    value = sum(value))

COI_pa<-COI_pa %>%
  mutate(across(where(is.numeric), ~ifelse(.x >0, 1, 0)))
```

back to wide format
```{r}
COI_pa_wide<-dcast(COI_pa, sample ~ Species )
```


```{r}
#remove first column
COI_pa_wide_x <- COI_pa_wide[ -c(1) ]
COI_pa_wide_x  <- COI_pa_wide_x [,-(which(colSums(COI_pa_wide_x ) == 0))]
```

Split up station, year, month
```{r}
COI_pa_wide_y<-transform(COI_pa_wide, Year = substr(sample, 1, 2), Month = substr(sample, 3, 4),Station=substr(sample, 5,7))
```

recode factors
```{r}
COI_pa_wide_y$Year <- recode(COI_pa_wide_y$Year, 
                                 "20" = "2020",
                         "19"="2019",
                         "18"="2018")

COI_pa_wide_y$Month <- recode(COI_pa_wide_y$Month, 
                                 "04" = "APR",
                         "07"="JUL",
                         "09"="SEP")

COI_pa_wide_y$Station <- recode(COI_pa_wide_y$Station, 
                                 "P40" = "P402")
COI_pa_wide_y$code<-paste(COI_pa_wide_y$Station,COI_pa_wide_y$Year,COI_pa_wide_y$Month,sep="-")
```

add in evironmental data
```{r}
COI_pa_wide_y <- merge(Env_subset,COI_pa_wide_y, by.x = "code", 
             by.y = "code", all.x = TRUE, all.y = TRUE)

#remove rows with NA
COI_pa_wide_y<-COI_pa_wide_y[!is.na(COI_pa_wide_y$"Acartia.longiremis"),]
COI_pa_wide_y<-COI_pa_wide_y[!is.na(COI_pa_wide_y$"Date"),]
COI_pa_wide_x<-COI_pa_wide_y[ , 15:325]
```



## PERMANOVA
```{r}
dist<-vegdist(COI_pa_wide_x, method='bray')
perm<-adonis2(dist~STATION_NO*Year.x*Month.x, data=COI_pa_wide_y, permutations = 999, method="bray")
perm
```

```{r}
COI_pa_wide_y$Station=as.factor(COI_pa_wide_y$Station)
adonis.pair(vegdist(COI_pa_wide_x),COI_pa_wide_y$Station)
```

```{r}
NMDSmodel <- metaMDS(COI_pa_wide_x, distance = "bray",trymax = 999)
NMDSmodel
```

get datascores
```{r}
data.scores<-scores(NMDSmodel,display="sites")
data.scores<- as.data.frame(data.scores)
```

add environmental variables back in
```{r}
colnames(COI_pa_wide_y)
data.scores$Month = COI_pa_wide_y$Month.x
data.scores$Year = COI_pa_wide_y$Year.x
data.scores$Station = COI_pa_wide_y$Station
data.scores$temp_deep = COI_pa_wide_y$CTDTMP_DEG_C_ITS90_Deep
data.scores$temp_surface = COI_pa_wide_y$CTDTMP_DEG_C_ITS90_Surface
data.scores$salinity_deep = COI_pa_wide_y$CTDSAL_PSS78_Deep
data.scores$salinity_surface = COI_pa_wide_y$CTDSAL_PSS78_Surface
data.scores$oxygen_deep = COI_pa_wide_y$OXYGEN_MG_L_1_Deep
data.scores$oxygen_surface = COI_pa_wide_y$OXYGEN_MG_L_1_Surface
data.scores$nitrate_deep = COI_pa_wide_y$NITRATE_UCOI_L_Deep
data.scores$nitrate_surface = COI_pa_wide_y$NITRATE_UCOI_L_Surface
```

plot datascores
```{r}
xx <- ggplot(data.scores, aes(x = NMDS1, y = NMDS2,colour = Station)) +geom_point(aes(colour = Station))+theme_bw()+theme(panel.grid.major = element_blank(),
                                                                                           panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))
xx
```

extract hulls
```{r}
P12 <- data.scores[data.scores$Station == "P12", ][chull(data.scores[data.scores$Station ==                                                           "P12", c("NMDS1", "NMDS2")]), ]  
P22 <- data.scores[data.scores$Station == "P22", ][chull(data.scores[data.scores$Station ==                                                           "P22", c("NMDS1", "NMDS2")]), ]  
P28 <- data.scores[data.scores$Station == "P28", ][chull(data.scores[data.scores$Station ==                                                           "P28", c("NMDS1", "NMDS2")]), ]
P38 <- data.scores[data.scores$Station == "P38", ][chull(data.scores[data.scores$Station ==                                                           "P38", c("NMDS1", "NMDS2")]), ]  
P4 <- data.scores[data.scores$Station == "P4", ][chull(data.scores[data.scores$Station ==                                                           "P4", c("NMDS1", "NMDS2")]), ]  
P402 <- data.scores[data.scores$Station == "P402", ][chull(data.scores[data.scores$Station ==                                                           "P402", c("NMDS1", "NMDS2")]), ]  
P8 <- data.scores[data.scores$Station == "P8", ][chull(data.scores[data.scores$Station ==                                                           "P8", c("NMDS1", "NMDS2")]), ]  
```

get hull data
```{r}
hull.data <- rbind(P12, P22, P28,P38,P4,P402,P8)  #combine grp.a and grp.b
hull.data
```
## environmental fitting
create environmetal fit table
```{r}
#check for correlation among environmental variables.
#remove nutrients other than nitrate because of autocorrelation

data.scores.cor<- data.scores[ -c(1:5,13) ]
cor_matrix <- cor(data.scores.cor)
corrplot(cor_matrix, method = "number", type = "lower", tl.pos = 'd')

Zoo.env <- data.frame(matrix(ncol = 0, nrow=52))
Zoo.env$temp_deep<-as.numeric(data.scores$temp_deep)
Zoo.env$temp_surface<-as.numeric(data.scores$temp_surface)
Zoo.env$salinity_deep<-as.numeric(data.scores$salinity_deep)
Zoo.env$salinity_surface<-as.numeric(data.scores$salinity_surface)
Zoo.env$oxygen_deep<-as.numeric(data.scores$oxygen_deep)
Zoo.env$oxygen_surface<-as.numeric(data.scores$oxygen_surface)

env <- envfit(NMDSmodel, Zoo.env, na.rm = TRUE,perm = 999)
env


env.scrs <- as.data.frame(scores(env, display = "vectors")) 
env.scrs <- cbind(env.scrs, Species = rownames(env.scrs), Pvalues = env$vectors$pvals, R_squared = env$vectors$r)
env.scrs <- subset(env.scrs, Pvalues < 0.05)
env.scrs <- subset(env.scrs, R_squared > 0.3)
env.scrs
```

## Species fitting
```{r}
vf <- envfit(NMDSmodel, COI_pa_wide_x, perm = 999)

spp.scrs <- as.data.frame(scores(vf, display = "vectors"))
spp.scrs <- cbind(spp.scrs, Species = rownames(spp.scrs))
####for ggplot
arrow_factor <- ordiArrowMul(vf)
spp.scrs <- as.data.frame(scores(vf, display = "vectors")) * arrow_factor
spp.scrs <- cbind(spp.scrs, Species = rownames(spp.scrs), Pvalues = vf$vectors$pvals, R_squared = vf$vectors$r)

# select significance similarly to `plot(vf, p.max = 0.01)`
spp.scrs <- subset(spp.scrs, Pvalues < 0.05)
spp.scrs <- subset(spp.scrs, R_squared > 0.3)
spp.scrs

```

```{r}
coi.data.scores<-data.scores
coi.data.scores$Station <- factor(coi.data.scores$Station, levels = c("P22", "P4", "P8","P28","P12","P402","P38"))
coi.data.scores$Station <- recode(coi.data.scores$Station, 
                                 "P22" = "Juan de Fuca",
                         "P4"="Whidbey Basin",
                         "P8"="HC Sill",
                         "P28"="Main Basin",
                         "P12"="Hood Canal 1",
                         "P402"="Hood Canal 2",
                         "P38"="South Sound")

hull.data$Station <- factor(hull.data$Station, levels = c("P22", "P4", "P8","P28","P12","P402","P38"))
hull.data$Station <- recode(hull.data$Station, 
                                 "P22" = "Juan de Fuca",
                         "P4"="Whidbey Basin",
                         "P8"="HC Sill",
                         "P28"="Main Basin",
                         "P12"="Hood Canal 1",
                         "P402"="Hood Canal 2",
                         "P38"="South Sound")
```

```{r,fig.width=7,fig.height=5, dpi=600}


Species_plot <- ggplot(coi.data.scores, aes(x = NMDS1, y = NMDS2)) +geom_point(aes(colour =  Station))+ scale_colour_manual(values=c("#F8766D","#C49A00","#53B400","#00C094","#00B6EB","#A69AFF","#FB61D7")) +
  scale_fill_manual(values=c("#F8766D","#C49A00","#53B400","#00C094","#00B6EB","#A69AFF","#FB61D7"))+theme_bw()+theme(panel.grid.major = element_blank(),
                                                                                           panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))+
  geom_segment(data = spp.scrs,size=0.4,
               aes(x = 0, xend = NMDS1, y = 0, yend = NMDS2),
               arrow = arrow(length = unit(0.2, "cm")), colour = "darkgrey")+
  geom_segment(data = env.scrs,size=0.4,
               aes(x = 0, xend = NMDS1, y = 0, yend = NMDS2),
               arrow = arrow(length = unit(0.2, "cm")), colour = "black")+ 
  geom_polygon(data=hull.data,aes(x=NMDS1,y=NMDS2,fill=Station,group=Station),alpha=0.20, size=0.1, linetype=1, colour="black") +
  annotate("text", label = "2D Stress: 0.2", x = 0.55, y = 0.67, size = 5, colour = "black")+
  geom_label_repel(data = spp.scrs, aes(x = NMDS1, y = NMDS2, label = Species),
                  size = 3, fontface="bold", fill="white", label.padding = unit(0.15, "lines"), box.padding = unit(0.16, "lines"), label.size = 0.05)
Species_plot
```


```{r,fig.width=7,fig.height=5, dpi=600}


Species_plot <- ggplot(coi.data.scores, aes(x = NMDS1, y = NMDS2)) +geom_point(aes(colour =  Station))+ scale_colour_manual(values=c("#F8766D","#C49A00","#53B400","#00C094","#00B6EB","#A69AFF","#FB61D7")) +
  scale_fill_manual(values=c("#F8766D","#C49A00","#53B400","#00C094","#00B6EB","#A69AFF","#FB61D7"))+theme_bw()+theme(panel.grid.major = element_blank(),
                                                                                           panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))+
  geom_segment(data = spp.scrs,size=0.4,
               aes(x = 0, xend = NMDS1, y = 0, yend = NMDS2),
               arrow = arrow(length = unit(0.2, "cm")), colour = "darkgrey")+
  geom_segment(data = env.scrs,size=0.4,
               aes(x = 0, xend = NMDS1, y = 0, yend = NMDS2),
               arrow = arrow(length = unit(0.2, "cm")), colour = "black")+ 
  geom_polygon(data=hull.data,aes(x=NMDS1,y=NMDS2,fill=Station,group=Station),alpha=0.20, size=0.1, linetype=1, colour="black") +
  annotate("text", label = "2D Stress: 0.2", x = 0.55, y = 0.67, size = 5, colour = "black")
Species_plot
```

save dendogram
```{r, eval=FALSE}
setwd("/Users/hailaschultz/Dropbox/Schultz_Dissertation/Data_Analysis/Schultz_dissertation-2/figures")
ggsave(filename = "PA_COI_nmds.png", plot = Species_plot, width = 7, height = 5.5, device='png', dpi=700)
```