---
title: "16_morph_nmds"
date: "2024-03-18"
output:
  html_document:
    toc: true
    toc_depth: 2
    toc_float: true
---
```{r}
Morph <- read.csv("/Users/hailaschultz/Dropbox/Schultz_Dissertation/Data_Analysis/Schultz_dissertation-2/data/Barcode_2018-2020/ASV_count_tables/Morphdensity_ASVs_counts.tsv", sep = "\t")

Morph_tax <- read.csv("/Users/hailaschultz/Dropbox/Schultz_Dissertation/Data_Analysis/Schultz_dissertation-2/data/Barcode_2018-2020/ASV2Taxonomy_Tables/morphology_asvTaxonomyTable.csv")
```

add taxonomy assignments to dataframe
```{r}
#rename column
Morph<-Morph %>% 
  rename(
    ASV = x,
    )

Morph <- merge(Morph_tax,Morph, by="ASV", all.x=FALSE)
```

Convert to present (1) vs absent (0)
```{r}
#remove non-species rows
Morph_complete<- Morph[complete.cases(Morph), ]

Morph_pa<-Morph_complete %>%
  mutate(across(where(is.numeric), ~ifelse(.x >0, 1, 0)))
unique(Morph_pa$Phylum)
```

```{r}
#remove excess columns including oblique tows
Morph_pa <- Morph_pa[ -c(1:7) ]
Morph_pa<-melt(Morph_pa, id.vars=c("Species"))
```

add up multiple lines per station
```{r}
Morph_pa <- Morph_pa %>%
  group_by(Species,variable) %>%
  summarise(
    value = sum(value))

Morph_pa<-Morph_pa %>%
  mutate(across(where(is.numeric), ~ifelse(.x >0, 1, 0)))
```
split code string up
```{r}
Morph_pa$variable<-str_split_fixed(Morph_pa$variable,"_",4)
```


remove oblique and other replicates
```{r}
Morph_pa<-subset(Morph_pa,variable[,4]!="Ob")
Morph_pa<-subset(Morph_pa,variable[,4]!="ObA")
Morph_pa<-subset(Morph_pa,variable[,4]!="ObB")
Morph_pa<-subset(Morph_pa,variable[,4]!="1B")
Morph_pa<-subset(Morph_pa,variable[,4]!="2B")
Morph_pa<-subset(Morph_pa,variable[,4]!="3B")
Morph_pa<-subset(Morph_pa,variable[,4]!="")


Morph_pa<-subset(Morph_pa,variable[,3]!="P105")
Morph_pa<-subset(Morph_pa,variable[,3]!="P123")
Morph_pa<-subset(Morph_pa,variable[,3]!="P136")
Morph_pa<-subset(Morph_pa,variable[,3]!="P381")
Morph_pa<-subset(Morph_pa,variable[,3]!="P7")

Morph_pa<-subset(Morph_pa,variable[,2]!="2010")
```

merge replicates
```{r}
Morph_pa$sample<-paste(Morph_pa$variable[,2],Morph_pa$variable[,3],sep="")

Morph_pa <- Morph_pa %>%
  group_by(Species,sample) %>%
  summarise(
    value = sum(value))

Morph_pa<-Morph_pa %>%
  mutate(across(where(is.numeric), ~ifelse(.x >0, 1, 0)))
```

back to wide format
```{r}
Morph_pa_wide<-dcast(Morph_pa, sample ~ Species )
```

```{r}
#remove first column
Morph_pa_wide_x <- Morph_pa_wide[ -c(1) ]
Morph_pa_wide_x  <- Morph_pa_wide_x [,-(which(colSums(Morph_pa_wide_x ) == 0))]
```

Split up station, year, month
```{r}
Morph_pa_wide_y<-transform(Morph_pa_wide, Year = substr(sample, 1, 2), Month = substr(sample, 3, 4),Station=substr(sample, 5,7))
```

recode factors
```{r}
Morph_pa_wide_y$Year <- recode(Morph_pa_wide_y$Year, 
                                 "20" = "2020",
                         "19"="2019",
                         "18"="2018")

Morph_pa_wide_y$Month <- recode(Morph_pa_wide_y$Month, 
                                 "04" = "APR",
                         "07"="JUL",
                         "09"="SEP")

Morph_pa_wide_y$Station <- recode(Morph_pa_wide_y$Station, 
                                 "P40" = "P402")
Morph_pa_wide_y$code<-paste(Morph_pa_wide_y$Station,Morph_pa_wide_y$Year,Morph_pa_wide_y$Month,sep="-")
```

add in evironmental data
```{r}
Morph_pa_wide_y <- merge(Env_subset,Morph_pa_wide_y, by.x = "code", 
             by.y = "code", all.x = TRUE, all.y = TRUE)

#remove rows with NA
Morph_pa_wide_y<-Morph_pa_wide_y[!is.na(Morph_pa_wide_y$"Acartia.longiremis"),]
Morph_pa_wide_y<-Morph_pa_wide_y[!is.na(Morph_pa_wide_y$"Date"),]
Morph_pa_wide_x<-Morph_pa_wide_y[ , 15:95]
```

## PERMANOVA
```{r}
dist<-vegdist(Morph_pa_wide_x, method='bray')
perm<-adonis2(dist~STATION_NO*Year.x*Month.x, data=Morph_pa_wide_y, permutations = 999, method="bray")
perm
```

```{r}
Morph_pa_wide_y$Station=as.factor(Morph_pa_wide_y$Station)
adonis.pair(vegdist(Morph_pa_wide_x),Morph_pa_wide_y$Station)
```

```{r}
NMDSmodel <- metaMDS(Morph_pa_wide_x, distance = "bray",trymax = 999)
NMDSmodel
```

get datascores
```{r}
data.scores<-scores(NMDSmodel,display="sites")
data.scores<- as.data.frame(data.scores)
```

add environmental variables back in
```{r}
colnames(Morph_pa_wide_y)
data.scores$Month = Morph_pa_wide_y$Month.x
data.scores$Year = Morph_pa_wide_y$Year.x
data.scores$Station = Morph_pa_wide_y$Station
data.scores$temp_deep = Morph_pa_wide_y$CTDTMP_DEG_C_ITS90_Deep
data.scores$temp_surface = Morph_pa_wide_y$CTDTMP_DEG_C_ITS90_Surface
data.scores$salinity_deep = Morph_pa_wide_y$CTDSAL_PSS78_Deep
data.scores$salinity_surface = Morph_pa_wide_y$CTDSAL_PSS78_Surface
data.scores$oxygen_deep = Morph_pa_wide_y$OXYGEN_MG_L_1_Deep
data.scores$oxygen_surface = Morph_pa_wide_y$OXYGEN_MG_L_1_Surface
data.scores$nitrate_deep = Morph_pa_wide_y$NITRATE_UMorph_L_Deep
data.scores$nitrate_surface = Morph_pa_wide_y$NITRATE_UMorph_L_Surface
```

plot datascores
```{r}
xx <- ggplot(data.scores, aes(x = NMDS1, y = NMDS2,colour = Station)) +geom_point(aes(colour = Station))+theme_bw()+theme(panel.grid.major = element_blank(),
                                                                                           panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))
xx
```
extract hulls
```{r}
P12 <- data.scores[data.scores$Station == "P12", ][chull(data.scores[data.scores$Station ==                                                           "P12", c("NMDS1", "NMDS2")]), ]  
P22 <- data.scores[data.scores$Station == "P22", ][chull(data.scores[data.scores$Station ==                                                           "P22", c("NMDS1", "NMDS2")]), ]  
P28 <- data.scores[data.scores$Station == "P28", ][chull(data.scores[data.scores$Station ==                                                           "P28", c("NMDS1", "NMDS2")]), ]
P38 <- data.scores[data.scores$Station == "P38", ][chull(data.scores[data.scores$Station ==                                                           "P38", c("NMDS1", "NMDS2")]), ]  
P4 <- data.scores[data.scores$Station == "P4", ][chull(data.scores[data.scores$Station ==                                                           "P4", c("NMDS1", "NMDS2")]), ]  
P402 <- data.scores[data.scores$Station == "P402", ][chull(data.scores[data.scores$Station ==                                                           "P402", c("NMDS1", "NMDS2")]), ]  
P8 <- data.scores[data.scores$Station == "P8", ][chull(data.scores[data.scores$Station ==                                                           "P8", c("NMDS1", "NMDS2")]), ]  
```

get hull data
```{r}
hull.data <- rbind(P12, P22, P28,P38,P4,P402,P8)  #combine grp.a and grp.b
hull.data
```


## environmental fitting
create environmetal fit table
```{r}
#check for correlation among environmental variables.
#remove nutrients other than nitrate because of autocorrelation

data.scores.cor<- data.scores[ -c(1:5,13) ]
cor_matrix <- cor(data.scores.cor)
corrplot(cor_matrix, method = "number", type = "lower", tl.pos = 'd')

Zoo.env <- data.frame(matrix(ncol = 0, nrow=53))
Zoo.env$temp_deep<-as.numeric(data.scores$temp_deep)
Zoo.env$temp_surface<-as.numeric(data.scores$temp_surface)
Zoo.env$salinity_deep<-as.numeric(data.scores$salinity_deep)
Zoo.env$salinity_surface<-as.numeric(data.scores$salinity_surface)
Zoo.env$oxygen_deep<-as.numeric(data.scores$oxygen_deep)
Zoo.env$oxygen_surface<-as.numeric(data.scores$oxygen_surface)

env <- envfit(NMDSmodel, Zoo.env, na.rm = TRUE,perm = 999)
env


env.scrs <- as.data.frame(scores(env, display = "vectors")) 
env.scrs <- cbind(env.scrs, Species = rownames(env.scrs), Pvalues = env$vectors$pvals, R_squared = env$vectors$r)
env.scrs <- subset(env.scrs, Pvalues < 0.05)
env.scrs <- subset(env.scrs, R_squared > 0.3)
env.scrs
```

## Species fitting
```{r}
vf <- envfit(NMDSmodel,  Morph_pa_wide_x, perm = 999)

spp.scrs <- as.data.frame(scores(vf, display = "vectors"))
spp.scrs <- cbind(spp.scrs, Species = rownames(spp.scrs))
####for ggplot
arrow_factor <- ordiArrowMul(vf)
spp.scrs <- as.data.frame(scores(vf, display = "vectors")) * arrow_factor
spp.scrs <- cbind(spp.scrs, Species = rownames(spp.scrs), Pvalues = vf$vectors$pvals, R_squared = vf$vectors$r)

# select significance similarly to `plot(vf, p.max = 0.01)`
spp.scrs <- subset(spp.scrs, Pvalues < 0.05)
spp.scrs <- subset(spp.scrs, R_squared > 0.3)
spp.scrs

```
```{r}
morph.data.scores<-data.scores
morph.data.scores$Station <- factor(morph.data.scores$Station, levels = c("P22", "P4", "P8","P28","P12","P402","P38"))
morph.data.scores$Station <- recode(morph.data.scores$Station, 
                                 "P22" = "Juan de Fuca",
                         "P4"="Whidbey Basin",
                         "P8"="HC Sill",
                         "P28"="Main Basin",
                         "P12"="Hood Canal 1",
                         "P402"="Hood Canal 2",
                         "P38"="South Sound")

hull.data$Station <- factor(hull.data$Station, levels = c("P22", "P4", "P8","P28","P12","P402","P38"))
hull.data$Station <- recode(hull.data$Station, 
                                 "P22" = "Juan de Fuca",
                         "P4"="Whidbey Basin",
                         "P8"="HC Sill",
                         "P28"="Main Basin",
                         "P12"="Hood Canal 1",
                         "P402"="Hood Canal 2",
                         "P38"="South Sound")
```


```{r,fig.width=7,fig.height=5, dpi=600}


Species_plot <- ggplot(morph.data.scores, aes(x = NMDS1, y = NMDS2)) +geom_point(aes(colour =  Station))+ scale_colour_manual(values=c("#F8766D","#C49A00","#53B400","#00C094","#00B6EB","#A69AFF","#FB61D7")) +
  scale_fill_manual(values=c("#F8766D","#C49A00","#53B400","#00C094","#00B6EB","#A69AFF","#FB61D7"))+theme_bw()+theme(panel.grid.major = element_blank(),
                                                                                           panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))+
  geom_segment(data = spp.scrs,size=0.4,
               aes(x = 0, xend = NMDS1, y = 0, yend = NMDS2),
               arrow = arrow(length = unit(0.2, "cm")), colour = "darkgrey")+
  geom_segment(data = env.scrs,size=0.4,
               aes(x = 0, xend = NMDS1, y = 0, yend = NMDS2),
               arrow = arrow(length = unit(0.2, "cm")), colour = "black")+ 
  geom_polygon(data=hull.data,aes(x=NMDS1,y=NMDS2,fill=Station,group=Station),alpha=0.20, size=0.1, linetype=1, colour="black") +
  annotate("text", label = "2D Stress: 0.2", x = 0.55, y = 0.67, size = 5, colour = "black")+
  geom_label_repel(data = spp.scrs, aes(x = NMDS1, y = NMDS2, label = Species),
                  size = 3, fontface="bold", fill="white", label.padding = unit(0.15, "lines"), box.padding = unit(0.16, "lines"), label.size = 0.05)
Species_plot
```

```{r,fig.width=7,fig.height=5, dpi=600}


Species_plot <- ggplot(morph.data.scores, aes(x = NMDS1, y = NMDS2)) +geom_point(aes(colour =  Station))+ scale_colour_manual(values=c("#F8766D","#C49A00","#53B400","#00C094","#00B6EB","#A69AFF","#FB61D7")) +
  scale_fill_manual(values=c("#F8766D","#C49A00","#53B400","#00C094","#00B6EB","#A69AFF","#FB61D7"))+theme_bw()+theme(panel.grid.major = element_blank(),
                                                                                           panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))+
  geom_segment(data = spp.scrs,size=0.4,
               aes(x = 0, xend = NMDS1, y = 0, yend = NMDS2),
               arrow = arrow(length = unit(0.2, "cm")), colour = "darkgrey")+
  geom_segment(data = env.scrs,size=0.4,
               aes(x = 0, xend = NMDS1, y = 0, yend = NMDS2),
               arrow = arrow(length = unit(0.2, "cm")), colour = "black")+ 
  geom_polygon(data=hull.data,aes(x=NMDS1,y=NMDS2,fill=Station,group=Station),alpha=0.20, size=0.1, linetype=1, colour="black") +
  annotate("text", label = "2D Stress: 0.2", x = 0.6, y = 0.67, size = 5, colour = "black")
Species_plot
```


save dendogram
```{r, eval=FALSE}
setwd("/Users/hailaschultz/Dropbox/Schultz_Dissertation/Data_Analysis/Schultz_dissertation-2/figures")
ggsave(filename = "PA_Morph_nmds.png", plot = Species_plot, width = 7, height = 5.5, device='png', dpi=700)
```