---
title: "17_total-zoop"
date: "2024-10-03"
output:
  html_document:
    toc: true
    toc_depth: 2
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,         # Display code chunks
  eval = TRUE,        # Evaluate code chunks
  warning = FALSE,     # Hide warnings
  message = FALSE,     # Hide messages
  fig.width = 10,       # Set plot width in inches
  fig.height = 6,      # Set plot height in inches
  fig.align = "center" # Align plots to the center
)
```

# Load Zooplankton Data
This dataset is from our WOAC zooplankton database and includes morphological abundance and biomass data for each species identified in each sample collected. Vertical and oblique tows are included and represent samples that are integrated through the water column.

Read data into R
```{r}
Zooplankton <- read.csv("/Users/hailaschultz/Dropbox/Schultz_Dissertation/Data_Analysis/Schultz_dissertation-2/data/WOAC_2014-2022_Species_Data_Keister_Lab.csv")
```
Read environmental data into R
```{r}
Environmental <- read.csv("/Users/hailaschultz/Dropbox/Schultz_Dissertation/Data_Analysis/Schultz_dissertation-2/data/WOAC_Chem_data_2014-2022_all-niskins.csv")
```

# Environmental Data Summary
recode months to match with zooplankton data
```{r}
unique(Environmental$Month)
Environmental$Month <- recode(Environmental$Month, 
                                 "Jul" = "JUL",
                         "Sep"="SEP",
                         "Oct"="SEP",
                         "Apr"="APR",
                         "May"="MAY",
                         "Nov"="NOV",
                         "Aug"="AUG",
                         "Jun"="JUL",
                         "OCT"="SEP")
#remove unnecessary sample dates
unique(Environmental$Date)
unique(Zooplankton$Date)
Environmental<-subset(Environmental,Date!="10/30/14")
Environmental<-subset(Environmental,Date!="10/31/14")
Environmental<-subset(Environmental,Date!="10/23/14")
Environmental<-subset(Environmental,Date!="10/29/14")
Environmental<-subset(Environmental,Date!="5/24/15")
Environmental<-subset(Environmental,Date!="11/18/15")
Environmental<-subset(Environmental,Date!="11/16/15")
Environmental<-subset(Environmental,Date!="11/17/15")
Environmental<-subset(Environmental,Date!="10/27/16")
Environmental<-subset(Environmental,Date!="5/3/17")
Environmental<-subset(Environmental,Date!="5/4/17")
Environmental<-subset(Environmental,Date!="10/18/17")
Environmental<-subset(Environmental,Date!="5/24/18")
Environmental<-subset(Environmental,Date!="10/18/18")
Environmental<-subset(Environmental,Date!="10/19/18")
Environmental<-subset(Environmental,Date!="5/23/19")
Environmental<-subset(Environmental,Date!="8/31/20")
Environmental<-subset(Environmental,Date!="10/1/21")
Environmental<-subset(Environmental,Date!="5/6/22")
Environmental<-subset(Environmental,Date!="10/14/22")
Environmental<-subset(Environmental,Date!="10/15/22")
Environmental<-subset(Environmental,Date!="9/29/21")
Environmental<-subset(Environmental,Date!="7/4/20")
Environmental<-subset(Environmental,Date!="9/30/20")
Environmental<-subset(Environmental,Date!="9/18/17")
```

Temp: summarize by depth category
```{r}
temp<-Environmental %>%
  group_by(Station,Date,Depth,Year,Month) %>%
  summarise(
    temp = mean(CTDTMP_DEG_C_ITS90))

#subset to surface samples
temp_surf<-subset(temp,Depth=="Surface")
#make unique column
temp_surf$code<-paste(temp_surf$Station, temp_surf$Year,temp_surf$Month,sep="-")
```


pH: remove missing rows and remove weird values
```{r}
pH<-Environmental %>% drop_na(pH)
pH <- subset(pH, pH != 2130.2)
```
summarize by depth category
```{r}
pH<-pH %>%
  group_by(Station,Date,Depth,Year,Month) %>%
  summarise(
    pH = mean(pH))
#subset to surface samples
pH_surf<-subset(pH,Depth=="Deep")
#make unique column
pH_surf$code<-paste(pH_surf$Station, pH_surf$Year,pH_surf$Month,sep="-")
```


O2: remove missing rows and remove weird values
```{r}
Oxygen<-Environmental %>% drop_na(O2.in.mg.l)
```

summarize by depth category
```{r}
Oxygen<-Oxygen %>%
  group_by(Station,Date,Depth,Year,Month) %>%
  summarise(
    Oxygen = mean(O2.in.mg.l))
#subset to surface samples
Oxygen_surf<-subset(Oxygen,Depth=="Deep")
#make unique column
Oxygen_surf$code<-paste(Oxygen_surf$Station, Oxygen_surf$Year,Oxygen_surf$Month,sep="-")
```

Merge environmental tables
```{r}
joined <- merge(temp_surf, pH_surf, by.x = "code", 
             by.y = "code", all.x = TRUE, all.y = TRUE)
joined <- merge(joined, Oxygen_surf, by.x = "code", 
             by.y = "code", all.x = TRUE, all.y = TRUE)
#remove excess columns
env_joined <- joined[ -c(8:12,14:18) ]
```

remove low quality zooplankton data
```{r}
unique(Zooplankton$Confidence)
Zoop_sub<-subset(Zooplankton,Confidence!="low- flow high")
Zoop_sub<-subset(Zoop_sub,Confidence!="LOW")
Zoop_sub<-subset(Zoop_sub,Confidence!="Low-low flow")
```


subset to stations sampled during WOAC cruises
```{r}
unique(Zoop_sub$Station)
Zoop_sub<-subset(Zoop_sub,Station!="P381")
Zoop_sub<-subset(Zoop_sub,Station!="P105")
Zoop_sub<-subset(Zoop_sub,Station!="P136")
Zoop_sub<-subset(Zoop_sub,Station!="P7")
Zoop_sub<-subset(Zoop_sub,Station!="P123")
Zoop_sub<-subset(Zoop_sub,Station!="P132")
unique(Zoop_sub$Station)
#recode p4
Zoop_sub$Station <- recode(Zoop_sub$Station, 
                                 "p4" = "P4")
#remove october and november samples
Zoop_sub<-subset(Zoop_sub,Sample.Month!="NOV")
Zoop_sub<-subset(Zoop_sub,Sample.Month!="OCT")
Zoop_sub<-subset(Zoop_sub,Sample.Month!="MAR")

#recode month names to match
unique(Zoop_sub$Sample.Month)
Zoop_sub$Month <- recode(Zoop_sub$Sample.Month, 
                                 "Jul" = "JUL",
                         "Sep"="SEP")
#exclude weird taxa
Zoop_sub<-subset(Zoop_sub,WOAC.Group!="Exclude")
```
get # per m2
```{r}
Zoop_sub$density_squared<-Zoop_sub$Density....m3.*Zoop_sub$Station.Depth..m.
```

## get total zoop abundance
add up multiple lines per station
```{r}
colnames(Zoop_sub)
#remove NA values in density
Zoop_sub<-Zoop_sub[!is.na(Zoop_sub$Density....m3.),]
Zoop_sub <- Zoop_sub %>%
  group_by(Sample.Code,Station,Basin,Sample.Year,Sample.Month) %>%
  summarise(
    zoop_density = sum(Density....m3.),
    zoop_density_squared=sum(density_squared))
```

```{r}
#make unique column
Zoop_sub$code<-paste(Zoop_sub$Station, Zoop_sub$Sample.Year,Zoop_sub$Sample.Month,sep="-")
unique(Zoop_sub$code)
```
# Combine env and zoop
```{r}
Zoop_env <- merge(env_joined,Zoop_sub, by.x = "code", 
             by.y = "code", all.x = TRUE, all.y = TRUE)
#remove excess columns
Zoop_env <- Zoop_env[ -c(11,13:14) ]
#remove rows with NA
Zoop_env<-Zoop_env[!is.na(Zoop_env$Year.x),]
#make year a factor
Zoop_env$Year.x<-as.factor(Zoop_env$Year.x)
```

distribution of zoop density
```{r}
colnames(Zoop_env)
ggplot(Zoop_env, aes(x=zoop_density_squared)) + geom_histogram()
ggplot(Zoop_env, aes(x=sqrt(zoop_density_squared))) + geom_histogram()
```


# scatterplot
```{r}
ggplot(Zoop_env, aes(x=temp, y=sqrt(zoop_density),color=Station.x)) + geom_point()+
    facet_wrap(~Month.x)+ 
  geom_smooth(method=lm, se=FALSE)+ylab("log(Total Zooplankton Density)")+xlab("Temperature")+ labs(color = "Station")
```
```{r}
model <- lm(sqrt(zoop_density) ~ pH, data = Zoop_env)

# Summary of the regression model
summary(model)

ggplot(Zoop_env, aes(x=pH, y=sqrt(zoop_density))) + geom_point()+ 
  geom_smooth(method=lm, se=FALSE)
```
```{r}
model <- lm(sqrt(zoop_density) ~ Oxygen, data = Zoop_env)

# Summary of the regression model
summary(model)
ggplot(Zoop_env, aes(x=Oxygen, y=sqrt(zoop_density))) + geom_point()+ 
  geom_smooth(method=lm, se=FALSE)
```


```{r}
ggplot(Zoop_env, aes(x=Oxygen, y=sqrt(zoop_density))) + geom_point()+ 
  geom_smooth(method=lm, se=FALSE)+theme_bw()+theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.line = element_line(colour = "black")) +
  annotate("text", 
           x = 7.15, y = 150, 
           label = "p-value = 0.13 ",
           hjust = 0, color = "black", size = 7)+xlab(expression(plain('Deep Oxygen') ~ plain('(mg/L)')))+ylab(expression(plain('sqrt (Total Zooplankton Abundance)') ~ plain('(#/' ~ m^2 ~ ')')))+
  theme(
    axis.text = element_text(size = 20), 
    axis.title = element_text(size = 20),  
    plot.margin = margin(20, 20, 20, 20))
```

```{r}
ggplot(Zoop_env, aes(x=temp, y=zoop_density, color=Station.x)) + geom_point()+
    facet_wrap(~Month.x)+ 
  geom_smooth(method=lm, se=FALSE)
```

```{r}
ggplot(Zoop_env, aes(x=temp, y=zoop_density, color=as.factor(Year.x))) + geom_point()+
    facet_wrap(~Month.x)
```

```{r}
ggplot(Zoop_env, aes(x=temp, y=log(zoop_density), color=Month.x)) + geom_point()+
    facet_wrap(~Station.x)+ 
  geom_smooth(method=lm, se=FALSE)
```

#boxplot
```{r}
ggplot(Zoop_env, aes(x=as.factor(Year.x), y=sqrt(zoop_density))) + 
    geom_boxplot(alpha=0.2) +
    facet_wrap(~Month.x)+ylab("sqrt(Total Zooplankton Density)")+xlab("Year")  +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),panel.background = element_blank(), axis.line = element_line(colour = "black"),panel.border = element_rect(colour = "black", fill=NA))
```
```{r}
ggplot(Zoop_env, aes(x=as.factor(Station.x), y=sqrt(zoop_density_squared))) + 
    geom_boxplot(alpha=0.2) +
    facet_wrap(~Month.x)+ylab("sqrt(Number of zooplankton/m2)")+xlab("Station")+
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),panel.background = element_blank(), axis.line = element_line(colour = "black"),panel.border = element_rect(colour = "black", fill=NA))
```

untransformed model
```{r}
Mod1 <- aov(zoop_density_squared ~ Station.x+Year.x+Month.x+Station.x:Year.x+Station.x:Month.x+Month.x:Year.x, data = Zoop_env)

summary(Mod1)

TukeyHSD(Mod1, which = "Station.x")
```
Station, month and year are significant, significant interactions with station and year

check normality assumptions
```{r}
ggplot(Zoop_env, aes(x=zoop_density_squared)) + geom_histogram()
qqnorm(Mod1$residuals)
qqline(Mod1$residuals)
shapiro.test(Mod1$residuals)
```
Histogram is right skewed, qqplot looks weird, and shapiro test is significant, need to transform

log transformation
```{r}
Mod2 <- aov(log(zoop_density) ~ Station.x+Year.x+Month.x+Station.x:Year.x+Station.x:Month.x+Month.x:Year.x, data = Zoop_env)

summary(Mod2)
ggplot(Zoop_env, aes(x=log(zoop_density))) + geom_histogram()
qqnorm(Mod2$residuals)
qqline(Mod2$residuals)
shapiro.test(Mod2$residuals)
TukeyHSD(Mod2, which = "Station.x")
```
histogram looks better, but shapiro wilk is still significang

square root
```{r}
Mod3 <- aov(sqrt(zoop_density_squared) ~ Station.x+Year.x+Month.x+Station.x:Year.x+Station.x:Month.x+Month.x:Year.x, data = Zoop_env)

summary(Mod3)
ggplot(Zoop_env, aes(x=sqrt(zoop_density))) + geom_histogram()
qqnorm(Mod3$residuals)
qqline(Mod3$residuals)
shapiro.test(Mod3$residuals)
TukeyHSD(Mod3, which = "Station.x")
TukeyHSD(Mod3, which = "Month.x")
TukeyHSD(Mod3, which = "Year.x")
```
square root looks best
- station, month, and year are significant
- no significant interactions
```{r}
Zoop_env$Station.x <- factor(Zoop_env$Station.x, levels = c("P22", "P8", "P12","P402","P28","P38","P4"))
Station_Total_zoop<-ggplot(Zoop_env, aes(x=as.factor(Station.x), y=sqrt(zoop_density_squared),fill=Station.x,color=Station.x)) + 
    geom_boxplot(alpha=0.2) +ylab(expression(plain('sqrt (Total Zooplankton Abundance)') ~ plain('(#/' ~ m^2 ~ ')')))+xlab("Station")+
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),panel.background = element_blank(), axis.line = element_line(colour = "black"),panel.border = element_rect(colour = "black", fill=NA)) +
  scale_fill_manual(values = c("P22"="#F8766D", "P8"="#53B400", "P12"="#00B6EB","P402"="#A69AFF","P28"="#00C094","P38"="#FB61D7","P4"="#C49A00"))+
scale_color_manual(values = c("P22"="#F8766D", "P8"="#53B400", "P12"="#00B6EB","P402"="#A69AFF","P28"="#00C094","P38"="#FB61D7","P4"="#C49A00"))+
  scale_x_discrete(labels = c("P22"="Strait of Juan de Fuca", "P8"="Hood Canal Sill", "P12"="Hood Canal 1","P402"="Hood Canal 2","P28"="Main Basin","P38"="South Sound","P4"="Whidbey Basin"))+
  theme(axis.text.y=element_text(size=15,colour="black"),
        legend.position="none",
        legend.text=element_text(size=15,colour="black"),
        legend.title=element_text(size=15,colour="black"),
        axis.title.y=element_text(size=15,colour="black"),
        axis.line = element_line(colour = "black"),
        axis.title.x=element_blank(),
        axis.text.x = element_text(angle = 90, hjust = 1,size=15,colour="black"))+
  geom_vline(xintercept = 4.5, linetype = "dashed", color = "black")+geom_vline(xintercept = 6.5, linetype = "dashed", color = "black")
Station_Total_zoop
```
save plot
```{r, eval=FALSE}
setwd("/Users/hailaschultz/Dropbox/Schultz_Dissertation/Data_Analysis/Schultz_dissertation-2/figures")
ggsave(filename = "Station_Total_zoop.png", plot = Station_Total_zoop, width = 200, height = 164, units="mm", device='png', dpi=600)

```

```{r}
ggplot(Zoop_env, aes(x=pH, y=sqrt(zoop_density_squared),color=Station.x)) + geom_point()+
    facet_wrap(~Month.x)+ 
  geom_smooth(method=lm, se=FALSE)+ylab("sqrt(Total Zooplankton Density)")+xlab("Temperature")+ labs(color = "Station")
```
```{r}
ggplot(Zoop_env, aes(x=Oxygen, y=sqrt(zoop_density_squared),color=Station.x)) + geom_point()+ 
  geom_smooth(method=lm, se=FALSE)+ylab("sqrt(Total Zooplankton Density)")+xlab("Oxygen")+ labs(color = "Station")+
    facet_wrap(~Month.x)
```



```{r}
ggplot(Zoop_env, aes(x=as.factor(Year.x), y=sqrt(zoop_density_squared))) + 
    geom_boxplot(alpha=0.2) +ylab("sqrt(Number of zooplankton/m2)")+xlab("Station")+
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),panel.background = element_blank(), axis.line = element_line(colour = "black"),panel.border = element_rect(colour = "black", fill=NA))
```


### April anova
```{r}
Zoop_env_APR<-subset(Zoop_env,Month.x=="APR")
```

log is best transformation
```{r}

Mod6 <- aov(log10(zoop_density) ~ Station.x+Year.x+Station.x:Year.x, data = Zoop_env_APR)

summary(Mod6)
ggplot(Zoop_env_APR, aes(x=log10(zoop_density))) + geom_histogram()
qqnorm(Mod6$residuals)
qqline(Mod6$residuals)
shapiro.test(Mod6$residuals)
TukeyHSD(Mod6, which = "Year.x")
TukeyHSD(Mod6, which = "Station.x")
```
Year and Station are significant

### July anova
```{r}
Zoop_env_JUL<-subset(Zoop_env,Month.x=="JUL")
```

square root transformed is best
```{r}
Mod6 <- aov(sqrt(zoop_density) ~ Station.x+Year.x+Station.x:Year.x, data = Zoop_env_JUL)

summary(Mod6)
ggplot(Zoop_env_JUL, aes(x=sqrt(zoop_density))) + geom_histogram()
qqnorm(Mod6$residuals)
qqline(Mod6$residuals)
shapiro.test(Mod6$residuals)
TukeyHSD(Mod6, which = "Year.x")
TukeyHSD(Mod6, which = "Station.x")
```
Year was significant 

### September anova
```{r}
Zoop_env_SEP<-subset(Zoop_env,Month.x=="SEP")
```

square root transformed is best
```{r}
Mod6 <- aov(sqrt(zoop_density) ~ Station.x+Year.x+Station.x:Year.x, data = Zoop_env_SEP)

summary(Mod6)
ggplot(Zoop_env_SEP, aes(x=sqrt(zoop_density))) + geom_histogram()
qqnorm(Mod6$residuals)
qqline(Mod6$residuals)
shapiro.test(Mod6$residuals)
TukeyHSD(Mod6, which = "Year.x")
```
square root is best, no year and station are not significant

```{r}
ggplot(Zoop_env, aes(x=as.factor(Station.x), y=log(zoop_density))) + 
    geom_boxplot(alpha=0.2) + facet_wrap(~Month.x)+ylab("log(Total Zooplankton Density)")+xlab("Station")
```

```{r}
ggplot(Zoop_env, aes(x=pH, y=log(zoop_density), color=Station.x)) + geom_point()+
    facet_wrap(~Month.x)+ 
  geom_smooth(method=lm, se=FALSE)
```

oxygen

```{r}
Zoop_env$Station.x <- factor(Zoop_env$Station.x, levels = c("P22", "P8", "P12","P402","P28","P38","P4"))
Station_Total_zoop<-ggplot(Zoop_env, aes(x=as.factor(Station.x), y=Oxygen,fill=Station.x,color=Station.x)) + 
    geom_boxplot(alpha=0.2) +ylab(expression(plain('Deep Oxygen') ~ plain('(mg/L)')))+xlab("Station")+
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),panel.background = element_blank(), axis.line = element_line(colour = "black"),panel.border = element_rect(colour = "black", fill=NA)) +
  scale_fill_manual(values = c("P22"="#F8766D", "P8"="#53B400", "P12"="#00B6EB","P402"="#A69AFF","P28"="#00C094","P38"="#FB61D7","P4"="#C49A00"))+
scale_color_manual(values = c("P22"="#F8766D", "P8"="#53B400", "P12"="#00B6EB","P402"="#A69AFF","P28"="#00C094","P38"="#FB61D7","P4"="#C49A00"))+
  scale_x_discrete(labels = c("P22"="Strait of Juan de Fuca", "P8"="Hood Canal Sill", "P12"="Hood Canal 1","P402"="Hood Canal 2","P28"="Main Basin","P38"="South Sound","P4"="Whidbey Basin"))+
  theme(axis.text.y=element_text(size=15,colour="black"),
        legend.position="none",
        axis.title.y=element_text(size=30,colour="black"),
        axis.line = element_line(colour = "black"),
        axis.title.x=element_blank(),
        axis.text.x = element_text(angle = 90, hjust = 1,size=20,colour="black"))+
  geom_vline(xintercept = 4.5, linetype = "dashed", color = "black")+geom_vline(xintercept = 6.5, linetype = "dashed", color = "black")
Station_Total_zoop
```
pH

```{r}
Zoop_env$Station.x <- factor(Zoop_env$Station.x, levels = c("P22", "P8", "P12","P402","P28","P38","P4"))
Station_Total_zoop<-ggplot(Zoop_env, aes(x=as.factor(Station.x), y=pH,fill=Station.x,color=Station.x)) + 
    geom_boxplot(alpha=0.2) +ylab("Deep pH")+xlab("Station")+
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),panel.background = element_blank(), axis.line = element_line(colour = "black"),panel.border = element_rect(colour = "black", fill=NA)) +
  scale_fill_manual(values = c("P22"="#F8766D", "P8"="#53B400", "P12"="#00B6EB","P402"="#A69AFF","P28"="#00C094","P38"="#FB61D7","P4"="#C49A00"))+
scale_color_manual(values = c("P22"="#F8766D", "P8"="#53B400", "P12"="#00B6EB","P402"="#A69AFF","P28"="#00C094","P38"="#FB61D7","P4"="#C49A00"))+
  scale_x_discrete(labels = c("P22"="Strait of Juan de Fuca", "P8"="Hood Canal Sill", "P12"="Hood Canal 1","P402"="Hood Canal 2","P28"="Main Basin","P38"="South Sound","P4"="Whidbey Basin"))+
  theme(axis.text.y=element_text(size=15,colour="black"),
        legend.position="none",
        axis.title.y=element_text(size=30,colour="black"),
        axis.line = element_line(colour = "black"),
        axis.title.x=element_blank(),
        axis.text.x = element_text(angle = 90, hjust = 1,size=20,colour="black"))+
  geom_vline(xintercept = 4.5, linetype = "dashed", color = "black")+geom_vline(xintercept = 6.5, linetype = "dashed", color = "black")
Station_Total_zoop
```


# GLM
## Examine Data
Histograms of Jellyfish Density and Copepod Density
```{r}
colnames(Zoop_env)
```

```{r}
hist(Zoop_env$zoop_density) 
hist(Zoop_env$temp) 
hist(Zoop_env$pH) 
hist(Zoop_env$Oxygen)
```
environmental variables are relatively normally distributed, but zoop density has log distribution

##investigate effect of year
```{r}
Zoop_env$Year.x<-as.factor(Zoop_env$Year.x)
Year<-glm(zoop_density~Year.x,data=Zoop_env,family = Gamma(link="log"))
summary(Year)
plot(sqrt(zoop_density)~Year.x,data=Zoop_env)
```

year seems to be significant


##investigate effect of Month
```{r}
Zoop_env$Month.x<-as.factor(Zoop_env$Month.x)
Month<-glm(zoop_density~Month.x,data=Zoop_env,family = Gamma(link="log"))
summary(Month)
plot(zoop_density~Month.x,data=Zoop_env)
```

month seems to be significant

##investigate effect of Station
```{r}
Zoop_env$Station.x<-as.factor(Zoop_env$Station.x)
Station<-glm(zoop_density~Station.x,data=Zoop_env,family = Gamma(link="log"))
summary(Station)
plot(sqrt(zoop_density)~Station.x,data=Zoop_env)
```
## Family Options
### Negative Binomial Distribution
```{r}
library(MASS)
#round to integers
Zoop_env$Zoop_density_round<-round(as.numeric(Zoop_env$zoop_density), 0)
#run model
Zoop_env_model1 <- glm.nb(Zoop_density_round~temp+pH+Oxygen, data = Zoop_env)
summary(Zoop_env_model1)
plot(Zoop_env_model1)
```

### Gaussian Distribution
```{r}
Zoop_env_model2 <- glm(Zoop_density_round~temp+pH+Oxygen, data = Zoop_env,family="gaussian")
summary(Zoop_env_model2)
plot(Zoop_env_model2)
```

worse than NB!

### Gamma distribution
```{r}
Zoop_env <- Zoop_env[complete.cases(Zoop_env), ]
Zoop_env_model3 <- glm(zoop_density~temp+pH+Oxygen, data = Zoop_env,family="Gamma"(link="log"))
summary(Zoop_env_model3)
plot(Zoop_env_model3)
```

### NB and gamma with log link are similar, choose NB

check for multicollinearity
```{r}
Data1 = subset(Zoop_env, select = c(temp,pH,Oxygen,zoop_density) )
pairs(Data1 , panel = panel.smooth)
```
oxygen and temp have slight relationship, remove oxygen

```{r}
Zoop_env_model4 <- glm.nb(Zoop_density_round~temp+pH+Year.x+Month.x+Station.x, data = Zoop_env,
            na.action = "na.fail")
summary(Zoop_env_model4)
plot(Zoop_env_model4)
```
temp not significant

get best model
```{r}
library(MuMIn)
allperms <- dredge(Zoop_env_model4)
mod.table1 <- data.frame(allperms)
mod.table1
```

station as random effect
```{r}
Zoop_env_model4 <- glmer.nb(Zoop_density_round~pH+Year.x+Month.x+(1|Station.x), data = Zoop_env,
            na.action = "na.fail")
summary(Zoop_env_model4)
plot(Zoop_env_model4)
```
get best model
```{r}
library(MuMIn)
allperms <- dredge(Zoop_env_model4)
mod.table1 <- data.frame(allperms)
mod.table1
```
best model: pH, month, and year are in best model

best model
```{r}
Zoop_env_model5<-glmer.nb(Zoop_density_round~pH+Month.x+Year.x+(1|Station.x), data = Zoop_env,
            na.action = "na.fail")
summary(Zoop_env_model5)
plot(Zoop_env_model5)
```
station and month as random effect (no year)
```{r}
Zoop_env_model6 <- glmer.nb(Zoop_density_round~temp+pH+(1|Month.x)+(1|Station.x), data = Zoop_env,
            na.action = "na.fail")
summary(Zoop_env_model6)
plot(Zoop_env_model6)
```

