---
title: "19_heatmaps"
date: "2024-10-22"
output:
  html_document:
    toc: true
    toc_depth: 2
    toc_float: true
editor_options: 
  chunk_output_type: console
---
```{r}
library(tidyr)
library(data.table)
library(vegan)
library(EcolUtils)
library(ggplot2)
library(corrplot)
library(ggrepel)
library(indicspecies)
library(dplyr)
library(reshape2)
library(lubridate)
```

# Zooplankton data
Read data into R
```{r}
Zooplankton <- read.csv("/Users/hailaschultz/Dropbox/Schultz_Dissertation/Data_Analysis/Schultz_dissertation-2/data/WOAC_2014-2022_Species_Data_Keister_Lab.csv")
```
remove low quality zooplankton data
```{r}
unique(Zooplankton$Confidence)
Zoop_sub<-subset(Zooplankton,Confidence!="low- flow high")
Zoop_sub<-subset(Zoop_sub,Confidence!="LOW")
Zoop_sub<-subset(Zoop_sub,Confidence!="Low-low flow")
```


subset to stations sampled during WOAC cruises
```{r}
unique(Zoop_sub$Station)
Zoop_sub<-subset(Zoop_sub,Station!="P381")
Zoop_sub<-subset(Zoop_sub,Station!="P105")
Zoop_sub<-subset(Zoop_sub,Station!="P136")
Zoop_sub<-subset(Zoop_sub,Station!="P7")
Zoop_sub<-subset(Zoop_sub,Station!="P123")
Zoop_sub<-subset(Zoop_sub,Station!="P132")
unique(Zoop_sub$Station)
#recode p4
Zoop_sub$Station <- recode(Zoop_sub$Station, 
                                 "p4" = "P4")
#remove october and november samples
Zoop_sub<-subset(Zoop_sub,Sample.Month!="NOV")
Zoop_sub<-subset(Zoop_sub,Sample.Month!="OCT")
Zoop_sub<-subset(Zoop_sub,Sample.Month!="MAR")

#recode month names to match
unique(Zoop_sub$Sample.Month)
Zoop_sub$Month <- recode(Zoop_sub$Sample.Month, 
                                 "Jul" = "JUL",
                         "Sep"="SEP")
#remove oblique samples
Zoop_sub<-subset(Zoop_sub,Tow.Type!="Oblique")
```


recode species
```{r}
Zoop_sub$Genus.species <- recode(Zoop_sub$Genus.species, 
                                 "Cancridae Lg" = "Cancridae",
                         "Cancridae Sm"="Cancridae",
                         "PSEUDOCALANUS Lg"="PSEUDOCALANUS",
                         "PSEUDOCALANUS Sm"="PSEUDOCALANUS",
                         "Aglantha digitale"="AGLANTHA DIGITALE",
                         "Clytia gregaria"="CLYTIA GREGARIA",
                         "Muggiaea atlantica"="MUGGIAEA ATLANTICA",
                         "Pleurobrachia bachei"="PLEUROBRACHIA BACHEI",
                       "AEQUOREA Victoria"="AEQUOREA VICTORIA")
#exclude weird taxa
Zoop_sub<-subset(Zoop_sub,WOAC.Group!="Exclude")
```

change date format
```{r}
#manually change weird dates
unique(Zoop_sub$Sample.Date)
Zoop_sub$Sample.Date <- recode(Zoop_sub$Sample.Date, 
                                 "07/12/2021" = "7/12/21",
                         "07/13/2021"="7/13/21",
                         "07/14/2021"="7/14/21",
                         "07/16/2021"="7/16/21",
                        "06/27/2022"="6/27/22",
                        "06/28/2022"="6/28/22",
                        "06/29/2022"="6/29/22")
# convert date format
Zoop_sub$Date.Convert <- mdy(Zoop_sub$Sample.Date)
```



# stacked barplot
sum by taxonomic group
```{r}
Zoop_group <- Zoop_sub %>%
  group_by(Sample.Code,Station,Basin,Sample.Year,Date.Convert,Sample.Month,Crustaceans.vs.Gelatinous,Broad.Group) %>%
  summarise(
    zoop_density = sum(Density....m3.))
```

combing year, date,station
```{r}
Zoop_group$Sample<-paste(Zoop_group$Sample.Month, sep="_",Zoop_group$Sample.Year,Zoop_group$Station)
```


stacked barplot
```{r}
ggplot(Zoop_group, aes(fill = Crustaceans.vs.Gelatinous, y = zoop_density, x = reorder(Sample, Date.Convert))) + 
  geom_bar(position = "stack", stat = "identity") + 
  theme(
    axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1),  # Adjusts x-axis text
    axis.title.x = element_text(margin = margin(t = 10))  # Adds margin to x-axis title
  ) + 
  facet_wrap(~ Station, scales = "free")
```

proportions
```{r}
ggplot(Zoop_group, aes(fill = Crustaceans.vs.Gelatinous, y = zoop_density, x = reorder(Sample, Date.Convert))) + 
  geom_bar(position = "fill", stat = "identity") + 
  theme(
    axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1),  # Adjusts x-axis text
    axis.title.x = element_text(margin = margin(t = 10))  # Adds margin to x-axis title
  ) + 
  facet_wrap(~ Station, scales = "free")
```
stacked barplot
```{r}
ggplot(Zoop_group, aes(fill = Broad.Group, y = zoop_density, x = reorder(Sample, Date.Convert))) + 
  geom_bar(position = "stack", stat = "identity") + 
  theme(
    axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1),  # Adjusts x-axis text
    axis.title.x = element_text(margin = margin(t = 10))  # Adds margin to x-axis title
  ) 
```

aggregated
```{r}
library(RColorBrewer)
nb.cols <- 36
mycolors <- colorRampPalette(brewer.pal(8, "Set2"))(nb.cols)
ggplot(Zoop_group, aes(fill = Broad.Group, y = zoop_density, x = reorder(Sample, Date.Convert))) + 
  geom_bar(position = "identity", stat = "identity") + 
  theme(
    axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1),  # Adjusts x-axis text
    axis.title.x = element_text(margin = margin(t = 10))  # Adds margin to x-axis title
  ) +
  scale_fill_manual(values = mycolors)
```

# Species Stacked Barplot

subset to copepods
```{r}
Zoop_copepod<-subset(Zoop_sub,Broad.Group=="Copepod")
```

subset to april
```{r}
Zoop_copepod<-subset(Zoop_copepod,Sample.Month=="APR")
```

```{r}
Zoop_copepod$Sample<-paste(Zoop_copepod$Sample.Month, sep="_",Zoop_copepod$Sample.Year,Zoop_copepod$Station)
```

sum by Genus.Species
```{r}
Zoop_copepod <- Zoop_copepod %>%
  group_by(Sample.Code,Sample,Station,Basin,Sample.Year,Date.Convert,Sample.Month,Crustaceans.vs.Gelatinous,Broad.Group,Genus.species) %>%
  summarise(
    zoop_density = sum(Density....m3.))
```

stacked barplot
```{r}
ggplot(Zoop_copepod, aes(fill = Genus.species, y = zoop_density, x = reorder(Sample, Date.Convert))) + 
  geom_bar(position = "stack", stat = "identity") + 
  theme(
    axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1),  # Adjusts x-axis text
    axis.title.x = element_text(margin = margin(t = 10))  # Adds margin to x-axis title
  ) + 
  facet_wrap(~ Station, scales = "free")
```

# Gelatinous Stacked Barplot

```{r}
Zoop_gelatinous<-subset(Zoop_sub,Crustaceans.vs.Gelatinous=="Gelatinous")
Zoop_gelatinous<-subset(Zoop_gelatinous,Genus.species!="PANDEIDAE")
```

subset to month
```{r}
Zoop_gelatinous<-subset(Zoop_gelatinous,Sample.Month=="APR")
```

```{r}
Zoop_gelatinous$Sample<-paste(Zoop_gelatinous$Sample.Month, sep="_",Zoop_gelatinous$Sample.Year,Zoop_gelatinous$Station)
```

sum by Genus.Species
```{r}
Zoop_gelatinous <- Zoop_gelatinous %>%
  group_by(Sample.Code,Station,Sample,Basin,Sample.Year,Date.Convert,Sample.Month,Crustaceans.vs.Gelatinous,Broad.Group,Genus.species) %>%
  summarise(
    zoop_density = sum(Density....m3.))
```

stacked barplot
```{r}
ggplot(Zoop_gelatinous, aes(fill = Genus.species, y = zoop_density, x = reorder(Sample, Date.Convert))) + 
  geom_bar(position = "stack", stat = "identity") + 
  theme(
    axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1),  # Adjusts x-axis text
    axis.title.x = element_text(margin = margin(t = 10))  # Adds margin to x-axis title
  ) + 
  facet_wrap(~ Station, scales = "free")
```

# Crustacean Stacked Barplot

subset to copepods
```{r}
Zoop_crustacean<-subset(Zoop_sub,Crustaceans.vs.Gelatinous=="Crustaceans")
```

subset to april
```{r}
Zoop_crustacean<-subset(Zoop_crustacean,Month=="APR")
```


sum by Genus.Species
```{r}
Zoop_crustacean <- Zoop_crustacean %>%
  group_by(Code,Station,Basin,Year,Date.Convert,Month,Crustaceans.vs.Gelatinous,Broad.Group,Genus.species) %>%
  summarise(
    zoop_density = sum(Density....m3.))
```

stacked barplot
```{r}
ggplot(Zoop_crustacean, aes(fill = Broad.Group, y = zoop_density, x = reorder(Code, Date.Convert))) + 
  geom_bar(position = "stack", stat = "identity") + 
  theme(
    axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1),  # Adjusts x-axis text
    axis.title.x = element_text(margin = margin(t = 10))  # Adds margin to x-axis title
  ) + 
  facet_wrap(~ Station, scales = "free")
```

# Mollusc Stacked Plot

subset to molluscs
```{r}
Zoop_mollusc<-subset(Zoop_sub,Crustaceans.vs.Gelatinous=="Molluscs")
```

combing year, date,station
```{r}
Zoop_mollusc$Sample<-paste(Zoop_mollusc$Sample.Month, sep="_",Zoop_mollusc$Sample.Year,Zoop_mollusc$Station)
```
sum by Genus.Species
```{r}
Zoop_mollusc <- Zoop_mollusc %>%
  group_by(Sample.Code,Station,Sample,Basin,Sample.Year,Date.Convert,Sample.Month,Crustaceans.vs.Gelatinous,Broad.Group,Genus.species) %>%
  summarise(
    zoop_density = sum(Density....m3.))
```



stacked barplot
```{r}
ggplot(Zoop_mollusc, aes(fill = Genus.species, y = zoop_density, x = reorder(Sample, Date.Convert))) + 
  geom_bar(position = "stack", stat = "identity") + 
  theme(
    axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1),  # Adjusts x-axis text
    axis.title.x = element_text(margin = margin(t = 10))  # Adds margin to x-axis title
  ) + 
  facet_wrap(~ Station, scales = "free")
```

stacked barplot aggregated
```{r}
Zoop_mollusc$Month_Yr <- format(as.Date(Zoop_mollusc$Date.Convert), "%Y-%m")
ggplot(Zoop_mollusc, aes(fill = Broad.Group, y = zoop_density, x = Month_Yr)) + 
  geom_bar(position = "stack", stat = "identity") + 
  theme(
    axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1),  # Adjusts x-axis text
    axis.title.x = element_text(margin = margin(t = 10))  # Adds margin to x-axis title
  ) + 
  facet_wrap(~ Sample.Month, scales = "free")
```
# heatmaps
add up multiple lines per station
```{r}
Zoop_sub <- Zoop_sub %>%
  group_by(Code,Station,Basin,Year,Month,Crustaceans.vs.Gelatinous,Broad.Group,Genus.species) %>%
  summarise(
    zoop_density = sum(Density....m3.))
```

combine larger group and species
```{r}
Zoop_sub$taxa<-paste(Zoop_sub$Crustaceans.vs.Gelatinous,Zoop_sub$Broad.Group,Zoop_sub$Genus.species,sep="_")
```

#make heatmaps
```{r}
# Get unique stations and months
stations <- unique(Zoop_sub$Station)
months <- unique(Zoop_sub$Month)

# Loop through each station and month
for (station in stations) {
  for (month in months) {
    # Subset data for the current station and month
    Zoop_subset <- subset(Zoop_sub, Month == month & Station == station)
    
    # Check if the subset is not empty
    if (nrow(Zoop_subset) > 0) {
      Zoop_sub_heat <- dcast(Zoop_subset, Code ~ taxa, value.var = "zoop_density")
      Zoop_sub_heat <- Zoop_sub_heat %>% replace(is.na(.), 0)

      # Convert to proportions
      Zoop_sub_heat_data <- subset(Zoop_sub_heat, select = -c(Code))

      Zoop_sub_heat_data1 <- as.matrix(Zoop_sub_heat_data)
      Zoop_sub_heat_data1 <- matrix(as.numeric(Zoop_sub_heat_data1), ncol = ncol(Zoop_sub_heat_data))
      rownames(Zoop_sub_heat_data1) <- Zoop_sub_heat$Code
      column <- colnames(Zoop_sub_heat_data)
      colnames(Zoop_sub_heat_data1) <- column

      # Check dimensions before creating heatmap
      if (nrow(Zoop_sub_heat_data1) >= 2 && ncol(Zoop_sub_heat_data1) >= 2) {
        # Specify the file name
        file_name <- paste0("heatmap_", station, "_", month, ".png")

        # Save heatmap as PNG with higher resolution
        png(file_name, width = 5000, height = 5000, res = 300)  # Adjust width and height as needed

        # Create heatmap
        heatmap(Zoop_sub_heat_data1, Colv = NA, show_row_names = TRUE, show_col_names = TRUE, mar = c(15, 15))
        

        dev.off()  # Close the device to save the file
      } else {
        message(paste("Skipping heatmap for", station, "in", month, "- not enough data."))
      }
    }
  }
}

```

#make normalized heatmaps
```{r}
library(dplyr)
library(reshape2)

# Get unique stations and months
stations <- unique(Zoop_sub$Station)
months <- unique(Zoop_sub$Month)

# Loop through each station and month
for (station in stations) {
  for (month in months) {
    # Subset data for the current station and month
    Zoop_subset <- subset(Zoop_sub, Month == month & Station == station)
    
    # Check if the subset is not empty
    if (nrow(Zoop_subset) > 0) {
      Zoop_sub_heat <- dcast(Zoop_subset, Code ~ taxa, value.var = "zoop_density")
      Zoop_sub_heat <- Zoop_sub_heat %>% replace(is.na(.), 0)

      # Convert to proportions
      Zoop_sub_heat_data <- subset(Zoop_sub_heat, select = -c(Code))

      Zoop_sub_heat_data1 <- as.matrix(Zoop_sub_heat_data)
      Zoop_sub_heat_data1 <- matrix(as.numeric(Zoop_sub_heat_data1), ncol = ncol(Zoop_sub_heat_data))
      rownames(Zoop_sub_heat_data1) <- Zoop_sub_heat$Code
      column <- colnames(Zoop_sub_heat_data)
      colnames(Zoop_sub_heat_data1) <- column

      # Check dimensions before creating heatmap
      if (nrow(Zoop_sub_heat_data1) >= 2 && ncol(Zoop_sub_heat_data1) >= 2) {
        # Specify the file name
        file_name <- paste0("normheatmap_", station, "_", month, ".png")

        # Save heatmap as PNG with higher resolution
        png(file_name, width = 5000, height = 5000, res = 300)  # Adjust width and height as needed

        # Create heatmap
        heatmap(Zoop_sub_heat_data1, Colv = NA, show_row_names = TRUE, show_col_names = TRUE, mar = c(15, 15))
        heatmap(Zoop_sub_heat_data1, scale = "column", show_row_names = TRUE, Colv = NA, show_col_names = TRUE, mar = c(15, 15))

        dev.off()  # Close the device to save the file
      } else {
        message(paste("Skipping heatmap for", station, "in", month, "- not enough data."))
      }
    }
  }
}

```


# Broad group heatmap
Subset to month and group
```{r}
Zoop_broad<-Zoop_sub
Zoop_broad<-subset(Zoop_broad,Month=="JUL")
Zoop_broad<-subset(Zoop_broad,Station=="P402")
```

add up multiple lines per station
```{r}
Zoop_broad[is.na(Zoop_broad)] <- 0
Zoop_broad <- Zoop_broad %>%
  group_by(Year,Crustaceans.vs.Gelatinous) %>%
  summarise(
    zoop_density = sum(Density....m3.))
```

wide format
```{r}
Zoop_broad_wide<-dcast(Zoop_broad, Year ~ Crustaceans.vs.Gelatinous )
```
make heatmap
```{r}
Zoop_broad_wide_col <-Zoop_broad_wide[, -c(1)]
Zoop_broad_wide1<- as.matrix(Zoop_broad_wide_col)
Zoop_broad_wide1 <- matrix(as.numeric(Zoop_broad_wide1), ncol =ncol(Zoop_broad_wide_col))
      rownames(Zoop_broad_wide1) <- Zoop_broad_wide$Year
      column <- colnames(Zoop_broad_wide_col)
      colnames(Zoop_broad_wide1) <- column
      
# Create heatmap
heatmap(Zoop_broad_wide1, Rowv=NA, Colv=NA,scale = "column",  show_row_names = TRUE, show_col_names = TRUE)

```



