---
title: "21_crustaceans_vs_gelatinous"
date: "2024-11-21"
output:
  html_document:
    toc: true
    toc_depth: 2
    toc_float: true
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,         # Display code chunks
  eval = TRUE,        # Evaluate code chunks
  warning = FALSE,     # Hide warnings
  message = FALSE,     # Hide messages
  fig.width = 10,       # Set plot width in inches
  fig.height = 6,      # Set plot height in inches
  fig.align = "center" # Align plots to the center
)
```

```{r}
library(lubridate)
```

# Load Zooplankton Data
This dataset is from our WOAC zooplankton database and includes morphological abundance and biomass data for each species identified in each sample collected. Vertical and oblique tows are included and represent samples that are integrated through the water column.

Read data into R
```{r}
Zooplankton <- read.csv("/Users/hailaschultz/Dropbox/Schultz_Dissertation/Data_Analysis/Schultz_dissertation-2/data/WOAC_2014-2022_Species_Data_Keister_Lab.csv")
```
Read environmental data into R
```{r}
Environmental <- read.csv("/Users/hailaschultz/Dropbox/Schultz_Dissertation/Data_Analysis/Schultz_dissertation-2/data/WOAC_Chem_data_2014-2022_all-niskins.csv")
```

# Environmental Data Summary
recode months to match with zooplankton data
```{r}
unique(Environmental$Month)
Environmental$Month <- recode(Environmental$Month, 
                                 "Jul" = "JUL",
                         "Sep"="SEP",
                         "Oct"="SEP",
                         "Apr"="APR",
                         "May"="MAY",
                         "Nov"="NOV",
                         "Aug"="AUG",
                         "Jun"="JUL",
                         "OCT"="SEP")
#remove unnecessary sample dates
unique(Environmental$Date)
unique(Zooplankton$Date)
Environmental<-subset(Environmental,Date!="10/30/14")
Environmental<-subset(Environmental,Date!="10/31/14")
Environmental<-subset(Environmental,Date!="10/23/14")
Environmental<-subset(Environmental,Date!="10/29/14")
Environmental<-subset(Environmental,Date!="5/24/15")
Environmental<-subset(Environmental,Date!="11/18/15")
Environmental<-subset(Environmental,Date!="11/16/15")
Environmental<-subset(Environmental,Date!="11/17/15")
Environmental<-subset(Environmental,Date!="10/27/16")
Environmental<-subset(Environmental,Date!="5/3/17")
Environmental<-subset(Environmental,Date!="5/4/17")
Environmental<-subset(Environmental,Date!="10/18/17")
Environmental<-subset(Environmental,Date!="5/24/18")
Environmental<-subset(Environmental,Date!="10/18/18")
Environmental<-subset(Environmental,Date!="10/19/18")
Environmental<-subset(Environmental,Date!="5/23/19")
Environmental<-subset(Environmental,Date!="8/31/20")
Environmental<-subset(Environmental,Date!="10/1/21")
Environmental<-subset(Environmental,Date!="5/6/22")
Environmental<-subset(Environmental,Date!="10/14/22")
Environmental<-subset(Environmental,Date!="10/15/22")
Environmental<-subset(Environmental,Date!="9/29/21")
Environmental<-subset(Environmental,Date!="7/4/20")
Environmental<-subset(Environmental,Date!="9/30/20")
Environmental<-subset(Environmental,Date!="9/18/17")
```

Temp: summarize by depth category
```{r}
temp<-Environmental %>%
  group_by(Station,Date,Depth,Year,Month) %>%
  summarise(
    temp = mean(CTDTMP_DEG_C_ITS90))

#subset to surface samples
temp_surf<-subset(temp,Depth=="Surface")
#make unique column
temp_surf$code<-paste(temp_surf$Station, temp_surf$Year,temp_surf$Month,sep="-")
```


pH: remove missing rows and remove weird values
```{r}
pH<-Environmental %>% drop_na(pH)
pH <- subset(pH, pH != 2130.2)
```
summarize by depth category
```{r}
pH<-pH %>%
  group_by(Station,Date,Depth,Year,Month) %>%
  summarise(
    pH = mean(pH))
#subset to surface samples
pH_surf<-subset(pH,Depth=="Deep")
#make unique column
pH_surf$code<-paste(pH_surf$Station, pH_surf$Year,pH_surf$Month,sep="-")
```


O2: remove missing rows and remove weird values
```{r}
Oxygen<-Environmental %>% drop_na(O2.in.mg.l)
```

summarize by depth category
```{r}
Oxygen<-Oxygen %>%
  group_by(Station,Date,Depth,Year,Month) %>%
  summarise(
    Oxygen = mean(O2.in.mg.l))
#subset to surface samples
Oxygen_surf<-subset(Oxygen,Depth=="Deep")
#make unique column
Oxygen_surf$code<-paste(Oxygen_surf$Station, Oxygen_surf$Year,Oxygen_surf$Month,sep="-")
```

Merge environmental tables
```{r}
joined <- merge(temp_surf, pH_surf, by.x = "code", 
             by.y = "code", all.x = TRUE, all.y = TRUE)
joined <- merge(joined, Oxygen_surf, by.x = "code", 
             by.y = "code", all.x = TRUE, all.y = TRUE)
#remove excess columns
env_joined <- joined[ -c(8:12,14:18) ]
```

remove low quality zooplankton data
```{r}
unique(Zooplankton$Confidence)
Zoop_sub<-subset(Zooplankton,Confidence!="low- flow high")
Zoop_sub<-subset(Zoop_sub,Confidence!="LOW")
Zoop_sub<-subset(Zoop_sub,Confidence!="Low-low flow")
```


subset to stations sampled during WOAC cruises
```{r}
unique(Zoop_sub$Station)
Zoop_sub<-subset(Zoop_sub,Station!="P381")
Zoop_sub<-subset(Zoop_sub,Station!="P105")
Zoop_sub<-subset(Zoop_sub,Station!="P136")
Zoop_sub<-subset(Zoop_sub,Station!="P7")
Zoop_sub<-subset(Zoop_sub,Station!="P123")
Zoop_sub<-subset(Zoop_sub,Station!="P132")
unique(Zoop_sub$Station)
#recode p4
Zoop_sub$Station <- recode(Zoop_sub$Station, 
                                 "p4" = "P4")
#remove october and november samples
Zoop_sub<-subset(Zoop_sub,Sample.Month!="NOV")
Zoop_sub<-subset(Zoop_sub,Sample.Month!="OCT")
Zoop_sub<-subset(Zoop_sub,Sample.Month!="MAR")

#recode month names to match
unique(Zoop_sub$Sample.Month)
Zoop_sub$Month <- recode(Zoop_sub$Sample.Month, 
                                 "Jul" = "JUL",
                         "Sep"="SEP")
#exclude weird taxa
Zoop_sub<-subset(Zoop_sub,WOAC.Group!="Exclude")
# convert date format
Zoop_sub$Date.Convert <- mdy(Zoop_sub$Sample.Date)
```
get # per m2
```{r}
Zoop_sub$density_squared<-Zoop_sub$Density....m3.*Zoop_sub$Station.Depth..m.
```
## get zoop abundance
add up multiple lines per station
```{r}
colnames(Zoop_sub)
#remove NA values in density
Zoop_sub<-Zoop_sub[!is.na(Zoop_sub$Density....m3.),]
Zoop_sub <- Zoop_sub %>%
  group_by(Sample.Code,Date.Convert,Station,Basin,Sample.Year,Sample.Month,Crustaceans.vs.Gelatinous) %>%
  summarise(
    zoop_density = sum(Density....m3.),
    zoop_density_squared=sum(density_squared))
```

```{r}
#make unique column
Zoop_sub$code<-paste(Zoop_sub$Station, Zoop_sub$Sample.Year,Zoop_sub$Sample.Month,sep="-")
unique(Zoop_sub$code)
```

# Combine env and zoop
```{r}
Zoop_env <- merge(env_joined,Zoop_sub, by.x = "code", 
             by.y = "code", all.x = TRUE, all.y = TRUE)
#remove excess columns
Zoop_env <- Zoop_env[ -c(12:15) ]
#remove rows with NA
Zoop_env<-Zoop_env[!is.na(Zoop_env$Year.x),]
#make year a factor
Zoop_env$Year.x<-as.factor(Zoop_env$Year.x)
```

```{r}
ggplot(Zoop_env, aes(x=as.factor(Year.x), y=log(zoop_density_squared))) + 
    geom_boxplot(alpha=0.2)+ylab("#/m2")+xlab("Station")+
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),panel.background = element_blank(), axis.line = element_line(colour = "black"),panel.border = element_rect(colour = "black", fill=NA))+facet_wrap(~Crustaceans.vs.Gelatinous, scale="free")
```

# Gelatinous to crustacean ratio
```{r}
CZ<-subset(Zoop_env, Zoop_env$Crustaceans.vs.Gelatinous %in% c("Crustaceans","Gelatinous"))
```

remove ethanol samples
```{r}
CZ<-CZ[- grep("EtOH", CZ$Sample.Code),]
```

```{r}
CZ_wide<- subset(CZ, select=-c(zoop_density_squared))
CZ_wide<-dcast(CZ_wide, code+Month.x+Year.x+Date.Convert+Station.x+temp+pH+Oxygen ~ Crustaceans.vs.Gelatinous )
```

```{r}
CZ_wide$Gelatinous[is.na(CZ_wide$Gelatinous)] <- 0
CZ_wide$GvsC<-CZ_wide$Gelatinous/CZ_wide$Crustaceans
```

```{r}
ggplot(CZ_wide, aes(x=GvsC)) + geom_histogram()
ggplot(CZ_wide, aes(x=temp)) + geom_histogram()
ggplot(CZ_wide, aes(x=log(GvsC))) + geom_histogram()
```

```{r}
CZ_wide_april<-subset(CZ_wide,Month.x=="SEP")
ggplot(CZ_wide_april, aes(x=temp, y=log(GvsC),color=Station.x))+ geom_point()+ 
  geom_smooth(method=lm, se=FALSE)
```
```{r}
CZ_wide$gvscp1<-CZ_wide$GvsC+0.00000001
model <- lm(gvscp1 ~ temp, data = CZ_wide)

# Summary of the regression model
summary(model)

r_squared <- summary(model)$r.squared
p_value <- summary(model)$coefficients[2, 4]
```

```{r}
ggplot(CZ_wide, aes(x=temp, y=log(GvsC)))+ geom_point()+ 
  geom_smooth(method=lm,se=FALSE)+theme_bw()+theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.line = element_line(colour = "black")) +
  annotate("text", 
           x = 18, y = -10, 
           label = paste0("R² = ", round(r_squared, 3), 
                          "\nP-value = ", signif(p_value, 3)),
           hjust = 0, color = "black", size = 7)+
  xlab("Temperature (°C)")+
  ylab("Gelatinous/Crustaceans")+
  theme(
    axis.text = element_text(size = 20), 
    axis.title = element_text(size = 20))
```




```{r}
ggplot(CZ_wide, aes(x=Oxygen, y=log(GvsC),color=Station.x))+ geom_point()+ 
  geom_smooth(method=lm, se=FALSE)
```

```{r}
ggplot(CZ_wide, aes(x=Oxygen, y=log(GvsC),color=Station.x))+ geom_point()+ 
  geom_smooth(method=lm,se=FALSE)
```

```{r}
ggplot(CZ_wide, aes(x=Oxygen, y=log(GvsC), color=Station.x))+
    facet_wrap(~Station.x, scales = "free") + geom_point()
```
```{r,fig.width=5,fig.height=5.5, dpi=600}
CZ_wide_april<-subset(CZ_wide,Month.x=="APR")
CZ_wide_april$Station.x <- factor(CZ_wide_april$Station.x, levels = c("P22", "P8", "P12","P402","P28","P38","P4"))
facet_labels<-c("P22"="Strait of Juan de Fuca", "P8"="Hood Canal Sill", "P12"="Hood Canal 1","P402"="Hood Canal 2","P28"="Main Basin","P38"="South Sound","P4"="Whidbey Basin")
ggplot(CZ_wide_april, aes(x=Date.Convert, y=GvsC,color=Station.x)) + 
    geom_point()+
    geom_line()+ylab("Gelatinous/Crustaceans")+xlab("Year")+
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),panel.background = element_blank(), axis.line = element_line(colour = "black"),panel.border = element_rect(colour = "black", fill=NA))+
    facet_wrap(~Station.x,ncol=2, scales = "free_y", labeller = as_labeller(facet_labels))+ scale_colour_manual(values=c("#F8766D","#53B400","#00B6EB","#A69AFF","#00C094","#FB61D7","#C49A00"))+
  scale_x_date(
    date_breaks = "1 year",   # Tick mark every year
    date_labels = "%Y"       # Format labels as "YYYY"
  ) +theme(axis.text.x = element_text(angle = 90, hjust = 1))
```
```{r,fig.width=4,fig.height=4, dpi=600}
ggplot(CZ_wide_april, aes(x=Date.Convert, y=temp,color=Station.x)) + 
    geom_point()+geom_line()+ylab("Temperature (C)")+xlab("Year")+
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),panel.background = element_blank(), axis.line = element_line(colour = "black"),panel.border = element_rect(colour = "black", fill=NA))+ scale_colour_manual(values=c("#F8766D","#53B400","#00B6EB","#A69AFF","#00C094","#FB61D7","#C49A00"))+
  scale_x_date(
    date_breaks = "1 year",   # Tick mark every year
    date_labels = "%Y"       # Format labels as "YYYY"
  )+theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

```{r}
ggplot(CZ_wide, aes(x=temp, y=GvsC, color=Month.x))+
    facet_wrap(~Station.x, scales = "free") + geom_point()
```
```{r}
ggplot(CZ_wide, aes(x=Oxygen, y=log(GvsC), color=Station.x))+
    facet_wrap(~Station.x, scales = "free") + geom_point()
```

```{r}
ggplot(CZ_wide, aes(x=as.factor(Date.Convert), y=GvsC, color=Station.x))+
    facet_wrap(~Station.x, scales = "free") + geom_bar(position = "stack", stat = "identity")+ 
  theme(
    axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1),  # Adjusts x-axis text
    axis.title.x = element_text(margin = margin(t = 10))  # Adds margin to x-axis title
  ) 
```


```{r}
CZ_wide$Station.x <- factor(CZ_wide$Station.x, levels = c("P22", "P8", "P12","P402","P28","P38","P4"))
ggplot(CZ_wide, aes(x=as.factor(Station.x), y=log(GvsC),color=Station.x,fill=Station.x)) + 
    geom_boxplot(alpha=0.2)+ylab("log(Gelatinous/Crustaceans)")+xlab("Station")+
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),panel.background = element_blank(), axis.line = element_line(colour = "black"),panel.border = element_rect(colour = "black", fill=NA))+
  scale_fill_manual(values = c("P22"="#F8766D", "P8"="#53B400", "P12"="#00B6EB","P402"="#A69AFF","P28"="#00C094","P38"="#FB61D7","P4"="#C49A00"))+
scale_color_manual(values = c("P22"="#F8766D", "P8"="#53B400", "P12"="#00B6EB","P402"="#A69AFF","P28"="#00C094","P38"="#FB61D7","P4"="#C49A00"))+
  scale_x_discrete(labels = c("P22"="Strait of Juan de Fuca", "P8"="Hood Canal Sill", "P12"="Hood Canal 1","P402"="Hood Canal 2","P28"="Main Basin","P38"="South Sound","P4"="Whidbey Basin"))+
  theme(axis.text.y=element_text(size=20,colour="black"),
        legend.position="none",
        axis.title.y=element_text(size=20,colour="black"),
        axis.line = element_line(colour = "black"),
        axis.title.x=element_blank(),
        axis.text.x = element_text(angle = 90, hjust = 1,size=15,colour="black"))+
  geom_vline(xintercept = 4.5, linetype = "dashed", color = "black")+geom_vline(xintercept = 6.5, linetype = "dashed", color = "black")
```
```{r}

Mod6 <- aov(log(gvscp1) ~ Station.x, data = CZ_wide)

summary(Mod6)
qqnorm(Mod6$residuals)
qqline(Mod6$residuals)
shapiro.test(Mod6$residuals)
TukeyHSD(Mod6, which = "Station.x")
```
```{r}

Mod6 <- aov(temp ~ Station.x, data = CZ_wide)

summary(Mod6)
qqnorm(Mod6$residuals)
qqline(Mod6$residuals)
shapiro.test(Mod6$residuals)
TukeyHSD(Mod6, which = "Station.x")
```
```{r}
CZ_wide$Station.x <- factor(CZ_wide$Station.x, levels = c("P22", "P8", "P12","P402","P28","P38","P4"))
ggplot(CZ_wide, aes(x=as.factor(Station.x), y=temp,color=Station.x,fill=Station.x)) + 
    geom_boxplot(alpha=0.2)+ylab("Temperature (°C)")+xlab("Station")+
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),panel.background = element_blank(), axis.line = element_line(colour = "black"),panel.border = element_rect(colour = "black", fill=NA))+
  scale_fill_manual(values = c("P22"="#F8766D", "P8"="#53B400", "P12"="#00B6EB","P402"="#A69AFF","P28"="#00C094","P38"="#FB61D7","P4"="#C49A00"))+
scale_color_manual(values = c("P22"="#F8766D", "P8"="#53B400", "P12"="#00B6EB","P402"="#A69AFF","P28"="#00C094","P38"="#FB61D7","P4"="#C49A00"))+
  scale_x_discrete(labels = c("P22"="Strait of Juan de Fuca", "P8"="Hood Canal Sill", "P12"="Hood Canal 1","P402"="Hood Canal 2","P28"="Main Basin","P38"="South Sound","P4"="Whidbey Basin"))+
  theme(axis.text.y=element_text(size=20,colour="black"),
        legend.position="none",
        axis.title.y=element_text(size=20,colour="black"),
        axis.line = element_line(colour = "black"),
        axis.title.x=element_blank(),
        axis.text.x = element_text(angle = 90, hjust = 1,size=15,colour="black"))+
  geom_vline(xintercept = 4.5, linetype = "dashed", color = "black")+geom_vline(xintercept = 6.5, linetype = "dashed", color = "black")
```

```{r}
ggplot(CZ_wide, aes(x=as.factor(Year.x), y=GvsC)) + 
    geom_boxplot(alpha=0.2)+ylab("gelatinous/crustacean")+xlab("Station")+
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),panel.background = element_blank(), axis.line = element_line(colour = "black"),panel.border = element_rect(colour = "black", fill=NA))+
    facet_wrap(~Station.x, scales = "free")
```


# Mollusc to crustacean ratio
```{r}
unique(Zoop_env$Crustaceans.vs.Gelatinous)
MC<-subset(Zoop_env, Zoop_env$Crustaceans.vs.Gelatinous %in% c("Crustaceans","Molluscs"))
```

remove ethanol samples
```{r}
MC<-MC[- grep("EtOH", MC$Sample.Code),]
```

```{r}
MC_wide<-dcast(MC, code+Month.x+Year.x+Date.Convert+Station.x+temp+pH+Oxygen ~ Crustaceans.vs.Gelatinous )
```

```{r}
MC_wide$MvsC<-MC_wide$Molluscs/MC_wide$Crustaceans
```

```{r}
ggplot(MC_wide, aes(x=MvsC)) + geom_histogram()
ggplot(MC_wide, aes(x=temp)) + geom_histogram()
ggplot(MC_wide, aes(x=log(MvsC))) + geom_histogram()
```

```{r}
ggplot(MC_wide, aes(x=pH, y=log(MvsC),color=Station.x))+ geom_point()+ 
  geom_smooth(method=lm, se=FALSE)
```

```{r}
ggplot(MC_wide, aes(x=as.factor(Month.x), y=log(MvsC))) + 
    geom_boxplot(alpha=0.2)+ylab("log(mollsc vs. crustacean)")+xlab("Station")+
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),panel.background = element_blank(), axis.line = element_line(colour = "black"),panel.border = element_rect(colour = "black", fill=NA))
```

```{r}
ggplot(MC_wide, aes(x=Date.Convert, y=temp,color=Station.x)) + 
    geom_point()+geom_line()+ylab("Surface temperature")+xlab("Year")+
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),panel.background = element_blank(), axis.line = element_line(colour = "black"),panel.border = element_rect(colour = "black", fill=NA))+
    facet_wrap(~Month.x, scales = "free")
```

```{r}
ggplot(MC_wide, aes(x=as.factor(Station.x), y=temp)) + 
    geom_boxplot(alpha=0.2)+ylab("temperature")+xlab("Station")+
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),panel.background = element_blank(), axis.line = element_line(colour = "black"),panel.border = element_rect(colour = "black", fill=NA))
```
```{r}
ggplot(MC_wide, aes(x=as.factor(Station.x), y=pH)) + 
    geom_boxplot(alpha=0.2)+ylab("pH")+xlab("Station")+
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),panel.background = element_blank(), axis.line = element_line(colour = "black"),panel.border = element_rect(colour = "black", fill=NA))+
    facet_wrap(~Month.x)
```
```{r}
ggplot(MC_wide, aes(x=as.factor(Station.x), y=Oxygen)) + 
    geom_boxplot(alpha=0.2)+ylab("Oxygen")+xlab("Station")+
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),panel.background = element_blank(), axis.line = element_line(colour = "black"),panel.border = element_rect(colour = "black", fill=NA))+
    facet_wrap(~Month.x)
```



```{r}
ggplot(MC_wide, aes(x=Date.Convert, y=Oxygen,color=Station.x)) + 
    geom_point()+geom_line()+ylab("Deep Oxygen")+xlab("Year")+
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),panel.background = element_blank(), axis.line = element_line(colour = "black"),panel.border = element_rect(colour = "black", fill=NA))+
    facet_wrap(~Month.x, scales = "free")
```

```{r}
ggplot(MC_wide, aes(x=Date.Convert, y=pH,color=Station.x)) + 
    geom_point()+geom_line()+ylab("Deep pH")+xlab("Year")+
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),panel.background = element_blank(), axis.line = element_line(colour = "black"),panel.border = element_rect(colour = "black", fill=NA))+
    facet_wrap(~Month.x, scales = "free")
```